### 2020-12-08 06_Problems/04_GroupsAgeDistribution/Propensity_score_matching.Rmd
Matching method implementation and regular meeting preparation

---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, include = F}
library(MatchIt)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(plyr)
library(cobalt)
library(broom)
library(kableExtra)
library(optmatch)
library(rlist)
library(lubridate)
```

```{r}
Sys.setlocale("LC_ALL", "C")
```

```{r}
wd = "/Users/mymac/Documents/research/CFI_Matching/"
setwd(wd)
```

```{r}
df = read.csv(paste0(wd,"/Propensity_score_matchingAllinOne_v2.csv"))
dim(df)
```

```{r}
df$LAD =rep(0, dim(df)[1])
df$LCX =rep(0, dim(df)[1])
df$RCA =rep(0, dim(df)[1])
for (i in 1:dim(df)[1]){
  if (df$Vessel[i] == 0){
    df$LAD[i] = 1
  }
  if (df$Vessel[i] == 1){
    df$LCX[i] = 1
  }
  if (df$Vessel[i] == 2){
    df$RCA[i] = 1
  }
}
df$LAD
df$LCX
df$RCA
```

```{r}
drops <- c("X","X.1", "X.2", "Sex", "Hypertension", "Vessel")
df = df[ , !(names(df) %in% drops)]
colnames(df)
```

```{r}
for (i in 1:dim(df)[1]){
  if (df$Group[i] == 0){
    df$Group[i] = 1
  }
  else {
    df$Group[i] = 0
  }

}
```

```{r, include = F}
df_cov = c(
  "CFI",
  "QCA",
  "Presence.of.AP",
  "LAD",
  "LCX",
  "RCA",
  "Age"
           )
length(df_cov)
```

```{r}
df$Group = as.numeric(df$Group)
```

```{r}
# df$Vessel <- as.factor(df$Vessel)
# df$Vessel
```

```{r, include = F}
sapply(df, class)
```

```{r}
sapply(df, summary)
```

```{r, include = F}
print("-----------------")
for (i in 1:length(df_cov)){
  print(df_cov[i])
  print(table(is.na(df[, df_cov[i]])))
  # for (j in 1:dim(df)[1]){
  #   if(is.na(df[j, df_cov[i]])){
  #     print(df[j, "center"])
  #   }
  #     else{
  #       next
  #     }
  # }
print("-----------------")
  }
```

```{r, include = F}
dim(df)
na_removed_from_df = function(covar) {
  na_var = numeric()
  for (i in 1:length(covar)){
    if ((df[, covar[i]]) %>% anyNA(recursive = FALSE) ==T){
      mess1 = paste0("NA were present in :", covar[i])
      print(mess1)
      na_var = c(na_var, which(is.na(df[,covar[i]]) == T))
    }
    else{
      next
    }
  }
  if(length(na_var) == 0){
    print("there was no NA found")
    df_new = df
  }
  else{
    mes = paste0("entries removed: ")
    mes2 = paste0(na_var, collapse = " ")
    mes_fin = paste0(mes, mes2)
    print(mes_fin)
    df_new = df[-na_var,]
  }
  return(df_new)
}
df_no_na = na_removed_from_df(df_cov)
dim(df_no_na)
```

```{r, include = F}
ttest = t.test(df_no_na$Age ~ df_no_na$Group)
# ttest = t.test(df$age_at_LAAC ~ df$device)
ttest
ttest$estimate[1]
ttest$estimate[2]
```

```{r}
# ttest_table_no_match = sapply(df_cov, function(v){
#   ttest = t.test(df_no_na[, v] ~ df_no_na$device_num)
#   p_value = round(ttest$p.value, digit =3)
#   estimate_asd_pfo = round(ttest$estimate[1], digit = 3)
#   estimate_punct = round(ttest$estimate[2], digit = 3)
#   ttest_table = c(p_value, estimate_asd_pfo, estimate_punct)
#   names(ttest_table) = c("p value", "estimate_asd_pfo", "estimate_punct")
#   return(ttest_table)
#   })
# ttest_table_no_match = t(as.data.frame(ttest_table_no_match))
# ttest_table_no_match = ttest_table_no_match[, c(2,3,1)]
# ttest_table_no_match
```

```{r, include = F}
# tab1 = kable(ttest_table_no_match, "latex", booktabs = T, col.names = c("estimate", "st. error", "p-value")) %>% 
# add_header_above(c("Unpaired t-test on variables on\namplatzer vs lambre - not matched" = 4)) %>%
# kable_styling(latex_options = c("striped", "hold_position"))
# kable_as_image(tab1, "/Users/mymac/Documents/research/Lambre_vs_Amplatzer/figures_and_tables/0_ttest_not_matched", file_format = "png")
```

```{r, include = F}
ps_df_not_matched = glm(f.build("Group", df_cov),
                      family = binomial(), data = df_no_na)
summa_ps_df_not_matched = summary(ps_df_not_matched)
df_not_matched_logit_coef = summa_ps_df_not_matched$coefficients
df_not_matched_logit_coef = round(df_not_matched_logit_coef, digit = 4)
df_not_matched_logit_coef = subset(df_not_matched_logit_coef, select = -c(3))
df_not_matched_logit_coef
```

```{r, include = F}
tab2 = kable(df_not_matched_logit_coef, "latex", booktabs = T, col.names = c("estimate", "st. error", "p-value")) %>% 
add_header_above(c("Logistic regression of variables on\namplatzer vs lambre - not matched" = 4)) %>%
# column_spec(1, width = "3cm") %>%
kable_styling(latex_options = c("striped", "hold_position"))
# kable_as_image(tab2, "/Users/mymac/Documents/research/Lambre_vs_Amplatzer/figures_and_tables/0_log_reg_not_matched", file_format = "png")
```

```{r, include = F}
prs_df = data.frame(pr_score = predict(ps_df_not_matched, type = "response"),
                    device = ps_df_not_matched$model$Group)
```

```{r, include = F}
labs = paste("actual procedure:", c("lambre", "amplatzer"))
prs_df %>%
  mutate(device = ifelse(device == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "white") +
  facet_wrap( ~ device) +
  xlab("Probability of Puncture") +
  theme_bw()
```

```{r}
df_mod_selec = data.frame(matrix(ncol = 4, nrow = 0))
col_names = c("model", "mean diff", "ratio", "caliper")
colnames(df_mod_selec) = col_names
df_mod_selec
```
```{r}
colnames(df_no_na)
```

```{r}
df_cov
```

```{r, echo = F}

#df_mod_selec = data.frame(matrix(ncol = 4, nrow = 0))
#col_names = c("model", "mean diff", "ratio", "caliper")
#colnames(df_mod_selec) = col_names
#df_mod_selec

col_names = c("model", "mean diff", "ratio", "caliper", "n_control_matched", "n_patients")
df_mod_selec = data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(df_mod_selec) = col_names
iter = 500
threshold = 0.60
ratio_list = c(1, 2, 3)
caliper_list = c(0.8, 0.6, 0.4, 0.2)
#ratio_list = c(2)
#caliper_list = c(0.4)
for (j in 1:length(ratio_list)){
  ratio_val = ratio_list[j]
  print(paste0("ratio: ", ratio_val))
  for (i in 1:length(caliper_list)){
      caliper_val = caliper_list[i]
      print(paste0("caliper: ", caliper_val))

    select_mod = data.frame(matrix(ncol = length(df_cov) + 2, nrow = iter))
    colnames(select_mod) = c(df_cov, "mean", "seed")
    dim(select_mod)
    for (i in 1:iter){
      set.seed(i)
      mod_match_logit = matchit(f.build("Group", df_cov), method = "nearest", distance = "logit", data = df_no_na, caliper = caliper_val, ratio = ratio_val)
      tabl = bal.tab(mod_match_logit)
      df_tbl_match_logit = data.frame(tabl$Balance$Diff.Adj)
      rownames(df_tbl_match_logit) = rownames(tabl$Balance)
      df_tbl_match_logit = rbind(df_tbl_match_logit, "Mean" = mean(abs(tabl$Balance$Diff.Adj)))
      df_tbl_match_logit = rbind(df_tbl_match_logit, "seed" = i)
      for (j in 2:dim(df_tbl_match_logit)[1]){
          select_mod[i,j-1] = df_tbl_match_logit[j,1]
      }
    }
    # create a new df without variable "mean" and "seed" for the function all()
    select_mod_red = select_mod[, -which(names(select_mod) %in% c("mean", "seed"))]

    # select model with all standardized difference < threshold
    select_index = c()
    for (i in 1:dim(select_mod_red)[1]){
      if(all(select_mod_red[i,] < threshold) == T){
        select_index = c(select_index, i)
      }
      else{
        next
      }
    }

    select_mod_ind = select_mod[select_index, ]

    golden_seed = select_mod_ind$seed[which(select_mod_ind$mean == min(select_mod_ind$mean))]
    set.seed(golden_seed)
    mod_match_logit = matchit(f.build("Group", df_cov), method = "nearest", distance = "logit", data = df_no_na, caliper = caliper_val, ratio = ratio_val, replace = FALSE)
    tabl = bal.tab(mod_match_logit)

    mean_diff = mean(abs(tabl$Balance$Diff.Adj))

    model_str = paste0("ratio: ", ratio_val, " - caliper ", caliper_val, ":1")

     df_mod_added = data.frame(model_str, mean_diff, ratio_val, caliper_val,  tabl$Observations$Treated[2], tabl$Observations$Control[2])
    #df_mod_added = data.frame(model_str, mean_diff, ratio_val, caliper_val)
    df_mod_selec = rbind(df_mod_selec, df_mod_added)
    df_mod_selec

    #df_observation = data.frame(rownames(tabl$Balance), tabl$Balance[,4])
    #colnames(df_balance) = c("variable", "Standardized difference")
    write.csv(tabl$Observations, paste0(wd, "Observations/observations_ratio_", ratio_val, "_caliper_", caliper_val,  ".csv"))

    df_balance = data.frame(rownames(tabl$Balance), tabl$Balance[,4])
    colnames(df_balance) = c("variable", "Standardized difference")
    write.csv(df_balance, paste0(wd, "balance/balance_ratio_", ratio_val, "_caliper_", caliper_val,  ".csv"), row.names = FALSE)

    love1 = love.plot(bal.tab(mod_match_logit), threshold = .1)
    ggsave(love1, file = paste0("love_ratio_", ratio_val, "_caliper_", caliper_val, ".png"), path = paste0(wd, "/figures_and_tables"))

    df_m_logit = match.data(mod_match_logit)
    for (i in 1:dim(df_m_logit)[1]){
      if (df_m_logit$Group[i] == 0){
        df_m_logit$Group[i] = 1
      }
      else {
        df_m_logit$Group[i] = 0
      }
    }

    ps_df_match_logit = glm(f.build("Group", df_cov), family = binomial(), data = df_m_logit)
    summa_ps_df_match_logit = summary(ps_df_match_logit)
    df_m_logit_coef = summa_ps_df_match_logit$coefficients
    df_m_logit_coef = round(df_m_logit_coef, digit = 4)
    df_m_logit_coef = subset(df_m_logit_coef, select = -c(3))
    df_m_logit_coef
    write.csv(df_m_logit_coef, paste0(wd, "lr_baseline_characteristcs_post_matching/logistic_regression_ratio_", ratio_val, "_caliper_", caliper_val, ".csv"))

    df_m_logit$Vessel =rep(0, dim(df_m_logit)[1])
    for (i in 1:dim(df_m_logit)[1]){
      if (df_m_logit$LAD[i] == 1){
        df_m_logit$Vessel[i] = 0
      }
      if (df_m_logit$LCX[i] == 1){
        df_m_logit$Vessel[i] = 1
      }
      if (df_m_logit$RCA[i] == 1){
        df_m_logit$Vessel[i] = 2
      }
    }
    df_m_logit$LAD <- NULL
    df_m_logit$LCX <- NULL
    df_m_logit$RCA <- NULL
    #df_m_logit$distance <- NULL
    df_m_logit$weights <- NULL

    write.csv(df_m_logit, paste0(wd, "matched_data/matched_data_ratio_", ratio_val, "_caliper_", caliper_val, ".csv"))
  }
}
colnames(df_mod_selec) = col_names
write.csv(df_mod_selec, paste0(wd, "/df_mod_selec.csv"))
```

### 2020-12-08 03_Scripts/09_GroupMatching.py

@@ -1,0 +1,179 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+
+
+# 01 Define covariance functions (NB: called variance if x = y)
+def Covariance(x,y):
+
+    if not len(x)==len(y):
+        print('Covariables vectors not of equal lengths')
+    else:
+        x_bar, y_bar = x.mean(), y.mean()
+        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
+        return Cov
+def CovarianceMatrix(Data, Covariables):
+    C = np.zeros((len(Covariables), len(Covariables)))
+    for i in range(len(Covariables)):
+        for j in range(len(Covariables)):
+            x = Data[Covariables[i]].values
+            y = Data[Covariables[j]].values
+            C[i, j] = Covariance(x, y)
+
+    return C
+def MahalanobisDistance(x,y,C):
+
+    XY = x-y
+    C_inv = np.linalg.inv(C)
+    Distance = np.sqrt(XY.T.dot(C_inv.dot(XY)))
+
+    return Distance
+def ComputeNewDistances(DistancesData, SortedIndices, Matches, ControlIndividual):
+
+    TestIndividuals = np.where(Matches == ControlIndividual)[0]
+    print('Match with test individuals n:')
+    print(TestIndividuals)
+
+    print('\nCompute new distances...\n')
+    DistancesIncrease = np.zeros(len(TestIndividuals))
+    NewControlIndividuals = np.zeros(len(TestIndividuals))
+    Index = 0
+    for TestIndividual in TestIndividuals:
+        print('Control individual n: %i' % ControlIndividual)
+        print('Test individual n: %i' % TestIndividual)
+
+        ControlIndividualIndex = np.where(SortedIndices[:, TestIndividual] == ControlIndividual)[0]
+        NewControlIndividual = SortedIndices[ControlIndividualIndex + 1, TestIndividual]
+        NewControlIndividuals[Index] = NewControlIndividual
+
+        print('New control individual n: %i' % NewControlIndividual)
+
+        Distance = DistancesData[TestIndividual].loc[ControlIndividual]
[…]

### 2021-01-11 03_Scripts/11_MatchedGroupFit.py
End of Holidays. Back to work

@@ -664,24 +664,25 @@ def DeterminationCoefficients(XData, YData, NParameters, NSamples):
 
 # 01 Load Data
 WorkingDirectory = os.getcwd()
-DataFolder = os.path.join(WorkingDirectory,'02_Data')
-MatchFolder = os.path.join(WorkingDirectory,'04_Results/00_Matching/Matches')
-TensorFolder = os.path.join(WorkingDirectory,'04_Results/')
-ResultFolder = os.path.join(WorkingDirectory,'04_Results/05_MatchedGroups')
+DataFolder = os.path.join(WorkingDirectory, '02_Data')
+MatchFolder = os.path.join(WorkingDirectory, '04_Results/00_Matching/Matches')
+TensorFolder = os.path.join(WorkingDirectory, '04_Results/')
+ResultFolder = os.path.join(WorkingDirectory, '04_Results/05_MatchedGroups')
 
 ## Load groups data
 ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('Scans.csv')]
 ScanLists.sort()
 
-HealthyGroup = pd.read_csv(os.path.join(DataFolder,ScanLists[0]))
-OIGroup = pd.read_csv(os.path.join(DataFolder,ScanLists[1]))
+HealthyGroup = pd.read_csv(os.path.join(DataFolder, ScanLists[0]))
+OIGroup = pd.read_csv(os.path.join(DataFolder, ScanLists[1]))
 
-OIGroup = OIGroup.dropna()     # Filter patients without data
+OIGroup = OIGroup.dropna()  # Filter patients without data
 OIGroup = OIGroup.reset_index()
-OIGroup = OIGroup[['Scan_ID_BSL','Sex','Age','OI_Type']]
-OIGroup.columns = ['Sample Number','Sex','Age','OI Type']
-OIGroup['Sex'] = OIGroup['Sex'].replace(['Female','Male'],['F','M'])
-OIGroup['OI Type'] = OIGroup['OI Type'].replace(['Type I','Type III','Type IV'],[1,3,4])
+OIGroup['Sample Number'] = OIGroup['Patient_ID'] + '_' + OIGroup['Scan_ID_BSL']
+OIGroup = OIGroup[['Sample Number', 'Sex', 'Age', 'OI_Type']]
+OIGroup.columns = ['Sample Number', 'Sex', 'Age', 'OI Type']
[…]

### 2021-01-11 03_Scripts/13_MatchedGroupStatistics.py

@@ -124,121 +124,224 @@ def QQPlot(DataValues, Alpha_CI, DataLabel='Data'):
     plt.close(Figure)
 
     return Variance
+def Covariance(x,y):
 
+    if not len(x)==len(y):
+        print('Covariables vectors not of equal lengths')
+    else:
+        x_bar, y_bar = x.mean(), y.mean()
+        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
+        return Cov
+def ShapiroWilkTest(DataValues):
+
+    print('!! Cross check results with scipy.stats.shapiro() !!')
+
+    # Data analysis
+    N = len(DataValues)
+    X_Bar = np.mean(DataValues)
+    S_X = np.std(DataValues,ddof=1)
+
+    if N < 3:
+        print('Sample vector must have at least 3 valid observations.')
+        return
+
+    elif N == 3:
+        w = np.array([np.sqrt(0.5),0,-np.sqrt(0.5)])
+
+    else:
+
+        if N > 5000:
+        print('Shapiro-Wilk statistic might be inaccurate due to large sample size ( > 5000).')
+
+        # Sort data to get the rank
+        Data_Sorted = np.zeros(N)
+        Data_Sorted += DataValues
+        Data_Sorted.sort()
+
+        # Expected values
+        m = NormalQuantile((np.arange(1,N+1) - 3/8) / (N + 0.25),0,1)
+
+        # Compute coefficients according to: https://math.stackexchange.com/questions/625359/how-are-the-values-in-the-shapiro-wilk-weight-table-calculated
+        C_N = m[-1]/np.sqrt(np.sum(m**2))
+        C_N1 = m[-2]/np.sqrt(np.sum(m**2))
+
+        # Polynomial vectors
+        p1 = np.array([-2.706056,4.434685,-2.071190,-0.147981,0.221157,C_N])
+        p2 = np.array([-3.582633,5.682633,-1.752461,-0.293762,0.042981,C_N1])
+
+        # Compute weights
+        u = 1/np.sqrt(N)
+        w = np.zeros(N)
+
+        for k in range(len(p1)):
+            w[-1] += p1[k] * u ** (5 - k)
+        w[0] = -w[-1]
+
+        if N >= 6:
+
[…]

### 2021-01-11 03_Scripts/14_TestsStatistics.py

@@ -1,0 +1,207 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+from scipy.stats import linregress
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width, suppress=True, formatter={'float_kind': '{:f}'.format})
+plt.rc('font', size=12)
+
+
+## Define functions
+def TwoSamples_tTest(x,y, SignificanceLevel=0.05):
+
+    # Analyze data
+    n = len(x)
+    m = len(y)
+    s_x = np.std(x,ddof=1)
+    s_y = np.std(y,ddof=1)
+    x_bar = np.mean(x)
+    y_bar = np.mean(y)
+
+    # Perform test statistic
+    DOFs = n+m-2
+    S_pool = np.sqrt(1/DOFs * ( (n-1)*s_x**2 + (m-1)*s_y**2 ))
+    T = (x_bar - y_bar) / (S_pool * np.sqrt(1/n + 1/m))
+
+    # Compute p value
+    from scipy.stats.distributions import t
+    if T >= 0:
+        p = 2 * (1-t.cdf(T,DOFs))
+    else:
+        p = 2 * t.cdf(T, DOFs)
+
+    # Compute confidence interval CI
+    T_Interval = np.array(t.interval(1-SignificanceLevel/2,DOFs))
+
+    # Compute CI for difference in means
+    MeansInterval =  (x_bar-y_bar) + T_Interval * S_pool * np.sqrt(1/n + 1/m)
+
+    return T, p, T_Interval, MeansInterval
+def PermutationTest(x,y,NRepetition=len(x)*len(y),SignificanceLevel=0.05):
+
+    # Analyze data
+    x_bar = np.mean(x)
+    y_bar = np.mean(y)
+
+    d = x_bar - y_bar
+
+    XData = pd.DataFrame({'Values':x,'Group':'Control'},index=range(len(x)))
+    YData = pd.DataFrame({'Values':y,'Group':'Test'},index=range(len(y)))
+
+    Pool = XData.append(YData,ignore_index=True)
+    N = len(Pool)
+
+    D = np.zeros(NRepetition)
+    for i in range(NRepetition):
+
+        n = np.random.randint(1,N-1)
+
+        SampleA = Pool.sample(n)
[…]

### 2021-01-11 03_Scripts/13_MatchedGroupStatistics.py
Implement statistical tests ad perform them for the different variables

@@ -119,7 +119,7 @@ def QQPlot(DataValues, Alpha_CI, DataLabel='Data'):
     plt.xlabel('Theoretical quantiles (-)')
     plt.ylabel('Empirical quantiles (-)')
     plt.ylim([Y_Min, Y_Max])
-    plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.15))
+    plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.15), prop={'size':10})
     plt.show()
     plt.close(Figure)
 
@@ -355,7 +355,7 @@ Axes.plot(SortedValues,KernelEstimator,color=(1,0,0),label='Kernel Density')
 Axes.plot(SortedValues,TheoreticalDistribution,linestyle='--',color=(0,0,0),label='Normal Distribution')
 plt.xlabel(Label)
 plt.ylabel('Density (-)')
-plt.legend(loc='upper center',ncol=3,bbox_to_anchor=(0.5,1.15))
+plt.legend(loc='upper center',ncol=3,bbox_to_anchor=(0.5,1.15), prop={'size':10})
 plt.show()
 plt.close(Figure)
 
@@ -369,7 +369,7 @@ Axes.plot(SortedValues,EmpiricalQuantiles, linestyle='none', marker='o', mew=0.5
 Axes.plot(SortedValues,TheoreticalQuantiles, linestyle='--', color=(1, 0, 0), label='Normal Distribution')
 plt.xlabel(Label)
 plt.ylabel('Quantile (-)')
-plt.legend(loc='upper center',ncol=2,bbox_to_anchor=(0.5,1.15))
+plt.legend(loc='upper center',ncol=2,bbox_to_anchor=(0.5,1.15), prop={'size':10})
 plt.show()
 plt.close(Figure)
 

### 2021-01-28 03_Scripts/13_MatchedGroupStatistics.py
Statistics: implementation of linear regression models and mixed-effect linear models

@@ -70,7 +70,7 @@ def NormalQuantile(p,u=0,s=1):
     z_p = u + s * np.sqrt(2) * InverseErrorFunction(2*p-1)
 
     return z_p
-def QQPlot(DataValues, Alpha_CI, DataLabel='Data'):
+def QQPlot(DataValues, Alpha_CI=0.95, DataLabel='Data'):
 
     ### Based on: https://www.tjmahr.com/quantile-quantile-plots-from-scratch/
     ### Itself based on Fox book: Fox, J. (2015)
@@ -112,7 +112,7 @@ def QQPlot(DataValues, Alpha_CI, DataLabel='Data'):
     Y_Min = Data_Sorted.min() - BorderSpace
     Y_Max = Data_Sorted.max() + BorderSpace
     Figure, Axes = plt.subplots(1, 1, figsize=(5.5, 4.5), dpi=100)
-    Axes.plot(TheoreticalQuantiles, Data_Sorted, linestyle='none', marker='o', mew=0.5, fillstyle='none', color=(0, 0, 0), label=Label)
+    Axes.plot(TheoreticalQuantiles, Data_Sorted, linestyle='none', marker='o', mew=0.5, fillstyle='none', color=(0, 0, 0), label=DataLabel)
     Axes.plot(Data_Space, Variance_Line, linestyle='--', color=(1, 0, 0), label='Variance :' + str(format(np.round(Variance, 2),'.2f')))
     Axes.plot(Data_Space, Variance_Line + Z_CI_Quantile * Data_SE, linestyle='--', color=(0, 0, 1), label=str(int(100*Alpha_CI)) + '% CI')
     Axes.plot(Data_Space, Variance_Line - Z_CI_Quantile * Data_SE, linestyle='--', color=(0, 0, 1))

### 2021-01-28 03_Scripts/16_Mixed_Models.py

@@ -1,0 +1,411 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.api as sm
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import t
+from scipy.stats.distributions import norm
+
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+## Define functions
+def QQPlot(DataValues, Alpha_CI=0.95, DataLabel='Data'):
+
+    ### Based on: https://www.tjmahr.com/quantile-quantile-plots-from-scratch/
+    ### Itself based on Fox book: Fox, J. (2015)
+    ### Applied Regression Analysis and Generalized Linear Models.
+    ### Sage Publications, Thousand Oaks, California.
+
+    # Data analysis
+    N = len(DataValues)
+    X_Bar = np.mean(DataValues)
+    S_X = np.std(DataValues)
+
+    # Sort data to get the rank
+    Data_Sorted = np.zeros(N)
+    Data_Sorted += DataValues
+    Data_Sorted.sort()
+
+    # Compute quantiles
+    EmpiricalQuantiles = np.arange(0.5, N + 0.5) / N
+    TheoreticalQuantiles = norm.ppf(EmpiricalQuantiles, X_Bar, S_X)
+    ZQuantiles = norm.ppf(EmpiricalQuantiles,0,1)
+
+    # Compute data variance
+    DataIQR = np.quantile(DataValues, 0.75) - np.quantile(DataValues, 0.25)
+    NormalIQR = np.sum(np.abs(norm.cdf(np.array([0.25, 0.75]), 0, 1)))
+    Variance = DataIQR / NormalIQR
+    Z_Space = np.linspace(min(ZQuantiles), max(ZQuantiles), 100)
+    Variance_Line = Z_Space * Variance + np.median(DataValues)
+
[…]

### 2021-01-28 03_Scripts/GetSValues.py

@@ -1,0 +1,275 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+plt.rc('font', size=12)
+
+
+def DyadicProduct(A,B):
+
+    if A.size == 3:
+
+        if isinstance(A[0], np.float):
+            C = np.zeros((3,3))
+        else:
+            C = sp.zeros(3)
+
+        for i in range(3):
+            for j in range(3):
+                C[i,j] = A[i]*B[j]
+
+        if not isinstance(A[0], np.float):
+            C = np.array(C)
+
+    elif A.size == 9:
+
+        if isinstance(A[0,0], np.float):
+            C = np.zeros((3,3,3,3))
+        else:
+            C = sp.zeros(9)
+
+        for i in range(3):
+            for j in range(3):
+                for k in range(3):
+                    for l in range(3):
+                        if isinstance(A[0, 0], np.float):
+                            C[i,j,k,l] = A[i,j] * B[k,l]
+                        else:
+                            C[3*i+j,3*k+l] = A[i,j] * B[k,l]
+
+        if not isinstance(A[0,0], np.float):
+            C = IsoMorphism99_3333(C)
+
+    else:
+        print('Matrices sizes mismatch')
+
+    return C
+def SymmetricProduct(A,B):
+
+    if isinstance(A[0, 0], np.float):
+        C = np.zeros((3, 3, 3, 3))
+    else:
+        C = sp.zeros(9)
+
+    for i in range(3):
+        for j in range(3):
+            for k in range(3):
+                for l in range(3):
+                    if isinstance(A[0, 0], np.float):
[…]

### 2021-01-28 03_Scripts/RCrosscheck.R

@@ -1,0 +1,77 @@
+# Title     : R crosscheck
+# Objective : Double check ANCOVA and mixed linear regression computations
+# Created by: mathieu
+# Created on: 21/01/2021
+
+# Load library and data
+library(carData)
+library(effects)
+library(car)
+Data <- read.table("/home/mathieu/Documents/Post-Msc/04_Results/06_ANCOVA/RData.csv", header=TRUE, sep=",")
+
+# Check data
+dim(Data)
+Data[sort(sample(5220, 12)),]
+
+# Build linear regression models for ANCOVA
+Model1.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy) - 1, data = Data)
+summary(Model1.mod)
+
+Model2.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
+                      + (Sii + Sij + Sjj):Sg - 1, data = Data)
+summary(Model2.mod)
+
+Model3.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
+                      + (Sii + Sij + Sjj):log(CV) - 1, data = Data)
+summary(Model3.mod)
+
+anova(Model1.mod,Model2.mod)
+anova(Model1.mod,Model3.mod)
+
+Model4.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
+                      + (Sii + Sij + Sjj):Sg
+                      + (Sii + Sij + Sjj):log(CV) - 1, data = Data)
+summary(Model4.mod)
+
+anova(Model2.mod,Model4.mod)
+anova(Model3.mod,Model4.mod)
+
+Model5.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
+                      + (Sii + Sij + Sjj):Sg
+                      + (Sii + Sij + Sjj):log(CV)
+                      + Sg:log(CV)- 1, data = Data)
+summary(Model5.mod)
+
+anova(Model4.mod,Model5.mod)
+
+# Check normality of residuals
+qqPlot(Model1.mod)
+
+# Perform ANCOVA
+anova(Model4.mod,Model3.mod)
+anova(Model4.mod,Model5.mod)
[…]

### 2021-02-02 03_Scripts/16_Mixed_Models.py
Presentation of mixed models results and beginning of ROI morphometry analysis

@@ -321,7 +321,7 @@ def PlotRegressionResults(Data,Y_Obs, Y_Fit, SE, R2, X, C_x, Alpha=0.95):
 
 # 01 Load Data
 WorkingDirectory = os.getcwd()
-ResultFolder = os.path.join(WorkingDirectory, '04_Results/06_ANCOVA')
+ResultFolder = os.path.join(WorkingDirectory, '04_Results/06_LinearRegression')
 Data = pd.read_csv(ResultFolder + '/RegressionData.csv')
 
 # 02 Set the linear regression parameters (log transformation)
@@ -340,8 +340,7 @@ MinBVTV = 0.1/3*2
 # 04 Filter data
 CVFilter = Data['CV'] < Threshold
 Windowing = Data['BVTV'] > MinBVTV
-FilteredData = Data[Windowing]
-FilteredData = Data
+FilteredData = Data[Windowing&CVFilter]
 
 
 # 05 Build and fit the model

### 2021-02-02 03_Scripts/17_RegressionResults_Comparison.py

@@ -1,0 +1,257 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib
+# matplotlib.use('WebAgg')
+import matplotlib.pyplot as plt
+
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+# Define functions
+def PlotValues(XYZ, Radii, NPoints=32):
+
+    ## New coordinate system
+    Q = np.eye(3)
+
+    ## Build data for fabric plotting
+    u = np.arange(0, 2 * np.pi + 2 * np.pi / NPoints, 2 * np.pi / NPoints)
+    v = np.arange(0, np.pi + np.pi / NPoints, np.pi / NPoints)
+    X1 = Radii[0][0] * np.outer(np.cos(u), np.sin(v)) + XYZ[0][0]
+    Y1 = Radii[0][1] * np.outer(np.sin(u), np.sin(v)) + XYZ[0][1]
+    Z1 = Radii[0][2] * np.outer(np.ones_like(u), np.cos(v)) + XYZ[0][2]
+    X2 = Radii[1][0] * np.outer(np.cos(u), np.sin(v)) + XYZ[1][0]
+    Y2 = Radii[1][1] * np.outer(np.sin(u), np.sin(v)) + XYZ[1][1]
+    Z2 = Radii[1][2] * np.outer(np.ones_like(u), np.cos(v)) + XYZ[1][2]
+    X3 = Radii[2][0] * np.outer(np.cos(u), np.sin(v)) + XYZ[2][0]
+    Y3 = Radii[2][1] * np.outer(np.sin(u), np.sin(v)) + XYZ[2][1]
+    Z3 = Radii[2][2] * np.outer(np.ones_like(u), np.cos(v)) + XYZ[2][2]
+
+    for i in range(len(X1)):
+        for j in range(len(X1)):
+            [X1[i, j], Y1[i, j], Z1[i, j]] = np.dot([X1[i, j], Y1[i, j], Z1[i, j]], Q)
+            [X2[i, j], Y2[i, j], Z2[i, j]] = np.dot([X2[i, j], Y2[i, j], Z2[i, j]], Q)
[…]

### 2021-02-02 03_Scripts/18_ROIMorphometry.py

@@ -1,0 +1,54 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+# 01 Define functions
+def Load_Itk(Filename):
+
+    # Reads the image using SimpleITK
+    Itkimage = sitk.ReadImage(Filename)
+
+    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
+    CT_Scan = sitk.GetArrayFromImage(Itkimage)
+
+    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
+    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
+
+    # Read the spacing along each dimension
+    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
+
+    # Read the dimension of the CT scan (in voxel)
+    Size = np.array(list(reversed(Itkimage.GetSize())))
+
+    return CT_Scan, Origin, Spacing, Size
+
+# 02 Set variables
+Path = '/home/mathieu/Documents/MscThesis/02_Data/FabricElasticityRelationship/1_ControlGroupNodaratis/B_Edit/04_ResizedROI/'
+File = '1C0001679_SEG_UNCOMPResized.mhd'
+
+# 03 Get matched individuals
+
+
+
+CT_Scan, Origin, Spacing, Size = Load_Itk(Path+File)
+np.unique(CT_Scan)
+
+Threshold = 50
+CT_Scan[CT_Scan < Threshold] = 1
+CT_Scan[CT_Scan > Threshold] = 2
+
+Image = sitk.GetImageFromArray(CT_Scan)
+sitk.WriteImage(Image,'Image.mhd')
+
+
+

### 2021-02-02 03_Scripts/RCrosscheck.R

@@ -7,7 +7,7 @@
 library(carData)
 library(effects)
 library(car)
-Data <- read.table("/home/mathieu/Documents/Post-Msc/04_Results/06_ANCOVA/RData.csv", header=TRUE, sep=",")
+Data <- read.table("/home/mathieu/Documents/Post-Msc/04_Results/06_LinearRegression/RData.csv", header=TRUE, sep=",")
 
 # Check data
 dim(Data)

### 2021-02-05 03_Scripts/00_ROI_Selection.py
Re-sort folder and re-run ROI analysis

@@ -1,0 +1,230 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+# 01 Define functions
+def Load_Itk(Filename):
+
+    # Reads the image using SimpleITK
+    Itkimage = sitk.ReadImage(Filename)
+
+    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
+    CT_Scan = sitk.GetArrayFromImage(Itkimage)
+
+    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
+    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
+
+    # Read the spacing along each dimension
+    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
+
+    # Read the dimension of the CT scan (in voxel)
+    Size = np.array(list(reversed(Itkimage.GetSize())))
+
+    return CT_Scan, Origin, Spacing, Size
+def SelectROI(Stack, VoxelsRange, VoxelMargin, ROINumber):
+
+    HalfStackHeight = int(Stack.shape[0] / 2)
+
+    ZMin, YMin, XMin = np.ceil(VoxelsRange / 2).astype('int') + VoxelMargin
+    ZMax, YMax, XMax = Stack.shape - np.ceil(VoxelsRange / 2).astype('int') - VoxelMargin
+
+    if ROINumber < 3:
+        Z, Y, X = np.random.randint([ZMin, YMin, XMin], [HalfStackHeight, YMax, XMax])
+    else:
+        Z, Y, X = np.random.randint([HalfStackHeight, YMin, XMin], [ZMax, YMax, XMax])
+
+    Z1, Y1, X1 = [Z, Y, X] - np.round(VoxelsRange / 2).astype('int')
+    Z2, Y2, X2 = [Z1, Y1, X1] + VoxelsRange
+
+    ROI = [Z, Y, X, Z1, Y1, X1, Z2, Y2, X2]
+
+    return ROI
+def ROISelection(Stack, VoxelsRange, VoxelMargin, ROINumber, TrabBone, CortBone):
+
[…]

### 2021-02-05 03_Scripts/01_Matching.py

@@ -1,0 +1,191 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+
+
+# 01 Define covariance functions (NB: called variance if x = y)
+def Covariance(x,y):
+
+    if not len(x)==len(y):
+        print('Covariables vectors not of equal lengths')
+    else:
+        x_bar, y_bar = x.mean(), y.mean()
+        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
+        return Cov
+def CovarianceMatrix(Data, Covariables):
+    C = np.zeros((len(Covariables), len(Covariables)))
+    for i in range(len(Covariables)):
+        for j in range(len(Covariables)):
+            x = Data[Covariables[i]].values
+            y = Data[Covariables[j]].values
+            C[i, j] = Covariance(x, y)
+
+    return C
+def MahalanobisDistance(x,y,C):
+
+    XY = x-y
+    C_inv = np.linalg.inv(C)
+    Distance = np.sqrt(XY.T.dot(C_inv.dot(XY)))
+
+    return Distance
+def ComputeNewDistances(DistancesData, SortedIndices, Matches, ControlIndividual):
+
+    TestIndividuals = np.where(Matches == ControlIndividual)[0]
+    print('Match with test individuals n:')
+    print(TestIndividuals)
+
+    print('\nCompute new distances...\n')
+    DistancesIncrease = np.zeros(len(TestIndividuals))
+    NewControlIndividuals = np.zeros(len(TestIndividuals))
+    Index = 0
+    for TestIndividual in TestIndividuals:
+        print('Control individual n: %i' % ControlIndividual)
+        print('Test individual n: %i' % TestIndividual)
+
+        ControlIndividualIndex = np.where(SortedIndices[:, TestIndividual] == ControlIndividual)[0]
+        NewControlIndividual = SortedIndices[ControlIndividualIndex + 1, TestIndividual]
+        NewControlIndividuals[Index] = NewControlIndividual
+
+        print('New control individual n: %i' % NewControlIndividual)
+
+        Distance = DistancesData[TestIndividual].loc[ControlIndividual]
[…]

### 2021-02-05 03_Scripts/01_ROISelection.py

@@ -1,220 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import SimpleITK as sitk
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-# 01 Define functions
-def Load_Itk(Filename):
-
-    # Reads the image using SimpleITK
-    Itkimage = sitk.ReadImage(Filename)
-
-    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
-    CT_Scan = sitk.GetArrayFromImage(Itkimage)
-
-    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
-    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
-
-    # Read the spacing along each dimension
-    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
-
-    # Read the dimension of the CT scan (in voxel)
-    Size = np.array(list(reversed(Itkimage.GetSize())))
-
-    return CT_Scan, Origin, Spacing, Size
-def SelectROI(Stack, VoxelsRange, VoxelMargin, ROINumber):
-
-    HalfStackHeight = int(Stack.shape[0] / 2)
-
-    ZMin, YMin, XMin = np.ceil(VoxelsRange / 2).astype('int') + VoxelMargin
-    ZMax, YMax, XMax = Stack.shape - np.ceil(VoxelsRange / 2).astype('int') - VoxelMargin
-
-    if ROINumber < 3:
-        Z, Y, X = np.random.randint([ZMin, YMin, XMin], [HalfStackHeight, YMax, XMax])
-    else:
-        Z, Y, X = np.random.randint([HalfStackHeight, YMin, XMin], [ZMax, YMax, XMax])
-
-    Z1, Y1, X1 = [Z, Y, X] - np.round(VoxelsRange / 2).astype('int')
-    Z2, Y2, X2 = [Z1, Y1, X1] + VoxelsRange
-
-    ROI = [Z, Y, X, Z1, Y1, X1, Z2, Y2, X2]
-
-    return ROI
-def ROISelection(Stack, VoxelsRange, VoxelMargin, ROINumber, TrabBone, CortBone):
-
[…]

### 2021-02-05 03_Scripts/02_EvaluateMatching.py

@@ -1,0 +1,163 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+
+
+# 01 Set paths
+WorkingDirectory = os.getcwd()
+DataFolder = os.path.join(WorkingDirectory,'02_Data')
+MatchingFolder = os.path.join(WorkingDirectory,'04_Results/01_Matching')
+MatchesFolder = os.path.join(MatchingFolder,'Matches')
+ResultsFolder = os.path.join(MatchingFolder,'Matching_Assessment')
+
+
+# 02 Load Data
+Matched_OI = pd.read_csv(os.path.join(MatchingFolder,'Matched_OI.csv'))
+Matched_Healthy = pd.read_csv(os.path.join(MatchingFolder,'Matched_Control.csv'))
+
+ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('Scans.csv')]
+ScanLists.sort()
+HealthyData = pd.read_csv(os.path.join(DataFolder,ScanLists[0]))
+OIData = pd.read_csv(os.path.join(DataFolder,ScanLists[1]))
+
+## Adapt labeling and values type
+OIData = OIData.dropna()     # Filter patients without data
+OIData = OIData.reset_index()
+OIData['Sex'] = OIData['Sex'].replace(['Female','Male'],['F','M'])
+OIData = OIData[['Scan_ID_BSL','Sex','Age']]
+OIData.columns = ['Sample Number','Sex','Age']
+OIData['Age'] = OIData['Age'].astype('int')
+
+## Binarize categorical values
+OIData['Sex'] = OIData['Sex'].replace(['F','M'],[1,0])
+HealthyData['Sex'] = HealthyData['Sex'].replace(['F','M'],[1,0])
+Matched_OI['Sex'] = Matched_OI['Sex'].replace(['F','M'],[1,0])
+Matched_Healthy['Sex'] = Matched_Healthy['Sex'].replace(['F','M'],[1,0])
+
+
+# 03 plot matching algorithm results
+Matches = [File for File in os.listdir(MatchesFolder) if File.endswith('.txt')]
+Matches.sort()
+for Match in Matches:
+
[…]

### 2021-02-05 03_Scripts/02_WriteROIAnalysisBatch.py

@@ -1,44 +1,0 @@
-import os
-import numpy as np
-import pandas as pd
-
-WorkingDirectory = os.getcwd()
-DataFolder = os.path.join(WorkingDirectory,'02_Data')
-
-DataSubFolders = [File for File in os.listdir(DataFolder) if os.path.isdir(os.path.join(DataFolder,File))]
-DataSubFolders.sort()
-
-DataGroup = 1   # 0 = Healthy group, 1 = OI group
-
-PipelineFile  = open(os.path.join(WorkingDirectory, '03_Scripts', '02_ROIAnalysisPipeline.txt'), 'r')
-PipelineText  = PipelineFile.read()
-
-# List samples
-MedtoolParametersFile = os.path.join(DataFolder, DataSubFolders[DataGroup][:-14] + 'ROI.csv')
-MedtoolParametersData = pd.read_csv(MedtoolParametersFile, sep=';')
-del MedtoolParametersData['Unnamed: 8']
-
-# Create the .bash script
-BatchFile = open(os.path.join(WorkingDirectory, '03_Scripts', '02_ROIAnalysisBatch.bash'),'w')
-
-# Set variables for files to remove
-LocalPath = 'C_FEA/03_JobResults/'
-FilesDict = {'*.com','*.dat','*.msg','*.par','*.pes','*.pmg','*.prt','*.sim','*.sta','*.odb'}
-
-# Write the analysis batch
-for Index in MedtoolParametersData.index:
-
-    ROIText = PipelineText
-
-    for Parameter in MedtoolParametersData.columns:
-        ROIParameter = MedtoolParametersData[Parameter].loc[Index]
-        ROIText = ROIText.replace(Parameter[1:],str(ROIParameter))
-
-    BatchFile.write(ROIText+'\n\n')
-
-    for FileType in FilesDict:
-        BatchFile.write('rm ' + LocalPath + str(FileType) + '\n')
-
-    BatchFile.write('\n\n\n\n\n')
-
-BatchFile.close()

### 2021-02-05 03_Scripts/03_TransformTensor.py

@@ -1,733 +1,0 @@
-import os
-import numpy as np
-import pandas as pd
-import matplotlib
-matplotlib.use('WebAgg')
-import matplotlib.pyplot as plt
-
-
-
-## 00 Initialization
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-## Define function
-def GetFabricInfos(FabricFile):
-    FabricFile = open(FabricFile, 'r')
-    FabricText = FabricFile.read()
-    FabricLines = FabricText.split('\n')
-
-    for Index in range(len(FabricLines)):
-
-        if 'BV/TV' in FabricLines[Index]:
-            BVTV = np.float(FabricLines[Index].split()[-1])
-
-        if 'Eigen values' in FabricLines[Index]:
-            m1 = np.float(FabricLines[Index].split()[5])
-            m2 = np.float(FabricLines[Index].split()[6])
-            m3 = np.float(FabricLines[Index].split()[7])
-
-            DA = np.round(max(m1,m2,m3)/min(m1,m2,m3),3)
-
-        if 'Eigen vector 1' in FabricLines[Index]:
-            m1Vector = np.array([np.float(FabricLines[Index].split()[6]),
-                                 np.float(FabricLines[Index].split()[7]),
-                                 np.float(FabricLines[Index].split()[8])])
-
-        if 'Eigen vector 2' in FabricLines[Index]:
-            m2Vector = np.array([np.float(FabricLines[Index].split()[6]),
-                                 np.float(FabricLines[Index].split()[7]),
[…]

### 2021-02-05 03_Scripts/04_TransformTensor.py

@@ -1,0 +1,734 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib
+matplotlib.use('WebAgg')
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+
+
+## Define function
+def GetFabricInfos(FabricFile):
+    FabricFile = open(FabricFile, 'r')
+    FabricText = FabricFile.read()
+    FabricLines = FabricText.split('\n')
+
+    for Index in range(len(FabricLines)):
+
+        if 'BV/TV' in FabricLines[Index]:
+            BVTV = np.float(FabricLines[Index].split()[-1])
+
+        if 'Eigen values' in FabricLines[Index]:
+            m1 = np.float(FabricLines[Index].split()[5])
+            m2 = np.float(FabricLines[Index].split()[6])
+            m3 = np.float(FabricLines[Index].split()[7])
+
+            DA = np.round(max(m1,m2,m3)/min(m1,m2,m3),3)
+
+        if 'Eigen vector 1' in FabricLines[Index]:
+            m1Vector = np.array([np.float(FabricLines[Index].split()[6]),
+                                 np.float(FabricLines[Index].split()[7]),
+                                 np.float(FabricLines[Index].split()[8])])
+
+        if 'Eigen vector 2' in FabricLines[Index]:
+            m2Vector = np.array([np.float(FabricLines[Index].split()[6]),
[…]

### 2021-02-05 03_Scripts/09_GroupMatching.py

@@ -1,179 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-
-
-# 01 Define covariance functions (NB: called variance if x = y)
-def Covariance(x,y):
-
-    if not len(x)==len(y):
-        print('Covariables vectors not of equal lengths')
-    else:
-        x_bar, y_bar = x.mean(), y.mean()
-        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
-        return Cov
-def CovarianceMatrix(Data, Covariables):
-    C = np.zeros((len(Covariables), len(Covariables)))
-    for i in range(len(Covariables)):
-        for j in range(len(Covariables)):
-            x = Data[Covariables[i]].values
-            y = Data[Covariables[j]].values
-            C[i, j] = Covariance(x, y)
-
-    return C
-def MahalanobisDistance(x,y,C):
-
-    XY = x-y
-    C_inv = np.linalg.inv(C)
-    Distance = np.sqrt(XY.T.dot(C_inv.dot(XY)))
-
-    return Distance
-def ComputeNewDistances(DistancesData, SortedIndices, Matches, ControlIndividual):
-
-    TestIndividuals = np.where(Matches == ControlIndividual)[0]
-    print('Match with test individuals n:')
-    print(TestIndividuals)
-
-    print('\nCompute new distances...\n')
-    DistancesIncrease = np.zeros(len(TestIndividuals))
-    NewControlIndividuals = np.zeros(len(TestIndividuals))
-    Index = 0
-    for TestIndividual in TestIndividuals:
-        print('Control individual n: %i' % ControlIndividual)
-        print('Test individual n: %i' % TestIndividual)
-
-        ControlIndividualIndex = np.where(SortedIndices[:, TestIndividual] == ControlIndividual)[0]
-        NewControlIndividual = SortedIndices[ControlIndividualIndex + 1, TestIndividual]
-        NewControlIndividuals[Index] = NewControlIndividual
-
-        print('New control individual n: %i' % NewControlIndividual)
-
-        Distance = DistancesData[TestIndividual].loc[ControlIndividual]
[…]

### 2021-02-05 03_Scripts/10_EvaluateMatching.py

@@ -1,162 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-
-
-# 01 Define functions
-
-
-
-# 02 Load Data
-WorkingDirectory = os.getcwd()
-DataFolder = os.path.join(WorkingDirectory,'02_Data')
-ResultFolder = os.path.join(WorkingDirectory,'04_Results/00_Matching/Matches')
-
-ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('Scans.csv')]
-ScanLists.sort()
-
-HealthyGroup = pd.read_csv(os.path.join(DataFolder,ScanLists[0]))
-OIGroup = pd.read_csv(os.path.join(DataFolder,ScanLists[1]))
-
-OIGroup = OIGroup.dropna()     # Filter patients without data
-OIGroup = OIGroup.reset_index()
-OIGroup['Sex'] = OIGroup['Sex'].replace(['Female','Male'],['F','M'])
-OIGroup = OIGroup[['Scan_ID_BSL','Sex','Age']]
-OIGroup.columns = ['Sample Number','Sex','Age']
-OIGroup['Age'] = OIGroup['Age'].astype('int')
-
-## Binarize categorical values
-OIGroup['Sex'] = OIGroup['Sex'].replace(['F','M'],[1,0])
-HealthyGroup['Sex'] = HealthyGroup['Sex'].replace(['F','M'],[1,0])
-
-
-# 03 plot matching algorithm results
-Matches = [File for File in os.listdir(ResultFolder) if File.endswith('.txt')]
-Matches.sort()
-
-for Match in Matches:
-
-    MatchingResults = pd.read_csv(os.path.join(ResultFolder,Match),header=None)
-    Occurrences = MatchingResults[0].value_counts()
-    Recurrences = Occurrences[Occurrences>1]
-
-    Figure, Axes = plt.subplots(1, 1, figsize=(5.5, 4.5), dpi=100)
-    Axes.plot(MatchingResults, linestyle='none', marker='o', fillstyle='none', color=(0, 0, 0), label='Matches')
-    for Recurrence in Recurrences.index:
[…]

### 2021-02-05 03_Scripts/18_ROIMorphometry.py

@@ -33,22 +33,48 @@ def Load_Itk(Filename):
     return CT_Scan, Origin, Spacing, Size
 
 # 02 Set variables
-Path = '/home/mathieu/Documents/MscThesis/02_Data/FabricElasticityRelationship/1_ControlGroupNodaratis/B_Edit/04_ResizedROI/'
-File = '1C0001679_SEG_UNCOMPResized.mhd'
+MatchingPath = '/home/mathieu/Documents/Post-Msc/04_Results/06_LinearRegression/'
+ROIsPath = '/home/mathieu/Documents/Post-Msc/04_Results/'
 
 # 03 Get matched individuals
+MatchedList = pd.read_csv(MatchingPath+'Data.csv')
+MatchedList = MatchedList[['ROI Number', 'Scan', 'Group', 'BVTV']]
 
+# 04 Segment resized ROI
+Index = 0
 
+for Index in MatchedList[MatchedList['Group']=='Test'].index:
 
-CT_Scan, Origin, Spacing, Size = Load_Itk(Path+File)
-np.unique(CT_Scan)
+    ROI = MatchedList['ROI Number'].loc[Index]
+    Scan = MatchedList['Scan'].loc[Index]
+    Group = MatchedList['Group'].loc[Index]
 
-Threshold = 50
-CT_Scan[CT_Scan < Threshold] = 1
-CT_Scan[CT_Scan > Threshold] = 2
+    if Group == 'Control':
+        GroupPath = os.path.join(ROIsPath,'01_HealthyTibiaXCT2Scans')
+    elif Group == 'Test':
+        GroupPath = os.path.join(ROIsPath, '02_OITibiaXCT2Scans')
 
-Image = sitk.GetImageFromArray(CT_Scan)
-sitk.WriteImage(Image,'Image.mhd')
+    ResizedROIsPath = os.path.join(GroupPath,'00_ROI/02_Resized_ROI/')
+    ResizedROI = str(ROI) + '_' + Scan + '_' + 'Resized.mhd'
+
+    CT_Scan, Origin, Spacing, Size = Load_Itk(ResizedROIsPath+ResizedROI)
+
+    Threshold = 50
+    CT_Scan[CT_Scan < Threshold] = 1
+    CT_Scan[CT_Scan > Threshold] = 2
+
+    # Check BVTV
+    BVTV = np.sum(CT_Scan-1) / (Size[0] * Size[1] * Size[2])
+
[…]

### 2021-02-11 03_Scripts/08_Morpho_NormalityCheck.py
Finish morphometry statistical analysis

@@ -1,0 +1,207 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.stats.distributions import norm, f, t, chi2
+from scipy.stats import shapiro
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+## Define functions
+def QQPlot(DataValues, Alpha_CI=0.95, DataLabel='Data', FigFile='QQPlot.png'):
+
+    ### Based on: https://www.tjmahr.com/quantile-quantile-plots-from-scratch/
+    ### Itself based on Fox book: Fox, J. (2015)
+    ### Applied Regression Analysis and Generalized Linear Models.
+    ### Sage Publications, Thousand Oaks, California.
+
+    # Data analysis
+    N = len(DataValues)
+    X_Bar = np.mean(DataValues)
+    S_X = np.std(DataValues,ddof=1)
+
+    # Sort data to get the rank
+    Data_Sorted = np.zeros(N)
+    Data_Sorted += DataValues
+    Data_Sorted.sort()
+
+    # Compute quantiles
+    EmpiricalQuantiles = np.arange(0.5, N + 0.5) / N
+    TheoreticalQuantiles = norm.ppf(EmpiricalQuantiles, X_Bar, S_X)
+    ZQuantiles = norm.ppf(EmpiricalQuantiles,0,1)
+
+    # Compute data variance
+    DataIQR = np.quantile(DataValues, 0.75) - np.quantile(DataValues, 0.25)
+    NormalIQR = np.sum(np.abs(norm.ppf(np.array([0.25, 0.75]), 0, 1)))
+    Variance = DataIQR / NormalIQR
+    Z_Space = np.linspace(min(ZQuantiles), max(ZQuantiles), 100)
+    Variance_Line = Z_Space * Variance + np.median(DataValues)
+
+    # Compute alpha confidence interval (CI)
+    Z_SE = np.sqrt(norm.cdf(Z_Space) * (1 - norm.cdf(Z_Space)) / N) / norm.pdf(Z_Space)
+    Data_SE = Z_SE * Variance
[…]

### 2021-02-19 03_Scripts/00_ROI_Selection.py
Perform adjustments according to meeting of 16.02

@@ -118,7 +118,7 @@ ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('.csv')]
 ScanLists.sort()
 
 # 03 Load Data
-DataGroup = 0   # 0 = Healthy group, 1 = OI group
+DataGroup = 1   # 0 = Healthy group, 1 = OI group
 Plots = False   # Plot the different ROI location
 
 ScansPath = os.path.join(DataFolder,DataSubFolders[DataGroup])
@@ -130,6 +130,7 @@ ScanList = pd.read_csv(os.path.join(DataFolder,ScanLists[DataGroup]),sep=',')
 ParametersDataFrame = pd.DataFrame()
 
 # perform for all scans
+Scan = Scans[0]
 for Scan in Scans:
 
     ScanFile = os.path.join(ScansPath,Scan)
@@ -175,6 +176,7 @@ for Scan in Scans:
     # Plot
     if Plots:
         Ratio = Size[2] / Size[1]
+        Index = 0
         for Index in ParametersDataFrame.index:
 
             GS = gridspec.GridSpec(1,2, width_ratios=[1,Ratio])
@@ -197,7 +199,7 @@ for Scan in Scans:
             for Plane in range(2):
 
                 if Plane == 0:
-                    Axes[Plane].imshow(CT_Scan[:, :, X]-TrabBone, cmap='bone', clim=[0,1])
+                    Axes[Plane].imshow(CT_Scan[:, :, X]-TrabBone+1, cmap='bone', clim=[0,2])
                     Axes[Plane].plot([Y1, Y2], [Z1, Z1], color=(0, 0, 1))
                     Axes[Plane].plot([Y1, Y2], [Z2, Z2], color=(0, 0, 1))
                     Axes[Plane].plot([Y1, Y1], [Z1, Z2], color=(0, 0, 1))
@@ -206,7 +208,7 @@ for Scan in Scans:
                     Axes[Plane].set_ylabel('Direction 3 (voxel)')
 
                 else:
-                    Axes[Plane].imshow(CT_Scan[:, Y, :]-TrabBone, cmap='bone', clim=[0,1])
+                    Axes[Plane].imshow(CT_Scan[:, Y, :]-TrabBone+1, cmap='bone', clim=[0,2])
                     Axes[Plane].plot([X1, X2], [Z1, Z1], color=(0, 0, 1))
[…]

### 2021-02-19 03_Scripts/08_Morpho_NormalityCheck.py

@@ -77,15 +77,18 @@ ResultFolder = os.path.join(WorkingDirectory, '04_Results/04_Morphometry_Statist
 Folders = ['BVTV',
            'SMI',
            'Tb Number',
-           'Tb Spacing',
            'Tb Thickness',
-           'Coefficient of Variation']
+           'Tb Spacing',
+           'Tb Inhomogeneity',
+           'Coefficient of Variation',
+           'Degree of Anisotropy']
 Results = pd.DataFrame()
 
 
 # 02 Set variable
-Variables = ['BV/TV','SMI','Mean Tb N','Mean Tb Sp','Mean Tb Th','Coefficient of Variation']
-#Variable = Variables[5]
+Variables = ['BV/TV','SMI','Mean Tb N','Mean Tb Th','Mean Tb Sp','Std Tb Sp',
+             'Coefficient of Variation','Degree of Anisotropy']
+# Variable = Variables[5]
 for Variable in Variables:
 
     Folder = Folders[Variables.index(Variable)]
@@ -94,7 +97,6 @@ for Variable in Variables:
         os.mkdir(os.path.join(ResultFolder,Folder))
 
     Groups = ['Healthy', 'OI']
-    #Group = Groups[0]
     for Group in Groups:
 
         if Group == 'Healthy':

### 2021-02-19 03_Scripts/09_Morpho_StatisticalTests.py

@@ -370,18 +370,3 @@ for Variable in Variables:
 
 StatisticalResults.to_csv(ResultFolder + '02_StatisticalTests.csv',index=False)
 print(StatisticalResults)
-
-
-## Convert trabecular measurement
-Variable = 'Mean Tb Sp'
-Variable = 'Mean Tb Th'
-Variable = 'Mean Tb N'
-x = ControlData.groupby('Scan')[Variable].median().values
-y = OIData.groupby('Scan')[Variable].median().values
-Resolution = 0.0607
-
-x.mean()
-x.mean()*Resolution
-
-y.mean()
-y.mean()*Resolution

### 2021-02-19 03_Scripts/11_LinearRegression.py

@@ -1,0 +1,957 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import norm, t, f
+from scipy.stats import shapiro
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+## Define functions
+def Z_test(x1, x2, Alpha=0.95):
+
+        ResultsTable = pd.DataFrame()
+
+        # Compute standard deviation and number of observation
+        S_x1 = x1.std(ddof=1)
+        S_x2 = x2.std(ddof=1)
+        N_x1 = len(x1)
+        N_x2 = len(x2)
+
+        # Test statistic and p value
+        Z = (x1.mean() - x2.mean()) / np.sqrt(S_x1 ** 2 / N_x1 + S_x2 ** 2 / N_x2)
+        p = 2 * (1 - norm.cdf(abs(Z)))
+
+        # Rejection range
+        MinValue = norm.ppf( (1 - Alpha) / 2)
+        MaxValue = norm.ppf(1 - (1 - Alpha) / 2)
+        RejectionRange = np.array([[-np.inf, round(MinValue,3)], [round(MaxValue,3), np.inf]])
+
+        Results = {'Test statistic': round(Z, 3),
+                   'p value': round(p, 9),
+                   'Significance level (%)': Alpha * 100,
+                   'Rejection range': RejectionRange}
+
+        ResultsTable = ResultsTable.append(Results,ignore_index=True)
+
+        return ResultsTable
+def Covariance(x,y):
+
+    if not len(x) == len(y):
+        print('Covariables vectors not of equal lengths')
+    else:
+        x_bar, y_bar = x.mean(), y.mean()
+        Cov = np.sum(((x - x_bar) * (y - y_bar)) / (len(x) - 1))
+        return Cov
+def FisherZTransformation(r):
+    z = 1/2 * np.log((1+r)/(1-r))
+    return z
[…]

### 2021-02-23 03_Scripts/11_LinearRegression.py
Update for weekly meeting

@@ -134,6 +134,10 @@ def ANCOVA(Observations, Models, ModelNames, ModelsType):
                 Models1.append(i)
             del ListA[0]
 
+    else:
+        Models0 = [1]
+        Models1 = [0]
+
     ## Build Analysis-of-Covariance table
     ANCOVATable = pd.DataFrame()
     for i in range(len(Models0)):
@@ -237,12 +241,12 @@ def ComputeCVModelConstants(Model,Alpha=0.95,k=0,l=0):
     Beta = B['LogCV']
 
     ## Get CV values for constant range computation
-    CV_Values = Model.model.exog[:, -1]
+    LogCV_Values = Model.model.exog[:, -1]
 
     ## Compute stiffness constants
-    Mu0 = np.exp(B['Sjj'] + CV_Values * Beta) / 2
-    Lambda0p = np.exp(B['Sij'] + CV_Values * Beta)
-    Lambda0 = np.exp(B['Sii'] + CV_Values * Beta) - 2 * Mu0
+    Mu0 = np.exp(B['Sjj'] + LogCV_Values * Beta) / 2
+    Lambda0p = np.exp(B['Sij'] + LogCV_Values * Beta)
+    Lambda0 = np.exp(B['Sii'] + LogCV_Values * Beta) - 2 * Mu0
 
     ## Compute CI for lambda0 (source:https://stats.stackexchange.com/questions/485636/how-do-you-get-confidence-intervals-for-interactions-of-variables)
     Interactions = ['Sii','Sjj','LogCV']
@@ -285,7 +289,7 @@ def ComputeCVModelConstants(Model,Alpha=0.95,k=0,l=0):
           str(int(round(np.exp(Mu0_o.mean()), 0))) + ']')
 
     ## Build data frame
-    Table = pd.DataFrame({'CV Values':CV_Values,
+    Table = pd.DataFrame({'LogCV Values':LogCV_Values,
                           'Lambda0': Lambda0,
                           'Lambda0 u': np.exp(L0_u),
                           'Lambda0 o': np.exp(L0_o),
@@ -306,15 +310,20 @@ def ComputeCVModelConstants(Model,Alpha=0.95,k=0,l=0):
               str(round(B_CI.loc['LogBVTV', 1], 3)) + ']')
         print('l: ' + str(round(l, 3)) +
[…]

### 2021-02-26 03_Scripts/00_ROI_Selection.py
Sort scripts again according to the step defined during the weekly meeting

@@ -129,7 +129,7 @@ ScanList = pd.read_csv(os.path.join(DataFolder,ScanLists[DataGroup]),sep=',')
 
 ParametersDataFrame = pd.DataFrame()
 
-# perform for all scans
+# Perform for all scans
 Scan = Scans[0]
 for Scan in Scans:
 

### 2021-02-26 03_Scripts/01_2_SelectNewROI.py

@@ -1,0 +1,270 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+
+
+# 01 Define functions
+def Load_Itk(Filename):
+
+    # Reads the image using SimpleITK
+    Itkimage = sitk.ReadImage(Filename)
+
+    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
+    CT_Scan = sitk.GetArrayFromImage(Itkimage)
+
+    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
+    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
+
+    # Read the spacing along each dimension
+    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
+
+    # Read the dimension of the CT scan (in voxel)
+    Size = np.array(list(reversed(Itkimage.GetSize())))
+
+    return CT_Scan, Origin, Spacing, Size
+def SelectROI(Stack, VoxelsRange, VoxelMargin, ROINumber):
+
+    HalfStackHeight = int(Stack.shape[0] / 2)
+
+    ZMin, YMin, XMin = np.ceil(VoxelsRange / 2).astype('int') + VoxelMargin
+    ZMax, YMax, XMax = Stack.shape - np.ceil(VoxelsRange / 2).astype('int') - VoxelMargin
+
+    if ROINumber < 3:
+        Z, Y, X = np.random.randint([ZMin, YMin, XMin], [HalfStackHeight, YMax, XMax])
+    else:
+        Z, Y, X = np.random.randint([HalfStackHeight, YMin, XMin], [ZMax, YMax, XMax])
+
+    Z1, Y1, X1 = [Z, Y, X] - np.round(VoxelsRange / 2).astype('int')
+    Z2, Y2, X2 = [Z1, Y1, X1] + VoxelsRange
[…]

### 2021-02-26 03_Scripts/01_3_TransformTensor.py

@@ -1,0 +1,734 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib
+matplotlib.use('WebAgg')
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+
+
+## Define function
+def GetFabricInfos(FabricFile):
+    FabricFile = open(FabricFile, 'r')
+    FabricText = FabricFile.read()
+    FabricLines = FabricText.split('\n')
+
+    for Index in range(len(FabricLines)):
+
+        if 'BV/TV' in FabricLines[Index]:
+            BVTV = np.float(FabricLines[Index].split()[-1])
+
+        if 'Eigen values' in FabricLines[Index]:
+            m1 = np.float(FabricLines[Index].split()[5])
+            m2 = np.float(FabricLines[Index].split()[6])
+            m3 = np.float(FabricLines[Index].split()[7])
+
+            DA = np.round(max(m1,m2,m3)/min(m1,m2,m3),3)
+
+        if 'Eigen vector 1' in FabricLines[Index]:
+            m1Vector = np.array([np.float(FabricLines[Index].split()[6]),
+                                 np.float(FabricLines[Index].split()[7]),
+                                 np.float(FabricLines[Index].split()[8])])
+
+        if 'Eigen vector 2' in FabricLines[Index]:
+            m2Vector = np.array([np.float(FabricLines[Index].split()[6]),
[…]

### 2021-02-26 03_Scripts/01_Matching.py

@@ -1,191 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-
-
-# 01 Define covariance functions (NB: called variance if x = y)
-def Covariance(x,y):
-
-    if not len(x)==len(y):
-        print('Covariables vectors not of equal lengths')
-    else:
-        x_bar, y_bar = x.mean(), y.mean()
-        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
-        return Cov
-def CovarianceMatrix(Data, Covariables):
-    C = np.zeros((len(Covariables), len(Covariables)))
-    for i in range(len(Covariables)):
-        for j in range(len(Covariables)):
-            x = Data[Covariables[i]].values
-            y = Data[Covariables[j]].values
-            C[i, j] = Covariance(x, y)
-
-    return C
-def MahalanobisDistance(x,y,C):
-
-    XY = x-y
-    C_inv = np.linalg.inv(C)
-    Distance = np.sqrt(XY.T.dot(C_inv.dot(XY)))
-
-    return Distance
-def ComputeNewDistances(DistancesData, SortedIndices, Matches, ControlIndividual):
-
-    TestIndividuals = np.where(Matches == ControlIndividual)[0]
-    print('Match with test individuals n:')
-    print(TestIndividuals)
-
-    print('\nCompute new distances...\n')
-    DistancesIncrease = np.zeros(len(TestIndividuals))
-    NewControlIndividuals = np.zeros(len(TestIndividuals))
-    Index = 0
-    for TestIndividual in TestIndividuals:
-        print('Control individual n: %i' % ControlIndividual)
-        print('Test individual n: %i' % TestIndividual)
-
-        ControlIndividualIndex = np.where(SortedIndices[:, TestIndividual] == ControlIndividual)[0]
-        NewControlIndividual = SortedIndices[ControlIndividualIndex + 1, TestIndividual]
-        NewControlIndividuals[Index] = NewControlIndividual
-
-        print('New control individual n: %i' % NewControlIndividual)
-
-        Distance = DistancesData[TestIndividual].loc[ControlIndividual]
[…]

### 2021-02-26 03_Scripts/02_1_Collect_StiffnessData.py

@@ -1,0 +1,324 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+## Define functions
+def Load_Itk(Filename):
+
+    # Reads the image using SimpleITK
+    Itkimage = sitk.ReadImage(Filename)
+
+    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
+    CT_Scan = sitk.GetArrayFromImage(Itkimage)
+
+    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
+    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
+
+    # Read the spacing along each dimension
+    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
+
+    # Read the dimension of the CT scan (in voxel)
+    Size = np.array(list(reversed(Itkimage.GetSize())))
+
+    return CT_Scan, Origin, Spacing, Size
+def ROIAnalysis(ROIFilePath):
+
+    # Load ROI
+    ImageArray, Origin, Spacing, Size = Load_Itk(ROIFilePath+'_Cleaned.mhd')
+
+    # Compute BV/TV for each subcube
+    RhoSub = np.zeros((2, 2, 2))
+    Indices = [0, np.int(Size[0] / 2), Size[0]]
+
+    for i in range(2):
+        for j in range(2):
+            for k in range(2):
+                ii = [Indices[i], Indices[i + 1]]
+                jj = [Indices[j], Indices[j + 1]]
+                kk = [Indices[k], Indices[k + 1]]
+
+                SubCube = ImageArray[ii[0]:ii[1], jj[0]:jj[1], kk[0]:kk[1]]
+
+                BV = np.sum(SubCube)
+                TV = SubCube.shape[0] * SubCube.shape[1] * SubCube.shape[2]
+                RhoSub[i, j, k] = BV / TV
+
[…]

### 2021-02-26 03_Scripts/02_2_General_LinearRegression.py

@@ -1,0 +1,477 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import norm, t, f
+from scipy.stats import shapiro
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+## Define functions
+def ANCOVA(Models, ModelNames, ModelsType):
+
+    ModelsResults = pd.DataFrame()
+
+    for Model in Models:
+
+        Index = list(Models).index(Model)
+
+        ## Compute values for ANCOVA
+        Y_Obs = Model.model.endog
+        E = Model.resid.values
+        RSS = np.sum(E ** 2)
+        SE = np.sqrt(RSS / Model.df_resid)
+        TSS = np.sum((Y_Obs - Y_Obs.mean()) ** 2)
+        RegSS = TSS - RSS
+        R2 = RegSS / TSS
+
+        if ModelsType == 'Standard':
+            DOF = Model.df_model
+        elif ModelsType == 'Mixed-Effect':
+            DOF = Model.df_modelwc
+
+        ModelResults = {'Model': ModelNames[Index],
+                        'Regression SS': RegSS,
+                        'Residual SS': RSS,
+                        'df': DOF,
+                        'R2': round(R2, 3),
+                        'SE': round(SE,3)}
+        ModelsResults = ModelsResults.append(ModelResults, ignore_index=True)
+    print('\n\nResults of the fitting:')
+    print(ModelsResults[['Model', 'df', 'Regression SS', 'R2', 'SE']])
+
+    ## Build lists of models to compare
+    NModels = len(Models)
+    Models0 = []
+    Models1 = []
+    if NModels > 2:
+        ListA = np.arange(1,NModels).tolist()
[…]

### 2021-02-26 03_Scripts/02_EvaluateMatching.py

@@ -1,163 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-
-
-# 01 Set paths
-WorkingDirectory = os.getcwd()
-DataFolder = os.path.join(WorkingDirectory,'02_Data')
-MatchingFolder = os.path.join(WorkingDirectory,'04_Results/01_Matching')
-MatchesFolder = os.path.join(MatchingFolder,'Matches')
-ResultsFolder = os.path.join(MatchingFolder,'Matching_Assessment')
-
-
-# 02 Load Data
-Matched_OI = pd.read_csv(os.path.join(MatchingFolder,'Matched_OI.csv'))
-Matched_Healthy = pd.read_csv(os.path.join(MatchingFolder,'Matched_Control.csv'))
-
-ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('Scans.csv')]
-ScanLists.sort()
-HealthyData = pd.read_csv(os.path.join(DataFolder,ScanLists[0]))
-OIData = pd.read_csv(os.path.join(DataFolder,ScanLists[1]))
-
-## Adapt labeling and values type
-OIData = OIData.dropna()     # Filter patients without data
-OIData = OIData.reset_index()
-OIData['Sex'] = OIData['Sex'].replace(['Female','Male'],['F','M'])
-OIData = OIData[['Scan_ID_BSL','Sex','Age']]
-OIData.columns = ['Sample Number','Sex','Age']
-OIData['Age'] = OIData['Age'].astype('int')
-
-## Binarize categorical values
-OIData['Sex'] = OIData['Sex'].replace(['F','M'],[1,0])
-HealthyData['Sex'] = HealthyData['Sex'].replace(['F','M'],[1,0])
-Matched_OI['Sex'] = Matched_OI['Sex'].replace(['F','M'],[1,0])
-Matched_Healthy['Sex'] = Matched_Healthy['Sex'].replace(['F','M'],[1,0])
-
-
-# 03 plot matching algorithm results
-Matches = [File for File in os.listdir(MatchesFolder) if File.endswith('.txt')]
-Matches.sort()
-for Match in Matches:
-
[…]

### 2021-02-26 03_Scripts/03_EvaluateMatching.py

@@ -1,0 +1,163 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+
+
+# 01 Set paths
+WorkingDirectory = os.getcwd()
+DataFolder = os.path.join(WorkingDirectory,'02_Data')
+MatchingFolder = os.path.join(WorkingDirectory,'04_Results/03_Individuals_Matching')
+MatchesFolder = os.path.join(MatchingFolder,'Matches')
+ResultsFolder = os.path.join(MatchingFolder,'Matching_Assessment')
+
+
+# 02 Load Data
+Matched_OI = pd.read_csv(os.path.join(MatchingFolder,'Matched_OI.csv'))
+Matched_Healthy = pd.read_csv(os.path.join(MatchingFolder,'Matched_Control.csv'))
+
+ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('Scans.csv')]
+ScanLists.sort()
+HealthyData = pd.read_csv(os.path.join(DataFolder,ScanLists[0]))
+OIData = pd.read_csv(os.path.join(DataFolder,ScanLists[1]))
+
+## Adapt labeling and values type
+OIData = OIData.dropna()     # Filter patients without data
+OIData = OIData.reset_index()
+OIData['Sex'] = OIData['Sex'].replace(['Female','Male'],['F','M'])
+OIData = OIData[['Scan_ID_BSL','Sex','Age']]
+OIData.columns = ['Sample Number','Sex','Age']
+OIData['Age'] = OIData['Age'].astype('int')
+
+## Binarize categorical values
+OIData['Sex'] = OIData['Sex'].replace(['F','M'],[1,0])
+HealthyData['Sex'] = HealthyData['Sex'].replace(['F','M'],[1,0])
+Matched_OI['Sex'] = Matched_OI['Sex'].replace(['F','M'],[1,0])
+Matched_Healthy['Sex'] = Matched_Healthy['Sex'].replace(['F','M'],[1,0])
+
+
+# 03 plot matching algorithm results
+Matches = [File for File in os.listdir(MatchesFolder) if File.endswith('.txt')]
+Matches.sort()
+for Match in Matches:
+
[…]

### 2021-02-26 03_Scripts/03_Individuals_Matching.py

@@ -1,0 +1,191 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+
+
+# 01 Define covariance functions (NB: called variance if x = y)
+def Covariance(x,y):
+
+    if not len(x)==len(y):
+        print('Covariables vectors not of equal lengths')
+    else:
+        x_bar, y_bar = x.mean(), y.mean()
+        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
+        return Cov
+def CovarianceMatrix(Data, Covariables):
+    C = np.zeros((len(Covariables), len(Covariables)))
+    for i in range(len(Covariables)):
+        for j in range(len(Covariables)):
+            x = Data[Covariables[i]].values
+            y = Data[Covariables[j]].values
+            C[i, j] = Covariance(x, y)
+
+    return C
+def MahalanobisDistance(x,y,C):
+
+    XY = x-y
+    C_inv = np.linalg.inv(C)
+    Distance = np.sqrt(XY.T.dot(C_inv.dot(XY)))
+
+    return Distance
+def ComputeNewDistances(DistancesData, SortedIndices, Matches, ControlIndividual):
+
+    TestIndividuals = np.where(Matches == ControlIndividual)[0]
+    print('Match with test individuals n:')
+    print(TestIndividuals)
+
+    print('\nCompute new distances...\n')
+    DistancesIncrease = np.zeros(len(TestIndividuals))
+    NewControlIndividuals = np.zeros(len(TestIndividuals))
+    Index = 0
+    for TestIndividual in TestIndividuals:
+        print('Control individual n: %i' % ControlIndividual)
+        print('Test individual n: %i' % TestIndividual)
+
+        ControlIndividualIndex = np.where(SortedIndices[:, TestIndividual] == ControlIndividual)[0]
+        NewControlIndividual = SortedIndices[ControlIndividualIndex + 1, TestIndividual]
+        NewControlIndividuals[Index] = NewControlIndividual
+
+        print('New control individual n: %i' % NewControlIndividual)
+
+        Distance = DistancesData[TestIndividual].loc[ControlIndividual]
[…]

### 2021-02-26 03_Scripts/03_WriteROIAnalysisBatch.py

@@ -1,112 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-
-
-# 01 Define paths
-WorkingDirectory = os.getcwd()
-DataFolder = os.path.join(WorkingDirectory,'02_Data')
-MatchingFolder = os.path.join(WorkingDirectory,'04_Results')
-ResultsFolder = os.path.join(WorkingDirectory,'04_Results/02_ROI_Analysis')
-
-
-# 02 Load data
-DataGroup = 1   # 0 = Healthy group, 1 = OI group
-
-DataSubFolders = [File for File in os.listdir(DataFolder) if os.path.isdir(os.path.join(DataFolder,File))]
-DataSubFolders.sort()
-MedtoolParametersFile = os.path.join(MatchingFolder, '00_ROI_Selection' , DataSubFolders[DataGroup][:-14] + 'ROI.csv')
-MedtoolParametersFile = os.path.join(MatchingFolder, '00_ROI_Selection' , '02_OIROI.csv')
-MedtoolParametersData = pd.read_csv(MedtoolParametersFile, sep=';')
-del MedtoolParametersData['Unnamed: 8']
-
-
-PipelineFile  = open(os.path.join(WorkingDirectory, '03_Scripts', '03_ROIAnalysisPipeline.txt'), 'r')
-PipelineText  = PipelineFile.read()
-
-if DataGroup == 0:
-    MatchingResults = pd.read_csv(os.path.join(MatchingFolder,'01_Matching','Matched_Control.csv'))
-elif DataGroup == 1:
-    MatchingResults = pd.read_csv(os.path.join(MatchingFolder,'01_Matching','Matched_OI.csv'))
-
-
-# 03 Filter matched individuals
-for Index in MedtoolParametersData.index:
-
-    Scan = MedtoolParametersData['$Scan'].loc[Index]
-
-    if DataGroup == 0:
-        ScanNumber = int(Scan[4:8])
-
-        if ScanNumber not in MatchingResults['Sample Number'].values:
-            MedtoolParametersData = MedtoolParametersData.drop(index=Index)
-
[…]

### 2021-02-26 03_Scripts/04_SelectNewROI.py

@@ -1,267 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import SimpleITK as sitk
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-# 01 Define functions
-def Load_Itk(Filename):
-
-    # Reads the image using SimpleITK
-    Itkimage = sitk.ReadImage(Filename)
-
-    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
-    CT_Scan = sitk.GetArrayFromImage(Itkimage)
-
-    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
-    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
-
-    # Read the spacing along each dimension
-    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
-
-    # Read the dimension of the CT scan (in voxel)
-    Size = np.array(list(reversed(Itkimage.GetSize())))
-
-    return CT_Scan, Origin, Spacing, Size
-def SelectROI(Stack, VoxelsRange, VoxelMargin, ROINumber):
-
-    HalfStackHeight = int(Stack.shape[0] / 2)
-
-    ZMin, YMin, XMin = np.ceil(VoxelsRange / 2).astype('int') + VoxelMargin
-    ZMax, YMax, XMax = Stack.shape - np.ceil(VoxelsRange / 2).astype('int') - VoxelMargin
-
-    if ROINumber < 3:
-        Z, Y, X = np.random.randint([ZMin, YMin, XMin], [HalfStackHeight, YMax, XMax])
-    else:
-        Z, Y, X = np.random.randint([HalfStackHeight, YMin, XMin], [ZMax, YMax, XMax])
-
-    Z1, Y1, X1 = [Z, Y, X] - np.round(VoxelsRange / 2).astype('int')
-    Z2, Y2, X2 = [Z1, Y1, X1] + VoxelsRange
-
-    ROI = [Z, Y, X, Z1, Y1, X1, Z2, Y2, X2]
-
-    return ROI
[…]

### 2021-02-26 03_Scripts/05_TransformTensor.py

@@ -1,734 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib
-matplotlib.use('WebAgg')
-import matplotlib.pyplot as plt
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-
-
-## Define function
-def GetFabricInfos(FabricFile):
-    FabricFile = open(FabricFile, 'r')
-    FabricText = FabricFile.read()
-    FabricLines = FabricText.split('\n')
-
-    for Index in range(len(FabricLines)):
-
-        if 'BV/TV' in FabricLines[Index]:
-            BVTV = np.float(FabricLines[Index].split()[-1])
-
-        if 'Eigen values' in FabricLines[Index]:
-            m1 = np.float(FabricLines[Index].split()[5])
-            m2 = np.float(FabricLines[Index].split()[6])
-            m3 = np.float(FabricLines[Index].split()[7])
-
-            DA = np.round(max(m1,m2,m3)/min(m1,m2,m3),3)
-
-        if 'Eigen vector 1' in FabricLines[Index]:
-            m1Vector = np.array([np.float(FabricLines[Index].split()[6]),
-                                 np.float(FabricLines[Index].split()[7]),
-                                 np.float(FabricLines[Index].split()[8])])
-
-        if 'Eigen vector 2' in FabricLines[Index]:
-            m2Vector = np.array([np.float(FabricLines[Index].split()[6]),
[…]

### 2021-02-26 03_Scripts/06_BinarizeROI.py

@@ -39,7 +39,7 @@ def Load_Itk(Filename):
 # 02 Set variables
 Group = 'OI'   # Healthy or OI
 WorkingDirectory = os.getcwd()
-DataPath = os.path.join(WorkingDirectory,'04_Results/02_ROI_Analysis')
+DataPath = os.path.join(WorkingDirectory,'04_Results/01_ROI_Analysis')
 ROIsPath = os.path.join(DataPath,'01_' + Group + '_ROIs')
 ResizedROIsPath = os.path.join(ROIsPath,'01_Resized_ROIs/')
 BinarizedROIsPath = os.path.join(ROIsPath,'02_Binarized_ROIs/')

### 2021-02-26 03_Scripts/10_Collect_StiffnessData.py

@@ -1,324 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import SimpleITK as sitk
-import matplotlib.pyplot as plt
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-plt.rc('font', size=12)
-
-## Define functions
-def Load_Itk(Filename):
-
-    # Reads the image using SimpleITK
-    Itkimage = sitk.ReadImage(Filename)
-
-    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
-    CT_Scan = sitk.GetArrayFromImage(Itkimage)
-
-    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
-    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
-
-    # Read the spacing along each dimension
-    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
-
-    # Read the dimension of the CT scan (in voxel)
-    Size = np.array(list(reversed(Itkimage.GetSize())))
-
-    return CT_Scan, Origin, Spacing, Size
-def ROIAnalysis(ROIFilePath):
-
-    # Load ROI
-    ImageArray, Origin, Spacing, Size = Load_Itk(ROIFilePath+'_Cleaned.mhd')
-
-    # Compute BV/TV for each subcube
-    RhoSub = np.zeros((2, 2, 2))
-    Indices = [0, np.int(Size[0] / 2), Size[0]]
-
-    for i in range(2):
-        for j in range(2):
-            for k in range(2):
-                ii = [Indices[i], Indices[i + 1]]
-                jj = [Indices[j], Indices[j + 1]]
-                kk = [Indices[k], Indices[k + 1]]
-
-                SubCube = ImageArray[ii[0]:ii[1], jj[0]:jj[1], kk[0]:kk[1]]
-
-                BV = np.sum(SubCube)
-                TV = SubCube.shape[0] * SubCube.shape[1] * SubCube.shape[2]
-                RhoSub[i, j, k] = BV / TV
-
[…]

### 2021-02-26 03_Scripts/Control_FilesValidity.py

@@ -1,0 +1,107 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+
+
+## Define function
+def GetFabricInfos(FabricFile):
+    FabricFile = open(FabricFile, 'r')
+    FabricText = FabricFile.read()
+    FabricLines = FabricText.split('\n')
+
+    for Index in range(len(FabricLines)):
+
+        if 'BV/TV' in FabricLines[Index]:
+            BVTV = np.float(FabricLines[Index].split()[-1])
+
+        if 'Eigen values' in FabricLines[Index]:
+            m1 = np.float(FabricLines[Index].split()[5])
+            m2 = np.float(FabricLines[Index].split()[6])
+            m3 = np.float(FabricLines[Index].split()[7])
+
+            DA = np.round(max(m1,m2,m3)/min(m1,m2,m3),3)
+
+        if 'Eigen vector 1' in FabricLines[Index]:
+            m1Vector = np.array([np.float(FabricLines[Index].split()[6]),
+                                 np.float(FabricLines[Index].split()[7]),
+                                 np.float(FabricLines[Index].split()[8])])
+
+        if 'Eigen vector 2' in FabricLines[Index]:
+            m2Vector = np.array([np.float(FabricLines[Index].split()[6]),
+                                 np.float(FabricLines[Index].split()[7]),
+                                 np.float(FabricLines[Index].split()[8])])
+
+        if 'Eigen vector 3' in FabricLines[Index]:
+            m3Vector = np.array([np.float(FabricLines[Index].split()[6]),
+                                 np.float(FabricLines[Index].split()[7]),
+                                 np.float(FabricLines[Index].split()[8])])
+
+    FabricInfosData = pd.DataFrame()
+
+    FabricInfos = {'BVTV':BVTV,'m1':m1,'m2':m2,'m3':m3,'DA':DA,
[…]

### 2021-02-26 03_Scripts/RCrosscheck.R

@@ -7,71 +7,32 @@
 library(carData)
 library(effects)
 library(car)
-Data <- read.table("/home/mathieu/Documents/Post-Msc/04_Results/06_LinearRegression/RData.csv", header=TRUE, sep=",")
+Data <- read.table("/home/mathieu/Documents/Post-Msc/04_Results/02_General_LinearRegression/Healthy_RData.csv", header=TRUE, sep=",")
 
 # Check data
 dim(Data)
-Data[sort(sample(5220, 12)),]
+Data[sort(sample(3528, 12)),]
 
 # Build linear regression models for ANCOVA
-Model1.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy) - 1, data = Data)
+Model1.mod <- lm(LogSxy ~ Sii + Sij + Sjj + LogBVTV + Logmxy - 1, data = Data)
 summary(Model1.mod)
 
-Model2.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
-                      + (Sii + Sij + Sjj):Sg - 1, data = Data)
+Model2.mod <- lm(LogSxy ~ Sii + Sij + Sjj + LogBVTV + Logmxy + LogCV - 1, data = Data)
 summary(Model2.mod)
 
-Model3.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
-                      + (Sii + Sij + Sjj):log(CV) - 1, data = Data)
-summary(Model3.mod)
-
 anova(Model1.mod,Model2.mod)
-anova(Model1.mod,Model3.mod)
-
-Model4.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
-                      + (Sii + Sij + Sjj):Sg
-                      + (Sii + Sij + Sjj):log(CV) - 1, data = Data)
-summary(Model4.mod)
-
-anova(Model2.mod,Model4.mod)
-anova(Model3.mod,Model4.mod)
-
-Model5.mod <- lm(log(Sxy) ~ Sii + Sij + Sjj + log(BVTV) + log(mxy)
-                      + (Sii + Sij + Sjj):Sg
-                      + (Sii + Sij + Sjj):log(CV)
[…]

### 2021-03-08 03_Scripts/02_1_Collect_StiffnessData.py
Latest version. Methods and results ready to be written for article

@@ -1,324 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import SimpleITK as sitk
-import matplotlib.pyplot as plt
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-plt.rc('font', size=12)
-
-## Define functions
-def Load_Itk(Filename):
-
-    # Reads the image using SimpleITK
-    Itkimage = sitk.ReadImage(Filename)
-
-    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
-    CT_Scan = sitk.GetArrayFromImage(Itkimage)
-
-    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
-    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
-
-    # Read the spacing along each dimension
-    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
-
-    # Read the dimension of the CT scan (in voxel)
-    Size = np.array(list(reversed(Itkimage.GetSize())))
-
-    return CT_Scan, Origin, Spacing, Size
-def ROIAnalysis(ROIFilePath):
-
-    # Load ROI
-    ImageArray, Origin, Spacing, Size = Load_Itk(ROIFilePath+'_Cleaned.mhd')
-
-    # Compute BV/TV for each subcube
-    RhoSub = np.zeros((2, 2, 2))
-    Indices = [0, np.int(Size[0] / 2), Size[0]]
-
-    for i in range(2):
-        for j in range(2):
-            for k in range(2):
-                ii = [Indices[i], Indices[i + 1]]
-                jj = [Indices[j], Indices[j + 1]]
-                kk = [Indices[k], Indices[k + 1]]
-
-                SubCube = ImageArray[ii[0]:ii[1], jj[0]:jj[1], kk[0]:kk[1]]
-
-                BV = np.sum(SubCube)
-                TV = SubCube.shape[0] * SubCube.shape[1] * SubCube.shape[2]
-                RhoSub[i, j, k] = BV / TV
-
[…]

### 2021-03-08 03_Scripts/02_1_Individuals_Matching.py

@@ -1,0 +1,193 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 320
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
+
+
+
+# 01 Define covariance functions (NB: called variance if x = y)
+def Covariance(x,y):
+
+    if not len(x)==len(y):
+        print('Covariables vectors not of equal lengths')
+    else:
+        x_bar, y_bar = x.mean(), y.mean()
+        Cov = np.sum(((x - x_bar)*(y - y_bar))/(len(x) - 1))
+        return Cov
+def CovarianceMatrix(Data, Covariables):
+    C = np.zeros((len(Covariables), len(Covariables)))
+    for i in range(len(Covariables)):
+        for j in range(len(Covariables)):
+            x = Data[Covariables[i]].values
+            y = Data[Covariables[j]].values
+            C[i, j] = Covariance(x, y)
+
+    return C
+def MahalanobisDistance(x,y,C):
+
+    XY = x-y
+    C_inv = np.linalg.inv(C)
+    Distance = np.sqrt(XY.T.dot(C_inv.dot(XY)))
+
+    return Distance
+def ComputeNewDistances(DistancesData, SortedIndices, Matches, ControlIndividual):
+
+    TestIndividuals = np.where(Matches == ControlIndividual)[0]
+    print('Match with test individuals n:')
+    print(TestIndividuals)
+
+    print('\nCompute new distances...\n')
+    DistancesIncrease = np.zeros(len(TestIndividuals))
+    NewControlIndividuals = np.zeros(len(TestIndividuals))
+    Index = 0
+    for TestIndividual in TestIndividuals:
+        print('Control individual n: %i' % ControlIndividual)
+        print('Test individual n: %i' % TestIndividual)
+
+        ControlIndividualIndex = np.where(SortedIndices[:, TestIndividual] == ControlIndividual)[0]
+        NewControlIndividual = SortedIndices[ControlIndividualIndex + 1, TestIndividual]
+        NewControlIndividuals[Index] = NewControlIndividual
+
+        print('New control individual n: %i' % NewControlIndividual)
+
+        Distance = DistancesData[TestIndividual].loc[ControlIndividual]
[…]

### 2021-03-08 03_Scripts/03_1_BinarizeROI.py

@@ -1,0 +1,118 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+
+# 01 Define functions
+def Load_Itk(Filename):
+
+    # Reads the image using SimpleITK
+    Itkimage = sitk.ReadImage(Filename)
+
+    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
+    CT_Scan = sitk.GetArrayFromImage(Itkimage)
+
+    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
+    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
+
+    # Read the spacing along each dimension
+    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
+
+    # Read the dimension of the CT scan (in voxel)
+    Size = np.array(list(reversed(Itkimage.GetSize())))
+
+    return CT_Scan, Origin, Spacing, Size
+
+
+
+# 02 Set variables
+Group = 'OI'   # Healthy or OI
+WorkingDirectory = os.getcwd()
+DataPath = os.path.join(WorkingDirectory,'04_Results/01_ROI_Analysis')
+ROIsPath = os.path.join(DataPath,'01_' + Group + '_ROIs')
+ResizedROIsPath = os.path.join(ROIsPath,'01_Resized_ROIs/')
+BinarizedROIsPath = os.path.join(ROIsPath,'02_Binarized_ROIs/')
+MatchingPath = os.path.join(WorkingDirectory, '04_Results/02_Individuals_Matching/')
+
+
+
+# 03 Binarized resized ROI and write file
+Resized_ROIs = [File for File in os.listdir(ResizedROIsPath) if File.endswith('.mhd')]
+MatchList = pd.read_csv(MatchingPath+'Matched_'+Group+'.csv')
+
+
+for Resized_ROI in Resized_ROIs:
+
+    if Group == 'Healthy':
+        Test = Resized_ROI[6:10] in list(MatchList['Sample Number'].astype('str'))
+
[…]

### 2021-03-08 03_Scripts/03_5_Morpho_Comparison.py

@@ -1,0 +1,107 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.api as sm
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+# 01 Set path and load data
+WorkingDirectory = os.getcwd()
+DataFolder = os.path.join(WorkingDirectory, '02_Data/')
+ResultFolder = os.path.join(WorkingDirectory, '04_Results/03_Morphometry_Statistics/')
+
+ControlMedtool = pd.read_csv(ResultFolder + '00_Healthy_Data.csv')
+OIMedtool = pd.read_csv(ResultFolder + '00_OI_Data.csv')
+ControlScanco = pd.read_csv(DataFolder + '01_HealthyTibiaXCT2Scans.csv')
+OIScanco = pd.read_csv(DataFolder + '02_OITibiaXCT2Scans.csv')
+
+
+# 02 Modify data frames and get medtool median values for each scan
+ControlScanco = ControlScanco.loc[:3,['Tb. BVTV','Tb. Thickness','Tb. N.','Tb. Sp.']]
+
+## build Scan ID columns for medtool results
+ControlMedtool['Scan_ID'] = ControlMedtool['Scan']
+OIMedtool['Patient_ID'] = OIMedtool['Scan']
+
+for Index in ControlMedtool.index:
+    ControlMedtool.loc[Index,'Scan_ID'] = ControlMedtool.loc[Index,'Scan_ID'][4:8]
+for Index in OIMedtool.index:
+    OIMedtool.loc[Index,'Patient_ID'] = OIMedtool.loc[Index,'Patient_ID'][:6]
+
+## Take median ROI values
+ControlMedtool = ControlMedtool[['Scan_ID','BV/TV','Mean Tb Th', 'Mean Tb Sp','Mean Tb N', 'Std Tb Sp']]
+OIMedtool = OIMedtool[['Patient_ID','BV/TV','Mean Tb Th', 'Mean Tb Sp','Mean Tb N', 'Std Tb Sp']]
[…]

### 2021-03-08 03_Scripts/03_EvaluateMatching.py

@@ -1,163 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-
-desired_width = 320
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:f}'.format})
-
-
-
-# 01 Set paths
-WorkingDirectory = os.getcwd()
-DataFolder = os.path.join(WorkingDirectory,'02_Data')
-MatchingFolder = os.path.join(WorkingDirectory,'04_Results/03_Individuals_Matching')
-MatchesFolder = os.path.join(MatchingFolder,'Matches')
-ResultsFolder = os.path.join(MatchingFolder,'Matching_Assessment')
-
-
-# 02 Load Data
-Matched_OI = pd.read_csv(os.path.join(MatchingFolder,'Matched_OI.csv'))
-Matched_Healthy = pd.read_csv(os.path.join(MatchingFolder,'Matched_Control.csv'))
-
-ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('Scans.csv')]
-ScanLists.sort()
-HealthyData = pd.read_csv(os.path.join(DataFolder,ScanLists[0]))
-OIData = pd.read_csv(os.path.join(DataFolder,ScanLists[1]))
-
-## Adapt labeling and values type
-OIData = OIData.dropna()     # Filter patients without data
-OIData = OIData.reset_index()
-OIData['Sex'] = OIData['Sex'].replace(['Female','Male'],['F','M'])
-OIData = OIData[['Scan_ID_BSL','Sex','Age']]
-OIData.columns = ['Sample Number','Sex','Age']
-OIData['Age'] = OIData['Age'].astype('int')
-
-## Binarize categorical values
-OIData['Sex'] = OIData['Sex'].replace(['F','M'],[1,0])
-HealthyData['Sex'] = HealthyData['Sex'].replace(['F','M'],[1,0])
-Matched_OI['Sex'] = Matched_OI['Sex'].replace(['F','M'],[1,0])
-Matched_Healthy['Sex'] = Matched_Healthy['Sex'].replace(['F','M'],[1,0])
-
-
-# 03 plot matching algorithm results
-Matches = [File for File in os.listdir(MatchesFolder) if File.endswith('.txt')]
-Matches.sort()
-for Match in Matches:
-
[…]

### 2021-03-08 03_Scripts/05_1_Collect_ROI_Data.py

@@ -1,0 +1,131 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+## Define functions
+def Load_Itk(Filename):
+
+    # Reads the image using SimpleITK
+    Itkimage = sitk.ReadImage(Filename)
+
+    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
+    CT_Scan = sitk.GetArrayFromImage(Itkimage)
+
+    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
+    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
+
+    # Read the spacing along each dimension
+    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
+
+    # Read the dimension of the CT scan (in voxel)
+    Size = np.array(list(reversed(Itkimage.GetSize())))
+
+    return CT_Scan, Origin, Spacing, Size
+def ROIAnalysis(ROIFilePath):
+
+    # Load ROI
+    ImageArray, Origin, Spacing, Size = Load_Itk(ROIFilePath)
+
+    # Compute BV/TV for each subcube
+    RhoSub = np.zeros((2, 2, 2))
+    Indices = [0, np.int(Size[0] / 2), Size[0]]
+
+    for i in range(2):
+        for j in range(2):
+            for k in range(2):
+                ii = [Indices[i], Indices[i + 1]]
+                jj = [Indices[j], Indices[j + 1]]
+                kk = [Indices[k], Indices[k + 1]]
+
+                SubCube = ImageArray[ii[0]:ii[1], jj[0]:jj[1], kk[0]:kk[1]]
+
+                BV = np.sum(SubCube)
+                TV = SubCube.shape[0] * SubCube.shape[1] * SubCube.shape[2]
+                RhoSub[i, j, k] = BV / TV
+
+    return np.std(RhoSub) / np.mean(RhoSub)
+def GetFabricInfos(FabricFile):
+    FabricFile = open(FabricFile, 'r')
[…]

### 2021-03-08 03_Scripts/05_3_EvaluateMatching.py

@@ -1,0 +1,88 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+# 01 Set paths
+WorkingDirectory = os.getcwd()
+DataFolder = os.path.join(WorkingDirectory,'04_Results/05_Variables_Matching/')
+ResultsFolder = os.path.join(DataFolder,'Matching_Assessment')
+
+
+# 02 Load Data
+HealthyData = pd.read_csv(DataFolder + '00_Healthy_Data.csv')
+OIData = pd.read_csv(DataFolder + '00_OI_Data.csv')
+Matched_OI = pd.read_csv(os.path.join(DataFolder,'01_Matched_OI.csv'))
+Matched_Healthy = pd.read_csv(os.path.join(DataFolder,'01_Matched_Healthy.csv'))
+
+
+
+# 03 Matching assessment - Numerical
+Covariates = ['BV/TV', 'Degree of Anisotropy']
+
+TotalMeanOI = OIData[Covariates].mean()
+TotalMeanControl = HealthyData[Covariates].mean()
+TotalStdOI = OIData[Covariates].std()
+
+TotalStandardBias = (TotalMeanOI-TotalMeanControl)/TotalStdOI
+
+MatchedMeanOI = Matched_OI[Covariates].mean()
+MatchedMeanControl = Matched_Healthy[Covariates].mean()
+MatchedStdOI= Matched_OI[Covariates].std()
+
+MatchedStandardBias = (MatchedMeanOI-MatchedMeanControl)/MatchedStdOI
+
+Colors = [(1,0,0),(0,0,1)]
+Labels = ['BV/TV','DA']
+
+Figure, Axes = plt.subplots(1, 1, figsize=(5.5, 4.5), dpi=100)
+for i in range(len(Covariates)):
+    Axes.plot([0,1], [TotalStandardBias[Covariates[i]], MatchedStandardBias[Covariates[i]]],
+              color=Colors[i])
+    Axes.plot([0], TotalStandardBias[Covariates[i]],
+              linestyle='none', marker='o',
[…]

### 2021-03-08 03_Scripts/06_BinarizeROI.py

@@ -1,86 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import SimpleITK as sitk
-import matplotlib.pyplot as plt
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-plt.rc('font', size=12)
-
-
-
-# 01 Define functions
-def Load_Itk(Filename):
-
-    # Reads the image using SimpleITK
-    Itkimage = sitk.ReadImage(Filename)
-
-    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
-    CT_Scan = sitk.GetArrayFromImage(Itkimage)
-
-    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
-    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
-
-    # Read the spacing along each dimension
-    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
-
-    # Read the dimension of the CT scan (in voxel)
-    Size = np.array(list(reversed(Itkimage.GetSize())))
-
-    return CT_Scan, Origin, Spacing, Size
-
-
-
-# 02 Set variables
-Group = 'OI'   # Healthy or OI
-WorkingDirectory = os.getcwd()
-DataPath = os.path.join(WorkingDirectory,'04_Results/01_ROI_Analysis')
-ROIsPath = os.path.join(DataPath,'01_' + Group + '_ROIs')
-ResizedROIsPath = os.path.join(ROIsPath,'01_Resized_ROIs/')
-BinarizedROIsPath = os.path.join(ROIsPath,'02_Binarized_ROIs/')
-
-
-
-# 03 Binarized resized ROI and write file
-Resized_ROIs = [File for File in os.listdir(ResizedROIsPath) if File.endswith('.mhd')]
-
-for Resized_ROI in Resized_ROIs:
-
-    CT_Scan, Origin, Spacing, Size = Load_Itk(ResizedROIsPath + Resized_ROI)
-
-    Threshold = 50
-    CT_Scan[CT_Scan < Threshold] = 1
[…]

### 2021-03-08 03_Scripts/07_Collect_MorphoData.py

@@ -1,162 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import SimpleITK as sitk
-import matplotlib.pyplot as plt
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-plt.rc('font', size=12)
-
-## Define functions
-def Load_Itk(Filename):
-
-    # Reads the image using SimpleITK
-    Itkimage = sitk.ReadImage(Filename)
-
-    # Convert the image to a  numpy array first and then shuffle the dimensions to get axis in the order z,y,x
-    CT_Scan = sitk.GetArrayFromImage(Itkimage)
-
-    # Read the origin of the ct_scan, will be used to convert the coordinates from world to voxel and vice versa.
-    Origin = np.array(list(reversed(Itkimage.GetOrigin())))
-
-    # Read the spacing along each dimension
-    Spacing = np.array(list(reversed(Itkimage.GetSpacing())))
-
-    # Read the dimension of the CT scan (in voxel)
-    Size = np.array(list(reversed(Itkimage.GetSize())))
-
-    return CT_Scan, Origin, Spacing, Size
-def ROIAnalysis(ROIFilePath):
-
-    # Load ROI
-    ImageArray, Origin, Spacing, Size = Load_Itk(ROIFilePath+'_Cleaned.mhd')
-
-    # Compute BV/TV for each subcube
-    RhoSub = np.zeros((2, 2, 2))
-    Indices = [0, np.int(Size[0] / 2), Size[0]]
-
-    for i in range(2):
-        for j in range(2):
-            for k in range(2):
-                ii = [Indices[i], Indices[i + 1]]
-                jj = [Indices[j], Indices[j + 1]]
-                kk = [Indices[k], Indices[k + 1]]
-
-                SubCube = ImageArray[ii[0]:ii[1], jj[0]:jj[1], kk[0]:kk[1]]
-
-                BV = np.sum(SubCube)
-                TV = SubCube.shape[0] * SubCube.shape[1] * SubCube.shape[2]
[…]

### 2021-03-08 03_Scripts/09_Morpho_StatisticalTests.py

@@ -1,372 +1,0 @@
-# 00 Initialization
-import os
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy.stats.distributions import norm, f, t, chi2
-from scipy.stats import shapiro
-
-desired_width = 500
-pd.set_option('display.max_rows', 100)
-pd.set_option('display.max_columns', desired_width)
-pd.set_option('display.width', desired_width)
-np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
-plt.rc('font', size=12)
-
-## Define functions
-def BartlettTest(x,y):
-
-    # Analyze data
-    k = 2
-    Ni = np.array([len(x),len(y)])
-    s_i = np.array([np.std(x,ddof=1),np.std(y,ddof=1)])
-
-    N = Ni.sum()
-    S_pool = 0
-    for i in range(k):
-        S_pool += (Ni[i] - 1) * s_i[i] ** 2
-    S_pool = S_pool / (N-k)
-
-    # Perform test statistics
-    Numerator1 = (N-k) * np.log(S_pool)
-
-    Numerator2 = 0
-    for i in range(k):
-        Numerator2 += (Ni[i] - 1) * np.log(s_i[i]**2)
-
-    Denominator = 0
-    for i in range(k):
-        Denominator += 1/(Ni[i]-1)
-    Denominator = (Denominator - 1/(N-k)) / (3*(k-1)) + 1
-
-    X2 = (Numerator1 - Numerator2) / Denominator
-
-    # Compute p value
-    pValue = 1-chi2.cdf(X2,k-1)
-
-    return  X2, pValue
-def BrownForsytheTest(x, y):
-
-    # Data analysis
-    K = 2
-    Ni = np.array([len(x),len(y)])
-    N = Ni.sum()
-
-    # Transform data
-    Zi = np.abs(np.array([x-np.median(x),y-np.median(y)]))
-    Z_Bar = np.concatenate([Zi[0],Zi[1]]).mean()
-    Zj = np.array([Zi[0].mean(),Zi[1].mean()])
-
-    # Compute test result
-    Nominator, Denominator = 0, 0
-    for i in range(K):
-        Nominator += Ni[i] * (Zj[i]-Z_Bar) ** 2
-
-        for j in range(Ni[i]):
-            Denominator += (Zi[i][j]-Zj[i]) ** 2
-
[…]

### 2021-03-08 03_Scripts/CVThresholdStudy.py

@@ -1,0 +1,107 @@
+# 00 Initialization
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+DataFrame = pd.DataFrame()
+NewData = {'CV Limit':0.2,
+           'N':50,
+          'Grouped L0 m':3744,
+          'Grouped L0 u':3098,
+          'Grouped L0 o':4525,
+          'Healthy L0 m': 3584,
+          'Healthy L0 u': 3152,
+          'Healthy L0 o': 4075,
+          'OI L0 m': 3851,
+          'OI L0 u': 3363,
+          'OI L0 o': 4410}
+DataFrame = DataFrame.append(NewData,ignore_index=True)
+NewData = {'CV Limit':0.25,
+           'N':76,
+          'Grouped L0 m':4205,
+          'Grouped L0 u':3563,
+          'Grouped L0 o':4963,
+          'Healthy L0 m': 3934,
+          'Healthy L0 u': 3463,
+          'Healthy L0 o': 4468,
+          'OI L0 m': 4496,
+          'OI L0 u': 3997,
+          'OI L0 o': 5056}
+DataFrame = DataFrame.append(NewData,ignore_index=True)
+NewData = {'CV Limit':0.3,
+           'N':96,
+          'Grouped L0 m':4298,
+          'Grouped L0 u':3706,
+          'Grouped L0 o':4986,
+          'Healthy L0 m': 3969,
+          'Healthy L0 u': 3563,
+          'Healthy L0 o': 4421,
+          'OI L0 m': 4704,
+          'OI L0 u': 4153,
+          'OI L0 o': 5329}
+DataFrame = DataFrame.append(NewData,ignore_index=True)
+NewData = {'CV Limit':0.35,
+           'N':109,
+          'Grouped L0 m':4160,
+          'Grouped L0 u':3622,
+          'Grouped L0 o':4777,
+          'Healthy L0 m': 3858,
+          'Healthy L0 u': 3468,
[…]

### 2021-03-08 03_Scripts/LinearModels.py

@@ -1,0 +1,883 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import norm, t, f
+from scipy.stats import shapiro
+
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+## Define functions
+def ANCOVA(Models, ModelNames, ModelsType):
+
+    ModelsResults = pd.DataFrame()
+
+    for Model in Models:
+
+        Index = list(Models).index(Model)
+
+        ## Compute values for ANCOVA
+        Y_Obs = Model.model.endog
+        E = Model.resid.values
+        RSS = np.sum(E ** 2)
+        SE = np.sqrt(RSS / Model.df_resid)
+        TSS = np.sum((Y_Obs - Y_Obs.mean()) ** 2)
+        RegSS = TSS - RSS
+        R2 = RegSS / TSS
+
+        if ModelsType == 'Standard':
+            DOF = Model.df_model
+        elif ModelsType == 'Mixed-Effect':
+            DOF = Model.df_modelwc
+
+        ModelResults = {'Model': ModelNames[Index],
+                        'Regression SS': RegSS,
+                        'Residual SS': RSS,
+                        'df': DOF,
+                        'R2': round(R2, 3),
+                        'SE': round(SE,3)}
+        ModelsResults = ModelsResults.append(ModelResults, ignore_index=True)
+    print('\n\nResults of the fitting:')
+    print(ModelsResults[['Model', 'df', 'Regression SS', 'R2', 'SE']])
+
+    ## Build lists of models to compare
+    NModels = len(Models)
+    Models0 = []
+    Models1 = []
+    if NModels > 2:
+        ListA = np.arange(1,NModels).tolist()
[…]

### 2021-03-08 03_Scripts/TestPlots.py

@@ -1,0 +1,214 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import norm, t, f
+from scipy.stats import shapiro
+
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+# 01 Set path and build data
+WorkingDirectory = os.getcwd()
+DataFolder = os.path.join(WorkingDirectory, '04_Results/02_General_LinearRegression/')
+ResultsFolder = os.path.join(WorkingDirectory, '04_Results/XX_ArticlePlots/OI_CV_Gif/')
+
+## Load data
+HealthyData = pd.read_csv(DataFolder + '00_Healthy_StiffnessData.csv')
+OIData = pd.read_csv(DataFolder + '00_OI_StiffnessData.csv')
+
+## Set group to analyze
+Group = 'OI'   # Healthy or OI
+
+if Group == 'Healthy':
+    Data = HealthyData
+elif Group == 'OI':
+    Data = OIData
+
+
+# 02 Build linear system for regressions
+Y_Elements = Data.columns[2:14]
+
+LinearSystem = pd.DataFrame()
+for Index in Data.index:
+    m1, m2, m3 = Data[['m1','m2','m3']].loc[Index]
+    BVTV = Data['BV/TV'].loc[Index]
+    CV = Data['Variation Coefficient'].loc[Index]
+    X_i = np.array([[1, 0, 0, np.log(BVTV), np.log(m1 ** 2), np.log(CV)],
+                    [0, 1, 0, np.log(BVTV), np.log(m1 * m2), np.log(CV)],
+                    [0, 1, 0, np.log(BVTV), np.log(m1 * m3), np.log(CV)],
+                    [0, 1, 0, np.log(BVTV), np.log(m2 * m1), np.log(CV)],
[…]

### 2021-04-08 09_Article/Article.tex
First article draft finished

%% 
%% Copyright 2019-2020 Elsevier Ltd
%% 
%% This file is part of the 'CAS Bundle'.
%% --------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'CAS Bundle' is
%% given in the file `manifest.txt'.
%% 
%% Template article for cas-dc documentclass for 
%% double column output.

%\documentclass[a4paper,fleqn,longmktitle]{cas-dc}
\documentclass[a4paper,fleqn]{DC_ArtStyle}

%\usepackage[authoryear,longnamesfirst]{natbib}
%\usepackage[authoryear]{natbib}
\usepackage[numbers]{natbib}
\usepackage{lipsum}
\usepackage{xcolor}
\usepackage[justification=centering]{caption}
\usepackage{subcaption}
\usepackage{siunitx}
\usepackage{array}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{float}
\usepackage{multicol}

\newcommand{\abbreviations}[1]{%
	\nonumnote{\textit{Abbreviations:\enspace}#1}}

\begin{document}
\let\WriteBookmarks\relax
\def\floatpagepagefraction{1}
\def\textpagefraction{.001}
\shorttitle{Fabric-Elasticity in Osteogenesis Imperfecta}
\shortauthors{Simon et~al.}

\title[mode = title]{Fabric-Elasticity Relationships in Osteogenesis Imperfecta (OI)}

% Autors
\author[1]{Mathieu Simon}
\ead{mathieu.simon@artorg.unibe.ch}

\author[1]{Michael Indermaur}

\author[1]{Denis Schenk}

\author[2,3]{Bettina Willie}

\author[1]{Philippe Zysset}

% Adresses
\address[1]{ARTORG Centre for Biomedical Engineering Research, University of Bern, Bern, Switzerland}

\address[2]{Shriners Hospital for Children, Montreal, Canada}

\address[3]{McGill University, Montreal, Canada}

% Abbreviations
\abbreviations{OI, osteogenesis imperfecta;
			   HR-pQCT, high resolution peripheral quantitative tomography;
			   BMD, bone mass density;
			   ROI, region of interest;
			   BV/TV, bone volume over total volume;
			   SMI, structure model index;
			   MIL, mean intercept length;
			   FEA, finite element analysis
			   KUBC, kinematic uniform boundary condition;
			   PMUBC, periodicity-compatible mixed uniform boundary condition.}

% Footnotes

\begin{abstract}
	Osteogenesis imperfecta (OI) is the most common inherited form of bone fragility. This disease, also called "brittle bone disease", is characterized by impaired synthesis of type I collagen leading to fragile bones. In this study, 50 OI diagnosed people were compared to 120 healthy controls using second generation high resolution peripheral computed tomography (HR-pQCT) scans of the distal tibia. Such scans are usually used in finite element analysis (FEA) for bone strength estimation and fracture risk assessment. The homogenized FE (hFE) scheme uses HR-pQCT bone volume fraction (BV/TV) and anisotropy information (fabric) and allows to assess bone strength within a reasonable computation time using fabric-elasticity relationships. Thus, the aim of this study is to investigate fabric-elasticity relationships in OI compared to healthy controls. For this, six regions of interest (ROIs) were selected per individual in a common anatomical region. A first age \& gender matching allowed to perform a morphometric analysis and compare the samples with literature. Then, stiffness tensors of ROIs was computed by FEA and multiple linear regression was performed on the Zysset-Curnier orthotropic model to compare groups using 5 parameters. This fit was first executed using the complete data sets separately. Then, data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity. A second fit was performed on these filtered data sets and a third fit was done on BV/TV \& fabric anisotropy (DA) matched data sets. This last matched data set was grouped to compare with literature \si{\micro}CT scans. Morphometric analysis shown results in agreement with literature, namely significant differences between groups for BV/TV, trabecular number (Tb N), trabecular spacing (Tb Sp), trabecular spacing standard deviation (Tb Sp SD) but no significant differences for trabecular thickness (Tb Th). The multiple linear regressions shown that OI stiffness range reach lower values than healthy controls and the fit tends of overestimate the stiffness at lower values. Filtering out high heterogeneous ROIs removes low stiffness ROIs and lead to similar fit quality for both OI and healthy group. The BV/TV \& DA matched data sets shown no significant differences between OI and healthy individuals and stiffness constants are lower at 61 \si{\micro}m resolution HR-pQCT scans than for higher resolution 36 \si{\micro}m \si{\micro}CT scans. Our main conclusions are that lower scan resolution lead to lower stiffness because the structure obtained is not as optimized as in reality. Finally, if trabecular bone is homogeneous enough ROIs inside ROI, there no reason to suppose different fabric-elasticity relationships between OI and healthy individual.
\end{abstract}

\begin{keywords}
Bone \sep
Elasticity \sep
Fabric \sep
Osteogenesis Imperfecta
\end{keywords}

\maketitle

\section{Introduction}

Osteogenesis imperfecta (OI) is the most common inherited form of bone fragility \cite{Tournis2018}. It was estimated to concern about 1/13'500 births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}. Therefore, OI is considered as a rare metabolic bone disorder. Sillence classification \cite{Sillence1979} allows to define four main types of OI based on the clinical phenotype:
\begin{itemize}
	\item Type I: less severe
	\item Type II: lethal at most shortly after birth
	\item Type III: most severe surviving form
	\item Type IV: intermediate severity
\end{itemize}
Also known as "brittle bone disease", OI is characterized by impaired synthesis of type I collagen, leading to such brittle and fragile bones \cite{LIM2017}.\\

Bone fragility in OI is complex and not totally understood. High resolution peripheral quantitative tomography (HR-pQCT) scans allow to perform \textit{in vivo} assessment of trabecular architecture and volumetric BMD in distal radius and tibia \cite{Boutroy2005}. Moreover, the bony microstructure obtained from HR-pQCT can be used for finite element analysis (FEA) to predict mechanical properties \cite{Boutroy2008}. Homogenized FE (hFE) is a FEA scheme including volume fraction (BV/TV) and anisotropy information (fabric) of trabecular bone from HR-pQCT which to assess bone strength within reasonable computation time \cite{Pahr2009}. High correlations were found between patient-specific hFE and mechanical compression experiment of cadaveric samples \cite{Varga2011,AriasMoreno2019}. Thus, it could be legitimate to use hFE for OI patients bone strength estimation and fracture risk assessment with the assumption of identical fabric-elasticity relationships between OI and healthy individuals. Therefore, the present study aims to compare healthy with OI diagnosed individuals and to draw conclusions about fabric-elasticity relationships.

\section{Methods}

\subsection{Subjects}
The healthy group include a total of 120 patients from a previous reproducibility study \cite{Schenk2020}. The sample is composed of 64 female and 56 male subjects aged between 20 and 92 years old with a median age of 26 [22 - 35] years. These subjects did not took any medication known to affect bone metabolism nor presented any prior osteoporosis fracture. The second group is composed of 35 female 15 male individuals leading to 50 OI diagnosed subjects. The youngest and oldest patients are 19 and 69 year old, respectively. The median age is 44 [33 - 55] years. 35 subjects were diagnosed with type I OI, 2 with type III, and 13 with type IV.

\subsection{HR-pQCT}
HR-pQCT scans (XtremeCTII, SCANCO Medical, Brütisellen,
Switzerland) were performed at the distal tibia on all patients from both groups. The OI subjects were scanned using the manufacturer's standard protocol. These scans were performed at the Shriners Hospital for Children and were shared to our group by the McGill University in Montreal. People of the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}. The main differences with the manufacturer's standard protocol are the following:
\begin{enumerate}
	\item The reference line is positioned at the proximal margin of the dense structure formed by the tibia plafond instead of  the subchondral endplate of the ankle joint (standard clinical section) \cite{Whittier2020}.
	\item Three stacks were scanned proximal to the reference line instead of one stack at 22.5 mm proximal to the reference line for standard clinical section \cite{Whittier2020}.
\end{enumerate}

These differences are shown in Figure \ref{01_ClinicalSections}. For both group, each stack consist of 168 voxels and a resolution of 61 \si{\micro}m in the three principal directions which lead to a thickness about 10.2 mm for each stack. Scanning settings were a voltage of 60 kVp, 900 μA, 100 ms integration time for the healthy group as well as for the OI group.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.225\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/01_OIClinicalSection}
		\caption{OI group}
		\label{01_OI}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.225\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/01_ControlClinicalSection}
		\caption{Healthy group}
		\label{01_Healthy}
	\end{subfigure}
	\caption{Clinical section scanned for both group}
	\label{01_ClinicalSections}
\end{figure}

\subsection{Image analysis}
The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, the segmented images were used for further analysis.\\

Six ROI were randomly selected in each scan. The conditions for ROI to be kept are no cortical bone inside and that it must contain trabecular bone. For the OI subjects, the stack was divided into two halves and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half. For the healthy people, the ROIs were selected in the more proximal stack uniquely, see Figure \ref{01_Healthy} stack n\textdegree 3. As for the OI subjects, the stack is halved and centers of three ROIs were selected in both half.\\

The ROI is defined as a cube of about 5.3 mm side length. This size was chosen to correspond to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013}. It was determined by \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to be the optimal size to obtain accurate FEA results with minimal computation efforts.\\

The ROI analysis was performed using medtool (v4.5; Dr. Pahr Ingenieurs e.U., Pfaffstätten, Austria). The morphological parameters analyzed are: BV/TV, SMI, trabecular number (Tb. N.), trabecular thickness (Tb. Th.), trabecular spacing (Tb. Sp.), and the standard deviation of the trabecular spacing (Tb. Sp. SD). Moreover, ROI fabric was evaluated. The fabric tensor $\mathbf{M}$ is a positive-definite second-order tensor. It is build as shown in Equation \ref{Eq201} below:

\begin{equation}
	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i}
	\label{Eq201}
\end{equation}

where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ \cite{Cowin1985,Harrigan1985}. The fabric is then independent of BV/TV and normalized with $tr(\mathbf{M}) = 3$. To evaluate this tensor, the MIL method was used \cite{Moreno2014}. The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI.\\

A mechanical analysis was performed using \textsc{ABAQUS 6.14}. Each voxel of the segmented ROI was converted to mesh using fully integrated linear brick elements (C3D8) with a stiffness $E$ of 10 000 MPa and a Poisson's ratio $\nu$ of 0.3. The simulation consisted of 6 independent load cases, 3 uni-axial and 3 shear cases, using KUBCs. KUBCs were used according to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} were it was determined to be better than PMUBC to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}. This model builds the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV $\rho$, fabric information $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$. The building of this tensor is shown in Equation \ref{Eq202}.

\begin{equation}
	\begin{split}
		\mathbb{S}(\rho,\mathbf{M}) & = \sum_{i=1}^{3} (\lambda_0 + 2\mu_0)\rho^k m_i^{2l} \mathbf{M}_i \otimes \mathbf{M}_i \\ & + \sum_{\substack{i,j=1\\i \neq j}}^{3} \lambda_0' \rho^k m_i^{l} m_j^{l} \mathbf{M}_i \otimes \mathbf{M}_j \\ & + \sum_{\substack{i,j=1\\i \neq j}}^{3} \mu_0 \rho^k m_i^{l} m_j^{l} \mathbf{M}_i \overline{\underline{\otimes}} \mathbf{M}_j
	\end{split}
	\label{Eq202}
\end{equation}

Where $\otimes$ and $\overline{\underline{\otimes}}$ are the dyadic and symmetric product of second order tensors, respectively. The Zysset-Curnier model is built with the assumption of orthotropy. However, the trabecular structure is not perfectly orthotropic. In order to assess the ROI heterogeneity a so-called coefficient of variation (CV) is computed as presented in \cite{Panyasantisuk2015}: the ROI is divided into eight identical subcubes, BV/TV is computed for each subcube and the CV is defined as the ratio between the standard deviation of these BV/TV and the mean value, see Equation \ref{Eq203} below:

\begin{equation}
	CV = \frac{std(BV/TV_{subcubes})}{mean(BV/TV_{subcubes})}
	\label{Eq203}
\end{equation}

\subsection{Statistics}
The height morphological parameters analyzed (BV/TV, Tb. N., Tb. Th., Tb. Sp., and Tb. Sp. SD, SMI, DA, and CV) were compared for both group. As the initial groups do not have similar distributions in quality of age and sex, a matching was performed leading to identical mean and median age as well as identical gender distribution. Then, the selection of statistical test to perform was executed as follow:
\begin{enumerate}
	\item For each parameter, the median value between the six ROI of the same individual was computed. The median was preferred over the mean because it is less influenced by outliers.
	\item Normality of the distribution was assessed with QQ plot and Shapiro-Wilk test.
	\item If normality assumption was met, Bartlett test for equal variances was performed. Otherwise, data were log transformed to try to achieve normal distribution. If even after log transformation, normality assumption was not met, original data space was kept and Brown-Forsythe test was applied to assess the equal variance assumption.
	\item According to the previous results, t-test was performed if normal distribution and equal variance were met. If only the equal variances assumption was met, Mann-Whitney test was preferred. Finally, if none of these conditions could be assumed, a non-parametric permutation test was performed.
\end{enumerate}
The general significance level was set to 95\% for all tests. Confidence intervals in means difference was computed for t-tested variables to confirm p values. As Mann-Whitney test is performed on the median, only corresponding p value is given. Finally, non-parametric permutation test is less powerful but give an empirical 95\% exclusion range and a p value. If the difference in means belong to this exclusion range, it can be stated that group means are different with 95\% certainty.

\subsection{Fit to model}
The stiffness tensors obtained from the mechanical simulations were transformed into fabric coordinate system and projected onto orthotropy, leading to 12 values. The resulting orthotropic stiffness tensors were then used to perform a multiple linear regression on the Zysset-Curnier model. Standard linear models assume independent and identically distributed (iid) variables. As this assumption is violated by the fact that six ROIs are analyzed by individual, a linear mixed-effect model was preferred. This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, takes into account the non-independence of ROIs from the same individual.

\begin{equation}
	y = X \beta + Z \delta + \epsilon \quad \text{with} \quad y = \ln(S_{rc})
	\label{Eq204}
\end{equation}

Where $S_{rc}$ is the $r$th row and $c$th column of the non-zero element of the orthotropic stiffness tensor $\mathbb{S}$ in Mandel notation \cite{MANDEL1965}, $X$ is a $12n$x$p$ design matrix containing the the BV/TV and fabric info of the $n$ ROIs and $\beta$ is a $p$x1 vector of fixed effects containing model parameters. $Z$ is a $12n$x$f$ design matrix which contains data with individual dependence and $\delta$ is a $f$x1 vector composed of random factors. Finally, $\epsilon$ is a $12n$x1 vector containing the regression residuals.\\

The linear regression was performed on both group (healthy and OI) separately. To improve the fit quality, the data sets were filtered. The aim here is to filter out ROIs whose are too far from the assumption of orthotropy. Therefore, analogously to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, a fixed threshold for the CV was used. To simplify comparison, the same value of 0.263 was fixed as exclusion criterion. Besides, the relation between BV/TV and CV was assessed using Spearman's correlation coefficient. Furthermore, to compare the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) between the groups, regression must be performed on identical value ranges. To do this, a matching was performed for BV/TV and DA to find corresponding control ROI for each OI in the filtered groups. Best correspondences were kept and duplicates were dropped. Finally, as the regression is performed in the log space, it is necessary to use identical exponent ($k$ and $l$) for both group, weighting identically BV/TV and DA between regressions, to compare stiffness constants. The exponents were determined by grouping healthy and OI for regression. Then a modified system is used to perform the fit on separated groups.\\

Another modification of the model is to add a regressor for the group variable (healthy or OI), i.e. add a column to the design matrix $X$ and a row to the parameter vector $\beta$. This modified model is compared to the original by analysis of covariance (ANCOVA) using the fixed-effects only to determine the statistical significance of the group. Implementation of this modification was performed according to \cite{Fox2016}. The detailed linear systems for each model discussed here are available in Appendix \ref{A1} and a summary of the data sets used for the different methods is shown in Table \ref{Table1}.\\

The regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}. Regression quality was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors ($NE$), see Equation \ref{Eq205} and \ref{Eq206}.

\begin{equation}
	R^2_{adj} = 1 - \frac{RSS}{TSS} \frac{(12n-1)}{(12n - p - 1)}
	\label{Eq205}
\end{equation}

Where RSS is the residual sum of squares and TSS is the total sum of squares i.e. sum of the square of the observations y.

\begin{equation}
	NE = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
	\label{Eq206}
\end{equation}

\begin{table*}[b]
	\centering
	\caption{Summary of the data set used for different methods}
	\label{Table1}
	\begin{tabular}{p{0.1\linewidth}*{2}{>{\centering\arraybackslash}p{0.075\linewidth}}*{2}{>{\centering\arraybackslash}p{0.075\linewidth}}*{2}{>{\centering\arraybackslash}p{0.075\linewidth}}*{2}{>{\centering\arraybackslash}p{0.075\linewidth}}}
		\toprule
		Data sets & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Age \& gender matched} & \multicolumn{2}{c}{CV filtered} & \multicolumn{2}{c}{BV/TV \& DA matched} \\
		\midrule
		Group & Healthy & OI & Healthy & OI & Healthy & OI & Healthy & OI \\
		Individuals & 120 & 50 & 28 & 28 & 119 & 38 & 58 & 33 \\
		ROIs & 720 & 300 & 168 & 168 & 603 & 115 & 83 & 83 \\
		\midrule
		Methods & \multicolumn{2}{c}{Fit to model} & \multicolumn{2}{c}{Statistics} & \multicolumn{2}{c}{Fit to model} & \multicolumn{2}{c}{Fit to model} \\
		\bottomrule
	\end{tabular}
\end{table*}

\section{Results}

\subsection{Morphological Analysis}
The results of morphological analysis are summarized in Table \ref{Table2}. In the present study, the individual matching allowed to have similar group distribution with 17 females and 11 males in each group. The mean age of matched healthy is 41 $\pm$ 14 and 41 $\pm$ 15 for the matched OI. BV/TV of healthy people is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01. Similarly, trabecular number is higher in the matched healthy group with a 95\% CI of [0.099, 0.285] and a corresponding p value <0.001. The trabecular thickness do not show significant differences between groups with a p value of 0.2. In the other hand, trabecular spacing is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$). Trabecular spacing SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$). SMI as well as degree of anisotropy are higher for matched OI than for healthy people with p values <0.001 and of 0.02, respectively. Finally, the log transformation of coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333].\\

In the table, absolute values and p values of test statistics are compared with literature. Regarding age, the present population is slightly younger than in other study but stay in a similar range. In a general way p values of the three other studies show significant differences for BV/TV, Tb N, Tb Sp, and Tb Sp SD and non significant differences for Tb Th. Absolute values of BV/TV, TB Th, Tb Sp, and Tb Sp SD are higher in the present study compared to literature. On the other hand, Tb N appears to be higher in the other studies than in the present one.

\begin{sidewaystable*}
	\centering
	\caption{Summary of morphological analysis and comparison with literature. Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians. The study of \cite{Kocijan2015} presents n.s. for non-significant p value test result.}
	\label{Table2}
	\begin{tabular}{cccccccccc}
		\toprule
		\multirow{2}{*}{Variable} & \multirow{2}{*}{Group} & \multicolumn{2}{c}{Present study} & \multicolumn{2}{c}{\citeauthor{Folkestad2012}\cite{Folkestad2012}} & \multicolumn{2}{c}{\citeauthor{Kocijan2015}\cite{Kocijan2015}} & \multicolumn{2}{c}{\citeauthor{Rolvien2018}\cite{Rolvien2018}} \\
		& & Values & p value & Values & p value & Values & p value & Values & p value \\
		\midrule

		\multirow{3}{*}{Age} & Healthy & 41 $\pm$ 14 &  & 54 (21-77) & & 44 (38-52) &  & 49 $\pm$ 16 &  \\
		& OI Type I & \multirow{2}{*}{41 $\pm$ 15} &  & 53 (21-77) &  & 42 (35-56) & & \multirow{2}{*}{46 $\pm$ 16} & \\
		&  OI Type III \& IV & &  &  &  & 48 (35-58) & & & \\[3ex]

		\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
		& OI Type I & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114) & <0.0001 & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
		&  OI Type III \& IV & & & & & 0.081 (0.056-0.092) & <0.0001 & & \\[3ex]

		\multirow{3}{*}{Tb N} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
		& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
		&  OI Type III \& IV & & & & & 0.89 (0.81-1.08) & <0.001 & & \\[3ex]

		\multirow{3}{*}{Tb Th} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
		& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
		&  OI Type III \& IV & & & & & 0.078 (0.068-0.092) & n.s. & & \\[3ex]

		\multirow{3}{*}{Tb Sp} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
		& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
		&  OI Type III \& IV & & & & & - &  & & \\[3ex]

		\multirow{3}{*}{Tb Sp SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
		& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
		&  OI Type III \& IV & & & & & 0.698 (0.511-0.890) & <0.001 & & \\[3ex]

		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & - &  & - & & - &  \\
		& OI Type I & \multirow{2}{*}{0.056 (0.015-0.079)} &  & - &  & - &  & - & \\
		&  OI Type III \& IV & & & & & - &  & & \\[3ex]

		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & - &  & - & & - &  \\
		& OI Type I & \multirow{2}{*}{2.018 (1.901-2.158)} &  & - &  & - &  & - & \\
		&  OI Type III \& IV & & & & & - &  & & \\[3ex]

		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & - &  & - & & - &  \\
		& OI Type I & \multirow{2}{*}{-1.178 $\pm$ 0.441} &  & - &  & - &  & - & \\
		&  OI Type III \& IV & & & & & - &  & & \\

		\bottomrule
	\end{tabular}
\end{sidewaystable*}

\subsection{Linear Regression}
Figure \ref{02_GeneralRegression} shows the result of regression for complete data sets of each group separately. The x axis represent the values of observed tensors from mechanical simulations. The y axis is the predicted values using the Zysset-Curnier model and the parameters obtained after performing the regression with linear mixed-effect model. The fitted line is represented by the dashed line. $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components. For the healthy group, see Figure \ref{02_Healthy}, the fit is performed on 720 ROIs leading to 8640 points. The $R^2_{adj}$ is slightly above 0.95 and the NE is of 18\% $\pm$ 10\%. The regression of the OI group could only be performed on 294 ROIs as it was not possible to find suitable ROIs for one patient. Nevertheless, it leads to 3528 points, a $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%. It can be noticed that, as the values of observed tensor decreases, points tends to be further apart of the diagonal (dashed line). Moreover, the points coming from the tensors with lowest values are only above the diagonal. One last observation which should be noted is the range covered by the observed tensors. The range of the OI group is wider than the healthy group and goes further in low observed values.\\

\begin{figure}[h!]
	\centering
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/02_GR_Healthy_LMM}
		\caption{Healthy group}
		\label{02_Healthy}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/02_GR_OI_LMM}
		\caption{OI group}
		\label{02_OI}
	\end{subfigure}
	\caption{Regression results using the fixed effects of the linear mixed-effect model on original data sets. $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components. The dashed line represents the fitted line.}
	\label{02_GeneralRegression}
\end{figure}

The CV in relation to BV/TV is shown in Figure \ref{02_CV_BVTV}. The OI data reach higher values of CV and stay at lower values of BV/TV than healthy data. Generally, the CV tends to increase with decreasing BV/TV. The Spearman coefficient is shown above the plot as value [95\% CI]. Its value is negative and strictly different from zero. Finally, the CV threshold value used to filter the data is represented by the dashed line. It can be observed that a relatively important part of OI data will be filtered out. On the other hand, relatively few healthy data will be removed by the filtering. Examples of extreme ROIs in terms of CV and BV/TV are available in Appendix \ref{A2}.\\

\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]
	{Pictures/03_CV_BVTV}
	\caption{Coefficient of variation in relation to BV/TV. Spearman correlation coefficient $\rho$ assess monotonic relation between two variable}
	\label{02_CV_BVTV}
\end{figure}

The regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}. After filtering, healthy group is reduced to 119 individuals and 603 ROIs. This leads to 7236 points for the regression. Results shown similar values as for the complete data set, namely a $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\%, see Figure \ref{04_Healthy}. The OI group lost more individuals after filtering leading to 38 people and 115 ROIs. Regression is performed on 1380 points then. Figure \ref{04_OI} presents results with a $R^2_{adj}$ rounded to 0.95 and a NE of 17\% $\pm$ 8\%.\\

\begin{figure}[h!]
	\centering
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/04_FR_Healthy_LMM}
		\caption{Healthy group}
		\label{04_Healthy}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=\textwidth]
		{Pictures/04_FR_OI_LMM}
		\caption{OI group}
		\label{04_OI}
	\end{subfigure}
	\caption{Regression results using the fixed effects of the linear mixed-effect model on filtered data sets. $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components. The dashed line represents the fitted line.}
	\label{04_FilteredRegression}
\end{figure}

Regression results after BV/TV \& DA ROI matching are shown in Table \ref{Table3}. The columns show which data set was used, the fives parameters of the Zysset-Curnier model and the assessment of fit quality. Grouping healthy and OI data together for regression lead to a $k$ of 1.91 and a $l$ of 0.95. Regression result shows a $R^2_{adj}$ of 0.94 and a NE of 18\% $\pm$ 9\%. Second and last rows show regression results using separated data sets and imposing the exponents $k$ and $l$. OI values are higher than healthy one. The increase is of 15\%, 1\%, and 2\% for $\lambda_0$, $\lambda_0'$, and $\mu_0$, respectively. The ANCOVA performed to quantify the group statistical significance shows a p value of 0.7.\\

\begin{table*}[b]
	\caption{Constants comparison between grouped and separated data sets. Values are presented as value [95\% CI] or mean $\pm$ standard deviation.}
	\label{Table3}
	\begin{tabular}{cccccccc}
		\toprule
		Data set & $\lambda_0$ & $\lambda_0'$ & $\mu_0$ & $k$ & $l$ & $R^2_{adj}$ & NE (\%) \\
		\midrule
		Grouped & 4626 [3902-5485] & 2692 [2473-2932] & 3537 [3246-3855] & 1.91 [1.86-1.95] & 0.95 [0.93-0.97] & 0.936 & 18 $\pm$ 9\\

		Healthy & 4322 [3855-4846] & 2684 [2535-2842] & 3509 [3307-3724] & 1.91 & 0.95 & 0.836 & 21 $\pm$ 10\\

		OI & 4980 [4359-5689] & 2722 [2547-2909] & 3594 [3355-3849] & 1.91 & 0.95 & 0.861 & 20 $\pm$ 10\\
		\bottomrule
	\end{tabular}
\end{table*}

Table \ref{Table4} shows results obtained compared to literature. \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs. Data sets of \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} show BV/TV ranges slightly higher than in the present study and the one of \citeauthor{Gross2013} \cite{Gross2013}. On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}. Setting the exponents $k$ and $l$ to the same values lead to lower stiffness constants for the observed data set than for the other studies.\\

\begin{table*}[b]
	\caption{Comparison with literature. N stands for the number of observations. Values are presented as computed value only or mean $\pm$ standard deviation.}
	\label{Table4}
	\begin{tabular}{lcccccccccc}
		\toprule
		Data set & N & BV/TV & DA & $\lambda_0$ & $\lambda_0'$ & $\mu_0$ & $k$ & $l$ & $R^2_{adj}$ & NE (\%) \\
		\midrule
		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 126 & 0.27 $\pm$ 0.08 & 1.57 $\pm$ 0.18 & 3306 & 2736 & 2837 & 1.55 & 0.82 & 0.984 & 8 $\pm$ 3\\

		Present study & 166 & 0.21 $\pm$ 0.08 & 2.03 $\pm$ 0.21 & 2559 & 1490 & 1957 & 1.55 & 0.82 & 0.815 & 25 $\pm$ 11\\[3ex]

		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 167 & 0.25 $\pm$ 0.08 & 1.54 $\pm$ 0.20 & 3841 & 3076 & 3115 & 1.60 & 0.99 & 0.983 & 14\\

		\citeauthor{Gross2013} \cite{Gross2013} & 264 & 0.19 $\pm$ 0.10 & 1.67 $\pm$ 0.34 & 4609 & 3692 & 3738 & 1.60 & 0.99 & 0.981 & 14\\

		Present study & 166 & 0.21 $\pm$ 0.08 & 2.03 $\pm$ 0.21 & 2817 & 1640 & 2154 & 1.60 & 0.99 & 0.826 & 22 $\pm$ 12\\
		\bottomrule
	\end{tabular}
\end{table*}

\section{Discussion \& Conclusions}
The age of matched individuals lies in the spectrum of the others studies which allow to compare the morphological values. The main explanation for differences between the absolute values of the present study compared to the others lies in the imaging system. \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed first generation Xtreme CT scans and the present study use image from XCTII. The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} showed that BV/TV, Tb Th, and Tb Sp is higher in second generation Xtreme CT and, on the other hand, Tb N is higher in first generation Xtreme CT. These results give confidence on the observed values. Another bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length. This conditions the Tb N and Tb Sp as they depend on the ROI size. Moreover, conditions imposed for ROI random selection can lead to further biased values, specially of OI patient, as it can not be empty of trabecular bone. The CV presenting the stronger significant difference between groups even given the low sample size show that heterogeneity is a main difference in OI patient compared to healthy individuals. Finally, the significant differences observed in BV/TV and DA even with age \& gender matched individuals justify the choice of a variable matching for fabric-elasticity relationships analysis.\\

Regression performed on original data sets shows $R^2_{adj}$ and NE in the expected range for the healthy group. Components of the stiffness tensors are distributed to both sides of the diagonal. On the other hand, regression of the OI data set presents lower $R^2_{adj}$ than such fit reach usually. The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation. These differences come from the ROIs presenting a low stiffness. It can be seen on the regression plot that when the stiffness term decrease about $10^0$ and under, the fit tends to overestimate the stiffness. This is due to the fact that ROI stiffness is highly impacted by BV/TV. Some low BV/TV ROIs do not have every side of the cube connected by bone leading to extremely low terms in the stiffness tensor. Trying to homogenize such ROI can lead to error of multiple order of magnitude, as observed on the plot. Therefore, a filtering is justified to compare fabric-elasticity relationships.\\

Figure \ref{02_CV_BVTV} presenting the CV in relation to BV/TV shows that it exist a tendency of CV to increase with a decreasing BV/TV. Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly becomes highly heterogeneous. A simple assumption about this relation is that it could be monotonic. Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation. As some higher BV/TV ROIs still present high CV, imposing a fixed threshold make sense for subsequent homogenization.\\

The fits performed on filtered data sets present direct effect of filtering. For the healthy group, the relatively small decrease of $R^2_{adj}$ (5\textperthousand) is negligible. On the other hand, NE presents an improved value decreasing by 2\%. These results are due to the filtering of point further away from the diagonal (better NE) and some points close to the diagonal leading to a smaller number of points ($R^2_{adj}$). For the OI group, filtering leads to a important improvement of the fit. Results become similar to the healthy group in terms of $R^2_{adj}$, NE, and the range of stiffness values. These results give confidence to the filtering procedure and are a first good point to the hypothesis of no different fabric-elasticity relationships between healthy and OI trabecular bone.\\

After BV/TV and DA matching, grouping the data sets together lead to similar $R^2_{adj}$ and NE as for individual filtered data set. This allows to give values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m. Imposing these values to perform the fit on individual match data set allow to highlight differences, if any, between healthy and OI trabecular bone. The relatively low differences for $\lambda_0'$ and $\mu_0$ once again provide arguments for similar relationships between the two groups. For $\lambda_0$ this relative difference being higher could rise some doubts about this similarity but 95\% CI intervals still show a common range. Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group shown a p value far above the 5\% significance level. With this statistical non-significance of the group and low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to suppose differences in fabric-elasticity relationships between healthy and OI diagnosed people. In FEA simulations, it is not possible to exclude part of the mesh because of high heterogeneity. Nevertheless, the error created by such ROIs can be negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\

Imposing $k$ and $l$ allows to estimate the effect of different image resolutions.  \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used scans with 18 \si{\micro}m spatial resolution and coarsened it to 36 \si{\micro}m. \citeauthor{Gross2013}\cite{Gross2013} shown that different anatomical locations lead to only slight differences. Comparing regression of filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the lower BV/TV and higher DA ranges and by the coarser resolution. Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones. Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges are well overlapped. DA is still higher in the present study and the stiffness constants remain lower than for the two other studies. Here, differences can be explained by resolution. The coarser structure resulting from XCTII can not be as optimized as the fine detailed structure obtained by \si{\micro}CT. Performing a fit on this less optimized structure lead to the observed lower stiffness constants. Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} show that lower spatial resolution lead to lower fit quality. Nevertheless, $l$ stands in the same range as for the two other studies meaning the relative weight of DA remains constant. On the other hand, the higher $k$ highlight an increased relative weight of BV/TV.\\

In conclusions, this work shown that, if the ROI is homogeneous enough, there no reason to suppose any differences in fabric-elasticity relationships between healthy and OI trabecular bone. The main limitations of this study are the definition of "homogeneous enough" and the fact that it is limited to tibiae XCTII scans. More investigations should be performed to improve the model for highly heterogeneous ROIs but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible. An idea could be to compute a mass gradient instead of CV to improve the model. Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution. Finally, it could be interesting to perform homogenization and fitting on identical samples using \si{\micro}CT and XCTII to quantify with more accuracy the resolution impact.

\section*{Acknowledgments}
The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the buildingi and interpretation of linear mixed-effect model. This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.

\appendix
\section{Linear Models}\label{A1}

The standard linear model has the form:

\begin{equation}
	\ln(S_{rc}) = X \beta + \epsilon
\end{equation}

Where $\epsilon$ is the vector of residuals. For one ROI, the system take the following form:

\begin{equation}
	\ln
	\begin{pmatrix}
		S_{11} \\
		S_{12} \\
		S_{13} \\
		S_{21} \\
		S_{22} \\
		S_{23} \\
		S_{31} \\
		S_{32} \\
		S_{33} \\
		S_{44} \\
		S_{55} \\
		S_{66} \\
	\end{pmatrix} = \begin{pmatrix}
		1 & 0 & 0 & \ln(\rho) & \ln(m_1^2) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_1 m_2) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_1 m_3) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_2 m_1) \\
		1 & 0 & 0 & \ln(\rho) & \ln(m_2^2) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_2 m_3) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_3 m_1) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_3 m_2) \\
		1 & 0 & 0 & \ln(\rho) & \ln(m_3^2) \\
		0 & 0 & 1 & \ln(\rho) & \ln(m_2 m_3) \\
		0 & 0 & 1 & \ln(\rho) & \ln(m_3 m_1) \\
		0 & 0 & 1 & \ln(\rho) & \ln(m_1 m_2) \\
	\end{pmatrix} \begin{pmatrix}
		\ln(\lambda^{*}) \\
		\ln(\lambda_0') \\
		\ln(\mu_0) \\
		k \\
		l \\
	\end{pmatrix} + \begin{pmatrix}
		\epsilon_{1} \\
		\epsilon_{2} \\
		\epsilon_{3} \\
		\epsilon_{4} \\
		\epsilon_{5} \\
		\epsilon_{6} \\
		\epsilon_{7} \\
		\epsilon_{8} \\
		\epsilon_{9} \\
		\epsilon_{10} \\
		\epsilon_{11} \\
		\epsilon_{12} \\
	\end{pmatrix}
\end{equation}

Where $\lambda^{*} = \lambda_0 + 2\mu_0$. Then, the mixed-effect model, which handles multiple measurement on the same individual, has the following general form:

\begin{equation}
	\ln(S_{rc}) = X \beta + Z \delta + \epsilon
\end{equation}

Where $Z$ is a design matrix composed of the observations which are correlated on the same individual and, in general, is a subset of $X$. In the present case, the stiffness variables ($\lambda_0$, $\lambda_0'$, and $\mu_0$) can vary between individuals but the hypothesis is that they all vary by an identical factor. Therefore, the design matrix $Z$ is composed of the addition of the three first columns of $X$ and the system for one ROI takes the following form:

\begin{equation}
	\begin{split}
	\ln
	\begin{pmatrix}
		S_{11} \\
		S_{12} \\
		S_{13} \\
		S_{21} \\
		S_{22} \\
		S_{23} \\
		S_{31} \\
		S_{32} \\
		S_{33} \\
		S_{44} \\
		S_{55} \\
		S_{66} \\
	\end{pmatrix} & = \begin{pmatrix}
		1 & 0 & 0 & \ln(\rho) & \ln(m_1^2) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_1 m_2) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_1 m_3) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_2 m_1) \\
		1 & 0 & 0 & \ln(\rho) & \ln(m_2^2) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_2 m_3) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_3 m_1) \\
		0 & 1 & 0 & \ln(\rho) & \ln(m_3 m_2) \\
		1 & 0 & 0 & \ln(\rho) & \ln(m_3^2) \\
		0 & 0 & 1 & \ln(\rho) & \ln(m_2 m_3) \\
		0 & 0 & 1 & \ln(\rho) & \ln(m_3 m_1) \\
		0 & 0 & 1 & \ln(\rho) & \ln(m_1 m_2) \\
	\end{pmatrix} \begin{pmatrix}
		\ln(\lambda^{*}) \\
		\ln(\lambda_0') \\
		\ln(\mu_0) \\
		k \\
		l \\
	\end{pmatrix}\\ & + \begin{pmatrix}
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
		1 \\
	\end{pmatrix}\begin{pmatrix}
	\delta \\
	\end{pmatrix} + \begin{pmatrix}
		\epsilon_{1} \\
		\epsilon_{2} \\
		\epsilon_{3} \\
		\epsilon_{4} \\
		\epsilon_{5} \\
		\epsilon_{6} \\
		\epsilon_{7} \\
		\epsilon_{8} \\
		\epsilon_{9} \\
		\epsilon_{10} \\
		\epsilon_{11} \\
		\epsilon_{12} \\
	\end{pmatrix}
	\end{split}
\end{equation}

As the linear regression is performed in the log space, it is necessary to impose the exponent $k$ and $l$ in order to compare the stiffness values between groups. The system is then modified as follow:

\begin{equation}
	\begin{split}
		\ln
		\begin{pmatrix}
			S_{11} \\
			S_{12} \\
			S_{13} \\
			S_{21} \\
			S_{22} \\
			S_{23} \\
			S_{31} \\
			S_{32} \\
			S_{33} \\
			S_{44} \\
			S_{55} \\
			S_{66} \\
		\end{pmatrix} - & \begin{pmatrix}
			\ln(\rho) & \ln(m_1^2) \\
			\ln(\rho) & \ln(m_1 m_2) \\
			\ln(\rho) & \ln(m_1 m_3) \\
			\ln(\rho) & \ln(m_2 m_1) \\
			\ln(\rho) & \ln(m_2^2) \\
			\ln(\rho) & \ln(m_2 m_3) \\
			\ln(\rho) & \ln(m_3 m_1) \\
			\ln(\rho) & \ln(m_3 m_2) \\
			\ln(\rho) & \ln(m_3^2) \\
			\ln(\rho) & \ln(m_2 m_3) \\
			\ln(\rho) & \ln(m_3 m_1) \\
			\ln(\rho) & \ln(m_1 m_2) \\
		\end{pmatrix} \begin{pmatrix}
			k \\
			l \\
		\end{pmatrix} = \\ & \begin{pmatrix}
			1 & 0 & 0 \\
			0 & 1 & 0 \\
			0 & 1 & 0 \\
			0 & 1 & 0 \\
			1 & 0 & 0 \\
			0 & 1 & 0 \\
			0 & 1 & 0 \\
			0 & 1 & 0 \\
			1 & 0 & 0 \\
			0 & 0 & 1 \\
			0 & 0 & 1 \\
			0 & 0 & 1 \\
		\end{pmatrix} \ln\begin{pmatrix}
			\lambda^{*} \\
			\lambda_0' \\
			\mu_0 \\
		\end{pmatrix} + \begin{pmatrix}
			\epsilon_{1} \\
			\epsilon_{2} \\
			\epsilon_{3} \\
			\epsilon_{4} \\
			\epsilon_{5} \\
			\epsilon_{6} \\
			\epsilon_{7} \\
			\epsilon_{8} \\
			\epsilon_{9} \\
			\epsilon_{10} \\
			\epsilon_{11} \\
			\epsilon_{12} \\
		\end{pmatrix}
	\end{split}
\end{equation}

Finally, a modification of the model is to add a regressor for the group variable. Using a grouped data set (healthy and OI), it allows to determine if the group is statistically significant using ANCOVA. In such case the system is written under the form:

\begin{equation}
	\begin{split}
		\ln
		\begin{pmatrix}
			S_{11} \\
			S_{12} \\
			S_{13} \\
			S_{21} \\
			S_{22} \\
			S_{23} \\
			S_{31} \\
			S_{32} \\
			S_{33} \\
			S_{44} \\
			S_{55} \\
			S_{66} \\
		\end{pmatrix} = &\begin{pmatrix}
			1 & 0 & 0 & \ln(\rho) & \ln(m_1^2) & S_g \\
			0 & 1 & 0 & \ln(\rho) & \ln(m_1 m_2) & S_g \\
			0 & 1 & 0 & \ln(\rho) & \ln(m_1 m_3) & S_g \\
			0 & 1 & 0 & \ln(\rho) & \ln(m_2 m_1) & S_g \\
			1 & 0 & 0 & \ln(\rho) & \ln(m_2^2) & S_g \\
			0 & 1 & 0 & \ln(\rho) & \ln(m_2 m_3) & S_g \\
			0 & 1 & 0 & \ln(\rho) & \ln(m_3 m_1) & S_g \\
			0 & 1 & 0 & \ln(\rho) & \ln(m_3 m_2) & S_g \\
			1 & 0 & 0 & \ln(\rho) & \ln(m_3^2) & S_g \\
			0 & 0 & 1 & \ln(\rho) & \ln(m_2 m_3) & S_g \\
			0 & 0 & 1 & \ln(\rho) & \ln(m_3 m_1) & S_g \\
			0 & 0 & 1 & \ln(\rho) & \ln(m_1 m_2) & S_g \\
		\end{pmatrix} \\ & \begin{pmatrix}
			\ln(\lambda^{*}) \\
			\ln(\lambda_0') \\
			\ln(\mu_0) \\
			k \\
			l \\
			\ln(\beta_{S_g})\\
		\end{pmatrix} + \begin{pmatrix}
			\epsilon_{1} \\
			\epsilon_{2} \\
			\epsilon_{3} \\
			\epsilon_{4} \\
			\epsilon_{5} \\
			\epsilon_{6} \\
			\epsilon_{7} \\
			\epsilon_{8} \\
			\epsilon_{9} \\
			\epsilon_{10} \\
			\epsilon_{11} \\
			\epsilon_{12} \\
		\end{pmatrix}
	\end{split}
\end{equation}

Where $S_g$ is coded using a summation constrain \cite{Fox2016}, meaning, $S_g = -1$ for the healthy group and $S_g = 1$ for the OI group.

\clearpage
\section{Extreme ROI Examples}\label{A2}

\vfill

\begin{minipage}{2.\linewidth}
	\centering
	\includegraphics[width=0.45\textwidth]
	{Pictures/A2_BVTV_Frontview}
	\centering
	\includegraphics[width=0.45\textwidth]
	{Pictures/A2_BVTV_Isoview}\\
	\captionof{figure}{ROI with maximum BV/TV observed. BV/TV: 0.59; CV:0.19;  DA:1.48. Left: front view; right: isometric view}
	\label{A2_MaxBVTV}
\end{minipage}

\vfill

\begin{minipage}{2.\linewidth}
	\centering
	\includegraphics[width=0.45\textwidth]
	{Pictures/A2_MinBVTV_Frontview}
	\centering
	\includegraphics[width=0.45\textwidth]
	{Pictures/A2_MinBVTV_Isoview}\\
	\captionof{figure}{ROI with minimum BV/TV after filtering. BV/TV: 0.04; CV:0.19;  DA:1.58. Left: front view; right: isometric view}
	\label{A2_MinBVTV}
\end{minipage}

\vfill

\begin{minipage}{2.\linewidth}
	\centering
	\includegraphics[width=0.45\textwidth]
	{Pictures/A2_CV_Frontview}
	\centering
	\includegraphics[width=0.45\textwidth]
	{Pictures/A2_CV_Isoview}\\
	\captionof{figure}{ROI with maximum CV observed. BV/TV: 0.02; CV:1.56;  DA:2.25. Left: front view; right: isometric view}
	\label{A2_MaxCV}
\end{minipage}

\vfill

\clearpage
% Loading bibliography database
\bibliographystyle{BibStyle}
\bibliography{Bibliography}

\end{document}

### 2021-04-08 03_Scripts/04_2_General_LinearRegression.py

@@ -75,7 +75,7 @@ def PlotRegressionResults(Model, Data, PlotTypes=['BV/TV', 'DA', 'Constants'], A
     Sjj = Y_Fit * np.array(X[:, 2].T)[0]
 
     ## Plots
-    DPI = 100
+    DPI = 500
     SMax = max(HealthyData[Y_Elements].max()) * 5
     SMin = min(OIData[Y_Elements].min()) / 5
     if 'BV/TV' in PlotTypes:
@@ -91,8 +91,8 @@ def PlotRegressionResults(Model, Data, PlotTypes=['BV/TV', 'DA', 'Constants'], A
         Axes.annotate(r'$R^2_{ajd}$: ' + format(round(R2adj, 3), '.3f'), xy=(0.65, 0.1), xycoords='axes fraction')
         Axes.annotate(r'$NE$ : ' + format(round(NE.mean(), 2), '.2f') + '$\pm$' + format(round(NE.std(), 2), '.2f'),
                       xy=(0.65, 0.025), xycoords='axes fraction')
-        Axes.set_xlabel('Observed $\mathbb{S}_{xy}$ (MPa)')
-        Axes.set_ylabel('Fitted $\mathbb{S}_{xy}$ (MPa)')
+        Axes.set_xlabel('Observed $\mathrm{\mathbb{S}}$ (MPa)')
+        Axes.set_ylabel('Fitted $\mathrm{\mathbb{S}}$ (MPa)')
         Axes.set_xlim([SMin,SMax])
         Axes.set_ylim([SMin,SMax])
         plt.xscale('log')
@@ -113,8 +113,8 @@ def PlotRegressionResults(Model, Data, PlotTypes=['BV/TV', 'DA', 'Constants'], A
         Axes.annotate(r'$R^2_{ajd}$: ' + format(round(R2adj, 3), '.3f'), xy=(0.65, 0.1), xycoords='axes fraction')
         Axes.annotate(r'$NE$ : ' + format(round(NE.mean(), 2), '.2f') + '$\pm$' + format(round(NE.std(), 2), '.2f'),
                       xy=(0.65, 0.025), xycoords='axes fraction')
-        Axes.set_xlabel('Observed $\mathbb{S}_{xy}$ (MPa)')
-        Axes.set_ylabel('Fitted $\mathbb{S}_{xy}$ (MPa)')
+        Axes.set_xlabel('Observed $\mathrm{\mathbb{S}}$ (MPa)')
[…]

### 2021-05-18 09_Article/Simon_2021_Bone_Draft.tex
Add motion artefact grading and BMD histograms

--- 2021-04-08 09_Article/Article.tex
@@
 \begin{abstract}
-	Osteogenesis imperfecta (OI) is the most common inherited form of bone fragility.
-This disease, also called "brittle bone disease", is characterized by impaired synthesis of type I collagen leading to fragile bones.
-In this study, 50 OI diagnosed people were compared to 120 healthy controls using second generation high resolution peripheral computed tomography (HR-pQCT) scans of the distal tibia.
-Such scans are usually used in finite element analysis (FEA) for bone strength estimation and fracture risk assessment.
-The homogenized FE (hFE) scheme uses HR-pQCT bone volume fraction (BV/TV) and anisotropy information (fabric) and allows to assess bone strength within a reasonable computation time using fabric-elasticity relationships.
-Thus, the aim of this study is to investigate fabric-elasticity relationships in OI compared to healthy controls.
-For this, six regions of interest (ROIs) were selected per individual in a common anatomical region.
-A first age \& gender matching allowed to perform a morphometric analysis and compare the samples with literature.
-Then, stiffness tensors of ROIs was computed by FEA and multiple linear regression was performed on the Zysset-Curnier orthotropic model to compare groups using 5 parameters.
-This fit was first executed using the complete data sets separately.
-Then, data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity.
-A second fit was performed on these filtered data sets and a third fit was done on BV/TV \& fabric anisotropy (DA) matched data sets.
-This last matched data set was grouped to compare with literature \si{\micro}CT scans.
-Morphometric analysis shown results in agreement with literature, namely significant differences between groups for BV/TV, trabecular number (Tb N), trabecular spacing (Tb Sp), trabecular spacing standard deviation (Tb Sp SD) but no significant differences for trabecular thickness (Tb Th).
-The multiple linear regressions shown that OI stiffness range reach lower values than healthy controls and the fit tends of overestimate the stiffness at lower values.
-Filtering out high heterogeneous ROIs removes low stiffness ROIs and lead to similar fit quality for both OI and healthy group.
-The BV/TV \& DA matched data sets shown no significant differences between OI and healthy individuals and stiffness constants are lower at 61 \si{\micro}m resolution HR-pQCT scans than for higher resolution 36 \si{\micro}m \si{\micro}CT scans.
-Our main conclusions are that lower scan resolution lead to lower stiffness because the structure obtained is not as optimized as in reality.
-Finally, if trabecular bone is homogeneous enough ROIs inside ROI, there no reason to suppose different fabric-elasticity relationships between OI and healthy individual.
+	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility.
+This disease, also called "brittle bone disease", is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass that lead to fragile bones.
+High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology in the weight-bearing distal tibia and the 3D reconstructions can be exploited in either micro-finite element ($\mu$FE) or homogenised finite element (hFE) analysis for bone strength estimation.
+The hFE scheme uses local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
+Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
+
+	In this study, the morphology of distal tibias of 50 OI diagnosed people were compared to 120 healthy controls using second generation HR-pQCT. 
+	Six regions of interest (ROIs) were selected per individual in a common anatomical region.
+A first age \& gender matching allowed to perform a morphometric analysis and compare the outcome with literature.
+Then, stiffness tensors of ROIs were computed by FEA and multiple linear regressions were performed on the Zysset-Curnier orthotropic model to compare the two groups using 5 parameters.
+After initial fits with all the samples of each group, the data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity.
+Second fits were performed on these filtered data sets and then additional fits were done on BV/TV \& fabric anisotropy (DA) matched data to detect statistical differences between the two groups.
+This full and filtered data were in turn compared with previous results from $\mu$CT reconstructions obtained in other anatomical locations.
+
+	In agreement with available literature, significantly lower BV/TV, trabecular number (Tb N), significantly higher trabecular spacing (Tb Sp), trabecular spacing standard deviation (Tb Sp SD), but no differences in trabecular thickness (Tb Th) were found between OI and controls.
+The stiffness of ROIs from OI bone reached lower values and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range.
+Filtering out highly heterogeneous ROIs removed these low stiffness ROIs and lead to similar correlation coefficients for both OI and healthy groups.
+Finally, the BV/TV \& DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
+Comparing with previous studies, the stiffness constants from 61 $\mu$m resolution HR-pQCT ROIs were lower than for 36 $\mu$m resolution $\mu$m CT images.
+
+	In conclusion, despite the reduced regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis.
+Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.
 \end{abstract}
@@
 
-Osteogenesis imperfecta (OI) is the most common inherited form of bone fragility \cite{Tournis2018}.
+Osteogenesis imperfecta (OI) is an inherited form of bone fragility \cite{Tournis2018}.
 It was estimated to concern about 1/13'500 births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
@@
 
-Bone fragility in OI is complex and not totally understood.
+Bone fragility in OI is complex and not totally understood despite the investigations at different hierarchical levels.
+Multiple studies show that DXA areal BMD (aBMD) tends to be lower in OI than compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.
+The microstructure is different as well.
+\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} shown that BV/TV and trabecular number in OI is lower than for healthy controls.
+Trabecular spacing and inhomogeneity are higher for OI diagnosed people but the trabecular thickness is not significantly different.
+At the ECM level, a recent study shown that OI bone tends to present higher modulus, ultimate stress and post-yield behavior than healthy bone in compression \cite{Indermaur2021}.\\
+
 High resolution peripheral quantitative tomography (HR-pQCT) scans allow to perform \textit{in vivo} assessment of trabecular architecture and volumetric BMD in distal radius and tibia \cite{Boutroy2005}.
@@
 High correlations were found between patient-specific hFE and mechanical compression experiment of cadaveric samples \cite{Varga2011,AriasMoreno2019}.
-Thus, it could be legitimate to use hFE for OI patients bone strength estimation and fracture risk assessment with the assumption of identical fabric-elasticity relationships between OI and healthy individuals.
-Therefore, the present study aims to compare healthy with OI diagnosed individuals and to draw conclusions about fabric-elasticity relationships.
+Thus, it could be legitimate to use hFE for OI patients bone strength estimation and fracture risk assessment.
+However, HR-pQCT-based FEA rely on fabric-elasticity relationships.
+Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI diagnosed individuals and to investigate the hypothesis of similar fabric-elasticity relationships.
 
@@
 \subsection{Subjects}
-The healthy group include a total of 120 patients from a previous reproducibility study \cite{Schenk2020}.
+The healthy group include a total of 120 patients from a previous reproducibility study performed at the University Department of osteoporosis in Bern \cite{Schenk2020}.
 The sample is composed of 64 female and 56 male subjects aged between 20 and 92 years old with a median age of 26 [22 - 35] years.
 These subjects did not took any medication known to affect bone metabolism nor presented any prior osteoporosis fracture.
-The second group is composed of 35 female 15 male individuals leading to 50 OI diagnosed subjects.
+The second group was scanned at the Shriners Hospital for Children and images were shared to our group by the McGill University in Montreal.
+This group is composed of 35 female 15 male individuals leading to 50 OI diagnosed subjects.
 The youngest and oldest patients are 19 and 69 year old, respectively.
@@
 Switzerland) were performed at the distal tibia on all patients from both groups.
-The OI subjects were scanned using the manufacturer's standard protocol.
-These scans were performed at the Shriners Hospital for Children and were shared to our group by the McGill University in Montreal.
-People of the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
-The main differences with the manufacturer's standard protocol are the following:
+People of the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020} whereas OI subjects were scanned using the manufacturer's standard protocol.
+The main differences of the in-house protocol with the manufacturer's standard protocol are the following:
 \begin{enumerate}
@@
 Scanning settings were a voltage of 60 kVp, 900 μA, 100 ms integration time for the healthy group as well as for the OI group.
+For the healthy group, motion artefacts of first, middle and last slice were graded.
+The scale used start from 1 (no motion artefacts) to 5 (extreme motion artefacts) as proposed by the manufacturer \cite{Pialat2012}.
+The final grade of each scan was defined as the highest slice grade.
+For the OI group, as the scan consist of one stack, only one grade is attributed using the same scale as for the control group.
+Scans were then processed independently of their quality grading.
+A summary of the scans grading densities is shown in Figure \ref{01_MotionArtefacts}.
 
@@
 	\centering
+	\begin{subfigure}[b]{0.225\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]
+		{Pictures/01_ControlClinicalSection}
+		\caption{Healthy group}
+		\label{01_Healthy}
+	\end{subfigure}
+	\hfill
 	\begin{subfigure}[b]{0.225\textwidth}
@@
 	\end{subfigure}
-	\hfill
-	\begin{subfigure}[b]{0.225\textwidth}
-		\centering
-		\includegraphics[width=\textwidth]
-		{Pictures/01_ControlClinicalSection}
-		\caption{Healthy group}
-		\label{01_Healthy}
-	\end{subfigure}
-	\caption{Clinical section scanned for both group}
+	\caption{\centering Clinical section scanned for both group}
 	\label{01_ClinicalSections}
@@
 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]
+	{Pictures/01_MotionArtefacts}
+	\caption{Summary of the motion artefacts grading. Histograms show density of each grade within both group.}
+	\label{01_MotionArtefacts}
+\end{figure}
+
 \subsection{Image analysis}
-The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, the segmented images were used for further analysis.\\
+The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, mask segmented images were used for further analysis.\\
 
@@
 
-The ROI is defined as a cube of about 5.3 mm side length.
-This size was chosen to correspond to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013}.
-It was determined by \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to be the optimal size to obtain accurate FEA results with minimal computation efforts.\\
+The ROI is defined as a cube of 5.3 mm side length.
+This size was chosen to correspond to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} which performed similar analysis with femur \si{\micro}CT scans.
+It was determined by \citeauthor{Zysset1998}\cite{Zysset1998} and \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to be the optimal size to obtain accurate FEA results.\\
 
-The ROI analysis was performed using medtool (v4.5; Dr.
+The morphological analysis of ROIs was performed using medtool (v4.5; Dr.
 Pahr Ingenieurs e.U., Pfaffstätten, Austria).
@@
 SD).
-Moreover, ROI fabric was evaluated.
+Moreover, ROI fabric was evaluated using MIL method \cite{Moreno2014}.
 The fabric tensor $\mathbf{M}$ is a positive-definite second-order tensor.
@@
 \begin{equation}
-	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i}
+	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i} = \sum_{i=1}^{3}{m_i \mathbf{m}_i \otimes \mathbf{m}_i}
 	\label{Eq201}
@@
 The fabric is then independent of BV/TV and normalized with $tr(\mathbf{M}) = 3$.
-To evaluate this tensor, the MIL method was used \cite{Moreno2014}.
-The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI.\\
+The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.\\
 
-A mechanical analysis was performed using \textsc{ABAQUS 6.14}.
-Each voxel of the segmented ROI was converted to mesh using fully integrated linear brick elements (C3D8) with a stiffness $E$ of 10 000 MPa and a Poisson's ratio $\nu$ of 0.3.
+After ROI cleaning, i.e.
+deletion of unconnected region of bone material, an homogenized mechanical analysis was performed using \textsc{ABAQUS 6.14}.
+Each voxel of the cleaned ROI was converted to mesh using fully integrated linear brick elements (C3D8) with a stiffness $E$ of 10 000 MPa and a Poisson's ratio $\nu$ of 0.3.
 The simulation consisted of 6 independent load cases, 3 uni-axial and 3 shear cases, using KUBCs.
-KUBCs were used according to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} were it was determined to be better than PMUBC to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
-This model builds the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV $\rho$, fabric information $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$.
-The building of this tensor is shown in Equation \ref{Eq202}.
+KUBCs were used according to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}.
+Unlike PMUBCs, KUBCs do not require to rotate the ROI into fabric coordinate system which decrease the image quality.
+This homogenization process allows to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
+This model builds the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric information $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$.
+The building of this tensor is shown in Equation \ref{Eq202}.\\
 
@@
 	\begin{split}
-		\mathbb{S}(\rho,\mathbf{M}) & = \sum_{i=1}^{3} (\lambda_0 + 2\mu_0)\rho^k m_i^{2l} \mathbf{M}_i \otimes \mathbf{M}_i \\ & + \sum_{\substack{i,j=1\\i \neq j}}^{3} \lambda_0' \rho^k m_i^{l} m_j^{l} \mathbf{M}_i \otimes \mathbf{M}_j \\ & + \sum_{\substack{i,j=1\\i \neq j}}^{3} \mu_0 \rho^k m_i^{l} m_j^{l} \mathbf{M}_i \overline{\underline{\otimes}} \mathbf{M}_j
+		&\mathbb{S}(\rho,\mathbf{M}) &=& \quad\sum_{i=1}^{3} \lambda_{ii} \mathbf{M}_i \otimes \mathbf{M}_i \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \lambda_{ij} \mathbf{M}_i \otimes \mathbf{M}_j \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \mu_{ij} \mathbf{M}_i \overline{\underline{\otimes}} \mathbf{M}_j \\
+		&\text{With} &\\
+		&\qquad\lambda_{ii} &=& \quad(\lambda_0 + 2\mu_0)\rho^k m_i^{2l} \\
+		&\qquad\lambda_{ij} &=& \quad\lambda_0' \rho^k m_i^{l} m_j^{l} \\
+		&\qquad\mu_{ij} &=& \quad\mu_0 \rho^k m_i^{l} m_j^{l} \\
 	\end{split}
@@
 Where $\otimes$ and $\overline{\underline{\otimes}}$ are the dyadic and symmetric product of second order tensors, respectively.
-The Zysset-Curnier model is built with the assumption of orthotropy.
-However, the trabecular structure is not perfectly orthotropic.
+The Zysset-Curnier model is built with the assumption of orthotropy and homogeneity.
+However, the trabecular structure is not perfectly homogeneous.
 In order to assess the ROI heterogeneity a so-called coefficient of variation (CV) is computed as presented in \cite{Panyasantisuk2015}: the ROI is divided into eight identical subcubes, BV/TV is computed for each subcube and the CV is defined as the ratio between the standard deviation of these BV/TV and the mean value, see Equation \ref{Eq203} below:
@@
 \subsection{Statistics}
-The height morphological parameters analyzed (BV/TV, Tb.
+The morphological parameters analyzed (BV/TV, Tb.
 N., Tb.
@@
 SD, SMI, DA, and CV) were compared for both group.
-As the initial groups do not have similar distributions in quality of age and sex, a matching was performed leading to identical mean and median age as well as identical gender distribution.
+As the initial groups do not have similar distributions of age and sex, a matching was performed leading to identical mean and median age as well as identical gender distribution.
 Then, the selection of statistical test to perform was executed as follow:
@@
 \subsection{Fit to model}
-The stiffness tensors obtained from the mechanical simulations were transformed into fabric coordinate system and projected onto orthotropy, leading to 12 values.
+The stiffness tensors obtained from the mechanical simulations were transformed into fabric coordinate system and projected onto orthotropy, leading to 12 components.
 The resulting orthotropic stiffness tensors were then used to perform a multiple linear regression on the Zysset-Curnier model.
@@
 This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, takes into account the non-independence of ROIs from the same individual.
+Detailed form of this model is presented in Appendix \ref{A1}.
 
@@
 To improve the fit quality, the data sets were filtered.
-The aim here is to filter out ROIs whose are too far from the assumption of orthotropy.
+The aim here is to filter out ROIs whose are too far from the assumption of homogeneity.
 Therefore, analogously to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, a fixed threshold for the CV was used.
@@
 	\centering
-	\caption{Summary of morphological analysis and comparison with literature.
+	\caption{Summary of the tibia ROIs morphological analysis and comparison with literature.
 Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians.
@@
 
-\subsection{Linear Regression}
+\subsection{Linear Regressions}
 Figure \ref{02_GeneralRegression} shows the result of regression for complete data sets of each group separately.
@@
 One last observation which should be noted is the range covered by the observed tensors.
-The range of the OI group is wider than the healthy group and goes further in low observed values.\\
+The range of the OI group is wider than the healthy group and ROIs with lower BV/TV present further low stiffness values.\\
 
@@
 On the other hand, relatively few healthy data will be removed by the filtering.
-Examples of extreme ROIs in terms of CV and BV/TV are available in Appendix \ref{A2}.\\
+Examples of extreme ROIs in terms of CV and BV/TV are shown in Appendix \ref{A2}.\\
 
@@
 \begin{table*}[b]
-	\caption{Constants comparison between grouped and separated data sets. Values are presented as value [95\% CI] or mean $\pm$ standard deviation.}
+	\caption{Constants obtained with BV/TV and DA matched data sets.
+Comparison is performed between grouped (N ROIs = 166) and separated data sets (N ROIs = 83).
+Values are presented as value [95\% CI] or mean $\pm$ standard deviation.}
 	\label{Table3}
@@
 \begin{table*}[b]
-	\caption{Comparison with literature. N stands for the number of observations. Values are presented as computed value only or mean $\pm$ standard deviation.}
+	\caption{Comparison with literature.
+N stands for the number of ROIs observed.
+Values are presented as computed value only or mean $\pm$ standard deviation.
+The present study shows values obtained with tibia XCTII scans.
+\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013} show values obtained with femur \si{\micro}CT scans.}
 	\label{Table4}
@@
 		\midrule
+		\multicolumn{11}{c}{\cellcolor[HTML]{D9D9D9}Filtered data sets}\\
+
 		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 126 & 0.27 $\pm$ 0.08 & 1.57 $\pm$ 0.18 & 3306 & 2736 & 2837 & 1.55 & 0.82 & 0.984 & 8 $\pm$ 3\\
 
-		Present study & 166 & 0.21 $\pm$ 0.08 & 2.03 $\pm$ 0.21 & 2559 & 1490 & 1957 & 1.55 & 0.82 & 0.815 & 25 $\pm$ 11\\[3ex]
+		\multirow{2}{*}{Present study} & 718 & 0.27 $\pm$ 0.09 & 1.93 $\pm$ 0.24 & 2510 & 1623 & 2056 & 1.55 & 0.82 & 0.832 & 21 $\pm$ 11\\
+		& 718 & 0.27 $\pm$ 0.09 & 1.93 $\pm$ 0.24 & 4775 & 3087 & 3910 & 1.99 & 0.85 & 0.949 & 17 $\pm$ 9\\[1ex]
 
+		\multicolumn{11}{c}{\cellcolor[HTML]{D9D9D9}Non-filtered data sets}\\
 		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 167 & 0.25 $\pm$ 0.08 & 1.54 $\pm$ 0.20 & 3841 & 3076 & 3115 & 1.60 & 0.99 & 0.983 & 14\\
@@
 
-		Present study & 166 & 0.21 $\pm$ 0.08 & 2.03 $\pm$ 0.21 & 2817 & 1640 & 2154 & 1.60 & 0.99 & 0.826 & 22 $\pm$ 12\\
+		\multirow{2}{*}{Present study} & 1014 & 0.23 $\pm$ 0.11 & 1.94 $\pm$ 0.26 & 2727 & 1655 & 2179 & 1.60 & 0.99 & 0.608 & 41 $\pm$ 182\\
+		& 1014 & 0.23 $\pm$ 0.11 & 1.94 $\pm$ 0.26 & 5323 & 3232 & 4254 & 2.02 & 0.91 & 0.915 & 31 $\pm$ 105\\
 		\bottomrule
@@
 
-\section{Discussion \& Conclusions}
+\section{Discussion}
 The age of matched individuals lies in the spectrum of the others studies which allow to compare the morphological values.
@@
 This is due to the fact that ROI stiffness is highly impacted by BV/TV.
-Some low BV/TV ROIs do not have every side of the cube connected by bone leading to extremely low terms in the stiffness tensor.
+Some low BV/TV ROIs do not have every side of the cube connected by bone leading to extremely low terms in the stiffness tensor, see Appendix \ref{A2}.
 Trying to homogenize such ROI can lead to error of multiple order of magnitude, as observed on the plot.
-Therefore, a filtering is justified to compare fabric-elasticity relationships.\\
+Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
+An alternative of CV filtering could be to compute the area ratio filled by bone on each of the six faces of the ROI to assess the ROI heterogeneity.\\
 
-Figure \ref{02_CV_BVTV} presenting the CV in relation to BV/TV shows that it exist a tendency of CV to increase with a decreasing BV/TV.
+Figure \ref{02_CV_BVTV} presenting the CV in relation to BV/TV shows that there is a tendency of CV to increase with a decreasing BV/TV.
 Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly becomes highly heterogeneous.
@@
 The fits performed on filtered data sets present direct effect of filtering.
-For the healthy group, the relatively small decrease of $R^2_{adj}$ (5\textperthousand) is negligible.
+For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (5\textperthousand) is negligible.
 On the other hand, NE presents an improved value decreasing by 2\%.
@@
 Results become similar to the healthy group in terms of $R^2_{adj}$, NE, and the range of stiffness values.
-These results give confidence to the filtering procedure and are a first good point to the hypothesis of no different fabric-elasticity relationships between healthy and OI trabecular bone.\\
+These results give confidence to the filtering procedure and are a first step to the hypothesis of no different fabric-elasticity relationships between healthy and OI trabecular bone.\\
 
@@
 Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group shown a p value far above the 5\% significance level.
-With this statistical non-significance of the group and low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to suppose differences in fabric-elasticity relationships between healthy and OI diagnosed people.
+With this statistical non-significance of the group and low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to suppose differences in fabric-elasticity relationships between healthy and OI trabecular bone.
 In FEA simulations, it is not possible to exclude part of the mesh because of high heterogeneity.
-Nevertheless, the error created by such ROIs can be negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\
+Nevertheless, the error created by such ROIs are negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\
 
 Imposing $k$ and $l$ allows to estimate the effect of different image resolutions.
-\citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used scans with 18 \si{\micro}m spatial resolution and coarsened it to 36 \si{\micro}m.
+\citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened it to 36 \si{\micro}m.
 \citeauthor{Gross2013}\cite{Gross2013} shown that different anatomical locations lead to only slight differences.
-Comparing regression of filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the lower BV/TV and higher DA ranges and by the coarser resolution.
+Comparing regression of filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution.
 Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones.
-Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges are well overlapped.
-DA is still higher in the present study and the stiffness constants remain lower than for the two other studies.
-Here, differences can be explained by resolution.
-The coarser structure resulting from XCTII can not be as optimized as the fine detailed structure obtained by \si{\micro}CT.
+Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges stay overlapped.
+As for filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies.
+Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants by resolution.
+The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explain this increase of DA.
+Lower stiffness constants are obtained because the coarser structure resulting from XCTII can not be as optimized as the fine detailed structure obtained by \si{\micro}CT.
+Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimized morphology of trabecular bone with a high fidelity.
+By decreasing the scan spatial resolution, the scanned structure becomes more bulky.
 Performing a fit on this less optimized structure lead to the observed lower stiffness constants.
@@
 
-In conclusions, this work shown that, if the ROI is homogeneous enough, there no reason to suppose any differences in fabric-elasticity relationships between healthy and OI trabecular bone.
 The main limitations of this study are the definition of "homogeneous enough" and the fact that it is limited to tibiae XCTII scans.
-More investigations should be performed to improve the model for highly heterogeneous ROIs but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible.
-An idea could be to compute a mass gradient instead of CV to improve the model.
-Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.
-Finally, it could be interesting to perform homogenization and fitting on identical samples using \si{\micro}CT and XCTII to quantify with more accuracy the resolution impact.
+The ROI homogeneity has an important impact on the analysis quality.
+As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for fitting.
+More investigations could be performed to improve the model for highly heterogeneous ROIs but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible.
+Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.\\
+
+In conclusions, the sample analyzed in the this study have similar morphology that data reported in the literature.
+We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROI is homogeneous enough.
+\citeauthor{Indermaur2021}\cite{Indermaur2021} shown that compressive behavior of OI bone tissue is similar to healthy control.
+If tensile and shearing behaviors agree, fabric-strength relationships will hold too.
+Therefore, OI bone fragility might mostly result from the decrease in BV/TV and loss of homogeneity in its organization.
 
 \section*{Acknowledgments}
-The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the buildingi and interpretation of linear mixed-effect model.
+The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII scans of OI individuals.
 This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.
 
+\clearpage
 \appendix
@@
 In the present case, the stiffness variables ($\lambda_0$, $\lambda_0'$, and $\mu_0$) can vary between individuals but the hypothesis is that they all vary by an identical factor.
-Therefore, the design matrix $Z$ is composed of the addition of the three first columns of $X$ and the system for one ROI takes the following form:
+Therefore, the design matrix $Z$ is composed of the addition of the three first columns of $X$ and the system for one ROI takes the following form:\\
 
@@
 Using a grouped data set (healthy and OI), it allows to determine if the group is statistically significant using ANCOVA.
-In such case the system is written under the form:
+In such case the system is written under the form:\\
 

### 2021-05-18 09_Article/Simon_2021_Bone_DraftPhZ.tex
Add motion artefact grading and BMD histograms

--- 2021-05-18 09_Article/Simon_2021_Bone_Draft.tex
@@
 Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.
+
 \end{abstract}
@@
 
-Osteogenesis imperfecta (OI) is an inherited form of bone fragility \cite{Tournis2018}.
+Osteogenesis Imperfecta (OI) is an inherited form of bone fragility \cite{Tournis2018}.
 It was estimated to concern about 1/13'500 births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
@@
 \subsection{Image analysis}
-The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, mask segmented images were used for further analysis.\\
+The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, mask segmented images were used for further analysis. \\
 
@@
 \subsection{Morphological Analysis}
-The results of morphological analysis are summarized in Table \ref{Table2}.
+The results of morphological analysis are summarised in Table \ref{Table2}.
 In the present study, the individual matching allowed to have similar group distribution with 17 females and 11 males in each group.
@@
 
-In conclusions, the sample analyzed in the this study have similar morphology that data reported in the literature.
-We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROI is homogeneous enough.
-\citeauthor{Indermaur2021}\cite{Indermaur2021} shown that compressive behavior of OI bone tissue is similar to healthy control.
-If tensile and shearing behaviors agree, fabric-strength relationships will hold too.
-Therefore, OI bone fragility might mostly result from the decrease in BV/TV and loss of homogeneity in its organization.
+In conclusions, the sample analysed in the this study have similar morphology than data reported in the literature.
+We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROIs are homogeneous enough.
+\citeauthor{Indermaur2021}\cite{Indermaur2021} showed that compressive behaviour of OI bone tissue is similar to healthy control.
+If tensile and shearing behaviours agree, fabric-strength relationships will hold too.
+Therefore, OI bone fragility might mostly result from the decrease in BV/TV and loss of homogeneity in its macroscopic organisation.
 
 \section*{Acknowledgments}
-The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII scans of OI individuals.
+The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the buildingi and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII scans of OI individuals.
 This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.

### 2021-05-18 03_Scripts/FindGradings.py

@@ -1,0 +1,79 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+# 01 Set variables
+WorkingDirectory = os.getcwd()
+DataPath = os.path.join(WorkingDirectory,'02_Data/')
+
+## Load files
+Data = pd.read_csv(DataPath + '01_HealthyTibiaXCT2Scans.csv')
+FullData = pd.read_excel(DataPath + 'Repro_UPAT_TOTAL_Denis_Cleaned_004_incl_nm_003.xlsx')
+MapTable = pd.read_csv(DataPath + 'MapTable.csv')
+
+# Find grading
+Gradings = pd.DataFrame()
+for i in Data.index:
+    Measurement = Data.loc[i,'Sample Number']
+    Folder = MapTable[MapTable['Measurement Number']==Measurement]['Measurement Folder'].values[0]
+    for j in range(1,4):
+        MeasN = 'MeasNo' + str(j)
+        GradN = 'Grading' + str(j)
+        Grading = FullData[FullData[MeasN] == Folder][GradN]
+        if len(Grading) > 0:
+            Gradings = Gradings.append({'Grading': int(Grading),'Sample Number':Measurement, 'Folder':Folder},ignore_index=True)
+
+# Add second table data
+FullData = pd.read_excel(DataPath + 'Norm_UPAT_TOTAL_hFE_corrected_Denis.xlsm')
+
+# Find grading
+for i in Data.index:
+    Measurement = Data.loc[i,'Sample Number']
+    Folder = MapTable[MapTable['Measurement Number']==Measurement]['Measurement Folder'].values[0]
+    MeasN = 'MeasNo1'
+    GradN = 'Quality1'
+    Grading = FullData[FullData[MeasN] == Folder][GradN]
+    if len(Grading) > 0:
+        print(i)
[…]

### 2021-06-02 09_Article/Simon_2021_Bone_Draft_V01.tex
Finish tBMD analysis and prepare SBMS presentation

--- 2021-05-18 09_Article/Simon_2021_Bone_Draft.tex
@@
 \def\textpagefraction{.001}
-\shorttitle{Fabric-Elasticity in Osteogenesis Imperfecta}
+\shorttitle{Fabric-Elasticity Relationships in Osteogenesis Imperfecta}
 \shortauthors{Simon et~al.}
 
-\title[mode = title]{Fabric-Elasticity Relationships in Osteogenesis Imperfecta (OI)}
+\title[mode = title]{Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals}
 
@@
 
-\author[2,3]{Bettina Willie}
+\author[2,3]{Bettina M. Willie}
 
@@
 
-\address[2]{Shriners Hospital for Children, Montreal, Canada}
-
-\address[3]{McGill University, Montreal, Canada}
+\address[2]{Research Centre, Shriners Hospital for Children, Montreal, Canada}
+
+\address[3]{Departement of Pediatric Surgery, McGill University, Montreal, Canada}
 
@@
 \abbreviations{OI, osteogenesis imperfecta;
-			   HR-pQCT, high resolution peripheral quantitative tomography;
+			   HR-pQCT, high resolution peripheral quantitative computed tomography;
 			   BMD, bone mass density;
@@
 \begin{abstract}
-	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility.
-This disease, also called "brittle bone disease", is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass that lead to fragile bones.
-High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology in the weight-bearing distal tibia and the 3D reconstructions can be exploited in either micro-finite element ($\mu$FE) or homogenised finite element (hFE) analysis for bone strength estimation.
-The hFE scheme uses local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
+	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility, also called "brittle bone disease".
+It is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass.
+High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology of extremities including the weight-bearing distal tibia.
+The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation.
+The hFE scheme uses homogenized local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
 Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
 
-	In this study, the morphology of distal tibias of 50 OI diagnosed people were compared to 120 healthy controls using second generation HR-pQCT. 
-	Six regions of interest (ROIs) were selected per individual in a common anatomical region.
-A first age \& gender matching allowed to perform a morphometric analysis and compare the outcome with literature.
-Then, stiffness tensors of ROIs were computed by FEA and multiple linear regressions were performed on the Zysset-Curnier orthotropic model to compare the two groups using 5 parameters.
-After initial fits with all the samples of each group, the data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity.
-Second fits were performed on these filtered data sets and then additional fits were done on BV/TV \& fabric anisotropy (DA) matched data to detect statistical differences between the two groups.
-This full and filtered data were in turn compared with previous results from $\mu$CT reconstructions obtained in other anatomical locations.
-
-	In agreement with available literature, significantly lower BV/TV, trabecular number (Tb N), significantly higher trabecular spacing (Tb Sp), trabecular spacing standard deviation (Tb Sp SD), but no differences in trabecular thickness (Tb Th) were found between OI and controls.
-The stiffness of ROIs from OI bone reached lower values and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range.
-Filtering out highly heterogeneous ROIs removed these low stiffness ROIs and lead to similar correlation coefficients for both OI and healthy groups.
-Finally, the BV/TV \& DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
-Comparing with previous studies, the stiffness constants from 61 $\mu$m resolution HR-pQCT ROIs were lower than for 36 $\mu$m resolution $\mu$m CT images.
+	In this study, the morphology of distal tibiae from 50 adults with OI were compared to 120 healthy controls using second generation HR-pQCT.
+Six cubic regions of interest (ROIs) were selected per individual in a common anatomical region.
+A first matching was performed by selecting similar individuals to obtain identical mean and median age and gender distribution in the OI and healthy control group.
+It allowed to perform a morphometric analysis and compare the outcome with literature.
+Then, stiffness tensors of ROIs were computed using hFEA and multiple linear regressions were performed on the Zysset-Curnier orthotropic model.
+The regressions allowed to compare the two groups using 5 parameters.
+An initial fit was performed on both the OI group and the healthy control group using all the ROIs extracted.
+Then, data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity and second fits were performed on these filtered data sets.
+These full and filtered data were in turn compared with previous results from \si{\micro}CT reconstructions obtained in other anatomical locations.
+Finally, the ROIs of both group were matched according to their BV/TV and fabric anisotropy (DA).
+Fits were performed again using these matched data to detect statistical differences between the two groups.
+
+	Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular spacing standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
+These results are in agreement to literature.
+The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range.
+The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and lead to similar correlation coefficients for both OI and healthy groups.
+Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
+Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.
 
@@
 
-Osteogenesis imperfecta (OI) is an inherited form of bone fragility \cite{Tournis2018}.
-It was estimated to concern about 1/13'500 births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
+Osteogenesis imperfecta (OI), also commonly known as "brittle bone disease", is an inherited form of bone fragility \cite{Tournis2018}.
+OI prevalence is estimated at about  1/13,500 of births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
 Therefore, OI is considered as a rare metabolic bone disorder.
-Sillence classification \cite{Sillence1979} allows to define four main types of OI based on the clinical phenotype:
+In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases.
+OI can be categorized according to disease severity \cite{Mortier2019} into:
 \begin{itemize}
-	\item Type I: less severe
-	\item Type II: lethal at most shortly after birth
+	\item Type I: mild
+	\item Type II: perinatally lethal
 	\item Type III: most severe surviving form
@@
 \end{itemize}
-Also known as "brittle bone disease", OI is characterized by impaired synthesis of type I collagen, leading to such brittle and fragile bones \cite{LIM2017}.\\
-
-Bone fragility in OI is complex and not totally understood despite the investigations at different hierarchical levels.
-Multiple studies show that DXA areal BMD (aBMD) tends to be lower in OI than compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.
-The microstructure is different as well.
-\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} shown that BV/TV and trabecular number in OI is lower than for healthy controls.
-Trabecular spacing and inhomogeneity are higher for OI diagnosed people but the trabecular thickness is not significantly different.
-At the ECM level, a recent study shown that OI bone tends to present higher modulus, ultimate stress and post-yield behavior than healthy bone in compression \cite{Indermaur2021}.\\
-
-High resolution peripheral quantitative tomography (HR-pQCT) scans allow to perform \textit{in vivo} assessment of trabecular architecture and volumetric BMD in distal radius and tibia \cite{Boutroy2005}.
-Moreover, the bony microstructure obtained from HR-pQCT can be used for finite element analysis (FEA) to predict mechanical properties \cite{Boutroy2008}.
-Homogenized FE (hFE) is a FEA scheme including volume fraction (BV/TV) and anisotropy information (fabric) of trabecular bone from HR-pQCT which to assess bone strength within reasonable computation time \cite{Pahr2009}.
-High correlations were found between patient-specific hFE and mechanical compression experiment of cadaveric samples \cite{Varga2011,AriasMoreno2019}.
-Thus, it could be legitimate to use hFE for OI patients bone strength estimation and fracture risk assessment.
-However, HR-pQCT-based FEA rely on fabric-elasticity relationships.
-Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI diagnosed individuals and to investigate the hypothesis of similar fabric-elasticity relationships.
+
+Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels.
+Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.
+\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well, namely bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls.
+Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI but the trabecular thickness (Tb.Th.) is not significantly different.
+At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\
+
+High resolution peripheral quantitative computed tomography (HR-pQCT) scans allow one to perform \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) in the distal radius and distal tibia \cite{Boutroy2005}.
+Moreover, the image obtained from HR-pQCT can be used for finite element analysis (FEA) to predict mechanical properties \cite{Boutroy2008}.
+Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from the HR-pQCT scan that can be used to assess bone strength within a reasonable computation time \cite{Pahr2009}.
+High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,AriasMoreno2019}.
+Thus, it could be legitimate to use hFE for OI patient's bone strength estimation and potentially fracture risk assessment.
+However, HR-pQCT-based FEA relies on fabric-elasticity relationships.
+Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to investigate the hypothesis of similar fabric-elasticity relationships.
 
@@
 
-\subsection{Subjects}
-The healthy group include a total of 120 patients from a previous reproducibility study performed at the University Department of osteoporosis in Bern \cite{Schenk2020}.
-The sample is composed of 64 female and 56 male subjects aged between 20 and 92 years old with a median age of 26 [22 - 35] years.
-These subjects did not took any medication known to affect bone metabolism nor presented any prior osteoporosis fracture.
-The second group was scanned at the Shriners Hospital for Children and images were shared to our group by the McGill University in Montreal.
-This group is composed of 35 female 15 male individuals leading to 50 OI diagnosed subjects.
-The youngest and oldest patients are 19 and 69 year old, respectively.
-The median age is 44 [33 - 55] years.
-35 subjects were diagnosed with type I OI, 2 with type III, and 13 with type IV.
+\subsection{Participants}
+The healthy group included a total of 120 patients from a previous reproducibility study performed at the University Department of osteoporosis in Bern \cite{Schenk2020}.
+The group was composed of 64 females and 56 males aged between 20 and 92 years old with a mean age of 32 $\pm$ 15 years.
+These subjects had not taken any medication known to affect bone metabolism nor presented with any prior osteoporosis fracture.
+The second group was scanned as part of the ASTEROID study at different locations in Canada, namely at the Shriners Hospital for Children-Canada.
+The study coordination was done by the McGill University in Montreal.
+This group was composed of 35 females and 15 males with confirmed diagnosis of OI Type I, III or IV.
+There were 35 patients diagnosed with OI type I, 2 with type III, and 13 with type IV.
+The participants of the OI group were aged between 19 and 69 years old with a mean age of 44 $\pm$ 14 years.
 
@@
 Switzerland) were performed at the distal tibia on all patients from both groups.
-People of the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020} whereas OI subjects were scanned using the manufacturer's standard protocol.
-The main differences of the in-house protocol with the manufacturer's standard protocol are the following:
-\begin{enumerate}
-	\item The reference line is positioned at the proximal margin of the dense structure formed by the tibia plafond instead of  the subchondral endplate of the ankle joint (standard clinical section) \cite{Whittier2020}.
-	\item Three stacks were scanned proximal to the reference line instead of one stack at 22.5 mm proximal to the reference line for standard clinical section \cite{Whittier2020}.
-\end{enumerate}
-
-These differences are shown in Figure \ref{01_ClinicalSections}.
-For both group, each stack consist of 168 voxels and a resolution of 61 \si{\micro}m in the three principal directions which lead to a thickness about 10.2 mm for each stack.
-Scanning settings were a voltage of 60 kVp, 900 μA, 100 ms integration time for the healthy group as well as for the OI group.
-For the healthy group, motion artefacts of first, middle and last slice were graded.
-The scale used start from 1 (no motion artefacts) to 5 (extreme motion artefacts) as proposed by the manufacturer \cite{Pialat2012}.
+Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
+Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
+On the other hand, participants in the OI group were scanned using the manufacturer's standard protocol i.e.
+the reference line is placed at the subchondral endplate of the ankle joint and one stack was scanned at 22.5 mm proximal to the reference line (standard clinical section) \cite{Whittier2020}, see Figure \ref{01_OI}.
+The scanned region according to the two different protocol are shown side by side in Figure \ref{01_ClinicalSections}.
+For both the healthy and the OI groups, each stack consisted of 168 slices and a voxel resolution of 61 \si{\micro}m in the three principal directions.
+This led to a thickness of roughly 10.2 mm for each stack.
+Standardized scanning settings were used (voltage of 60 kVp, 900 μA, 100 ms integration time) for the healthy group as well as for the OI group.
+For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale of 1 (no motion artefacts) to 5 (extreme motion artefacts), as proposed by \citeauthor{Pialat2012} \cite{Pialat2012}.
 The final grade of each scan was defined as the highest slice grade.
-For the OI group, as the scan consist of one stack, only one grade is attributed using the same scale as for the control group.
+For the OI group, as the scan consisted of one stack, only one grade was attributed using the same scale as for the control group.
 Scans were then processed independently of their quality grading.
-A summary of the scans grading densities is shown in Figure \ref{01_MotionArtefacts}.
+A summary of the scans grading is shown in Figure \ref{01_MotionArtefacts}.
 
@@
 \subsection{Image analysis}
-The HR-pQCT scans were evaluated using the manufacturer's standard protocol. Then, mask segmented images were used for further analysis.\\
-
-Six ROI were randomly selected in each scan.
-The conditions for ROI to be kept are no cortical bone inside and that it must contain trabecular bone.
-For the OI subjects, the stack was divided into two halves and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half.
-For the healthy people, the ROIs were selected in the more proximal stack uniquely, see Figure \ref{01_Healthy} stack n\textdegree 3.
-As for the OI subjects, the stack is halved and centers of three ROIs were selected in both half.\\
-
-The ROI is defined as a cube of 5.3 mm side length.
-This size was chosen to correspond to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} which performed similar analysis with femur \si{\micro}CT scans.
-It was determined by \citeauthor{Zysset1998}\cite{Zysset1998} and \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to be the optimal size to obtain accurate FEA results.\\
-
-The morphological analysis of ROIs was performed using medtool (v4.5; Dr.
+The HR-pQCT scans were evaluated using the manufacturer's standard protocol.
+Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking) and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}).
+Then, mask segmented images were used for further analysis.\\
+
+In general, six cubic ROIs were selected at random position in each scan in a defined area.
+Each ROI had to contain trabecular bone, but no cortical bone.
+For the healthy group, the ROIs were selected in the more proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at the same anatomical location as for the OI group.
+Then for both the healthy group and the OI group, the stack was divided into two halves and the ROIs were selected to have the centres of three ROIs in the proximal half and three in the distal half.
+For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone.
+This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\
+
+The ROI was defined as a cube of 5.3 mm side length.
+This size was in agreement with the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013}, who performed similar analysis with femur \si{\micro}CT scans.
+It was determined by \citeauthor{Zysset1998}\cite{Zysset1998} and \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to allow one to have a relative homogeneity of trabecular tissue within the ROI leading to accurate \si{\micro}FE results with a minimal computational cost.\\
+
+After ROI cleaning, i.e.
+deletion of unconnected region of bone material, the morphological analysis of ROIs was performed using medtool (v4.5; Dr.
 Pahr Ingenieurs e.U., Pfaffstätten, Austria).
-The morphological parameters analyzed are: BV/TV, SMI, trabecular number (Tb.
-N.), trabecular thickness (Tb.
-Th.), trabecular spacing (Tb.
-Sp.), and the standard deviation of the trabecular spacing (Tb.
-Sp.
-SD).
-Moreover, ROI fabric was evaluated using MIL method \cite{Moreno2014}.
-The fabric tensor $\mathbf{M}$ is a positive-definite second-order tensor.
-It is build as shown in Equation \ref{Eq201} below:
+The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular spacing (Tb.Sp.SD).
+Moreover, ROI tissue bone mineral density (tBMD) and fabric was evaluated.
+The fabric tensor $\mathbf{M}$ was computed using mean intercept length (MIL) method \cite{Moreno2014}.
+It is a positive-definite second-order tensor built as shown in Equation \ref{Eq201} below:
 
@@
 where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ \cite{Cowin1985,Harrigan1985}.
-The fabric is then independent of BV/TV and normalized with $tr(\mathbf{M}) = 3$.
-The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.\\
-
-After ROI cleaning, i.e.
-deletion of unconnected region of bone material, an homogenized mechanical analysis was performed using \textsc{ABAQUS 6.14}.
-Each voxel of the cleaned ROI was converted to mesh using fully integrated linear brick elements (C3D8) with a stiffness $E$ of 10 000 MPa and a Poisson's ratio $\nu$ of 0.3.
-The simulation consisted of 6 independent load cases, 3 uni-axial and 3 shear cases, using KUBCs.
-KUBCs were used according to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}.
-Unlike PMUBCs, KUBCs do not require to rotate the ROI into fabric coordinate system which decrease the image quality.
-This homogenization process allows to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
-This model builds the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric information $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$.
-The building of this tensor is shown in Equation \ref{Eq202}.\\
+The fabric tensor is independent of BV/TV and normalized with $tr(\mathbf{M}) = 3$.
+The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.
+Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualization of its fabric tensor.\\
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth, trim= 0 0 0 100]
+	{Pictures/01_FabricExample}
+	\caption{Typical ROI with the visualization of its fabric tensor using MIL method.
+Eigenvectors of the fabric tensor define its orientation and eigenvalues set lengths of the ellipsoid radii.
+DA is the ratio between the highest and the lowest eigenvalue.}
+	\label{01_FabricExample}
+\end{figure}
+
+A \si{\micro}FE mechanical analysis was performed using \textsc{ABAQUS 6.14}.
+In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach.
+Then, a stiffness $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned.
+The homogenization process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}.
+Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require one to rotate the ROI into fabric coordinate system.
+Such rotation would potentially decrease image quality.
+The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
+This model builds the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\
 
@@
 Where $\otimes$ and $\overline{\underline{\otimes}}$ are the dyadic and symmetric product of second order tensors, respectively.
+To express the stiffness tensor obtained from the homogenization process with the Zysset-Curnier model, it had to be transformed into the fabric coordinate system using a coordinates transformation formula (see Equation \ref{Eq203p}) and projected onto orthotropy, leading to 12 components.
+
+\begin{equation}
+	\mathbb{S}_{ijkl}' = Q_{im}Q_{jn}Q_{ko}Q_{lp} \mathbb{S}_{mnop}
+	\label{Eq203p}
+\end{equation}
+
+Where $\mathbb{S}'$ and $\mathbb{S}$ are the transformed and the original stiffness tensor respectively and $Q$ is the orthogonal matrix that maps the original coordinate system into the new one (fabric).
 The Zysset-Curnier model is built with the assumption of orthotropy and homogeneity.
 However, the trabecular structure is not perfectly homogeneous.
-In order to assess the ROI heterogeneity a so-called coefficient of variation (CV) is computed as presented in \cite{Panyasantisuk2015}: the ROI is divided into eight identical subcubes, BV/TV is computed for each subcube and the CV is defined as the ratio between the standard deviation of these BV/TV and the mean value, see Equation \ref{Eq203} below:
-
-\begin{equation}
-	CV = \frac{std(BV/TV_{subcubes})}{mean(BV/TV_{subcubes})}
+In order to assess the ROI heterogeneity, a coefficient of variation (CV) is computed according to \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}: the ROI is divided into eight identical sub-cubes and BV/TV is computed for each of them.
+The CV is defined as the ratio between the standard deviation of these BV/TV and the mean value (Equation \ref{Eq203}).
+
+\begin{equation}
+	CV = \frac{SD(BV/TV_{subcubes})}{mean(BV/TV_{subcubes})}
 	\label{Eq203}
@@
 \subsection{Statistics}
-The morphological parameters analyzed (BV/TV, Tb.
-N., Tb.
-Th., Tb.
-Sp., and Tb.
-Sp.
-SD, SMI, DA, and CV) were compared for both group.
-As the initial groups do not have similar distributions of age and sex, a matching was performed leading to identical mean and median age as well as identical gender distribution.
-Then, the selection of statistical test to perform was executed as follow:
-\begin{enumerate}
-	\item For each parameter, the median value between the six ROI of the same individual was computed. The median was preferred over the mean because it is less influenced by outliers.
-	\item Normality of the distribution was assessed with QQ plot and Shapiro-Wilk test.
-	\item If normality assumption was met, Bartlett test for equal variances was performed.
-Otherwise, data were log transformed to try to achieve normal distribution.
-If even after log transformation, normality assumption was not met, original data space was kept and Brown-Forsythe test was applied to assess the equal variance assumption.
-	\item According to the previous results, t-test was performed if normal distribution and equal variance were met.
-If only the equal variances assumption was met, Mann-Whitney test was preferred.
-Finally, if none of these conditions could be assumed, a non-parametric permutation test was performed.
-\end{enumerate}
+The analyzed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI groups.
+As the groups do not have similar distributions of age and sex, a matching was performed by selecting similar individuals leading to identical mean and median age as well as identical gender distribution.
+For each parameter, the median value between the six ROIs from the same individual was computed.
+The median was preferred over the mean because it is less influenced by outliers.
+Normality of the distribution was assessed with QQ plot and Shapiro-Wilk test.
+CV had to be log-transformed to meet normal distribution assumption.
+Then, equal variances was assessed using Bartlett test or Brown-Forsythe test according to the normality distribution of the data.
+According to the normality and equal variances assumptions, t-test, Mann-Whitney test or a non-parametric permutation test was performed.
 The general significance level was set to 95\% for all tests.
-Confidence intervals in means difference was computed for t-tested variables to confirm p values.
-As Mann-Whitney test is performed on the median, only corresponding p value is given.
-Finally, non-parametric permutation test is less powerful but give an empirical 95\% exclusion range and a p value.
+Confidence intervals was computed for t-tested variables to confirm a difference in both groups means.
+As Mann-Whitney tests are performed on the median, only the corresponding p-value is presented.
+Finally, non-parametric permutation tests are less powerful but give an empirical 95\% exclusion range and a p-value.
 If the difference in means belong to this exclusion range, it can be stated that group means are different with 95\% certainty.
 
-\subsection{Fit to model}
-The stiffness tensors obtained from the mechanical simulations were transformed into fabric coordinate system and projected onto orthotropy, leading to 12 components.
-The resulting orthotropic stiffness tensors were then used to perform a multiple linear regression on the Zysset-Curnier model.
+\subsection{Linear Regression}
+The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model.
 Standard linear models assume independent and identically distributed (iid) variables.
-As this assumption is violated by the fact that six ROIs are analyzed by individual, a linear mixed-effect model was preferred.
-This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, takes into account the non-independence of ROIs from the same individual.
-Detailed form of this model is presented in Appendix \ref{A1}.
+As this assumption was violated by the fact that six ROIs were analyzed per individual, a linear mixed-effect model was preferred.
+This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual.
+A more detailed form of this model is presented in Appendix \ref{A1}.
 
@@
 $Z$ is a $12n$x$f$ design matrix which contains data with individual dependence and $\delta$ is a $f$x1 vector composed of random factors.
-Finally, $\epsilon$ is a $12n$x1 vector containing the regression residuals.\\
-
-The linear regression was performed on both group (healthy and OI) separately.
+Finally, $\epsilon$ is a $12n$x1 vector containing the regression residuals.
+As $k$ and $l$ are exponents, the regression was performed on the log space.\\
+
+The linear regression was performed on both group (heal\-thy and OI) separately.
 To improve the fit quality, the data sets were filtered.
-The aim here is to filter out ROIs whose are too far from the assumption of homogeneity.
+The aim here was to filter out ROIs violating the assumption of homogeneity.
 Therefore, analogously to the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, a fixed threshold for the CV was used.
-To simplify comparison, the same value of 0.263 was fixed as exclusion criterion.
-Besides, the relation between BV/TV and CV was assessed using Spearman's correlation coefficient.
-Furthermore, to compare the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) between the groups, regression must be performed on identical value ranges.
-To do this, a matching was performed for BV/TV and DA to find corresponding control ROI for each OI in the filtered groups.
+To simplify comparison, we used the same value 0.263 as exclusion criterion \cite{Panyasantisuk2015}.
+Then, the relation between BV/TV and CV was assessed using Spearman's correlation coefficient.
+To compare the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) between the groups, regression must be performed on identical value ranges.
+Therefore, a matching was performed for BV/TV and DA to find corresponding control ROIs for each OI in the filtered groups.
 Best correspondences were kept and duplicates were dropped.
-Finally, as the regression is performed in the log space, it is necessary to use identical exponent ($k$ and $l$) for both group, weighting identically BV/TV and DA between regressions, to compare stiffness constants.
+Finally, as the regression was performed in the log space, slight differences in the exponents ($k$ and $l$) would lead to important variation of the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) so it was necessary to use identical exponents  for both groups, weighting identically BV/TV and DA between regressions.
 The exponents were determined by grouping healthy and OI for regression.
-Then a modified system is used to perform the fit on separated groups.\\
-
-Another modification of the model is to add a regressor for the group variable (healthy or OI), i.e.
+Then a modified system was used to perform the fit on separated groups, see Appendix \ref{A1}, Equation \ref{EqA11}.\\
+
+Another modification of the model was to add a regressor for the group variable (healthy or OI), i.e.
 add a column to the design matrix $X$ and a row to the parameter vector $\beta$.
@@
 Implementation of this modification was performed according to \cite{Fox2016}.
+A similar mixed-effect model was used to analyze the relation between tBMD and BV/TV and the significance of the group, see Appendix \ref{A1} Equation \ref{EqA14}.
+The model used the BV/TV and the group (healthy or OI) as fixed variables and the individual as random variable.
+Moreover, to test the hypothesis of no interaction between the BV/TV and the group, i.e.
+the group has no significant influence on the tBMD versus BV/TV slope, the model was modified to add the interaction regressor (BV/TV x Group).
 The detailed linear systems for each model discussed here are available in Appendix \ref{A1} and a summary of the data sets used for the different methods is shown in Table \ref{Table1}.\\
@@
 The regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}.
-Regression quality was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors ($NE$), see Equation \ref{Eq205} and \ref{Eq206}.
+Regression quality for the tBMD analysis was assessed using the Pearson correlation coefficient ($R^2$) and the standard error of the estimate (SE).
+Regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}.
 
@@
 
-Where RSS is the residual sum of squares and TSS is the total sum of squares i.e. sum of the square of the observations y.
-
-\begin{equation}
-	NE = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
+Where RSS is the residual sum of squares and TSS is the total sum of squares i.e. sum of the square of the observations y. $n$ is the number of ROIs and $p$ the number of parameters.
+
+\begin{equation}
+	\text{NE} = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
 	\label{Eq206}
@@
 		Group & Healthy & OI & Healthy & OI & Healthy & OI & Healthy & OI \\
-		Individuals & 120 & 50 & 28 & 28 & 119 & 38 & 58 & 33 \\
-		ROIs & 720 & 300 & 168 & 168 & 603 & 115 & 83 & 83 \\
+		Individuals & 120 & 49 & 28 & 28 & 119 & 38 & 57 & 32 \\
+		ROIs & 720 & 294 & 168 & 168 & 603 & 117 & 82 & 82 \\
 		\midrule
-		Methods & \multicolumn{2}{c}{Fit to model} & \multicolumn{2}{c}{Statistics} & \multicolumn{2}{c}{Fit to model} & \multicolumn{2}{c}{Fit to model} \\
+		Methods & \multicolumn{2}{c}{Linear regression} & \multicolumn{2}{c}{Statistics} & \multicolumn{2}{c}{Linear regression} & \multicolumn{2}{c}{Linear regression} \\
 		\bottomrule
@@
 The results of morphological analysis are summarized in Table \ref{Table2}.
-In the present study, the individual matching allowed to have similar group distribution with 17 females and 11 males in each group.
+In the present study, the individual matching allowed us to have similar group distributions with 17 females and 11 males in each group.
 The mean age of matched healthy is 41 $\pm$ 14 and 41 $\pm$ 15 for the matched OI.
-BV/TV of healthy people is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01.
+BV/TV of healthy individuals is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01.
 Similarly, trabecular number is higher in the matched healthy group with a 95\% CI of [0.099, 0.285] and a corresponding p value <0.001.
 The trabecular thickness do not show significant differences between groups with a p value of 0.2.
-In the other hand, trabecular spacing is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$).
-Trabecular spacing SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$).
+On the other hand, trabecular separation is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$).
+Trabecular separation SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$).
 SMI as well as degree of anisotropy are higher for matched OI than for healthy people with p values <0.001 and of 0.02, respectively.
@@
 
-In the table, absolute values and p values of test statistics are compared with literature.
-Regarding age, the present population is slightly younger than in other study but stay in a similar range.
-In a general way p values of the three other studies show significant differences for BV/TV, Tb N, Tb Sp, and Tb Sp SD and non significant differences for Tb Th.
-Absolute values of BV/TV, TB Th, Tb Sp, and Tb Sp SD are higher in the present study compared to literature.
-On the other hand, Tb N appears to be higher in the other studies than in the present one.
+In Table \ref{Table2}, absolute values and p values of test statistics are compared with literature.
+Regarding age, the present population is slightly younger than in the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} but stay in a similar range.
+In a general way p values of the three other studies (\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018}) show significant differences for BV/TV, Tb.N., Tb.Sp., and Tb.Sp.SD and non significant differences for Tb.Th.
+Absolute values of BV/TV, TB.Th., Tb.Sp., and Tb.Sp.SD are higher in the present study compared to literature.
+On the other hand, Tb.N.
+appears to be higher in the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} than in the present one.
 
@@
 
-		\multirow{3}{*}{Tb N} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
+		\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
 		& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
@@
 
-		\multirow{3}{*}{Tb Th} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
+		\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
 		& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
@@
 
-		\multirow{3}{*}{Tb Sp} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
+		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
 		& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
@@
 
-		\multirow{3}{*}{Tb Sp SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
+		\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
 		& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
@@
 
-\subsection{Linear Regressions}
-Figure \ref{02_GeneralRegression} shows the result of regression for complete data sets of each group separately.
-The x axis represent the values of observed tensors from mechanical simulations.
+\subsection{Linear Regression with Original Data Sets}
+Figure \ref{02_GeneralRegression} shows the results of the regression analysis for the complete data sets of each group separately.
+The x axis represent the values of observed tensors from \si{\micro}FE simulations.
 The y axis is the predicted values using the Zysset-Curnier model and the parameters obtained after performing the regression with linear mixed-effect model.
-The fitted line is represented by the dashed line.
+The fitted line is represented by the dashed line, indicating the theoretical perfect correlation.
 $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components.
@@
 The $R^2_{adj}$ is slightly above 0.95 and the NE is of 18\% $\pm$ 10\%.
-The regression of the OI group could only be performed on 294 ROIs as it was not possible to find suitable ROIs for one patient.
-Nevertheless, it leads to 3528 points, a $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%.
-It can be noticed that, as the values of observed tensor decreases, points tends to be further apart of the diagonal (dashed line).
+The regression analysis for the OI group led to 3528 points, a $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%.
+It can be noticed that, as the values of the observed tensor decreases, points tends to be further apart of the diagonal (dashed line).
 Moreover, the points coming from the tensors with lowest values are only above the diagonal.
 One last observation which should be noted is the range covered by the observed tensors.
-The range of the OI group is wider than the healthy group and ROIs with lower BV/TV present further low stiffness values.\\
+The range of the OI group is wider than the healthy group and ROIs with lower BV/TV present further low stiffness values.
+The values of these ROIs tensors components tends to be overestimated by the fit.\\
 
@@
 
+\subsection{Filtering}
 The CV in relation to BV/TV is shown in Figure \ref{02_CV_BVTV}.
@@
 The regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}.
-After filtering, healthy group is reduced to 119 individuals and 603 ROIs.
-This leads to 7236 points for the regression.
-Results shown similar values as for the complete data set, namely a $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\%, see Figure \ref{04_Healthy}.
+After filtering, the healthy group was reduced to 119 individuals and 603 ROIs.
+This led to 7236 points for the regression.
+Results show similar values to the complete data set, namely a $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\%, see Figure \ref{04_Healthy}.
 The OI group lost more individuals after filtering leading to 38 people and 115 ROIs.
@@
 
-Regression results after BV/TV \& DA ROI matching are shown in Table \ref{Table3}.
+\subsection{BV/TV and DA Matching}
+Regression results after BV/TV and DA ROI matching are shown in Table \ref{Table3}.
 The columns show which data set was used, the fives parameters of the Zysset-Curnier model and the assessment of fit quality.
@@
 Comparison is performed between grouped (N ROIs = 166) and separated data sets (N ROIs = 83).
-Values are presented as value [95\% CI] or mean $\pm$ standard deviation.}
+Values are presented as value [95\% CI] or mean $\pm$ standard deviation.
+Values in gray were imposed in the regression.}
 	\label{Table3}
@@
 		\midrule
-		Grouped & 4626 [3902-5485] & 2692 [2473-2932] & 3537 [3246-3855] & 1.91 [1.86-1.95] & 0.95 [0.93-0.97] & 0.936 & 18 $\pm$ 9\\
-
-		Healthy & 4322 [3855-4846] & 2684 [2535-2842] & 3509 [3307-3724] & 1.91 & 0.95 & 0.836 & 21 $\pm$ 10\\
-
-		OI & 4980 [4359-5689] & 2722 [2547-2909] & 3594 [3355-3849] & 1.91 & 0.95 & 0.861 & 20 $\pm$ 10\\
+		Grouped & 4626 [3892-5494] & 2695 [2472-2937] & 3541 [3246-3862] & 1.91 [1.86-1.95] & 0.95 [0.93-0.97] & 0.936 & 19 $\pm$ 9\\
+
+		Healthy & 4318 [3844-4851] & 2685 [2533-2845] & 3512 [3306-3731] & \textcolor{gray}{1.91} & \textcolor{gray}{0.95} & 0.835 & 21 $\pm$ 10\\
+
+		OI & 4983 [4345-5716] & 2727 [2547-2921] & 3600 [3355-3863] & \textcolor{gray}{1.91} & \textcolor{gray}{0.95} & 0.860 & 20 $\pm$ 10\\
 		\bottomrule
@@
 On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}.
-Setting the exponents $k$ and $l$ to the same values lead to lower stiffness constants for the observed data set than for the other studies.\\
+Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set than for the other studies \cite{Gross2013,Panyasantisuk2015}.\\
 
@@
 Values are presented as computed value only or mean $\pm$ standard deviation.
-The present study shows values obtained with tibia XCTII scans.
-\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013} show values obtained with femur \si{\micro}CT scans.}
+The present study shows values obtained with ROIs of tibia XCTII scans of healthy and OI individuals pooled together.
+\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013} show values obtained with ROIs of femur \si{\micro}CT scans of healthy individuals only.
+Values in gray were imposed in the regression.}
 	\label{Table4}
@@
 
-		\multirow{2}{*}{Present study} & 718 & 0.27 $\pm$ 0.09 & 1.93 $\pm$ 0.24 & 2510 & 1623 & 2056 & 1.55 & 0.82 & 0.832 & 21 $\pm$ 11\\
-		& 718 & 0.27 $\pm$ 0.09 & 1.93 $\pm$ 0.24 & 4775 & 3087 & 3910 & 1.99 & 0.85 & 0.949 & 17 $\pm$ 9\\[1ex]
+		\multirow{2}{*}{Present study} & 720 & 0.27 $\pm$ 0.09 & 1.94 $\pm$ 0.24 & 2507 & 1620 & 2052 & \textcolor{gray}{1.55} & \textcolor{gray}{0.82} & 0.832 & 21 $\pm$ 11\\
+		& 720 & 0.27 $\pm$ 0.09 & 1.94 $\pm$ 0.24 & 4778 & 3087 & 3911 & 1.99 & 0.85 & 0.949 & 17 $\pm$ 9\\[1ex]
 
 		\multicolumn{11}{c}{\cellcolor[HTML]{D9D9D9}Non-filtered data sets}\\
-		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 167 & 0.25 $\pm$ 0.08 & 1.54 $\pm$ 0.20 & 3841 & 3076 & 3115 & 1.60 & 0.99 & 0.983 & 14\\
+		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 167 & 0.25 $\pm$ 0.08 & 1.54 $\pm$ 0.20 & 3841 & 3076 & 3115 & \textcolor{gray}{1.60} & \textcolor{gray}{0.99} & 0.983 & 14\\
 
@@
 
-		\multirow{2}{*}{Present study} & 1014 & 0.23 $\pm$ 0.11 & 1.94 $\pm$ 0.26 & 2727 & 1655 & 2179 & 1.60 & 0.99 & 0.608 & 41 $\pm$ 182\\
-		& 1014 & 0.23 $\pm$ 0.11 & 1.94 $\pm$ 0.26 & 5323 & 3232 & 4254 & 2.02 & 0.91 & 0.915 & 31 $\pm$ 105\\
+		\multirow{2}{*}{Present study} & 1014 & 0.23 $\pm$ 0.11 & 1.94 $\pm$ 0.26 & 2738 & 1662 & 2187 & \textcolor{gray}{1.60} & \textcolor{gray}{0.99} & 0.622 & 40 $\pm$ 177\\
+		& 1014 & 0.23 $\pm$ 0.11 & 1.94 $\pm$ 0.26 & 5020 & 3047 & 4010 & 1.98 & 0.91 & 0.916 & 30 $\pm$ 113\\
 		\bottomrule
@@
 
+The analysis of tBMD in relation to BV/TV is shown in Figure \ref{02_tBMD}.
+The t-test performed on the tBMD distributions led to a p-value <0.0001 and a 95\% CI of [-35,-13].
+The linear regression performed using the linear mixed-effects model without BV/TV and group interaction provided a slope of 223 [153,292] (value [95\% CI]).
+The intercept was 534 [518,550] and the group variable led to a value of 14 [7,20].
+The prediction using the fixed effects only led to a $R^2$ of 0.23 and a SE of 33.
+Then, using the linear mixed-effect model with the interaction regressor (BV/TV x Group), this added variable presented a value of 3 [-66,72] and a p-value of 0.94.
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]
+	{Pictures/05_tBMDvsBVTV}
+	\caption{tBMD in relation to BV/TV. The fitted lines are obtained using the fixed effects of the linear mixed-effect model and fixing the group variable.}
+	\label{02_tBMD}
+\end{figure}
+
 \section{Discussion}
-The age of matched individuals lies in the spectrum of the others studies which allow to compare the morphological values.
+Osteogenesis imperfecta is an inherited form of bone fra\-gi\-li\-ty with a severity going from mild to perinatally lethal.
+This study aim to confirm that fabric-elasticity relationships in OI trabecular bone are similar than in healthy conditions, encouraging the use of HR-pQCT scans for fracture risk assessment.
+To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.\\
+
+The age of matched individuals lies in the spectrum of the others studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} which allow one to compare the morphological values.
 The main explanation for differences between the absolute values of the present study compared to the others lies in the imaging system.
-\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed first generation Xtreme CT scans and the present study use image from XCTII.
-The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} showed that BV/TV, Tb Th, and Tb Sp is higher in second generation Xtreme CT and, on the other hand, Tb N is higher in first generation Xtreme CT.
-These results give confidence on the observed values.
+\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed first generation Xtreme CT scans with a voxel size of 82 \si{\micro}m and the present study used a second generation scanner, XCTII, with a voxel size of 61 \si{\micro}m.
+The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} investigated differences between the two scanner types.
+They showed that BV/TV, Tb.Th., and Tb.Sp.
+is higher in second generation Xtreme CT and, on the other hand, Tb.N.
+is higher in first generation Xtreme CT.
+These results give confidence to the observed values.
 Another bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length.
-This conditions the Tb N and Tb Sp as they depend on the ROI size.
-Moreover, conditions imposed for ROI random selection can lead to further biased values, specially of OI patient, as it can not be empty of trabecular bone.
-The CV presenting the stronger significant difference between groups even given the low sample size show that heterogeneity is a main difference in OI patient compared to healthy individuals.
-Finally, the significant differences observed in BV/TV and DA even with age \& gender matched individuals justify the choice of a variable matching for fabric-elasticity relationships analysis.\\
+This conditions the Tb.N.
+and Tb.Sp.
+as they depend on the ROI size.
+Moreover, conditions imposed for ROI random selection can lead to further biased values, especially for OI patients, as it can not be empty of trabecular bone.
+The CV presenting the stronger significant difference between groups even given the low sample size show that heterogeneity is a main difference in OI patients compared to healthy individuals.
+Finally, the significant differences observed in BV/TV and DA even with age and gender matched individuals justify the choice of a variable matching for fabric-elasticity relationships analysis.\\
 
@@
 These differences come from the ROIs presenting a low stiffness.
-It can be seen on the regression plot that when the stiffness term decrease about $10^0$ and under, the fit tends to overestimate the stiffness.
+It can be seen on the regression plot (Figure \ref{02_OI}) that when the stiffness term decreases lower than $10^0$ MPa, the fit tends to overestimate the stiffness.
 This is due to the fact that ROI stiffness is highly impacted by BV/TV.
@@
 Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation.
-As some higher BV/TV ROIs still present high CV, imposing a fixed threshold make sense for subsequent homogenization.\\
+As some higher BV/TV ROIs still present high CV, imposing a fixed threshold make sense for subsequent homogenization.
+However, the value of this threshold actually results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\
 
 The fits performed on filtered data sets present direct effect of filtering.
-For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (5\textperthousand) is negligible.
+For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) is negligible.
 On the other hand, NE presents an improved value decreasing by 2\%.
-These results are due to the filtering of point further away from the diagonal (better NE) and some points close to the diagonal leading to a smaller number of points ($R^2_{adj}$).
-For the OI group, filtering leads to a important improvement of the fit.
+These results are due to the filtering of points further away from the diagonal (better NE) and some points close to the diagonal leading to a smaller number of points ($R^2_{adj}$).
+For the OI group, filtering leads to a important improvement in the fit.
 Results become similar to the healthy group in terms of $R^2_{adj}$, NE, and the range of stiffness values.
-These results give confidence to the filtering procedure and are a first step to the hypothesis of no different fabric-elasticity relationships between healthy and OI trabecular bone.\\
-
-After BV/TV and DA matching, grouping the data sets together lead to similar $R^2_{adj}$ and NE as for individual filtered data set.
-This allows to give values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m.
-Imposing these values to perform the fit on individual match data set allow to highlight differences, if any, between healthy and OI trabecular bone.
+These results give confidence to the filtering procedure and are a first step to the hypothesis of there being no difference in fabric-elasticity relationships between healthy and OI trabecular bone.\\
+
+After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for individual filtered data sets.
+This allows one to give values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m.
+Imposing these values to perform the fit on individual matched data sets allows us to highlight differences, if any, between healthy and OI trabecular bone.
 The relatively low differences for $\lambda_0'$ and $\mu_0$ once again provide arguments for similar relationships between the two groups.
-For $\lambda_0$ this relative difference being higher could rise some doubts about this similarity but 95\% CI intervals still show a common range.
+For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but 95\% CI intervals still show a common range.
 Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group shown a p value far above the 5\% significance level.
@@
 
-Imposing $k$ and $l$ allows to estimate the effect of different image resolutions.
+Imposing $k$ and $l$ allows one to estimate the effect of different image resolutions.
 \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened it to 36 \si{\micro}m.
@@
 Performing a fit on this less optimized structure lead to the observed lower stiffness constants.
-Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} show that lower spatial resolution lead to lower fit quality.
+Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality.
 Nevertheless, $l$ stands in the same range as for the two other studies meaning the relative weight of DA remains constant.
@@
 
+The analysis of tBMD gives interesting outputs as well.
+Although BV/TV and DA are in the same range because of the matching, the t-test reveal a higher tBMD in OI trabecular bone than in healthy condition with 95\% certainty and a very high significance level.
+The coefficients obtained from the linear mixed-effects model show that there is a relation between tBMD and BV/TV as zero is not included in the slope CI.
+This result can have different origins.
+From a biological point of view, the remodeling process leads to a mineralization gradient from the core of the trabecula to the outer surface.
+As trabecular thickness decreases with BV/TV, this means that with a lower BV/TV the core of trabeculae could be less mineralized.
+The second explanation for this slope comes from the scanning.
+Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV.
+Nevertheless, the former (biological) explanation is expect to have a less significant impact than the latter (scanning).
+Regarding the other coefficients of the linear regression, the CI of group variable exclude zero as well leading to the conclusion that we have 95\% certainty that the intercept is different depending on the group.
+This could be explained by the bisphosphonate treatment that OI patients receive.
+Effectively, bisphosphonate is aimed to freeze the remodeling process which lead to higher mineralization of the bone.
+According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest to add a correction accounting for the tBMD in FE simulations to catch the higher modulus, ultimate stress and post-yield behavior of OI bone compared to healthy bone at the ECM level.
+Regarding the interaction between BV/TV and the group, the p-value obtained show a high non-significance meaning that the slopes of both OI and healthy groups are the same.
+This could be visualized in Figure \ref{02_tBMD} where the fitted lines are obtained using the fixed effects of the model and the group variable if fixed.
+In this plot, performed with the interaction term (BV/TV x Group) the fitted lines appear to be quasi-parallel.
+
 The main limitations of this study are the definition of "homogeneous enough" and the fact that it is limited to tibiae XCTII scans.
+Moreover, having only one patient with OI type III where we could extract ROIs does not allows to do statistics.
+Effectively as those patients are in wheelchair, it could be interesting to analyze the impact of this condition on the weight bearing tibia.
 The ROI homogeneity has an important impact on the analysis quality.
 As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for fitting.
-More investigations could be performed to improve the model for highly heterogeneous ROIs but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible.
-Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.\\
-
-In conclusions, the sample analyzed in the this study have similar morphology that data reported in the literature.
+More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible.
+Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.
+Another limitation is that the scans were performed on different devices which were not cross-calibrated.
+However, as they are the same model, it is expected to have a minor impact.\\
+
+In conclusions, the sample analyzed in the this study have similar morphology to data reported in the literature.
 We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROI is homogeneous enough.
@@
 If tensile and shearing behaviors agree, fabric-strength relationships will hold too.
-Therefore, OI bone fragility might mostly result from the decrease in BV/TV and loss of homogeneity in its organization.
+Therefore, OI trabecular bone can explain part of the bone fragility by the decrease in BV/TV and loss of homogeneity in its trabecular organization.
 
@@
 This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.
-
-\clearpage
+Bettina M.
+Willie is supported by the Shriners Hospital for Children and the FRQS Programme de bourses de chercheur.
+
+%\clearpage
 \appendix
@@
 	\end{split}
+	\label{EqA11}
 \end{equation}
@@
 Where $S_g$ is coded using a summation constrain \cite{Fox2016}, meaning, $S_g = -1$ for the healthy group and $S_g = 1$ for the OI group.
+This model is adapted into a linear mixed-effect model to analyze the relation between tBMD and BV/TV and the effect of the group.
+
+\begin{equation}
+	\begin{split}
+		\text{tBMD} = \begin{pmatrix}
+			1 & \rho & S_g \\
+		\end{pmatrix} \begin{pmatrix}
+			\beta_1 \\
+			\beta_2 \\
+			\beta_3 \\
+		\end{pmatrix} + \begin{pmatrix}
+			1 & 1 \\
+		\end{pmatrix}\begin{pmatrix}
+			\delta_1 \\
+			\delta_2 \\
+		\end{pmatrix} + \epsilon
+	\end{split}
+	\label{EqA14}
+\end{equation}
+
+Where $\delta_1$ and $\delta_2$ represent the intercept and the slope for each different individual respectively.
+To test the hypothesis of no interaction between the BV/TV and the group, i.e.
+the group has no significant influence on the tBMD versus BV/TV slope, the previous model was modified to add the interaction regressor.
+
+\begin{equation}
+	\begin{split}
+		\text{tBMD} = \begin{pmatrix}
+			1 & \rho & S_g & \rho S_g\\
+		\end{pmatrix} \begin{pmatrix}
+			\beta_1 \\
+			\beta_2 \\
+			\beta_3 \\
+			\beta_4 \\
+		\end{pmatrix} + \begin{pmatrix}
+			1 & 1 \\
+		\end{pmatrix}\begin{pmatrix}
+			\delta_1 \\
+			\delta_2 \\
+		\end{pmatrix} + \epsilon
+	\end{split}
+\end{equation}
 

### 2021-06-02 12_SBMS_Presentation/SBMS_Simon_et_al.tex
Finish tBMD analysis and prepare SBMS presentation

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (01/07/19)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------
%	PACKAGES AND THEMES		-----------------------------------
%----------------------------------------------------------------

\documentclass[xcolor=table,11pt]{beamer}

\mode<presentation> {

\usetheme{Frankfurt}
\usecolortheme{dove}
\usefonttheme{serif}

}
\usepackage{newtxtext,newtxmath}
\usepackage{graphicx}
\usepackage{booktabs} 
\usepackage{subfig}
\usepackage{pgf}
\usepackage{multirow}
\usepackage{appendixnumberbeamer}
\usepackage{bookmark}
\usepackage{siunitx}
\usepackage{animate}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{pifont}
\usepackage{caption}
\captionsetup{skip=0pt,belowskip=0pt}

%----------------------------------------------------------------
%	GENERAL OPTIONS 	-----------------------------------------
%----------------------------------------------------------------

% Set template options
\setbeamertemplate{section in toc}{\inserttocsectionnumber.~\inserttocsection}
\setbeamertemplate{frametitle}{\vspace*{1em}\insertframetitle}
\setbeamertemplate{enumerate items}[default]
\setbeamercolor{section in head/foot}{fg=white, bg=black}

% Headline
\makeatletter
\setbeamertemplate{headline}
{%
  \pgfuseshading{beamer@barshade}%
    \vskip-5ex%
  \begin{beamercolorbox}[ignorebg,ht=2.25ex,dp=3.75ex]{section in head/foot}
  \insertsectionnavigationhorizontal{\paperwidth}{\hskip0pt plus1fill}{\hskip0pt plus1fill}
  \end{beamercolorbox}%
  \ifbeamer@sb@subsection%
    \begin{beamercolorbox}[ignorebg,ht=2.125ex,dp=1.125ex,%
      leftskip=.3cm,rightskip=.3cm plus1fil]{subsection in head/foot}
      \usebeamerfont{subsection in head/foot}\insertsubsectionhead
    \end{beamercolorbox}%
  \fi%
}%
\makeatother

% Footline
\makeatletter
\setbeamertemplate{footline}
{
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,left]{section in head/foot}%
    \usebeamerfont{author in head/foot}\hspace{10pt}\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{section in head/foot}%
    \usebeamerfont{title in head/foot}\insertshorttitle
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{section in head/foot}%
    \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}
    \insertframenumber{}\hspace*{2em}
  \end{beamercolorbox}}%
  \vskip0pt%
}
\makeatother

% Add logo
\logo{\pgfputat{\pgfxy(0,7)}{\includegraphics[width=0.1\paperwidth]{Pictures/00_Unibe_Logo}}}

% Table settings
\renewcommand{\arraystretch}{2}
\captionsetup{labelformat=empty,labelsep=none}
\definecolor{Gray}{gray}{0.9}

% Define highlitghing command
\makeatletter
\let\HL\hl
\renewcommand\hl{%
	\let\set@color\beamerorig@set@color
	\let\reset@color\beamerorig@reset@color
	\HL}
\makeatother

% Add overview at each begin of section
%\AtBeginSection[]
%{
%	\begin{frame}
%		\frametitle{Overview}
%		\tableofcontents[currentsection]
%	\end{frame}
%}

\renewcommand{\arraystretch}{1.2}

%----------------------------------------------------------------
%	TITLE PAGE 	-------------------------------------------------
%----------------------------------------------------------------

\title[Fabric-Elasticity Relationships in OI]{
\uppercase{Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals}
}

\author[Simon et al.]{Simon M.\inst{a} \and Indermaur M.\inst{a} \and Schenk D.\inst{a} \and Mahdi T.\inst{b,c} \and Willie B. M.\inst{b,c} \and Zysset P.\inst{a}}
\institute{\inst{a}ARTORG Center of Biomedical Engineering Research\\University of Bern \and \inst{b}Departement of Pediatric Surgery\\McGill University, Montreal \and \inst{c}Research Centre\\Shriners Hospital for Children, Montreal
\medskip
}
\date{June 8, 2021}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\begin{frame}
	\frametitle{Overview}
	\tableofcontents
\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\section{Introdution}
\begin{frame}
	\frametitle{Osteogenesis Imperfecta (OI)}
	Metabolic bone disorder
	\begin{itemize}
		\item Prevalence \textasciitilde 1/13'500 birth
		\item Mutations in genes encoding collagen type I
		\item Different Types: I, II, III, and IV
		\item Increased bone fragility $\rightarrow$ \textit{Brittle bone disease}
		\item Blue sclera, hearing loss
	\end{itemize}
	\vfill
	Treatment
	\begin{itemize}
		\item Gold standard: bisphosphonate
		\item Nailing
		\item Physiotherapy
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Study Objective}
	\fontsize{10pt}{10pt}\selectfont
	HR-pQCT scans...
	\begin{itemize}
		\item Distal skeleton \textit{in vivo} assessment
		\item Microstructure analysis
	\end{itemize}\vfill
	Homogenized FE (hFE)...
	\begin{itemize}
		\item Bone strength estimation based on HR-pQCT scans
		\item High correlations between hFE and experiment
	\end{itemize}\vfill
	But...
	\begin{itemize}
		\item HR-pQCT-based FEA relies on fabric-elasticity relationships
	\end{itemize}\vfill
	Therefore...
	\begin{itemize}
		\item Compare trabecular bone microstructure
		\item Investigate fabric-elasticity relationships in OI
	\end{itemize}
\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\section{Methods}
\subsection{HR-pQCT Scans}
\begin{frame}
	\frametitle{HR-pQCT Scans}
	\begin{columns}
		\column{.35\textwidth}
		Healthy group:
		\begin{itemize}
			\item 120 participants
			\item In-house protocol
		\end{itemize}\vspace{1em}

		OI group:
		\begin{itemize}
			\item 50 participants
			\item Standard protocol
		\end{itemize}

		\column{.2\textwidth}
		\centering		
		\includegraphics[width=1\linewidth,trim=200 22 200 0]
		{Pictures/01_ControlClinicalSection2}\\
		Healthy

		\column{.2\textwidth}
		\centering
		\includegraphics[width=1\linewidth,trim=200 0 200 22]
		{Pictures/01_OIClinicalSection2}\\
		OI

	\end{columns}
\end{frame}

%----------------------------------------------------------------

\subsection{Image Analysis}
\begin{frame}
	\frametitle{Image Analysis}
	\begin{columns}
		\column{.5\textwidth}
		Region of Interest (ROI)
		\begin{itemize}
			\item 6 cubes of 5.3mm
		\end{itemize}\vspace{1em}
		Morphometric analysis
		\begin{itemize}
			\item Trabecular morphology\\
				  Bone volume fraction (BV/TV)\\
			\item Fabric tensor $\mathbf{M}$\\
				  Degree of anisotropy (DA)
			\item Coefficient of variation (CV)
		\end{itemize}

		\column{.4\textwidth}
		\centering
		\includegraphics[width=1\linewidth,trim=100 0 100 0]
		{Pictures/01_FabricExample}\\
		Fabric tensor
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Fit to Model}
	\begin{columns}
		\column{.5\textwidth}
		Stiffness tensor $\mathbb{S}$
		\begin{itemize}
			\item \si{\micro}FE homogenization
			\item Transformation and projection
		\end{itemize}\vspace{1em}
		Linear Regression
		\begin{itemize}
			\item Linear mixed-effects model\\
			\item Fit to Zysset-Curnier model\\
			$\mathbb{S}$ \textasciitilde \textcolor{blue}{$\lambda_{0}$}, \textcolor{blue}{$\lambda_{0}$'}, \textcolor{blue}{$\mu_{0}$}, $m_{i}m_{j}^{\textcolor{red}{k}}$, BV/TV$^{\textcolor{red}{l}}$
			\item Log space for \textcolor{red}{$k$} and \textcolor{red}{$l$}
			\item $R_{adj}^{2}$ and NE for fit quality
		\end{itemize}

		\column{.4\textwidth}
		\centering
		\includegraphics[width=1\linewidth,trim=100 0 100 0]
		{Pictures/01_StiffnessExample}\\
		Stiffness tensor
	\end{columns}
\end{frame}

\begin{frame}
	\frametitle{Healthy vs OI Comparison}
	\begin{columns}[t]
		\column{.5\textwidth}
		Morphometric analysis
		\begin{itemize}
			\item Age and gender matching
		\end{itemize}

		\column{.5\textwidth}
		Linear Regression
		\begin{itemize}
			\item Fit using original data sets
			\item CV filtering
			\item Fit using filtered data sets
			\item BV/TV and DA matching
		\end{itemize}
	\end{columns}

	\vfill

	\begin{table}[b]
		\centering
		\caption{Summary of the data set used for different comparison\vspace{-1em}}
		\label{Table1}\resizebox{1.\linewidth}{!}{\begin{tabular}{p{0.15\linewidth}*{2}{>{\centering\arraybackslash}p{0.1\linewidth}}*{2}{>{\centering\arraybackslash}p{0.15\linewidth}}*{2}{>{\centering\arraybackslash}p{0.1\linewidth}}*{2}{>{\centering\arraybackslash}p{0.15\linewidth}}}
				\toprule
				Data sets & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Age \& gender matched} & \multicolumn{2}{c}{CV filtered} & \multicolumn{2}{c}{BV/TV \& DA matched} \\
				\midrule
				Group & Healthy & OI & Healthy & OI & Healthy & OI & Healthy & OI \\
				Individuals & 120 & 49 & 28 & 28 & 119 & 38 & 57 & 32 \\
				ROIs & 720 & 294 & 168 & 168 & 603 & 117 & 82 & 82 \\
				\midrule
				Methods & \multicolumn{2}{c}{Linear regression} & \multicolumn{2}{c}{Morphometric} & \multicolumn{2}{c}{Linear regression} & \multicolumn{2}{c}{Linear regression} \\
				\bottomrule
		\end{tabular}}
	\end{table}

\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\section{Results and Discussion}
\subsection{Morphological Analysis}
\begin{frame}
	\frametitle{Morphometric Analysis}
	\begin{table}[]
		\centering
		\caption{Summary of morphometric analysis\vspace{-1em}}
		\label{Table2}\resizebox{1.\linewidth}{!}{
		\begin{tabular}{lcccccccc}
				\toprule
				Variable & BV/TV & Tb.N. & Tb.Th. & Tb.Sp. & Tb.Sp.SD & SMI & DA & Ln(CV) \\
				Healthy ? OI & > & > & = & < & < & < & < & < \\
				p value & <0.01 & <0.001 & 0.2 & 0.01 & 0.02 & <0.001 & 0.02 & <0.0001 \\
				\bottomrule
		\end{tabular}}
	\end{table}\vfill
	Results in agreement with:
	\begin{itemize}
		\item \cite{p9}
		\item \cite{p10}
		\item \cite{p11}
	\end{itemize}
\end{frame}

%----------------------------------------------------------------

\subsection{Linear Regression and Filtering}
\begin{frame}
	\frametitle{Linear Regression - Original Data Sets}
	\begin{itemize}
		\item Healthy results in the expected range
		\item OI fit overestimation for low $\mathbb{S}$ components (low BV/TV)
	\end{itemize}
	\begin{columns}
		\column{0.5\textwidth}
		\centering
		\includegraphics[width=1.\linewidth]{Pictures/02_GR_Healthy_LMM}\\
		Healthy group
		\column{0.5\textwidth}
		\centering
		\includegraphics[width=1.\linewidth]{Pictures/02_GR_OI_LMM}\\
		OI group
	\end{columns}	
\end{frame}

\begin{frame}
	\frametitle{Linear Regression - CV Filtering}
	\fontsize{10pt}{10pt}\selectfont
	\begin{columns}
		\column{0.5\textwidth}
		Homogeneity assumption
		\begin{itemize}
			\item High CV decreases fit quality
		\end{itemize}

		\vspace{10pt}

		CV is related to BV/TV
		\begin{itemize}
			\item Remove high CV ROIs
		\end{itemize}

		\vspace{10pt}

		Remove high CV ROIs
		\begin{itemize}
			\item Improve fit quality
			\item Increase results accuracy
		\end{itemize}

		\column{0.5\textwidth}
		\centering
		\includegraphics[width=1.\linewidth]{Pictures/03_CV_BVTV_ROI}\\
		CV in relation to BV/TV
	\end{columns}	
\end{frame}

\begin{frame}
	\frametitle{Linear Regression - Filtered Data Sets}
	\begin{itemize}
		\item \textasciitilde 1/6 healthy ROIs removed $\rightarrow$ Results in the same range
		\item \textasciitilde 6/10 OI ROIs removed \hspace{0.52cm}$\rightarrow$ Results similar to healthy group
	\end{itemize}
	\begin{columns}
		\column{0.5\textwidth}
		\centering
		\includegraphics[width=1.\linewidth]{Pictures/04_FR_Healthy_LMM}\\
		Healthy group
		\column{0.5\textwidth}
		\centering
		\includegraphics[width=1.\linewidth]{Pictures/04_FR_OI_LMM}\\
		OI group
	\end{columns}	
\end{frame}

%----------------------------------------------------------------

\subsection{BV/TV and DA Matching}

\begin{frame}
	\frametitle{Linear Regression - BV/TV and DA Matching}

	Regression conditions:
	\begin{itemize}
		\item Matching allow to fit over same range
		\item Identical $k$ and $l$ for healthy and OI comparison
	\end{itemize}

	\vfill

	\begin{table}[]
		\centering
		\caption{Summary of BV/TV and DA matched data set linear regression\vspace{-1em}}
		\label{Table3}\resizebox{1.\linewidth}{!}{
		\begin{tabular}{cccccccc}
				\toprule
				Data set & $\lambda_0$ & $\lambda_0'$ & $\mu_0$ & $k$ & $l$ & $R^2_{adj}$ & NE (\%) \\
				\midrule
				Grouped & 4626 [3892-5494] & 2695 [2472-2937] & 3541 [3246-3862] & 1.91 [1.86-1.95] & 0.95 [0.93-0.97] & 0.936 & 19 $\pm$ 9\\

				Healthy & 4318 [3844-4851] & 2685 [2533-2845] & 3512 [3306-3731] & \textcolor{gray}{1.91} & \textcolor{gray}{0.95} & 0.835 & 21 $\pm$ 10\\

				OI & 4983 [4345-5716] & 2727 [2547-2921] & 3600 [3355-3863] & \textcolor{gray}{1.91} & \textcolor{gray}{0.95} & 0.860 & 20 $\pm$ 10\\
				\bottomrule
		\end{tabular}}
	\end{table}

	\vfill

	$\rightarrow$ Overlapping stiffness values: Similar fabric-elasticity relationships
\end{frame}

\begin{frame}
	\frametitle{BV/TV and DA Matching - Tissue BMD (tBMD)}
	\fontsize{10pt}{10pt}\selectfont
	\begin{columns}
		\column{0.5\textwidth}

		Results
		\begin{itemize}
			\item tBMD OI > Healthy
			\item tBMD related to BV/TV\\
			$\rightarrow$ Partial volume effect
		\end{itemize}

		\vspace{10pt}

		\cite{p12} tBMD
		\begin{itemize}
			\item Modulus
			\item Ultimate stress
			\item Post-yield behavior
		\end{itemize}

		\vspace{10pt}

		$\rightarrow$ Use tBMD in hFE

		\column{0.5\textwidth}
		\centering
		\includegraphics[width=1.\linewidth]{Pictures/05_tBMDvsBVTV}\\
		tBMD in relation to BV/TV
	\end{columns}
\end{frame}

%----------------------------------------------------------------

\subsection{Comparison with Other Resolutions}

\begin{frame}
	\frametitle{Comparison with Other Resolutions}
	\fontsize{10pt}{10pt}\selectfont

	\begin{table}[]
		\centering
		\caption{Summary of studies compared\vspace{-1em}}
		\label{Table4}\resizebox{1.\linewidth}{!}{
			\begin{tabular}{lccc}
				\toprule
				Study & Scanned site & Scan & Resolution \\
				\midrule

				\cite{p13} & Multiple sites & \si{\micro}CT & 18\si{\micro}m $\rightarrow$ 36\si{\micro}m \\

				\cite{p7} & Proximal femur & \si{\micro}CT & 18\si{\micro}m $\rightarrow$ 36\si{\micro}m \\

				Present study & Distal tibia & HR-pQCT & 61\si{\micro}m \\
				\bottomrule
		\end{tabular}}
	\end{table}

	\vfill

	\begin{columns}[t]
		\column{0.5\textwidth}
		Higher DA
		\begin{itemize}
			\item Distal tibia vs proximal femur
		\end{itemize}

		\vspace{7pt}

		Lower stiffness constants ($\lambda_0$, $\lambda_0'$, $\mu_0$)
		\begin{itemize}
			\item 61\si{\micro}m vs 36\si{\micro}m
		\end{itemize}

		\vspace{7pt}

		Lower fit quality ($R^2_{adj}$ and NE)
		\begin{itemize}
			\item 61\si{\micro}m vs 36\si{\micro}m
		\end{itemize}

		\column{0.5\textwidth}
		Higher $k$
		\begin{itemize}
			\item Higher weight of BV/TV
		\end{itemize}

		\vspace{7pt}

		Similar $l$
		\begin{itemize}
			\item Similar weight of DA
		\end{itemize}

	\end{columns}
\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\section{Conclusions}
\begin{frame}
	\frametitle{Conclusions}

	Bone samples...
	\begin{itemize}
		\item Similar as reported in literature
	\end{itemize}

	\vfill

	Fabric-elasticity relationships...
	\begin{itemize}
		\item Similar in OI and healthy tibial trabecular bone
	\end{itemize}

	\vfill

	But tBMD higher in OI...
	\begin{itemize}
		\item \cite{p12} $\rightarrow$ use tBMD in hFE
	\end{itemize}

	\vfill

	Moreover high CV in OI...
	\begin{itemize}
		\item Homogenization process should be adapted
	\end{itemize}

\end{frame}

\begin{frame}
	Thank you for your attention !	
\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------
\appendix

\section{References}
\begin{frame}
	\frametitle{Osteogenesis Imperfecta}
	\footnotesize{
		\begin{thebibliography}{99}
			\setbeamertemplate{bibliography item}[triangle]

			\bibitem[Mortier et al. 2019]{p1} Mortier et al. (2019)
			\newblock Nosology and classification of genetic skeletal disorders: 2019 revision
			\newblock \textit{American Journal of Medical Genetics, Part A} (179), 2393-2419

			\bibitem[Tournis 2018]{p2} Tournis, S., Dede, A.D. (2018)
			\newblock Osteogenesis imperfecta A clinical update
			\newblock \textit{Metabolism: Clinical and Experimental} (80), 27-37

			\bibitem[Lim et al. 2017]{p3}  Lim, J., Grafe, I., Alexander, S., Lee, B. (2017)
			\newblock Genetic causes and mechanisms of osteogenesis imperfecta
			\newblock \textit{Bone} (102), 40-49

		\end{thebibliography}
	}
\end{frame}

\begin{frame}
	\frametitle{HR-pQCT Scans Protocols}
	\footnotesize{
		\begin{thebibliography}{99}
			\setbeamertemplate{bibliography item}[triangle]

			\bibitem[Schenk et al. 2020]{p4}  Schenk, D., Mathis, A., Lippuner, K., Zysset, P. (2020)
			\newblock In vivo repeatability of homogenized finite element analysis based on multiple HR-pQCT sections for assessment of distal radius and tibia	strength
			\newblock \textit{Bone} (141)

			\bibitem[Whittier et al. 2020]{p5} Whittier et al. (2020)
			\newblock Guidelines for the assessment of bone density and microarchitecture in vivo using high-resolution peripheral quantitative computed tomography
			\newblock \textit{Osteoporosis International} (31), 1607-1627

		\end{thebibliography}
	}
\end{frame}

\begin{frame}
	\frametitle{Region of Interest Analysis}
	\footnotesize{
		\begin{thebibliography}{99}
			\setbeamertemplate{bibliography item}[triangle]

			\bibitem[Daszkiewicz et al. 2017]{p6} Daszkiewicz, K., Maquer, G., Zysset, P.K. (2017)
			\newblock The effective elastic properties of human trabecular bone may be approximated using micro-finite element analyses of embedded volume elements
			\newblock \textit{Biomechanics and Modeling in Mechanobiology} (16), 731-742

			\bibitem[Panyasantisuk et al. 2015]{p7} Panyasantisuk, J., Pahr, D.H., Gross, T., Zysset, P.K. (2015)
			\newblock Comparison of Mixed and Kinematic Uniform Boundary Conditions in Homogenized Elasticity of Femoral Trabecular Bone Using Microfinite Element Analyses
			\newblock \textit{Journal of Biomechanical Engineering} (137), 1-7

			\bibitem[Zysset 1995]{p8}  Zysset, P.K., Curnier, A. (1995)
			\newblock An alternative model for anisotropic
			elasticity based on fabric tensors
			\newblock \textit{Mechanics of Materials} (21), 243-250

		\end{thebibliography}
	}
\end{frame}

\begin{frame}
	\frametitle{Morphometric Analysis}
	\footnotesize{
		\begin{thebibliography}{99}
			\setbeamertemplate{bibliography item}[triangle]

			\bibitem[Folkestad et al. 2012]{p9} Folkestad, L., Hald, J.D., Hansen, S., Gram, J., Langdahl, B., Abra-
			hamsen, B., Brixen, K. (2012)
			\newblock Bone geometry, density, and microarchitecture in the distal radius and tibia in adults with osteogenesis imperfecta type i assessed by high-resolution pQCT
			\newblock \textit{Journal of Bone and Mineral Research} (27), 1405-1412

			\bibitem[Kocijan et al. 2015]{p10} Kocijan, R., Muschitz, C., Haschka, J., Hans, D., Nia, A., Geroldinger, A., Ardelt, M., Wakolbinger, R., Resch, H. (2015)
			\newblock Bone structure assessed by HR-pQCT, TBS and DXL in adult patients with different types of osteogenesis imperfecta
			\newblock \textit{Osteoporosis International
			} (26), 2431-2440

			\bibitem[Rolvien et al. 2018]{p11} Rolvien, T., Stürznickel, J., Schmidt, F.N., Butscheidt, S., Schmidt, T., Busse, B., Mundlos, S., Schinke, T., Kornak, U., Amling, M., Oheim, R. (2018)
			\newblock Comparison of Bone Microarchitecture Between
			Adult Osteogenesis Imperfecta and Early-Onset Osteoporosis
			\newblock \textit{Calcified Tissue International} (103), 512-521

		\end{thebibliography}
	}
\end{frame}

\begin{frame}
	\frametitle{Other References}
	\footnotesize{
		\begin{thebibliography}{99}
			\setbeamertemplate{bibliography item}[triangle]

			\bibitem[Indermaur et al. 2021]{p12}  Indermaur, M. et al. (2021)
			\newblock Compressive Strength of Iliac Bone ECM Is Not Reduced in Osteogenesis Imperfecta and Increases With Mineralization
			\newblock \textit{Journal of Bone and Mineral Research}

			\bibitem[Gross et al. 2013]{p13}  Gross, T., Pahr, D.H., Zysset, P.K. (2013)
			\newblock Morphology elasticity relationships using decreasing fabric information of human trabecular bone from three major anatomical locations
			\newblock \textit{ Biomechanics and Modeling in Mechanobiology}, 793–800

		\end{thebibliography}
	}
\end{frame}

%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\end{document}

### 2021-06-02 03_Scripts/00_ROI_Selection.py

@@ -130,7 +130,7 @@ ScanList = pd.read_csv(os.path.join(DataFolder,ScanLists[DataGroup]),sep=',')
 ParametersDataFrame = pd.DataFrame()
 
 # Perform for all scans
-Scan = Scans[0]
+Scan = Scans[25]
 for Scan in Scans:
 
     ScanFile = os.path.join(ScansPath,Scan)
@@ -228,5 +228,5 @@ for Scan in Scans:
 # Save Parameters for Medtool
 ColsToInt = ['$XPos', '$YPos', '$ZPos', '$XRange', '$YRange', '$ZRange', '$ROINumber']
 ParametersDataFrame[ColsToInt] = ParametersDataFrame[ColsToInt].astype(int)
-ParametersDataFrame.to_csv(os.path.join(ResultsFolder, DataSubFolders[DataGroup][:-14] + 'ROI.csv'),
+ParametersDataFrame.to_csv(os.path.join(ResultsFolder, DataSubFolders[DataGroup][:-14] + 'ROI2.csv'),
                            sep=';', line_terminator=';\n', index=False)

### 2021-06-02 03_Scripts/04_1_Collect_StiffnessData.py

@@ -260,7 +260,7 @@ def ComplianceTensor(EngineeringConstants, EigenValues=[1,1,1], Rho=0, EigenVect
 WorkingDirectory = os.getcwd()
 DataFolder = os.path.join(WorkingDirectory, '04_Results/01_ROI_Analysis/')
 
-Group = 'Healthy'   # Healthy or OI
+Group = 'OI'   # Healthy or OI
 
 StiffnessFolder = os.path.join(DataFolder, '03_' + Group + '_TransformedTensors/')
 ROIFolder = os.path.join(DataFolder,'01_' + Group + '_ROIs/00_Cleaned_ROIs/')

### 2021-06-02 03_Scripts/04_2_General_LinearRegression.py

@@ -25,7 +25,7 @@ def LikelihoodRatioTest(Model1, Model2, Delta_DOF):
     p = 1 - chi2.cdf(LRT, Delta_DOF)
 
     return p
-def PlotRegressionResults(Model, Data, PlotTypes=['BV/TV', 'DA', 'Constants'], Alpha=0.95, Random=True):
+def PlotRegressionResults(Model, Data, PlotTypes=['BV/TV', 'DA', 'Constants'], Alpha=0.95, Random=True, Colors=[(0,0,1),(0,1,0),(1,0,0)]):
 
     ## Get data from the model
     Y_Obs = np.exp(Model.model.endog)
@@ -130,18 +130,18 @@ def PlotRegressionResults(Model, Data, PlotTypes=['BV/TV', 'DA', 'Constants'], A
     if 'Constants' in PlotTypes:
         Figure, Axes = plt.subplots(1, 1, figsize=(5.5, 4.5), dpi=DPI, sharey=True, sharex=True)
         Axes.plot(Y_Obs, Sii,
-                  color=(0, 0, 1), linestyle='none', marker='o', label=r'$\lambda_{ii}$')
+                  color=Colors[0], linestyle='none', marker='s', label=r'$\lambda_{ii}$')
         Axes.plot(Y_Obs, Sij,
-                  color=(0, 1, 0), linestyle='none', marker='o', label=r'$\lambda_{ij}$')
+                  color=Colors[1], linestyle='none', marker='o', label=r'$\lambda_{ij}$')
         Axes.plot(Y_Obs, Sjj,
-                  color=(1, 0, 0), linestyle='none', marker='o', label=r'$\mu_{ij}$')
+                  color=Colors[2], linestyle='none', marker='^', label=r'$\mu_{ij}$')
         # Axes.plot(np.sort(Line), np.sort(CI_Line_u), color=(0.4, 0.4, 0.4), linestyle='--')
         # Axes.plot(np.sort(Line), np.sort(CI_Line_o), color=(0.4, 0.4, 0.4), linestyle='--')
         Axes.plot(Line, Line, color=(0, 0, 0), linestyle='--')
         Axes.annotate(r'N ROIs   : ' + str(NROIs), xy=(0.3, 0.1), xycoords='axes fraction')
[…]

### 2021-06-02 03_Scripts/05_1_Collect_ROI_Data.py

@@ -99,7 +99,7 @@ def GetFabricInfos(FabricFile):
 
 
 # 01 Set variables
-Group = 'Healthy'   # Healthy or OI
+Group = 'OI'   # Healthy or OI
 WorkingDirectory = os.getcwd()
 DataPath = os.path.join(WorkingDirectory,'04_Results/01_ROI_Analysis')
 MorphoPath =  os.path.join(DataPath,'04_' + Group + '_Morphometry/')

### 2021-06-02 03_Scripts/A1_Healthy_BMD_Histograms.py

@@ -196,6 +196,7 @@ Matched_Healthy = pd.read_csv(os.path.join(MatchingPath,'01_Matched_Healthy.csv'
 
 
 # 02 Load files
+ROI_Data = pd.read_csv(ResultsDirectory+'00_Shifts.csv')
 for Index in Matched_Healthy.index:
 
     Scan = Matched_Healthy.loc[Index,'Scan']
@@ -204,19 +205,44 @@ for Index in Matched_Healthy.index:
     ## Load scan
     CT_Scan, Origin, Spacing, Size = Load_Itk(ScanPath+Scan[:8]+'.mhd')
 
+    ## Scale scan to BMD values using AIM header
+    AIMFile = open(ScanPath+Scan[:8]+'.AIM', 'rb')
+
+    BinInts = AIMFile.read(32 * 4)
+    AIM_ints = struct.unpack('=32i', BinInts)
+    if int(AIM_ints[5]) == 16:
+        Header = AIMFile.read(AIM_ints[2])
+    else:
+        Header = AIMFile.read(AIM_ints[8])
+
+    ## Read AIM header
+    Header = re.sub(b'(?i) +', b' ', Header)
+    Header = Header.split(b'\n')
+    Header.pop(0); Header.pop(0); Header.pop(0); Header.pop(0)
+    Scaling = None; Slope = None; Intercept = None
+    for Line in Header:
+        if Line.find(b'Scaled by factor') > -1:
+            Scaling = float(Line.split(b' ')[-1])
+        if Line.find(b'Density: intercept') > -1:
+            Intercept = float(Line.split(b' ')[-1])
+        if Line.find(b'Density: slope') > -1:
+            Slope = float(Line.split(b' ')[-1])
+
+    BMD_Scan = CT_Scan / Scaling * Slope + Intercept
+
     ## Load segmented ROI
     ROI_File = str(ROINumber) + '_' + Scan + '_Cleaned.mhd'
     ROI_Scan, ROI_Origin, ROI_Spacing, ROI_Size = Load_Itk(ROIPath+ROI_File)
 
 
-
     # 03 Extract BMD ROI
     ROI_Offset = np.array(ROI_Origin / ROI_Spacing).astype('int')
[…]

### 2021-06-02 03_Scripts/A1_OI_BMD_Histograms.py

@@ -4,6 +4,9 @@ import numpy as np
 import pandas as pd
 import SimpleITK as sitk
 import matplotlib.pyplot as plt
+import struct
+import re
+from scipy.stats.distributions import norm
 
 
 desired_width = 500
@@ -185,24 +188,50 @@ def WriteMHD(ImageArray, Spacing, Origin, Path, FileName, PixelType='uint'):
 # 01 Set paths
 WorkingDirectory = os.getcwd()
 MatchingPath = os.path.join(WorkingDirectory,'04_Results/05_Variables_Matching')
-ScanPath = os.path.join(WorkingDirectory,'10_Additionnal/01_OI_BMD/01_Scans/')
+ScanPath = os.path.join(WorkingDirectory,'10_Additionnal/02_OI_BMD/01_Scans/')
 ROIPath = os.path.join(WorkingDirectory,'04_Results/01_ROI_Analysis/01_OI_ROIs/00_Cleaned_ROIs/')
-ResultsDirectory = os.path.join(WorkingDirectory,'10_Additionnal/01_OI_BMD/02_ROIs/')
+ResultsDirectory = os.path.join(WorkingDirectory,'10_Additionnal/02_OI_BMD/02_ROIs/')
 
 ## Initialization
-BinsValues = np.arange(0,1.5e4,300)
-Histograms = pd.DataFrame(columns=BinsValues[:-1])
 Matched_OI = pd.read_csv(os.path.join(MatchingPath,'01_Matched_OI.csv'))
 
 
+
+
 # 02 Load files
+ROI_Data = pd.read_csv(ResultsDirectory+'00_Shifts.csv')
 for Index in Matched_OI.index:
 
     Scan = Matched_OI.loc[Index,'Scan']
     ROINumber = Matched_OI.loc[Index,'ROI Number']
 
     ## Load scan
-    CT_Scan, Origin, Spacing, Size = Load_Itk(ScanPath + Scan[:6] + '/' + Scan[7:] +'.mhd')
+    CT_Scan, Origin, Spacing, Size = Load_Itk(ScanPath + Scan[:6] + '/' + Scan[7:15] +'_UNCOMP.mhd')
+
+    ## Scale scan to BMD values using AIM header
+    AIMFile = open(ScanPath + Scan[:6] + '/' + Scan[7:15] +'_UNCOMP.AIM','rb')
+
+    BinInts = AIMFile.read(32 * 4)
[…]

### 2021-06-02 03_Scripts/A1_tBMD_Evaluation.py

@@ -1,0 +1,703 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import SimpleITK as sitk
+import matplotlib.pyplot as plt
+import struct
+import re
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import norm, chi2, t, f
+from scipy.stats import shapiro
+
+
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+## Define functions
+def QQPlot(DataValues, Alpha_CI=0.95, DataLabel='Data', FigFile='QQPlot.png'):
+
+    ### Based on: https://www.tjmahr.com/quantile-quantile-plots-from-scratch/
+    ### Itself based on Fox book: Fox, J. (2015)
+    ### Applied Regression Analysis and Generalized Linear Models.
+    ### Sage Publications, Thousand Oaks, California.
+
+    # Data analysis
+    N = len(DataValues)
+    X_Bar = np.mean(DataValues)
+    S_X = np.std(DataValues,ddof=1)
+
+    # Sort data to get the rank
+    Data_Sorted = np.zeros(N)
+    Data_Sorted += DataValues
+    Data_Sorted.sort()
+
+    # Compute quantiles
+    EmpiricalQuantiles = np.arange(0.5, N + 0.5) / N
+    TheoreticalQuantiles = norm.ppf(EmpiricalQuantiles, X_Bar, S_X)
+    ZQuantiles = norm.ppf(EmpiricalQuantiles,0,1)
+
+    # Compute data variance
+    DataIQR = np.quantile(DataValues, 0.75) - np.quantile(DataValues, 0.25)
+    NormalIQR = np.sum(np.abs(norm.ppf(np.array([0.25, 0.75]), 0, 1)))
+    Variance = DataIQR / NormalIQR
+    Z_Space = np.linspace(min(ZQuantiles), max(ZQuantiles), 100)
+    Variance_Line = Z_Space * Variance + np.median(DataValues)
+
[…]

### 2021-06-02 03_Scripts/RCrosscheck.R

@@ -8,6 +8,8 @@ library(carData)
 library(effects)
 library(car)
 Data <- read.table("/home/mathieu/Documents/Post-Msc/04_Results/04_General_LinearRegression/OI_RData.csv", header=TRUE, sep=",")
+Data <- read.table("/home/mathieu/Documents/Post-Msc/Test.csv", header=TRUE, sep=",")
+
 
 # Check data
 dim(Data)
@@ -31,6 +33,21 @@ LMM1 = lmer(LogSxy ~ Sii + Sij + Sjj + LogBVTV + Logmxy - 1 + (IF-1|Scan),
 summary(LMM1)
 
 LMM2 = lmer(LogSxy ~ Sii + Sij + Sjj + LogBVTV + Logmxy + LogCV - 1 + (IF-1|Scan),
+           REML = TRUE,
+           data = Data)
+summary(LMM2)
+
+anova(LMM2,LMM1)
+
+# Mixed-linear model computation
+library(lme4)
+
+LMM1 = lmer(tBMD ~ BVTV + Group + BVTV*Group + (1|Scan),
+           REML = TRUE,
+           data = Data)
+summary(LMM1)
+
+LMM2 = lmer(tBMD ~ BVTV + Group + (1|Scan),
            REML = TRUE,
            data = Data)
 summary(LMM2)

### 2021-06-28 09_Article/Simon_2021_Bone_Draft_V01.tex
Second article co-authors revision

@@
 
+\author[2,3]{Mahdi Tabatabaei}
+
 \author[2,3]{Bettina M. Willie}
@@
 The general significance level was set to 95\% for all tests.
-Confidence intervals was computed for t-tested variables to confirm a difference in both groups means.
+Confidence intervals were computed for t-tested variables to quantify the difference in both groups means.
 As Mann-Whitney tests are performed on the median, only the corresponding p-value is presented.
@@
 \subsection{Morphological Analysis}
-The results of morphological analysis are summarized in Table \ref{Table2}.
-In the present study, the individual matching allowed us to have similar group distributions with 17 females and 11 males in each group.
-The mean age of matched healthy is 41 $\pm$ 14 and 41 $\pm$ 15 for the matched OI.
-BV/TV of healthy individuals is higher than BV/TV of OI group with a 95\% CI of [0.016, 0.101] and p value <0.01.
-Similarly, trabecular number is higher in the matched healthy group with a 95\% CI of [0.099, 0.285] and a corresponding p value <0.001.
-The trabecular thickness do not show significant differences between groups with a p value of 0.2.
-On the other hand, trabecular separation is higher in matched OI group than in healthy individuals with a p value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$).
-Trabecular separation SD present to be higher in OI patients than in matched healthy with a p value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$).
-SMI as well as degree of anisotropy are higher for matched OI than for healthy people with p values <0.001 and of 0.02, respectively.
-Finally, the log transformation of coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333].\\
-
-In Table \ref{Table2}, absolute values and p values of test statistics are compared with literature.
-Regarding age, the present population is slightly younger than in the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} but stay in a similar range.
-In a general way p values of the three other studies (\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018}) show significant differences for BV/TV, Tb.N., Tb.Sp., and Tb.Sp.SD and non significant differences for Tb.Th.
-Absolute values of BV/TV, TB.Th., Tb.Sp., and Tb.Sp.SD are higher in the present study compared to literature.
+The results of the morphological analysis are summarized in Table \ref{Table2}.
+The individual matching for age and sex allowed us to have similar group distributions with 17 females and 11 males in each group.
+The mean age of matched healthy individuals is 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched OI individuals.
+BV/TV of healthy individuals is higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
+Similarly, trabecular number is higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001.
+The trabecular thickness does not show significant differences between groups.
+On the other hand, permutation test performed for trabecular separation showed that trabecular separation is higher in mat\-ched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$).
+Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$).
+SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively.
+Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\
+
+Table \ref{Table2} compares absolute values and p-values to literature.
+The present population age is fairly consistent with the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018}.
+The three other studies,  \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018}, show significant differences for BV/TV, Tb.N., Tb.Sp., and Tb.Sp.SD and no significant differences for Tb.Th.
+In the present study, absolute values of BV/TV, TB.Th., Tb.Sp., and Tb.Sp.SD seem to be higher compared to literature.
 On the other hand, Tb.N.
-appears to be higher in the other studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} than in the present one.
+appears to be lower compared to studies in literature \cite{Folkestad2012,Kocijan2015,Rolvien2018}.
 
@@
 \subsection{Linear Regression with Original Data Sets}
-Figure \ref{02_GeneralRegression} shows the results of the regression analysis for the complete data sets of each group separately.
-The x axis represent the values of observed tensors from \si{\micro}FE simulations.
-The y axis is the predicted values using the Zysset-Curnier model and the parameters obtained after performing the regression with linear mixed-effect model.
+Figure \ref{02_GeneralRegression} shows the results of the linear regression analysis of each group separately, between the values of the observed stiffness tensors from \si{\micro}FE simulations and the predicted values using the Zysset-Curnier model\cite{Zysset1995} and the parameters obtained after performing the regression with linear mixed-effect model.
 The fitted line is represented by the dashed line, indicating the theoretical perfect correlation.
-$\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components.
-For the healthy group, see Figure \ref{02_Healthy}, the fit is performed on 720 ROIs leading to 8640 points.
+For the healthy group (Figure \ref{02_Healthy}) the fit is performed on 720 ROIs leading to 8640 data points.
 The $R^2_{adj}$ is slightly above 0.95 and the NE is of 18\% $\pm$ 10\%.
-The regression analysis for the OI group led to 3528 points, a $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%.
-It can be noticed that, as the values of the observed tensor decreases, points tends to be further apart of the diagonal (dashed line).
-Moreover, the points coming from the tensors with lowest values are only above the diagonal.
-One last observation which should be noted is the range covered by the observed tensors.
-The range of the OI group is wider than the healthy group and ROIs with lower BV/TV present further low stiffness values.
-The values of these ROIs tensors components tends to be overestimated by the fit.\\
+The regression analysis for the OI group (Figure \ref{02_OI}) performed on 294 ROIs led to 3528 data points, an $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%.
+It can be noticed that, as the values of the observed stiffness tensor decreases, data points tend to be further apart from the diagonal (dashed line).
+Moreover, data the points from the stiffness tensors with lowest values are exclusively above the diagonal.
+The range of stiffness tensors of the OI group is wider compared to the one of the healthy group and ROIs with lower BV/TV present lower stiffness values.
+The values of these ROIs stiffness tensors components tends to be overestimated by the fit.\\
 
@@
 The CV in relation to BV/TV is shown in Figure \ref{02_CV_BVTV}.
-The OI data reach higher values of CV and stay at lower values of BV/TV than healthy data.
+The OI data reached higher CV values and lower BV/TV values compared to healthy data.
 Generally, the CV tends to increase with decreasing BV/TV.
@@
 It can be observed that a relatively important part of OI data will be filtered out.
-On the other hand, relatively few healthy data will be removed by the filtering.
-Examples of extreme ROIs in terms of CV and BV/TV are shown in Appendix \ref{A2}.\\
+On the other hand, relatively few healthy data gets removed.
+3D representation of extreme ROIs in terms of CV and BV/TV are shown in Appendix \ref{A2}.\\
 
@@
 The regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}.
-After filtering, the healthy group was reduced to 119 individuals and 603 ROIs.
-This led to 7236 points for the regression.
-Results show similar values to the complete data set, namely a $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\%, see Figure \ref{04_Healthy}.
-The OI group lost more individuals after filtering leading to 38 people and 115 ROIs.
-Regression is performed on 1380 points then.
-Figure \ref{04_OI} presents results with a $R^2_{adj}$ rounded to 0.95 and a NE of 17\% $\pm$ 8\%.\\
+After filtering, the healthy group was reduced to 119 individuals and 603 ROIs resulting in 7236 data points, an $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\% (Figure \ref{04_Healthy}).
+In the OI group, more individuals were filtered leading to 38 people, 115 ROIs, and 1380 data points (Figure \ref{04_OI}).
+This resulted in an $R^2_{adj}$ close to 0.95 and a NE of 17\% $\pm$ 8\%.\\
 
@@
 Regression results after BV/TV and DA ROI matching are shown in Table \ref{Table3}.
-The columns show which data set was used, the fives parameters of the Zysset-Curnier model and the assessment of fit quality.
+The columns show the used data set, the fives parameters of the Zysset-Curnier model ($\lambda_0$, $\lambda_0'$, $\mu_0$, $k$, and $l$) and the assessment of fit quality ($R^2_{adj}$ and NE).
 Grouping healthy and OI data together for regression lead to a $k$ of 1.91 and a $l$ of 0.95.
 Regression result shows a $R^2_{adj}$ of 0.94 and a NE of 18\% $\pm$ 9\%.
-Second and last rows show regression results using separated data sets and imposing the exponents $k$ and $l$.
-OI values are higher than healthy one.
+The second and the last row show regression results using separated data sets and imposing the exponents $k$ and $l$ to fixed values.
+OI stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) are higher than healthy one.
 The increase is of 15\%, 1\%, and 2\% for $\lambda_0$, $\lambda_0'$, and $\mu_0$, respectively.
@@
 On the other hand, DA is higher in the present study than for \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013}.
-Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set than for the other studies \cite{Gross2013,Panyasantisuk2015}.\\
+Setting the exponents $k$ and $l$ to the same values led to lower stiffness constants for the observed data set compared to the other studies \cite{Gross2013,Panyasantisuk2015}.\\
 
@@
 The linear regression performed using the linear mixed-effects model without BV/TV and group interaction provided a slope of 223 [153,292] (value [95\% CI]).
-The intercept was 534 [518,550] and the group variable led to a value of 14 [7,20].
+The intercept value was 534 [518,550] and the group variable led to a value of 14 [7,20].
 The prediction using the fixed effects only led to a $R^2$ of 0.23 and a SE of 33.
@@
 
-The age of matched individuals lies in the spectrum of the others studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} which allow one to compare the morphological values.
-The main explanation for differences between the absolute values of the present study compared to the others lies in the imaging system.
-\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed first generation Xtreme CT scans with a voxel size of 82 \si{\micro}m and the present study used a second generation scanner, XCTII, with a voxel size of 61 \si{\micro}m.
+As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare morphological parameters.
+The imaging system explains most of the differences between the absolute morphological values of the present study compared to the others.
+\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed their measurements on first generation XCT scanners with a voxel size of 82 \si{\micro}m, while we have used a second generation XCT with a voxel size of 61 \si{\micro}m.
 The work from \citeauthor{Agarwal2016}\cite{Agarwal2016} investigated differences between the two scanner types.
 They showed that BV/TV, Tb.Th., and Tb.Sp.
-is higher in second generation Xtreme CT and, on the other hand, Tb.N.
-is higher in first generation Xtreme CT.
-These results give confidence to the observed values.
+are higher in second generation XCT scanners and Tb.N.
+is lower compared to first generation XCT.
+These results give confidence in our observed values.
 Another bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length.
@@
 as they depend on the ROI size.
-Moreover, conditions imposed for ROI random selection can lead to further biased values, especially for OI patients, as it can not be empty of trabecular bone.
-The CV presenting the stronger significant difference between groups even given the low sample size show that heterogeneity is a main difference in OI patients compared to healthy individuals.
-Finally, the significant differences observed in BV/TV and DA even with age and gender matched individuals justify the choice of a variable matching for fabric-elasticity relationships analysis.\\
-
-Regression performed on original data sets shows $R^2_{adj}$ and NE in the expected range for the healthy group.
+Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain a portion of trabecular bone.
+Even with the low sample size (2x28 individuals), the statistic tests have shown significant differences between groups with the more significant being for the CV values.
+The CV values show that heterogeneity of OI trabecular bone is higher compared to healthy control and the more discriminant parameter.
+Finally, the significant differences observed in BV/TV and DA even with matched age and gender justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the fit must be performed on identical ranges to obtain comparable values.\\
+
+The linear regressions performed in this study on original data sets showed $R^2_{adj}$ and NE in the expected range (i.
+e.
+slightly lower than \citeauthor{Gross2013}\cite{Gross2013} and \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}) for the healthy group.
 Components of the stiffness tensors are distributed to both sides of the diagonal.
-On the other hand, regression of the OI data set presents lower $R^2_{adj}$ than such fit reach usually.
+On the other hand, the linear regressions performed in this study using the OI original data set presents lower $R^2_{adj}$ and higher NE than such fit reach usually \cite{Gross2013,Panyasantisuk2015}.
 The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation.
-These differences come from the ROIs presenting a low stiffness.
-It can be seen on the regression plot (Figure \ref{02_OI}) that when the stiffness term decreases lower than $10^0$ MPa, the fit tends to overestimate the stiffness.
-This is due to the fact that ROI stiffness is highly impacted by BV/TV.
-Some low BV/TV ROIs do not have every side of the cube connected by bone leading to extremely low terms in the stiffness tensor, see Appendix \ref{A2}.
-Trying to homogenize such ROI can lead to error of multiple order of magnitude, as observed on the plot.
+These differences come from ROIs presenting a low stiffness.
+The regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to $10^0$ MPa and lower, the fit tends to overestimate the stiffness.
+This is because ROI stiffness is highly dependent on BV/TV values.
+Some ROIs with low BV/TV don't have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor, see Appendix \ref{A2}.
+Trying to homogenize such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
 Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
-An alternative of CV filtering could be to compute the area ratio filled by bone on each of the six faces of the ROI to assess the ROI heterogeneity.\\
-
-Figure \ref{02_CV_BVTV} presenting the CV in relation to BV/TV shows that there is a tendency of CV to increase with a decreasing BV/TV.
-Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly becomes highly heterogeneous.
+An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.\\
+
+Figure \ref{02_CV_BVTV} presents the CV in relation to BV/TV.
+It shows a tendency of CV values to increase with decreasing BV/TV values.
+Effectively, if the quantity of material inside the ROI decreases, the distribution homogeneity of this mass is more sensitive and therefore can quickly become highly heterogeneous.
 A simple assumption about this relation is that it could be monotonic.
 Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation.
-As some higher BV/TV ROIs still present high CV, imposing a fixed threshold make sense for subsequent homogenization.
+As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenization seem feasible.
 However, the value of this threshold actually results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\
 
-The fits performed on filtered data sets present direct effect of filtering.
-For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) is negligible.
-On the other hand, NE presents an improved value decreasing by 2\%.
-These results are due to the filtering of points further away from the diagonal (better NE) and some points close to the diagonal leading to a smaller number of points ($R^2_{adj}$).
-For the OI group, filtering leads to a important improvement in the fit.
-Results become similar to the healthy group in terms of $R^2_{adj}$, NE, and the range of stiffness values.
-These results give confidence to the filtering procedure and are a first step to the hypothesis of there being no difference in fabric-elasticity relationships between healthy and OI trabecular bone.\\
-
-After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for individual filtered data sets.
-This allows one to give values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m.
-Imposing these values to perform the fit on individual matched data sets allows us to highlight differences, if any, between healthy and OI trabecular bone.
-The relatively low differences for $\lambda_0'$ and $\mu_0$ once again provide arguments for similar relationships between the two groups.
-For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but 95\% CI intervals still show a common range.
-Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group shown a p value far above the 5\% significance level.
-With this statistical non-significance of the group and low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to suppose differences in fabric-elasticity relationships between healthy and OI trabecular bone.
+The linear regressions performed on filtered data sets present direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values.
+For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) compared to the linear regression using the unfiltered data set is negligible.
+On the other hand, NE values are decreased by 2\% and therefore improved.
+The filtering eliminates data points further away from the diagonal (better NE) and other data points close to the diagonal leading to a smaller number of points (modifying $R^2_{adj}$).
+For the OI group, filtering leads to an important improvement of the linear regression (i.e.
+higher $R^2_{adj}$ and lower NE).
+$R^2_{adj}$, NE, and the range of stiffness values are almost at the level of the healthy group.
+These results give confidence to the filtering procedure and are a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\
+
+After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for the individual filtered data sets.
+This allows one to determine values for $k$ and $l$ for the tibia at a spatial resolution of 61 \si{\micro}m.
+Imposing these values to perform the linear regression on data sets of the matched individuals allows us to highlight differences, if any, between healthy and OI trabecular bone.
+The relatively low differences for $\lambda_0'$ and $\mu_0$ once again supports the hypothesis for similar fabric-elasticity relationships between healthy and OI trabecular bone.
+For $\lambda_0$, this relative difference being higher could rise some doubts about this similarity, but the 95\% CI intervals still show a common range which almost include both the $\lambda_0$ of OI and healthy linear regressions.
+Moreover, ANCOVA performed comparing the original formulation and the one with addition of a regressor for the group showed a p-value far above the 5\% significance level.
+With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
 In FEA simulations, it is not possible to exclude part of the mesh because of high heterogeneity.
-Nevertheless, the error created by such ROIs are negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\
+Nevertheless, the error created by such ROIs is negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\
 
 Imposing $k$ and $l$ allows one to estimate the effect of different image resolutions.
-\citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened it to 36 \si{\micro}m.
-\citeauthor{Gross2013}\cite{Gross2013} shown that different anatomical locations lead to only slight differences.
-Comparing regression of filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution.
+\citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened them to 36 \si{\micro}m.
+\citeauthor{Gross2013}\cite{Gross2013} showed that different anatomical locations lead to only slight differences.
+Comparing regression of the filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution.
 Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones.
-Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges stay overlapped.
-As for filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies.
-Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants by resolution.
-The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explain this increase of DA.
-Lower stiffness constants are obtained because the coarser structure resulting from XCTII can not be as optimized as the fine detailed structure obtained by \si{\micro}CT.
+Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges overlap.
+As for the filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies.
+Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants as a result of the different image resolutions.
+The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explains this increase of DA.
+Lower stiffness constants are obtained because the coarser structure resulting from XCTII can't be as optimized as the fine detailed structure obtained by \si{\micro}CT.
 Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimized morphology of trabecular bone with a high fidelity.
-By decreasing the scan spatial resolution, the scanned structure becomes more bulky.
-Performing a fit on this less optimized structure lead to the observed lower stiffness constants.
+By decreasing the scan spatial resolution, the scanned structure becomes bulkier.
+Performing a linear regression on this less optimized structure leads to the observed lower stiffness constants.
 Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality.
-Nevertheless, $l$ stands in the same range as for the two other studies meaning the relative weight of DA remains constant.
-On the other hand, the higher $k$ highlight an increased relative weight of BV/TV.\\
+Nevertheless, $l$ stays in the same range as for the two other studies \cite{Gross2013,Panyasantisuk2015}, meaning the relative weight of DA remains constant.
+On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.\\
 
@@
 As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for fitting.
-More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models can be negligible.
-Similar study could be performed using XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.
+More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models could be negligible if the proportion of such low stiffness ROIs stays low.
+A similar study could be performed on XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.
 Another limitation is that the scans were performed on different devices which were not cross-calibrated.
@@
 
-In conclusions, the sample analyzed in the this study have similar morphology to data reported in the literature.
-We find no differences in fabric-elasticity relationships between healthy and OI trabecular bone if the ROI is homogeneous enough.
-\citeauthor{Indermaur2021}\cite{Indermaur2021} shown that compressive behavior of OI bone tissue is similar to healthy control.
-If tensile and shearing behaviors agree, fabric-strength relationships will hold too.
-Therefore, OI trabecular bone can explain part of the bone fragility by the decrease in BV/TV and loss of homogeneity in its trabecular organization.
+In conclusions, the samples analyzed in the present study had similar morphology compared to data reported in the literature.
+We couldn't find differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
+with a CV lower than 0.263.
+\citeauthor{Indermaur2021}\cite{Indermaur2021} could show that the compressive behaviour of OI bone tissue is similar to the one of healthy control at the ECM level.
+If the tensile and shearing behaviour is similar as well, fabric-strength relationships will hold too.
+Therefore, OI trabecular bone can explain part of the bone fragility by the decrease in BV/TV and the loss of homogeneity in its trabecular organization.
 

### 2021-06-28 12_SBMS_Presentation/SBMS_Simon_et_al.tex
Second article co-authors revision

@@
 		\end{itemize}\vspace{1em}
-		Linear Regression
+		Multiple Linear Regression
 		\begin{itemize}
-			\item Linear mixed-effects model\\
 			\item Fit to Zysset-Curnier model\\
-			$\mathbb{S}$ \textasciitilde \textcolor{blue}{$\lambda_{0}$}, \textcolor{blue}{$\lambda_{0}$'}, \textcolor{blue}{$\mu_{0}$}, $m_{i}m_{j}^{\textcolor{red}{k}}$, BV/TV$^{\textcolor{red}{l}}$
+			$\mathbb{S}$ \textasciitilde \textcolor{blue}{$\lambda_{0}$}, \textcolor{blue}{$\lambda_{0}$'}, \textcolor{blue}{$\mu_{0}$}, $m_{i}^{\textcolor{red}{k}}m_{j}^{\textcolor{red}{k}}$, BV/TV$^{\textcolor{red}{l}}$
 			\item Log space for \textcolor{red}{$k$} and \textcolor{red}{$l$}
@@
 			\item Age and gender matching
+		\end{itemize}
+
+		\vspace{17.5pt}
+
+		Tissue BMD (tBMD)
+		\begin{itemize}
+			\item BV/TV and DA matching
 		\end{itemize}
@@
 		\item Matching allow to fit over same range
-		\item Identical $k$ and $l$ for healthy and OI comparison
+		\item Identical $k$ and $l$ for stiffness comparison
 	\end{itemize}
@@
 
+\begin{frame}[noframenumbering]
+	\frametitle{Linear Regression - BV/TV and DA Matching}
+
+	Regression conditions:
+	\begin{itemize}
+		\item Matching allow to fit over same range
+		\item Identical $k$ and $l$ for stiffness comparison
+	\end{itemize}
+
+	\vfill
+
+	\begin{table}[]
+		\centering
+		\caption{Summary of BV/TV and DA matched data set linear regression\vspace{-1em}}
+		\label{Table4}\resizebox{1.\linewidth}{!}{
+			\begin{tabular}{cccccccc}
+				\toprule
+				Data set & $\lambda_0$ & $\lambda_0'$ & $\mu_0$ & $k$ & $l$ & $R^2_{adj}$ & NE (\%) \\
+				\midrule
+				Grouped & 4626 [3892-5494] & 2695 [2472-2937] & 3541 [3246-3862] & 1.91 [1.86-1.95] & 0.95 [0.93-0.97] & 0.936 & 19 $\pm$ 9\\
+
+				Healthy & 4318 [\hl{3844-4851}] & 2685 [\hl{2533-2845}] & 3512 [\hl{3306-3731}] & \textcolor{gray}{1.91} & \textcolor{gray}{0.95} & 0.835 & 21 $\pm$ 10\\
+
+				OI & 4983 [\hl{4345-5716}] & 2727 [\hl{2547-2921}] & 3600 [\hl{3355-3863}] & \textcolor{gray}{1.91} & \textcolor{gray}{0.95} & 0.860 & 20 $\pm$ 10\\
+				\bottomrule
+		\end{tabular}}
+	\end{table}
+
+	\vfill
+
+	$\rightarrow$ Overlapping stiffness values: Similar fabric-elasticity relationships
+\end{frame}
+
 \begin{frame}
-	\frametitle{BV/TV and DA Matching - Tissue BMD (tBMD)}
+	\frametitle{BV/TV and DA Matching - tBMD}
 	\fontsize{10pt}{10pt}\selectfont
@@
 
-\subsection{Comparison with Other Resolutions}
-
-\begin{frame}
-	\frametitle{Comparison with Other Resolutions}
-	\fontsize{10pt}{10pt}\selectfont
-
-	\begin{table}[]
-		\centering
-		\caption{Summary of studies compared\vspace{-1em}}
-		\label{Table4}\resizebox{1.\linewidth}{!}{
-			\begin{tabular}{lccc}
-				\toprule
-				Study & Scanned site & Scan & Resolution \\
-				\midrule
-
-				\cite{p13} & Multiple sites & \si{\micro}CT & 18\si{\micro}m $\rightarrow$ 36\si{\micro}m \\
-
-				\cite{p7} & Proximal femur & \si{\micro}CT & 18\si{\micro}m $\rightarrow$ 36\si{\micro}m \\
-
-				Present study & Distal tibia & HR-pQCT & 61\si{\micro}m \\
-				\bottomrule
-		\end{tabular}}
-	\end{table}
-
-	\vfill
-
-	\begin{columns}[t]
-		\column{0.5\textwidth}
-		Higher DA
-		\begin{itemize}
-			\item Distal tibia vs proximal femur
-		\end{itemize}
-
-		\vspace{7pt}
-
-		Lower stiffness constants ($\lambda_0$, $\lambda_0'$, $\mu_0$)
-		\begin{itemize}
-			\item 61\si{\micro}m vs 36\si{\micro}m
-		\end{itemize}
-
-		\vspace{7pt}
-
-		Lower fit quality ($R^2_{adj}$ and NE)
-		\begin{itemize}
-			\item 61\si{\micro}m vs 36\si{\micro}m
-		\end{itemize}
-
-		\column{0.5\textwidth}
-		Higher $k$
-		\begin{itemize}
-			\item Higher weight of BV/TV
-		\end{itemize}
-
-		\vspace{7pt}
-
-		Similar $l$
-		\begin{itemize}
-			\item Similar weight of DA
-		\end{itemize}
-
-	\end{columns}
-\end{frame}
+%\subsection{Comparison with Other Resolutions}
+%
+%\begin{frame}
+%	\frametitle{Comparison with Other Resolutions}
+%	\fontsize{10pt}{10pt}\selectfont
+%	
+%	\begin{table}[]
+%		\centering
+%		\caption{Summary of studies compared\vspace{-1em}}
+%		\label{Table4}\resizebox{1.\linewidth}{!}{
+%			\begin{tabular}{lccc}
+%				\toprule
+%				Study & Scanned site & Scan & Resolution \\
+%				\midrule
+%				
+%				\cite{p13} & Multiple sites & \si{\micro}CT & 18\si{\micro}m $\rightarrow$ 36\si{\micro}m \\
+%				
+%				\cite{p7} & Proximal femur & \si{\micro}CT & 18\si{\micro}m $\rightarrow$ 36\si{\micro}m \\
+%				
+%				Present study & Distal tibia & HR-pQCT & 61\si{\micro}m \\
+%				\bottomrule
+%		\end{tabular}}
+%	\end{table}
+%
+%	\vfill
+%
+%	\begin{columns}[t]
+%		\column{0.5\textwidth}
+%		Higher DA
+%		\begin{itemize}
+%			\item Distal tibia vs proximal femur
+%		\end{itemize}
+%	
+%		\vspace{7pt}
+%		
+%		Lower stiffness constants ($\lambda_0$, $\lambda_0'$, $\mu_0$)
+%		\begin{itemize}
+%			\item 61\si{\micro}m vs 36\si{\micro}m
+%		\end{itemize}
+%	
+%		\vspace{7pt}
+%		
+%		Lower fit quality ($R^2_{adj}$ and NE)
+%		\begin{itemize}
+%			\item 61\si{\micro}m vs 36\si{\micro}m
+%		\end{itemize}
+%	
+%		\column{0.5\textwidth}
+%		Higher $k$
+%		\begin{itemize}
+%			\item Higher weight of BV/TV
+%		\end{itemize}
+%	
+%		\vspace{7pt}
+%		
+%		Similar $l$
+%		\begin{itemize}
+%			\item Similar weight of DA
+%		\end{itemize}
+%	
+%	\end{columns}
+%\end{frame}
 

### 2021-06-28 09_Article/Simon_2021_Bone_Draft_V01.tex
Correct .gitignore

@@
 \begin{abstract}
-	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility, also called "brittle bone disease".
+	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility.
 It is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass.
-High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology of extremities including the weight-bearing distal tibia.
+High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia.
 The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation.
@@
 
-	In this study, the morphology of distal tibiae from 50 adults with OI were compared to 120 healthy controls using second generation HR-pQCT.
+	In the present study, the morphology of distal tibiae from 50 adults with OI were compared to 120 healthy controls using second generation HR-pQCT.
 Six cubic regions of interest (ROIs) were selected per individual in a common anatomical region.
-A first matching was performed by selecting similar individuals to obtain identical mean and median age and gender distribution in the OI and healthy control group.
+A first matching between OI and healthy control group was performed by selecting similar individuals to obtain identical mean and median age and gender distribution.
 It allowed to perform a morphometric analysis and compare the outcome with literature.
-Then, stiffness tensors of ROIs were computed using hFEA and multiple linear regressions were performed on the Zysset-Curnier orthotropic model.
-The regressions allowed to compare the two groups using 5 parameters.
-An initial fit was performed on both the OI group and the healthy control group using all the ROIs extracted.
-Then, data were filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity and second fits were performed on these filtered data sets.
+Then, stiffness tensors of ROIs were computed using hFE and multiple linear regressions were performed with the Zysset-Curnier orthotropic model.
+The regressions allowed to compare the two groups using five parameters.
+An initial fit was performed on both the OI group and the healthy control group using all extracted ROIs.
+Then, data was filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity and additionnal linear regressions were performed on these filtered data sets.
 These full and filtered data were in turn compared with previous results from \si{\micro}CT reconstructions obtained in other anatomical locations.
-Finally, the ROIs of both group were matched according to their BV/TV and fabric anisotropy (DA).
-Fits were performed again using these matched data to detect statistical differences between the two groups.
+Finally, the ROIs of both groups were matched according to their BV/TV and fabric anisotropy (DA).
+Linear regressions were performed using these matched data to detect statistical differences between the two groups.
 
@@
 
-Osteogenesis imperfecta (OI), also commonly known as "brittle bone disease", is an inherited form of bone fragility \cite{Tournis2018}.
-OI prevalence is estimated at about  1/13,500 of births, less severe forms being not accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
+Osteogenesis imperfecta (OI), commonly known as brittle bone disease, is an inherited form of bone fragility \cite{Tournis2018}.
+OI prevalence is estimated at about  1/13,500 of births, less severe forms not being accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
 Therefore, OI is considered as a rare metabolic bone disorder.
 In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases.
-OI can be categorized according to disease severity \cite{Mortier2019} into:
+OI can be categorized according to disease severity \cite{Mortier2019}, the most familiar forms are:
 \begin{itemize}

### 2021-06-30 09_Article/Simon_2021_Bone_Draft_V02.tex

--- 2021-06-28 09_Article/Simon_2021_Bone_Draft_V01.tex
@@
 In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases.
-OI can be categorized according to disease severity \cite{Mortier2019}, the most familiar forms are:
+OI can be categorized according to disease severity \cite{Mortier2019}, where most familiar forms are:
 \begin{itemize}
@@
 Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.
-\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well, namely bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls.
-Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI but the trabecular thickness (Tb.Th.) is not significantly different.
+\citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well.
+Bone volume fraction (BV/TV) and trabecular number (Tb.N.) in OI bone is lower than for healthy controls.
+Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different.
 At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\

### 2021-07-02 09_Article/Simon_2021_Bone_Draft_V02.tex
Article corrections

@@
 Trabecular separation (Tb.Sp.) and inhomogeneity (Tb.Sp.SD) are higher for individuals with OI compared to healthy control, and the trabecular thickness (Tb.Th.) is not significantly different.
-At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\
-
-High resolution peripheral quantitative computed tomography (HR-pQCT) scans allow one to perform \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) in the distal radius and distal tibia \cite{Boutroy2005}.
-Moreover, the image obtained from HR-pQCT can be used for finite element analysis (FEA) to predict mechanical properties \cite{Boutroy2008}.
-Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from the HR-pQCT scan that can be used to assess bone strength within a reasonable computation time \cite{Pahr2009}.
+At the ECM level, a recent study showed that, in compression, OI bone tends to present higher modulus, ultimate stress and post-yield behavior (ultimate stress and ultimate strain) than healthy bone, mostly affected by the higher degree of mineralization of OI bone \cite{Indermaur2021}.\\
+
+High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}.
+Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}.
+Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
 High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,AriasMoreno2019}.
-Thus, it could be legitimate to use hFE for OI patient's bone strength estimation and potentially fracture risk assessment.
-However, HR-pQCT-based FEA relies on fabric-elasticity relationships.
+Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
+However, HR-pQCT-based hFE relies on fabric-elasticity relationships.
 Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to investigate the hypothesis of similar fabric-elasticity relationships.

### 2021-07-02 09_Article/Simon_2021_Bone_Draft_V02.tex
Finish article corrections

@@
 Then, stiffness tensors of ROIs were computed using hFE and multiple linear regressions were performed with the Zysset-Curnier orthotropic model.
-The regressions allowed to compare the two groups using five parameters.
+The linear regressions allowed to compare the two groups using five parameters.
 An initial fit was performed on both the OI group and the healthy control group using all extracted ROIs.
@@
 
-	In conclusion, despite the reduced regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis.
+	In conclusion, despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis.
 Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.
@@
 These subjects had not taken any medication known to affect bone metabolism nor presented with any prior osteoporosis fracture.
-The second group was scanned as part of the ASTEROID study at different locations in Canada, namely at the Shriners Hospital for Children-Canada.
+The second group was scanned as part of the ASTEROID study (NCT number: NCT03118570) at different locations, namely at the Shriners Hospital for Children-Canada.
 The study coordination was done by the McGill University in Montreal.
@@
 The scanned region according to the two different protocol are shown side by side in Figure \ref{01_ClinicalSections}.
-For both the healthy and the OI groups, each stack consisted of 168 slices and a voxel resolution of 61 \si{\micro}m in the three principal directions.
+For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m.
 This led to a thickness of roughly 10.2 mm for each stack.
 Standardized scanning settings were used (voltage of 60 kVp, 900 μA, 100 ms integration time) for the healthy group as well as for the OI group.
-For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale of 1 (no motion artefacts) to 5 (extreme motion artefacts), as proposed by \citeauthor{Pialat2012} \cite{Pialat2012}.
+For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts).
 The final grade of each scan was defined as the highest slice grade.
@@
 
-In general, six cubic ROIs were selected at random position in each scan in a defined area.
+In general, six cubic ROIs were selected at random position in each scan in a predefined area.
 Each ROI had to contain trabecular bone, but no cortical bone.
-For the healthy group, the ROIs were selected in the more proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at the same anatomical location as for the OI group.
-Then for both the healthy group and the OI group, the stack was divided into two halves and the ROIs were selected to have the centres of three ROIs in the proximal half and three in the distal half.
+For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at rougthly the same anatomical location as for the OI group.
+Then for both the healthy group and the OI group, the stack was divided into two halves and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
 For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone.
@@
 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=\linewidth]
+	{Pictures/01_ROISelection_Example}
+	\caption{Middle slices example of ROIs selection for HR-pQCT reconstruction of a healthy tibia.
+Cortical bone is in white and trabecular bone in gray.
+The stacks are separated by red lines and blue squares represent the ROIs.
+The stack used for analysis is higlighted in green and the dashed red line represents the middle of the stack.}
+	\label{01_ROISelection}
+\end{figure}
+
 The ROI was defined as a cube of 5.3 mm side length.
-This size was in agreement with the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013}, who performed similar analysis with femur \si{\micro}CT scans.
-It was determined by \citeauthor{Zysset1998}\cite{Zysset1998} and \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to allow one to have a relative homogeneity of trabecular tissue within the ROI leading to accurate \si{\micro}FE results with a minimal computational cost.\\
+This size agreed with the work of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013}, who performed similar analysis with femur \si{\micro}CT scans.
+It was determined by \citeauthor{Zysset1998}\cite{Zysset1998} and \citeauthor{Daszkiewicz2017}\cite{Daszkiewicz2017} to allow a relative homogeneity of trabecular tissue within the ROI leading to accurate \si{\micro}FE results and a minimal computational cost.\\
 
 After ROI cleaning, i.e.
-deletion of unconnected region of bone material, the morphological analysis of ROIs was performed using medtool (v4.5; Dr.
+deletion of bone material regions unconnected to the main structure, the morphological analysis of ROIs was performed using medtool (v4.5; Dr.
 Pahr Ingenieurs e.U., Pfaffstätten, Austria).
 The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular spacing (Tb.Sp.SD).
-Moreover, ROI tissue bone mineral density (tBMD) and fabric was evaluated.
+Moreover, ROI tissue mineral density (TMD) and fabric was evaluated.
 The fabric tensor $\mathbf{M}$ was computed using mean intercept length (MIL) method \cite{Moreno2014}.
-It is a positive-definite second-order tensor built as shown in Equation \ref{Eq201} below:
+It is a positive-definite second-order tensor computed as shown below.
 
@@
 
-A \si{\micro}FE mechanical analysis was performed using \textsc{ABAQUS 6.14}.
+A \si{\micro}FE analysis was performed using \textsc{ABAQUS 6.14}.
 In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach.
-Then, a stiffness $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned.
+Then, a stiffness $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element.
 The homogenization process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}.
-Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require one to rotate the ROI into fabric coordinate system.
+Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system.
 Such rotation would potentially decrease image quality.
 The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
-This model builds the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\
+This model compute the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\
 
@@
 However, the trabecular structure is not perfectly homogeneous.
-In order to assess the ROI heterogeneity, a coefficient of variation (CV) is computed according to \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}: the ROI is divided into eight identical sub-cubes and BV/TV is computed for each of them.
+To assess the ROI heterogeneity, a coefficient of variation (CV) is computed according to \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}: the ROI is divided into eight identical sub-cubes and BV/TV is computed for each of them.
 The CV is defined as the ratio between the standard deviation of these BV/TV and the mean value (Equation \ref{Eq203}).
@@
 
-\subsection{Statistics}
-The analyzed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI groups.
+\subsection{Morpholigical Analysis}
+The analyzed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI group.
+
+\subsubsection{Participants Matching}
 As the groups do not have similar distributions of age and sex, a matching was performed by selecting similar individuals leading to identical mean and median age as well as identical gender distribution.
+
+\subsubsection{Statistical Tests}
 For each parameter, the median value between the six ROIs from the same individual was computed.
@@
 According to the normality and equal variances assumptions, t-test, Mann-Whitney test or a non-parametric permutation test was performed.
+
+\subsubsection{Statistical Significance}
 The general significance level was set to 95\% for all tests.
@@
 $Z$ is a $12n$x$f$ design matrix which contains data with individual dependence and $\delta$ is a $f$x1 vector composed of random factors.
-Finally, $\epsilon$ is a $12n$x1 vector containing the regression residuals.
-As $k$ and $l$ are exponents, the regression was performed on the log space.\\
-
+Finally, $\epsilon$ is a $12n$x1 vector containing the linear regression residuals.
+As $k$ and $l$ are exponents, the linear regression was performed on the log space.\\
+
+\subsubsection{Data Filtering}
 The linear regression was performed on both group (heal\-thy and OI) separately.
@@
 Then, the relation between BV/TV and CV was assessed using Spearman's correlation coefficient.
-To compare the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) between the groups, regression must be performed on identical value ranges.
+
+\subsubsection{ROIs Matching}
+To compare the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) between the groups, linear regression must be performed on identical value ranges.
 Therefore, a matching was performed for BV/TV and DA to find corresponding control ROIs for each OI in the filtered groups.
 Best correspondences were kept and duplicates were dropped.
-Finally, as the regression was performed in the log space, slight differences in the exponents ($k$ and $l$) would lead to important variation of the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) so it was necessary to use identical exponents  for both groups, weighting identically BV/TV and DA between regressions.
-The exponents were determined by grouping healthy and OI for regression.
+Finally, as the linear regression was performed in the log space, slight differences in the exponents ($k$ and $l$) would lead to important variation of the stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) so it was necessary to use identical exponents  for both groups, weighting identically BV/TV and DA between linear regressions.
+The exponents were determined by grouping healthy and OI for linear regression.
 Then a modified system was used to perform the fit on separated groups, see Appendix \ref{A1}, Equation \ref{EqA11}.\\
 
+\subsubsection{Model Modifications}
 Another modification of the model was to add a regressor for the group variable (healthy or OI), i.e.
@@
 Implementation of this modification was performed according to \cite{Fox2016}.
-A similar mixed-effect model was used to analyze the relation between tBMD and BV/TV and the significance of the group, see Appendix \ref{A1} Equation \ref{EqA14}.
+A similar mixed-effect model was used to analyze the relation between TMD and BV/TV and the significance of the group, see Appendix \ref{A1} Equation \ref{EqA14}.
 The model used the BV/TV and the group (healthy or OI) as fixed variables and the individual as random variable.
 Moreover, to test the hypothesis of no interaction between the BV/TV and the group, i.e.
-the group has no significant influence on the tBMD versus BV/TV slope, the model was modified to add the interaction regressor (BV/TV x Group).
+the group has no significant influence on the TMD versus BV/TV slope, the model was modified to add the interaction regressor (BV/TV x Group).
 The detailed linear systems for each model discussed here are available in Appendix \ref{A1} and a summary of the data sets used for the different methods is shown in Table \ref{Table1}.\\
 
-The regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}.
-Regression quality for the tBMD analysis was assessed using the Pearson correlation coefficient ($R^2$) and the standard error of the estimate (SE).
-Regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}.
+The linear regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}.
+Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient ($R^2$) and the standard error of the estimate (SE).
+Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}.
 
@@
 The results of the morphological analysis are summarized in Table \ref{Table2}.
-The individual matching for age and sex allowed us to have similar group distributions with 17 females and 11 males in each group.
-The mean age of matched healthy individuals is 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched OI individuals.
-BV/TV of healthy individuals is higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
-Similarly, trabecular number is higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001.
+The individual matching for age and sex allowed to have similar group distributions with 17 females and 11 males in each group.
+The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively.
+BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
+Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001.
 The trabecular thickness does not show significant differences between groups.
-On the other hand, permutation test performed for trabecular separation showed that trabecular separation is higher in mat\-ched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$ ,-0.384] $\cup$ [0.421,$\infty$).
-Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$ ,-0.232] $\cup$ [0.251,$\infty$).
+On the other hand, permutation test performed for trabecular separation showed that it is higher in mat\-ched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$).
+Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$, -0.232] $\cup$ [0.251, $\infty$).
 SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively.
@@
 \subsection{Linear Regression with Original Data Sets}
-Figure \ref{02_GeneralRegression} shows the results of the linear regression analysis of each group separately, between the values of the observed stiffness tensors from \si{\micro}FE simulations and the predicted values using the Zysset-Curnier model\cite{Zysset1995} and the parameters obtained after performing the regression with linear mixed-effect model.
-The fitted line is represented by the dashed line, indicating the theoretical perfect correlation.
+Figure \ref{02_GeneralRegression} shows the results of the linear regression analysis of each group separately, between the values of the observed stiffness tensors from \si{\micro}FE simulations and the predicted values using the Zysset-Curnier model\cite{Zysset1995} and the parameters obtained after performing the linear regression with linear mixed-effect model.
+The dashed line represents the 1:1 correlation.
 For the healthy group (Figure \ref{02_Healthy}) the fit is performed on 720 ROIs leading to 8640 data points.
 The $R^2_{adj}$ is slightly above 0.95 and the NE is of 18\% $\pm$ 10\%.
-The regression analysis for the OI group (Figure \ref{02_OI}) performed on 294 ROIs led to 3528 data points, an $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%.
+The linear regression analysis for the OI group (Figure \ref{02_OI}) performed on 294 ROIs led to 3528 data points, an $R^2_{adj}$ close to 0.85 and a NE of 62\% $\pm$ 233\%.
 It can be noticed that, as the values of the observed stiffness tensor decreases, data points tend to be further apart from the diagonal (dashed line).
@@
 The range of stiffness tensors of the OI group is wider compared to the one of the healthy group and ROIs with lower BV/TV present lower stiffness values.
-The values of these ROIs stiffness tensors components tends to be overestimated by the fit.\\
+The values of these ROIs stiffness tensors components tend to be overestimated by the fit.\\
 
@@
 	\end{subfigure}
-	\caption{Regression results using the fixed effects of the linear mixed-effect model on original data sets.
+	\caption{Linear regression results using the fixed effects of the linear mixed-effect model on original data sets.
 $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components.
-The dashed line represents the fitted line.}
+The dashed line represents the 1:1 correlation.}
 	\label{02_GeneralRegression}
@@
 The CV in relation to BV/TV is shown in Figure \ref{02_CV_BVTV}.
-The OI data reached higher CV values and lower BV/TV values compared to healthy data.
+The OI samples reached higher CV values and lower BV/TV values compared to healthy ones.
 Generally, the CV tends to increase with decreasing BV/TV.
@@
 	{Pictures/03_CV_BVTV}
-	\caption{Coefficient of variation in relation to BV/TV. Spearman correlation coefficient $\rho$ assess monotonic relation between two variable}
+	\caption{Coefficient of variation in relation to BV/TV. Spearman correlation coefficient $\rho$ assess monotonic relation between two variables}
 	\label{02_CV_BVTV}
@@
 
-The regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}.
+The linear regression results of the filtered data are presented in Figure \ref{04_FilteredRegression}.
 After filtering, the healthy group was reduced to 119 individuals and 603 ROIs resulting in 7236 data points, an $R^2_{adj}$ close to 0.95 and a NE of 16\% $\pm$ 8\% (Figure \ref{04_Healthy}).
-In the OI group, more individuals were filtered leading to 38 people, 115 ROIs, and 1380 data points (Figure \ref{04_OI}).
-This resulted in an $R^2_{adj}$ close to 0.95 and a NE of 17\% $\pm$ 8\%.\\
+In the OI group, more individuals were filtered, leading to 38 individuals and, 115 ROIs resulting in 1380 data points (Figure \ref{04_OI}), an $R^2_{adj}$ close to 0.95 and a NE of 17\% $\pm$ 8\%.\\
 
@@
 	\end{subfigure}
-	\caption{Regression results using the fixed effects of the linear mixed-effect model on filtered data sets.
+	\caption{Linear regression results using the fixed effects of the linear mixed-effect model on filtered data sets.
 $\lambda_{ii}$ stands for the diagonal terms of normal components of $\mathbb{S}$ in Mandel notation\cite{MANDEL1965}, $\lambda_{ij}$ for the off-diagonal terms of normal components, and $\mu_{ij}$ for the shear components.
-The dashed line represents the fitted line.}
+The dashed line represents the 1:1 correlation.}
 	\label{04_FilteredRegression}
@@
 
-\subsection{BV/TV and DA Matching}
-Regression results after BV/TV and DA ROI matching are shown in Table \ref{Table3}.
-The columns show the used data set, the fives parameters of the Zysset-Curnier model ($\lambda_0$, $\lambda_0'$, $\mu_0$, $k$, and $l$) and the assessment of fit quality ($R^2_{adj}$ and NE).
-Grouping healthy and OI data together for regression lead to a $k$ of 1.91 and a $l$ of 0.95.
-Regression result shows a $R^2_{adj}$ of 0.94 and a NE of 18\% $\pm$ 9\%.
-The second and the last row show regression results using separated data sets and imposing the exponents $k$ and $l$ to fixed values.
-OI stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) are higher than healthy one.
+\subsection{ROIs Matching}
+Linear regression results after BV/TV and DA ROI matching are shown in Table \ref{Table3}.
+The columns show the used data set, the fives parameters of the Zysset-Curnier model ($\lambda_0$, $\lambda_0'$, $\mu_0$, $k$, and $l$) and the assessment of linear regression quality ($R^2_{adj}$ and NE).
+Grouping healthy and OI data together for linear regression leads to a $k$ of 1.91 and a $l$ of 0.95, an $R^2_{adj}$ of 0.94 and a NE of 18\% $\pm$ 9\%.
+The second and the last row show linear regression results using the individual data sets and setting the exponents $k$ and $l$ to fixed values.
+OI stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) are consistantly higher than healthy one.
 The increase is of 15\%, 1\%, and 2\% for $\lambda_0$, $\lambda_0'$, and $\mu_0$, respectively.
-The ANCOVA performed to quantify the group statistical significance shows a p value of 0.7.\\
+The ANCOVA performed to quantify the group statistical significance shows a p value of 0.7, meaning the difference in stiffness constants is not significant.\\
 
@@
 Values are presented as value [95\% CI] or mean $\pm$ standard deviation.
-Values in gray were imposed in the regression.}
+Values in gray were fixed in the linear regression.}
 	\label{Table3}
@@
 
+\subsection{Comparison with literature}
 Table \ref{Table4} shows results obtained compared to literature.
@@
 \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} and \citeauthor{Gross2013} \cite{Gross2013} show values obtained with ROIs of femur \si{\micro}CT scans of healthy individuals only.
-Values in gray were imposed in the regression.}
+Values in gray were imposed in the linear regression.}
 	\label{Table4}
@@
 
-		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 126 & 0.27 $\pm$ 0.08 & 1.57 $\pm$ 0.18 & 3306 & 2736 & 2837 & 1.55 & 0.82 & 0.984 & 8 $\pm$ 3\\
+		\citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015} & 126 & 0.27 $\pm$ 0.08 & 1.57 $\pm$ 0.18 & 3306 & 2736 & 2837 & 1.55 & 0.82 & 0.984 & 8 $\pm$ 3\\[1ex]
 
@@
 
-		\citeauthor{Gross2013} \cite{Gross2013} & 264 & 0.19 $\pm$ 0.10 & 1.67 $\pm$ 0.34 & 4609 & 3692 & 3738 & 1.60 & 0.99 & 0.981 & 14\\
+		\citeauthor{Gross2013} \cite{Gross2013} & 264 & 0.19 $\pm$ 0.10 & 1.67 $\pm$ 0.34 & 4609 & 3692 & 3738 & 1.60 & 0.99 & 0.981 & 14\\[1ex]
 
@@
 
-The analysis of tBMD in relation to BV/TV is shown in Figure \ref{02_tBMD}.
-The t-test performed on the tBMD distributions led to a p-value <0.0001 and a 95\% CI of [-35,-13].
-The linear regression performed using the linear mixed-effects model without BV/TV and group interaction provided a slope of 223 [153,292] (value [95\% CI]).
-The intercept value was 534 [518,550] and the group variable led to a value of 14 [7,20].
+The analysis of TMD in relation to BV/TV is shown in Figure \ref{02_TMD}.
+The t-test performed on the TMD distributions led to a p-value <0.0001 and a 95\% CI of [-35,-13] mgHA/cm\textsuperscript{3}.
+The linear regression performed using the linear mixed-effects model without BV/TV and group interaction (i.e.
+TMD \textasciitilde \space BVTV + Group) provided a slope of 223 [153,292] mgHA/cm\textsuperscript{3} (value [95\% CI]).
+The intercept value was 534 [518,550] mgHA/cm\textsuperscript{3} and the group variable led to a value of 14 [7,20] mgHA/cm\textsuperscript{3}.
 The prediction using the fixed effects only led to a $R^2$ of 0.23 and a SE of 33.
-Then, using the linear mixed-effect model with the interaction regressor (BV/TV x Group), this added variable presented a value of 3 [-66,72] and a p-value of 0.94.
+Then, using the linear mixed-effect model with the interaction regressor (i.e.
+TMD \textasciitilde BVTV + Group + BV/TV x Group), this added variable presented a value of 3 [-66,72]mgHA/cm\textsuperscript{3} and a p-value of 0.94.
 
@@
 	\includegraphics[width=\linewidth]
-	{Pictures/05_tBMDvsBVTV}
-	\caption{tBMD in relation to BV/TV. The fitted lines are obtained using the fixed effects of the linear mixed-effect model and fixing the group variable.}
-	\label{02_tBMD}
+	{Pictures/05_TMDvsBVTV}
+	\caption{TMD in relation to BV/TV. The fitted lines are obtained using the fixed effects of the linear mixed-effect model and fixing the group variable.}
+	\label{02_TMD}
 \end{figure}
@@
 Osteogenesis imperfecta is an inherited form of bone fra\-gi\-li\-ty with a severity going from mild to perinatally lethal.
-This study aim to confirm that fabric-elasticity relationships in OI trabecular bone are similar than in healthy conditions, encouraging the use of HR-pQCT scans for fracture risk assessment.
+This study aims to confirm that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment.
 To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.\\
@@
 These results give confidence in our observed values.
-Another bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length.
+Another potential bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length.
 This conditions the Tb.N.
@@
 as they depend on the ROI size.
-Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain a portion of trabecular bone.
-Even with the low sample size (2x28 individuals), the statistic tests have shown significant differences between groups with the more significant being for the CV values.
-The CV values show that heterogeneity of OI trabecular bone is higher compared to healthy control and the more discriminant parameter.
-Finally, the significant differences observed in BV/TV and DA even with matched age and gender justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the fit must be performed on identical ranges to obtain comparable values.\\
+Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain a certain portion of trabecular bone.
+Even with the low sample size (2x28 individuals), the statistical tests have shown significant differences between groups, with the CV presenting the most significant difference.
+The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
+Finally, the significant differences observed in BV/TV and DA, even with matched age and gender, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.\\
 
@@
 Components of the stiffness tensors are distributed to both sides of the diagonal.
-On the other hand, the linear regressions performed in this study using the OI original data set presents lower $R^2_{adj}$ and higher NE than such fit reach usually \cite{Gross2013,Panyasantisuk2015}.
+On the other hand, the linear regressions performed in this study using the original OI data set presents lower $R^2_{adj}$ and higher NE than comparable linear regressions in literature \cite{Gross2013,Panyasantisuk2015}.
 The important value of NE and its standard deviation shows that the fitted stiffness can deviate significantly from the observation.
 These differences come from ROIs presenting a low stiffness.
-The regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to $10^0$ MPa and lower, the fit tends to overestimate the stiffness.
+The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to $10^0$ MPa and lower, the linear regression tends to overestimate the stiffness.
 This is because ROI stiffness is highly dependent on BV/TV values.
-Some ROIs with low BV/TV don't have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor, see Appendix \ref{A2}.
+Some ROIs with low BV/TV do not have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor (see Appendix \ref{A2}).
 Trying to homogenize such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
@@
 As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenization seem feasible.
-However, the value of this threshold actually results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\
-
-The linear regressions performed on filtered data sets present direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values.
+However, the value of this threshold results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\
+
+The linear regressions performed on the filtered data sets present a direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values.
 For the healthy group (N=603), the relatively small decrease of $R^2_{adj}$ (0.5\%) compared to the linear regression using the unfiltered data set is negligible.
@@
 \citeauthor{Gross2013}\cite{Gross2013} showed that different anatomical locations lead to only slight differences.
-Comparing regression of the filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution.
+Comparing linear regression of the filtered data set of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} with the present study, the lower stiffness constants observed can be explained partially by the higher DA range and by the coarser resolution.
 Differences of $R^2_{adj}$ and NE come from the imposition of $k$ and $l$ to a different value than the optimal ones.
-Then, comparing regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges overlap.
+Then, comparing linear regression results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, \citeauthor{Gross2013}\cite{Gross2013}, and the present study, BV/TV ranges overlap.
 As for the filtered data sets, DA is higher in the present study and the stiffness constants remain lower than for the two other studies.
@@
 
-The analysis of tBMD gives interesting outputs as well.
-Although BV/TV and DA are in the same range because of the matching, the t-test reveal a higher tBMD in OI trabecular bone than in healthy condition with 95\% certainty and a very high significance level.
-The coefficients obtained from the linear mixed-effects model show that there is a relation between tBMD and BV/TV as zero is not included in the slope CI.
-This result can have different origins.
+The analysis of TMD gives interesting outputs as well.
+The t-test reveal a higher TMD in OI trabecular bone than in healthy condition with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
+The linear regression slope CI being strictly different from zero shows that is exist a relation between TMD and BV/TV.
+This can have different origins.
 From a biological point of view, the remodeling process leads to a mineralization gradient from the core of the trabecula to the outer surface.
-As trabecular thickness decreases with BV/TV, this means that with a lower BV/TV the core of trabeculae could be less mineralized.
+As trabecular thickness decreases with BV/TV, this means that a lower BV/TV could lead to a lower mineralization of the core of trabeculae.
 The second explanation for this slope comes from the scanning.
 Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV.
-Nevertheless, the former (biological) explanation is expect to have a less significant impact than the latter (scanning).
-Regarding the other coefficients of the linear regression, the CI of group variable exclude zero as well leading to the conclusion that we have 95\% certainty that the intercept is different depending on the group.
+The former (biological) explanation is expected to have a less significant impact than the latter (scanning).
+Regarding the other coefficients of the linear regression, the CI of group variable excludes zero as well, meaning that the intercept depends on the group.
 This could be explained by the bisphosphonate treatment that OI patients receive.
-Effectively, bisphosphonate is aimed to freeze the remodeling process which lead to higher mineralization of the bone.
-According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest to add a correction accounting for the tBMD in FE simulations to catch the higher modulus, ultimate stress and post-yield behavior of OI bone compared to healthy bone at the ECM level.
-Regarding the interaction between BV/TV and the group, the p-value obtained show a high non-significance meaning that the slopes of both OI and healthy groups are the same.
-This could be visualized in Figure \ref{02_tBMD} where the fitted lines are obtained using the fixed effects of the model and the group variable if fixed.
+Effectively, bisphosphonate is aimed to slow down the remodeling process which lead to higher mineralization of the bone.
+According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest adding a correction accounting for the TMD in FE simulations to catch the higher modulus, ultimate stress and post-yield behavior of OI bone compared to healthy bone at the ECM level.
+Regarding the interaction between BV/TV and the group, the p-value obtained showed a high non-significance meaning that the slopes of both OI and healthy groups are the same.
+This could be visualized in Figure \ref{02_TMD} where the fitted lines are obtained using the fixed effects of the model and the group variable if fixed.
 In this plot, performed with the interaction term (BV/TV x Group) the fitted lines appear to be quasi-parallel.
 
-The main limitations of this study are the definition of "homogeneous enough" and the fact that it is limited to tibiae XCTII scans.
+The main limitations of this study are the definition of "homogeneous enough" (i.e.
+low CV and DA) and the fact that it is limited to tibiae XCTII scans.
 Moreover, having only one patient with OI type III where we could extract ROIs does not allows to do statistics.
-Effectively as those patients are in wheelchair, it could be interesting to analyze the impact of this condition on the weight bearing tibia.
+Effectively as those patients are less active, it could be interesting to analyze the impact of this condition on the weight bearing tibia.
 The ROI homogeneity has an important impact on the analysis quality.
-As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for fitting.
+As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for linear regression analysis.
 More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models could be negligible if the proportion of such low stiffness ROIs stays low.
-A similar study could be performed on XCTII radii scans to confirm the low differences between anatomical locations for coarser resolution.
-Another limitation is that the scans were performed on different devices which were not cross-calibrated.
-However, as they are the same model, it is expected to have a minor impact.\\
+A similar study could be performed on radius XCTII reconstrauctions to confirm the low differences between anatomical locations for coarser resolution.
+Another limitation is that the measurements were performed on different devices which were not cross-calibrated.
+However, as they are the same model, this is expected to have a minor impact.\\
 
@@
 \section*{Acknowledgments}
-The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII scans of OI individuals.
+The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII reconstructions of OI individuals.
 This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.
@@
 
-Where $\epsilon$ is the vector of residuals. For one ROI, the system take the following form:
+Where $\epsilon$ is the vector of residuals. For one ROI, the system takes the following form:
 
@@
 Where $S_g$ is coded using a summation constrain \cite{Fox2016}, meaning, $S_g = -1$ for the healthy group and $S_g = 1$ for the OI group.
-This model is adapted into a linear mixed-effect model to analyze the relation between tBMD and BV/TV and the effect of the group.
+This model is adapted into a linear mixed-effect model to analyze the relation between TMD and BV/TV and the effect of the group.
 
@@
 	\begin{split}
-		\text{tBMD} = \begin{pmatrix}
+		\text{TMD} = \begin{pmatrix}
 			1 & \rho & S_g \\
@@
 To test the hypothesis of no interaction between the BV/TV and the group, i.e.
-the group has no significant influence on the tBMD versus BV/TV slope, the previous model was modified to add the interaction regressor.
+the group has no significant influence on the TMD versus BV/TV slope, the previous model was modified to add the interaction regressor.
 
@@
 	\begin{split}
-		\text{tBMD} = \begin{pmatrix}
+		\text{TMD} = \begin{pmatrix}
 			1 & \rho & S_g & \rho S_g\\

### 2021-07-02 03_Scripts/00_ROI_Selection.py

@@ -1,4 +1,5 @@
 # 00 Initialization
+
 import os
 import numpy as np
 import pandas as pd
@@ -118,8 +119,8 @@ ScanLists = [File for File in os.listdir(DataFolder) if File.endswith('.csv')]
 ScanLists.sort()
 
 # 03 Load Data
-DataGroup = 1   # 0 = Healthy group, 1 = OI group
-Plots = False   # Plot the different ROI location
+DataGroup = 0   # 0 = Healthy group, 1 = OI group
+Plots = True   # Plot the different ROI location
 
 ScansPath = os.path.join(DataFolder,DataSubFolders[DataGroup])
 Scans = [File for File in os.listdir(ScansPath) if File.endswith('.mhd')]
@@ -230,3 +231,97 @@ ColsToInt = ['$XPos', '$YPos', '$ZPos', '$XRange', '$YRange', '$ZRange', '$ROINu
 ParametersDataFrame[ColsToInt] = ParametersDataFrame[ColsToInt].astype(int)
 ParametersDataFrame.to_csv(os.path.join(ResultsFolder, DataSubFolders[DataGroup][:-14] + 'ROI2.csv'),
                            sep=';', line_terminator=';\n', index=False)
+
+
+
+
+Scan = Scans[12]
+ScanFile = os.path.join(ScansPath,Scan)
+CT_Scan, Origin, Spacing, Size = Load_Itk(ScanFile)
+
+CubeSide = 5.3    # Side length of cubic ROI (mm)
+VoxelsRange = np.round(CubeSide / Spacing).astype('int')
+
+# 04 Select stack
+if DataGroup == 0:
+    StackHeight = int(Size[0] / 3)
+    SelectedStack = CT_Scan[2*StackHeight:,:,:]
+    SoftTissue, TrabBone, CortBone = np.unique(SelectedStack)
+elif DataGroup == 1:
+    StackHeight = int(Size[0])
+    SelectedStack = CT_Scan
+    SoftTissue, TrabBone, CortBone = np.unique(SelectedStack)
+
+
+# 05 Select 6 ROIs
+VoxelMargin = 5
+ParametersDataFrame = pd.DataFrame()
+for ROINumber in range(6):
+
+    Z1, Y1, X1 = ROISelection(SelectedStack, VoxelsRange, VoxelMargin, ROINumber, TrabBone, CortBone)
+
+    if DataGroup == 0:
[…]

### 2021-07-21 09_Article/Simon_2021_Bone_Draft_V03.tex
Article corrections

--- 2021-07-02 09_Article/Simon_2021_Bone_Draft_V02.tex
@@
 
-\author[2,3]{Mahdi Tabatabaei}
-
-\author[2,3]{Bettina M. Willie}
+\author[2,3,4]{Seyedmahdi Hosseinitabatabaei}
+
+\author[2,3,4]{Bettina M. Willie}
 
@@
 
-\address[3]{Departement of Pediatric Surgery, McGill University, Montreal, Canada}
+\address[3]{Department of Pediatric Surgery, McGill University, Montreal, Canada}
+
+\address[4]{Department of Biomedical Engineering, McGill University, Montreal, Canada}
 
@@
 
+%
+%
+%
+% ABSTRACT
+%
+%
+%
+
 \begin{abstract}
-	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility.
-It is characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass.
+	Osteogenesis Imperfecta (OI) is an inherited form of bone fragility characterised by impaired synthesis of type I collagen, altered trabecular bone architecture and reduced bone mass.
 High resolution peripheral computed tomography (HR-pQCT) is a powerful method to investigate bone morphology at peripheral sites including the weight-bearing distal tibia.
 The resulting 3D reconstructions can be used as a basis of micro-finite element (\si{\micro}FE) or homogenised finite element (hFE) models for bone strength estimation.
-The hFE scheme uses homogenized local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
+The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
 Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
@@
 Six cubic regions of interest (ROIs) were selected per individual in a common anatomical region.
-A first matching between OI and healthy control group was performed by selecting similar individuals to obtain identical mean and median age and gender distribution.
-It allowed to perform a morphometric analysis and compare the outcome with literature.
-Then, stiffness tensors of ROIs were computed using hFE and multiple linear regressions were performed with the Zysset-Curnier orthotropic model.
-The linear regressions allowed to compare the two groups using five parameters.
+A first matching between OI and healthy control group was performed by selecting similar individuals to obtain identical mean and median age and sex distribution.
+It allowed us to perform a first morphometric analysis and compare the outcome with literature.
+Then, stiffness tensors of the ROIs were computed using  \si{\micro}FE and multiple linear regressions were performed with the Zysset-Curnier orthotropic fabric-elasticity model.
+%The linear regressions allowed to compare the two groups using five parameters.
+
 	An initial fit was performed on both the OI group and the healthy control group using all extracted ROIs.
-Then, data was filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing the ROI heterogeneity and additionnal linear regressions were performed on these filtered data sets.
+Then, data was filtered according to a fixed threshold for a defined coefficient of variation (CV) assessing ROI heterogeneity and additional linear regressions were performed on these filtered data sets.
 These full and filtered data were in turn compared with previous results from \si{\micro}CT reconstructions obtained in other anatomical locations.
-Finally, the ROIs of both groups were matched according to their BV/TV and fabric anisotropy (DA).
+Finally, the ROIs of both groups were matched according to their BV/TV and degree of anisotropy (DA).
 Linear regressions were performed using these matched data to detect statistical differences between the two groups.
 
-	Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular spacing standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
-These results are in agreement to literature.
+	Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
+These results are in agreement with previous studies.
 The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range.
-The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and lead to similar correlation coefficients for both OI and healthy groups.
+The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups.
 Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
@@
 	In conclusion, despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis.
-Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.
+Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections.
 \end{abstract}
@@
 
+%
+%
+%
+% INTRO
+%
+%
+%
 \section{Introduction}
@@
 OI prevalence is estimated at about  1/13,500 of births, less severe forms not being accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
-Therefore, OI is considered as a rare metabolic bone disorder.
+Regardless, OI is considered as a rare metabolic bone disorder.
 In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases.
@@
 Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
-High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,AriasMoreno2019}.
+High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}.
 Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
 However, HR-pQCT-based hFE relies on fabric-elasticity relationships.
-Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to investigate the hypothesis of similar fabric-elasticity relationships.
+Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.
+
+%
+%
+%
+% METHODS
+%
+%
+%
 
@@
 \subsection{HR-pQCT}
-HR-pQCT scans (XtremeCTII, SCANCO Medical, Brütisellen,
+HR-pQCT scans (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
 Switzerland) were performed at the distal tibia on all patients from both groups.
@@
 This led to a thickness of roughly 10.2 mm for each stack.
-Standardized scanning settings were used (voltage of 60 kVp, 900 μA, 100 ms integration time) for the healthy group as well as for the OI group.
+Standardized scanning settings were used (voltage of 60 kVp, 900 $\mu$A, 100 ms integration time) for the healthy group as well as for the OI group.
 For the healthy group, motion artefacts of first, middle and last slice (i.e slices number 1, 252, and 504) were graded on a scale proposed by \citeauthor{Pialat2012} \cite{Pialat2012} of 1 (no motion artefacts) to 5 (extreme motion artefacts).
@@
 	\end{subfigure}
-	\caption{\centering Clinical section scanned for both group}
+	\caption{Clinical section scanned for both group}
 	\label{01_ClinicalSections}
@@
 Cortical bone is in white and trabecular bone in gray.
-The stacks are separated by red lines and blue squares represent the ROIs.
-The stack used for analysis is higlighted in green and the dashed red line represents the middle of the stack.}
+The stacks are separated by red lines and cyan squares represent the ROIs.
+The stack used for analysis is higlighted in cyan and the dashed rose line represents the middle of the stack.}
 	\label{01_ROISelection}
@@
 deletion of bone material regions unconnected to the main structure, the morphological analysis of ROIs was performed using medtool (v4.5; Dr.
-Pahr Ingenieurs e.U., Pfaffstätten, Austria).
-The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular spacing (Tb.Sp.SD).
+Pahr Ingenieurs e.U., Pfaffst\"{a}tten, Austria).
+The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular separation (Tb.Sp.SD).
 Moreover, ROI tissue mineral density (TMD) and fabric was evaluated.
@@
 \begin{equation}
-	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i} = \sum_{i=1}^{3}{m_i \mathbf{m}_i \otimes \mathbf{m}_i}
+	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i} = \sum_{i=1}^{3}{m_i (\mathbf{m}_i \otimes \mathbf{m}_i})
 	\label{Eq201}
@@
 where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ \cite{Cowin1985,Harrigan1985}.
-The fabric tensor is independent of BV/TV and normalized with $tr(\mathbf{M}) = 3$.
+The fabric tensor is independent of BV/TV and normalised with $tr(\mathbf{M}) = 3$.
 The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.
-Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualization of its fabric tensor.\\
+Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualisation of its fabric tensor.\\
 
@@
 Then, a stiffness $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element.
-The homogenization process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}.
+The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}.
 Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system.
@@
 		&\mathbb{S}(\rho,\mathbf{M}) &=& \quad\sum_{i=1}^{3} \lambda_{ii} \mathbf{M}_i \otimes \mathbf{M}_i \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \lambda_{ij} \mathbf{M}_i \otimes \mathbf{M}_j \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \mu_{ij} \mathbf{M}_i \overline{\underline{\otimes}} \mathbf{M}_j \\
-		&\text{With} &\\
+		&\text{with} &\\
 		&\qquad\lambda_{ii} &=& \quad(\lambda_0 + 2\mu_0)\rho^k m_i^{2l} \\
@@
 Where $\otimes$ and $\overline{\underline{\otimes}}$ are the dyadic and symmetric product of second order tensors, respectively.
-To express the stiffness tensor obtained from the homogenization process with the Zysset-Curnier model, it had to be transformed into the fabric coordinate system using a coordinates transformation formula (see Equation \ref{Eq203p}) and projected onto orthotropy, leading to 12 components.
-
-\begin{equation}
-	\mathbb{S}_{ijkl}' = Q_{im}Q_{jn}Q_{ko}Q_{lp} \mathbb{S}_{mnop}
+To express the stiffness tensor obtained from the homogenisation process with the Zysset-Curnier model, it had to be transformed into the fabric coordinate system using a coordinates transformation formula (see Equation \ref{Eq203p}) and projected onto orthotropy, leading to 12 components.
+
+\begin{equation}
+	S_{ijkl}' = Q_{im}Q_{jn}Q_{ko}Q_{lp} S_{mnop}
 	\label{Eq203p}
@@
 
-Where $\mathbb{S}'$ and $\mathbb{S}$ are the transformed and the original stiffness tensor respectively and $Q$ is the orthogonal matrix that maps the original coordinate system into the new one (fabric).
-The Zysset-Curnier model is built with the assumption of orthotropy and homogeneity.
-However, the trabecular structure is not perfectly homogeneous.
+Where $\mathbb{S}'$ and $\mathbb{S}$ are the transformed and the original stiffness tensor respectively and $\mathbf{Q}$ is the orthogonal matrix that maps the original coordinate system into the new one (fabric).
+The Zysset-Curnier model is built with the assumption of orthotropy and homogeneity of the ROI.
+However, the natural trabecular structure is not perfectly homogeneous.
 To assess the ROI heterogeneity, a coefficient of variation (CV) is computed according to \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}: the ROI is divided into eight identical sub-cubes and BV/TV is computed for each of them.
-The CV is defined as the ratio between the standard deviation of these BV/TV and the mean value (Equation \ref{Eq203}).
-
-\begin{equation}
-	CV = \frac{SD(BV/TV_{subcubes})}{mean(BV/TV_{subcubes})}
+The CV is defined as the ratio between the standard deviation of these $\mathrm{BV/TV_i}$ and the mean value (Equation \ref{Eq203}).
+
+\begin{equation}
+	\mathrm{CV} = \frac{\mathrm{SD\;(BV/TV_i)}}{\mathrm{Mean\;(BV/TV_i)}}
 	\label{Eq203}
@@
 
-\subsection{Morpholigical Analysis}
-The analyzed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI group.
+\subsection{Morphological Analysis}
+The analysed morphological parameters (BV/TV, Tb.N., Tb.Th., Tb.Sp., and Tb.Sp.SD, SMI, DA, and CV) were compared between the healthy and the OI group.
 
 \subsubsection{Participants Matching}
-As the groups do not have similar distributions of age and sex, a matching was performed by selecting similar individuals leading to identical mean and median age as well as identical gender distribution.
+As the groups do not have similar distributions of age and sex, a matching was performed.
+The matching algorithm computed the Mahalanobis distance between participants with age and sex as covariates.
+Then, best correspondences between OI and healthy participants were kept and duplicates were dropped (i.e.
+no replacement)\cite{Stuart2010}.
+This matching led to identical mean and median age as well as identical sex distribution.
 
@@
 Standard linear models assume independent and identically distributed (iid) variables.
-As this assumption was violated by the fact that six ROIs were analyzed per individual, a linear mixed-effect model was preferred.
+As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred.
 This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual.
@@
 
-Where $S_{rc}$ is the $r$th row and $c$th column of the non-zero element of the orthotropic stiffness tensor $\mathbb{S}$ in Mandel notation \cite{MANDEL1965}, $X$ is a $12n$x$p$ design matrix containing the the BV/TV and fabric info of the $n$ ROIs and $\beta$ is a $p$x1 vector of fixed effects containing model parameters.
-$Z$ is a $12n$x$f$ design matrix which contains data with individual dependence and $\delta$ is a $f$x1 vector composed of random factors.
-Finally, $\epsilon$ is a $12n$x1 vector containing the linear regression residuals.
+Where $S_{rc}$ is the $r$th row and $c$th column of the non-zero element of the orthotropic stiffness tensor $\mathbb{S}$ in Mandel notation \cite{MANDEL1965}, $X$ is a $12 n \times p$ design matrix containing the the BV/TV and fabric info of the $n$ ROIs and $\beta$ is a $p \times 1$ vector of fixed effects containing model parameters.
+$Z$ is a $12 n \times f$ design matrix which contains data with individual dependence and $\delta$ is a $f \times 1$ vector composed of random factors.
+Finally, $\epsilon$ is a $12n \times 1$ vector containing the linear regression residuals.
 As $k$ and $l$ are exponents, the linear regression was performed on the log space.\\
@@
 \begin{equation}
-	R^2_{adj} = 1 - \frac{RSS}{TSS} \frac{(12n-1)}{(12n - p - 1)}
+	R^2_{adj} = 1 - \frac{\mathrm{RSS}}{\mathrm{TSS}} \frac{(12n-1)}{(12n - p - 1)}
 	\label{Eq205}
@@
 	\centering
-	\caption{Summary of the data set used for different methods}
+	\caption{Summary of the number of ROIs used for the different steps of the study.}
 	\label{Table1}
@@
 		\toprule
-		Data sets & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Age \& gender matched} & \multicolumn{2}{c}{CV filtered} & \multicolumn{2}{c}{BV/TV \& DA matched} \\
+		Data sets & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Age \& sex matched} & \multicolumn{2}{c}{CV filtered} & \multicolumn{2}{c}{BV/TV \& DA matched} \\
 		\midrule
@@
 
+%
+%
+%
+% RESULTS
+%
+%
+%
+
 \section{Results}
-
+%
 \subsection{Morphological Analysis}
-The results of the morphological analysis are summarized in Table \ref{Table2}.
+The results of the morphological analysis are summarised in Table \ref{Table2}.
 The individual matching for age and sex allowed to have similar group distributions with 17 females and 11 males in each group.
@@
 The trabecular thickness does not show significant differences between groups.
-On the other hand, permutation test performed for trabecular separation showed that it is higher in mat\-ched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$).
+On the other hand, permutation test performed for trabecular separation showed that it is higher in matched OI group compared to healthy individuals with a p-value of 0.01 and an exclusion range of ($-\infty$, -0.384] $\cup$ [0.421, $\infty$).
 Trabecular separation SD is higher in OI individuals compared to matched healthy individuals with a p-value of 0.02 and an exclusion range of ($-\infty$, -0.232] $\cup$ [0.251, $\infty$).
 SMI as well as the degree of anisotropy are higher for matched OI than for healthy people with p-values <0.001 and of 0.02, respectively.
-Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [−0.757, −0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\
+Finally, the log transformation of the coefficient of variation gives the stronger difference in means with a p value <0.0001 and a 95\% CI of [-0.757, -0.333] where CV is higher in matched OI individuals compared to matched healthy individuals.\\
 
@@
 The second and the last row show linear regression results using the individual data sets and setting the exponents $k$ and $l$ to fixed values.
-OI stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) are consistantly higher than healthy one.
+OI stiffness constants ($\lambda_0$, $\lambda_0'$, and $\mu_0$) are consistently higher than healthy one.
 The increase is of 15\%, 1\%, and 2\% for $\lambda_0$, $\lambda_0'$, and $\mu_0$, respectively.
@@
 Values are presented as value [95\% CI] or mean $\pm$ standard deviation.
-Values in gray were fixed in the linear regression.}
+Values in grey were fixed in the linear regression.}
 	\label{Table3}
@@
 
+%
+%
+%
+% DISCUSSION
+%
+%
+%
+
 \section{Discussion}
-Osteogenesis imperfecta is an inherited form of bone fra\-gi\-li\-ty with a severity going from mild to perinatally lethal.
+Osteogenesis imperfecta is an inherited form of bone fra\-gility with a severity going from mild to perinatally lethal.
 This study aims to confirm that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment.
-To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.\\
-
-As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare morphological parameters.
-The imaging system explains most of the differences between the absolute morphological values of the present study compared to the others.
+To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
+\\
+As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare the morphological parameters.
+The CT imaging system explains most of the differences between the absolute morphological values of the present study compared to the others.
 \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have performed their measurements on first generation XCT scanners with a voxel size of 82 \si{\micro}m, while we have used a second generation XCT with a voxel size of 61 \si{\micro}m.
@@
 is lower compared to first generation XCT.
-These results give confidence in our observed values.
+These findings are in full agreement with our observations.
 Another potential bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length.
@@
 as they depend on the ROI size.
-Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain a certain portion of trabecular bone.
+Moreover, conditions imposed for random ROI selection can lead to further biased values, especially for OI patients, as the ROI must contain at least some portion of trabecular bone.
 Even with the low sample size (2x28 individuals), the statistical tests have shown significant differences between groups, with the CV presenting the most significant difference.
 The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
-Finally, the significant differences observed in BV/TV and DA, even with matched age and gender, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.\\
-
+Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
+\\
 The linear regressions performed in this study on original data sets showed $R^2_{adj}$ and NE in the expected range (i.
@@
 Some ROIs with low BV/TV do not have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor (see Appendix \ref{A2}).
-Trying to homogenize such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
+Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
 Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
-An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.\\
-
+An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.
+\\
 Figure \ref{02_CV_BVTV} presents the CV in relation to BV/TV.
@@
 Pearson's correlation coefficient being strictly negative confirms a negative monotonic relation.
-As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenization seem feasible.
-However, the value of this threshold results from an optimization process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be subject to more investigations.\\
-
+As some ROIs with higher BV/TV still present high CV values, imposing a fixed threshold for subsequent homogenisation seems feasible.
+However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further.
+\\
 The linear regressions performed on the filtered data sets present a direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values.
@@
 $R^2_{adj}$, NE, and the range of stiffness values are almost at the level of the healthy group.
-These results give confidence to the filtering procedure and are a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\
-
+These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\
 After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for the individual filtered data sets.
@@
 With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-In FEA simulations, it is not possible to exclude part of the mesh because of high heterogeneity.
-Nevertheless, the error created by such ROIs is negligible as this concerns ROIs with extremely low stiffness leading to a minor impact on the full model.\\
-
-Imposing $k$ and $l$ allows one to estimate the effect of different image resolutions.
+In hFE analysis of bones, it is not possible to exclude part of the mesh because of high heterogeneity.
+Nevertheless, the error created by such ROIs is negligible as this concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.\\
+Imposing $k$ and $l$ allows one to compare the other regressions parameters and estimate the effect of different image resolutions.
 \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and \citeauthor{Gross2013}\cite{Gross2013} both used femur scans with 18 \si{\micro}m spatial resolution and coarsened them to 36 \si{\micro}m.
@@
 Here, differences in DA can mainly be explained by the different anatomical location and differences in stiffness constants as a result of the different image resolutions.
-The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explains this increase of DA.
-Lower stiffness constants are obtained because the coarser structure resulting from XCTII can't be as optimized as the fine detailed structure obtained by \si{\micro}CT.
-Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimized morphology of trabecular bone with a high fidelity.
+The distal tibia, unlike the proximal femur, is mainly loaded in one direction which explains this higher DA.
+Lower stiffness constants are obtained because the coarser structure resulting from XCTII cannot be as optimised as the fine detailed structure obtained by \si{\micro}CT for the same BV/TV.
+Effectively, the architecture resulting from \si{\micro}CT scans can reproduce the optimised morphology of trabecular bone with higher fidelity.
 By decreasing the scan spatial resolution, the scanned structure becomes bulkier.
-Performing a linear regression on this less optimized structure leads to the observed lower stiffness constants.
+Performing a linear regression on this less optimised structure leads to the observed lower stiffness constants.
 Finally, the comparison between $R^2_{adj}$ and NE of current study without imposing $k$ and $l$ and the ones of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and  \citeauthor{Gross2013}\cite{Gross2013} shows that lower spatial resolution leads to lower fit quality.
 Nevertheless, $l$ stays in the same range as for the two other studies \cite{Gross2013,Panyasantisuk2015}, meaning the relative weight of DA remains constant.
-On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.\\
-
+On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.
+\\
 The analysis of TMD gives interesting outputs as well.
 The t-test reveal a higher TMD in OI trabecular bone than in healthy condition with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
-The linear regression slope CI being strictly different from zero shows that is exist a relation between TMD and BV/TV.
+The linear regression slope CI being strictly different from zero shows that a relation exists between TMD and BV/TV.
 This can have different origins.
-From a biological point of view, the remodeling process leads to a mineralization gradient from the core of the trabecula to the outer surface.
-As trabecular thickness decreases with BV/TV, this means that a lower BV/TV could lead to a lower mineralization of the core of trabeculae.
+From a biological point of view, a mineralisation gradient develops between the core of the trabecula undergoing secondary mineralisation and the outer surface subjected to remodelling and deposition of younger, less mineralised matrix.
+As trabecular thickness decreases with BV/TV, this means that a lower BV/TV could lead to a lower mineralization of the core of the trabeculae.
 The second explanation for this slope comes from the scanning.
-Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV.
-The former (biological) explanation is expected to have a less significant impact than the latter (scanning).
+Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV as the ratio of bone surface to bone volume BS/BV decreases.
+The former (biological) explanation is expected to have a lower impact than the latter (scanning).
 Regarding the other coefficients of the linear regression, the CI of group variable excludes zero as well, meaning that the intercept depends on the group.
 This could be explained by the bisphosphonate treatment that OI patients receive.
-Effectively, bisphosphonate is aimed to slow down the remodeling process which lead to higher mineralization of the bone.
-According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest adding a correction accounting for the TMD in FE simulations to catch the higher modulus, ultimate stress and post-yield behavior of OI bone compared to healthy bone at the ECM level.
+Effectively, bisphosphonates are aimed to slow down the remodelling process which lead to higher (secondary) mineralization of the bone.
+According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest adding a correction accounting for the TMD in FE simulations to catch the higher modulus, ultimate stress and post-yield behaviour of OI bone compared to healthy bone at the ECM level.
 Regarding the interaction between BV/TV and the group, the p-value obtained showed a high non-significance meaning that the slopes of both OI and healthy groups are the same.
-This could be visualized in Figure \ref{02_TMD} where the fitted lines are obtained using the fixed effects of the model and the group variable if fixed.
+This could be visualised in Figure \ref{02_TMD} where the fitted lines are obtained using the fixed effects of the model and the group variable if fixed.
 In this plot, performed with the interaction term (BV/TV x Group) the fitted lines appear to be quasi-parallel.
-
+\\
 The main limitations of this study are the definition of "homogeneous enough" (i.e.
-low CV and DA) and the fact that it is limited to tibiae XCTII scans.
+low CV) and the fact that it is limited to  XCTII reconstructions of the distal tibia.
 Moreover, having only one patient with OI type III where we could extract ROIs does not allows to do statistics.
-Effectively as those patients are less active, it could be interesting to analyze the impact of this condition on the weight bearing tibia.
-The ROI homogeneity has an important impact on the analysis quality.
-As proposed earlier, ROI homogeneity could be assessed in another way to be able to propose a more precise ROI filtering for linear regression analysis.
-More investigations could be performed to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROI with low stiffness the impact on FEA models could be negligible if the proportion of such low stiffness ROIs stays low.
-A similar study could be performed on radius XCTII reconstrauctions to confirm the low differences between anatomical locations for coarser resolution.
-Another limitation is that the measurements were performed on different devices which were not cross-calibrated.
-However, as they are the same model, this is expected to have a minor impact.\\
-
-In conclusions, the samples analyzed in the present study had similar morphology compared to data reported in the literature.
-We couldn't find differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
-with a CV lower than 0.263.
-\citeauthor{Indermaur2021}\cite{Indermaur2021} could show that the compressive behaviour of OI bone tissue is similar to the one of healthy control at the ECM level.
-If the tensile and shearing behaviour is similar as well, fabric-strength relationships will hold too.
-Therefore, OI trabecular bone can explain part of the bone fragility by the decrease in BV/TV and the loss of homogeneity in its trabecular organization.
+%As those patients are less active, it could be interesting to analyse the impact of this condition on the weight bearing tibia.
+ROI homogeneity is a necessary condition underlying homogenisation of mechanical properties and the use of fabric-elasticity relationships. %has an important impact on the analysis quality. 
+As proposed earlier, ROI homogeneity could be assessed with an alternative ROI filtering and more efforts could be invested to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROIs with extremely low stiffness, the impact on hFE models should remain negligible.
+%if the proportion of such ROIs stays low.
+A similar study could be performed on radius XCTII reconstructions to generalise the above findings.
+Another limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated.
+However, as they are the same model (XCTII) and the quality control measurement ensures $\pm$ 1\% precision on BMD, this is expected to have a minor impact.
+\\
+
+In conclusion, the OI trabecular bone from the distal tibia analysed in the present study had similar morphology compared to those reported in the literature.
+We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
+with a CV lower than 0.263 and were matched for BV/TV and DA.
+\citeauthor{Indermaur2021} \cite{Indermaur2021} found that the compressive behaviour of OI bone extracellular matrix is not inferior to healthy control at the ECM level.
+Similar results for tensile properties at the extracellular matrix level would indicate that fabric-strength relationships may be similar as well.
+Under this hypothesis, fragility of OI bone should mainly be explained by thin cortices and low BV/TV of the trabecular compartment.
+
+%
+%
+%
+% ACKNOWLEGMENTS
+%
+%
+%
 
 \section*{Acknowledgments}
-The authors acknowledge Christina Wapp from the ARTORG Center for Biomedical Engineering Research for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII reconstructions of OI individuals.
-This work was internally funded by the ARTORG Center for Biomedical Engineering Research and Mereo BioPharma.
+The authors acknowledge Christina Wapp from the ARTORG Center for her contribution to the building and interpretation of linear mixed-effect model and Mereo BioPharma for sharing the tibia XCTII reconstructions of OI individuals.
+This work was internally funded by the ARTORG Center and Mereo BioPharma.
 Bettina M.

### 2021-07-21 09_Article/Submission/Simon_2021_Bone_Draft.tex
Article corrections

--- 2021-07-21 09_Article/Simon_2021_Bone_Draft_V03.tex
@@
 
-\author[2,3,4]{Seyedmahdi Hosseinitabatabaei}
-
-\author[2,3,4]{Bettina M. Willie}
+\author[2,3]{Mahdi Tabatabaei}
+
+\author[2,3]{Bettina M. Willie}
 
@@
 
-\address[3]{Department of Pediatric Surgery, McGill University, Montreal, Canada}
-
-\address[4]{Department of Biomedical Engineering, McGill University, Montreal, Canada}
+\address[3]{Departement of Pediatric Surgery, McGill University, Montreal, Canada}
 
@@
 Six cubic regions of interest (ROIs) were selected per individual in a common anatomical region.
-A first matching between OI and healthy control group was performed by selecting similar individuals to obtain identical mean and median age and sex distribution.
-It allowed us to perform a first morphometric analysis and compare the outcome with literature.
+A first matching between OI and healthy control group was performed by selecting similar individuals to obtain identical mean and median age and gender distribution.
+It allowed to perform a first morphometric analysis and compare the outcome with literature.
 Then, stiffness tensors of the ROIs were computed using  \si{\micro}FE and multiple linear regressions were performed with the Zysset-Curnier orthotropic fabric-elasticity model.
@@
 
-	Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
+	Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular spacing standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
 These results are in agreement with previous studies.
 The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range.
-The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups.
+The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and lead to similar correlation coefficients for both OI and healthy groups.
 Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
@@
 	In conclusion, despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis.
-Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections.
+Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections or parts.
 \end{abstract}
@@
 OI prevalence is estimated at about  1/13,500 of births, less severe forms not being accounted in this estimation as they are recognized later in life \cite{Lindahl2015}.
-Regardless, OI is considered as a rare metabolic bone disorder.
+Therefore, OI is considered as a rare metabolic bone disorder.
 In most cases, OI is caused by mutations in genes encoding type I collagen (COL1A1 and COL1A2), leading to brittle and fragile bones \cite{LIM2017}, as well as deformed geometry and size in some cases.
@@
 Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
-High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}.
+High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,AriasMoreno2019}.
 Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
@@
 Pahr Ingenieurs e.U., Pfaffst\"{a}tten, Austria).
-The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular separation (Tb.Sp.SD).
+The morphological parameters analyzed were: BV/TV, structural model index (SMI), trabecular number (Tb.N.), trabecular thickness (Tb.Th.), trabecular separation (Tb.Sp.), and the standard deviation of the trabecular spacing (Tb.Sp.SD).
 Moreover, ROI tissue mineral density (TMD) and fabric was evaluated.
@@
 As the groups do not have similar distributions of age and sex, a matching was performed.
-The matching algorithm computed the Mahalanobis distance between participants with age and sex as covariates.
+The matching algorithm computed the Mahalanobis distance between participants with age and gender as covariates.
 Then, best correspondences between OI and healthy participants were kept and duplicates were dropped (i.e.
 no replacement)\cite{Stuart2010}.
-This matching led to identical mean and median age as well as identical sex distribution.
+This matching led to identical mean and median age as well as identical gender distribution.
 
@@
 		\toprule
-		Data sets & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Age \& sex matched} & \multicolumn{2}{c}{CV filtered} & \multicolumn{2}{c}{BV/TV \& DA matched} \\
+		Data sets & \multicolumn{2}{c}{Original} & \multicolumn{2}{c}{Age \& gender matched} & \multicolumn{2}{c}{CV filtered} & \multicolumn{2}{c}{BV/TV \& DA matched} \\
 		\midrule
@@
 The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
-Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
+Finally, the significant differences observed in BV/TV and DA, even with matched age and gender, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
 \\

### 2021-07-21 03_Scripts/A1_tBMD_Evaluation.py

@@ -505,23 +505,6 @@ if pValue > 5/100:
               + str(np.quantile(y, 0.75).round(3)) + ']')
         print('\nMann-Whitney test statistic results')
         print('p value: ' + str(np.round(p, 9)))
-else:
-
-    d, RejectionRange, p = PermutationTest(x,y,Variable=Variable)
-
-    print('Variances are not equal')
-    print('\nHealthy group')
-    print('Mean: ' + str(np.round(x.mean(), 3)))
-    print('Std: ' + str(np.round(x.std(ddof=1), 3)))
-    print('OI group')
-    print('Mean: ' + str(np.round(y.mean(), 3)))
-    print('Std: ' + str(np.round(y.std(ddof=1), 3)))
-    print('\nPermutation test statistic results')
-    print('p value: ' + str(np.round(p, 9)))
-    print('95% CI rejection range : ('
-          + str(RejectionRange[0].round(3))
-          + '] U ['
-          + str(RejectionRange[1].round(3)) + ')')
 
 
 BoxData = pd.concat([Healthy_Data['Mean tBMD'],OI_Data['Mean tBMD']],axis=1)
@@ -635,13 +618,6 @@ R2, SE = PlotRegressionResults(Data_LMM, Data, Alpha=0.95, Random=False)
 
 
 ## ANCOVA
-Data_LM_1 = smf.ols("tBMD ~ BVTV + Group + BVTV*Group", data=Data).fit()
-Data_LM_2 = smf.ols("tBMD ~ BVTV + Group", data=Data).fit()
-Data_LM_3 = smf.ols("tBMD ~ BVTV", data=Data).fit()
-
-Models = [Data_LM_1,Data_LM_2,Data_LM_3]
-ModelsResults, ANCOVATable = ANCOVA(Data,Models,Random=False)
-
 def ANCOVA(Data,Models,Random):
 
     ModelsResults = pd.DataFrame()
@@ -696,6 +672,13 @@ def ANCOVA(Data,Models,Random):
     print(ANCOVA[['Source','Models contrasted','Sum of Squares','df','F','p']])
 
     return ModelsResults, ANCOVA
+
+Data_LM_1 = smf.ols("tBMD ~ BVTV + Group + BVTV*Group", data=Data).fit()
+Data_LM_2 = smf.ols("tBMD ~ BVTV + Group", data=Data).fit()
[…]

### 2021-07-22 09_Article/Simon_2021_Bone_Draft_V03.tex
Article corrections

@@
 These subjects had not taken any medication known to affect bone metabolism nor presented with any prior osteoporosis fracture.
-The second group was scanned as part of the ASTEROID study (NCT number: NCT03118570) at different locations, namely at the Shriners Hospital for Children-Canada.
-The study coordination was done by the McGill University in Montreal.
+The second group was scanned as part of the ASTEROID study (NCT number: NCT03118570) at 6 HR-pQCT (XtremeCT II) imaging sites in North America.
+The study coordination was done by the team at Shriners Hospital for Children-Canada/McGill University in Montreal.
 This group was composed of 35 females and 15 males with confirmed diagnosis of OI Type I, III or IV.
@@
 \subsection{HR-pQCT}
-HR-pQCT scans (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
-Switzerland) were performed at the distal tibia on all patients from both groups.
+HR-pQCT imaging (XtremeCTII, SCANCO Medical, Br\"{u}tisellen,
+Switzerland) was performed at the distal tibia on all patients from both groups.
 Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
 Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
-On the other hand, participants in the OI group were scanned using the manufacturer's standard protocol i.e.
-the reference line is placed at the subchondral endplate of the ankle joint and one stack was scanned at 22.5 mm proximal to the reference line (standard clinical section) \cite{Whittier2020}, see Figure \ref{01_OI}.
+On the other hand, for participants in the OI group, the reference line is placed at the subchondral endplate of the ankle joint.
+The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}.
+Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol, see Figure \ref{01_OI}.
 The scanned region according to the two different protocol are shown side by side in Figure \ref{01_ClinicalSections}.
@@
 Each ROI had to contain trabecular bone, but no cortical bone.
-For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at rougthly the same anatomical location as for the OI group.
+For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at roughly the same anatomical location as for the OI group.
 Then for both the healthy group and the OI group, the stack was divided into two halves and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
@@
 	{Pictures/01_ROISelection_Example}
-	\caption{Middle slices example of ROIs selection for HR-pQCT reconstruction of a healthy tibia.
-Cortical bone is in white and trabecular bone in gray.
-The stacks are separated by red lines and cyan squares represent the ROIs.
-The stack used for analysis is higlighted in cyan and the dashed rose line represents the middle of the stack.}
+	\caption{Example of randomly selected cubic ROIs (cyan squares) on a healthy bone in two planar views (left and right).
+Different stacks are separated by red lines, the stack used for analysis is highlighted in cyan, and the middle slice of the stack is represented by the dashed rose line.
+Cortical bone is in white and trabecular bone in gray.}
 	\label{01_ROISelection}
@@
 	{Pictures/01_FabricExample}
-	\caption{Typical ROI with the visualization of its fabric tensor using MIL method.
+	\caption{Typical ROI with the visualization of its fabric tensor (colored ellipsoid) using MIL method.
 Eigenvectors of the fabric tensor define its orientation and eigenvalues set lengths of the ellipsoid radii.
@@
 In brief, each voxel of the cleaned ROI was converted to a fully integrated linear brick elements (C3D8) using a direct voxel conversion approach.
-Then, a stiffness $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element.
+Then, a Young's modulus $E$ of 10,000 MPa and a Poisson's ratio $\nu$ of 0.3 were assigned to each element.
 The homogenisation process consisted of 6 independent simulations of different load cases, 3 uni-axial and 3 simple shear cases, using kinematic uniform boundary conditions (KUBCs) \cite{Panyasantisuk2015}.
@@
 The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
-This model compute the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\
+This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\
 
@@
 \subsubsection{Statistical Significance}
-The general significance level was set to 95\% for all tests.
+The general confidence level was set to 95\% for all tests.
 Confidence intervals were computed for t-tested variables to quantify the difference in both groups means.
@@
 The linear regression was performed using the \textsc{statsmodels} package from \textsc{Python 3.6}.
-Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient ($R^2$) and the standard error of the estimate (SE).
+Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient squared ($R^2$) and the standard error of the estimate (SE).
 Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}.
@@
 It can be noticed that, as the values of the observed stiffness tensor decreases, data points tend to be further apart from the diagonal (dashed line).
-Moreover, data the points from the stiffness tensors with lowest values are exclusively above the diagonal.
+Moreover, the data points from the stiffness tensors with lowest values are exclusively above the diagonal.
 The range of stiffness tensors of the OI group is wider compared to the one of the healthy group and ROIs with lower BV/TV present lower stiffness values.
@@
 Another potential bias is introduced by the fact that the present study analyses the median values of six cubic ROIs with 5.3 mm side length.
-This conditions the Tb.N.
+This influences the Tb.N.
 and Tb.Sp.
@@
 The analysis of TMD gives interesting outputs as well.
-The t-test reveal a higher TMD in OI trabecular bone than in healthy condition with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
+The t-test reveal a higher TMD in OI trabecular bone than in healthy condition, agreeing with \citeauthor{Boyde1999}\cite{Boyde1999}, with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
 The linear regression slope CI being strictly different from zero shows that a relation exists between TMD and BV/TV.
@@
 low CV) and the fact that it is limited to  XCTII reconstructions of the distal tibia.
-Moreover, having only one patient with OI type III where we could extract ROIs does not allows to do statistics.
+Moreover, having only one patient with OI type III where we could extract ROIs does not allows to perform statistics.
 %As those patients are less active, it could be interesting to analyse the impact of this condition on the weight bearing tibia.
@@
 A similar study could be performed on radius XCTII reconstructions to generalise the above findings.
-Another limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated.
+Another limitation is the anatomical site scanned in control and OI group.
+The section scanned for the control group was at a fixed distance.
+On the other hand, the section scanned for the OI group was placed at a distance relative to the total length of the tibia.
+Nevertheless, this relative distance is aimed to be at the same location that the standard clinical section on average.
+Moreover, this study investigates mainly the fabric-elasticity relationships of trabecular bone and ROIs were matched to compare only architectural differences between healthy and OI trabecular bone.
+A final limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated.
 However, as they are the same model (XCTII) and the quality control measurement ensures $\pm$ 1\% precision on BMD, this is expected to have a minor impact.

### 2021-07-22 03_Scripts/A1_tBMD_Evaluation.py

@@ -2,10 +2,7 @@
 import os
 import numpy as np
 import pandas as pd
-import SimpleITK as sitk
 import matplotlib.pyplot as plt
-import struct
-import re
 import statsmodels.formula.api as smf
 from scipy.stats.distributions import norm, chi2, t, f
 from scipy.stats import shapiro

### 2021-07-28 09_Article/Simon_2021_Bone_Draft_V03.tex
Submission of the manuscript

@@
 Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
-On the other hand, for participants in the OI group, the reference line is placed at the subchondral endplate of the ankle joint.
-The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}.
-Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol, see Figure \ref{01_OI}.
-The scanned region according to the two different protocol are shown side by side in Figure \ref{01_ClinicalSections}.
+On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.
+The mid-point of the scanned volume of interest was located at a distance of 7\% of total tibia length from the reference line \cite{GHASEMZADEH2017,Bonaretti2017}, see Figure \ref{01_OI}.
+Then, the scans were acquired according to the manufacturer's standard in vivo acquisition protocol.
+The scanned regions according to these two different protocols coincide approximately and are shown side by side in Figure \ref{01_ClinicalSections}.
 For both the healthy and the OI groups, each stack consisted of 168 slices and an isotropic voxel resolution of 61 \si{\micro}m.
@@
 A similar study could be performed on radius XCTII reconstructions to generalise the above findings.
-Another limitation is the anatomical site scanned in control and OI group.
-The section scanned for the control group was at a fixed distance.
-On the other hand, the section scanned for the OI group was placed at a distance relative to the total length of the tibia.
-Nevertheless, this relative distance is aimed to be at the same location that the standard clinical section on average.
-Moreover, this study investigates mainly the fabric-elasticity relationships of trabecular bone and ROIs were matched to compare only architectural differences between healthy and OI trabecular bone.
+Another minor limitation is the slightly different protocol for selection of the anatomical site scanned in the control and OI groups.
+The section scanned for the control group was at a fixed distance, while the section scanned for the OI group was placed at a distance relative to the total length of the tibia.
+Nevertheless, as shown in Figure \ref{01_ClinicalSections}, the locations compare well and the ROIs were matched for BV/TV and DA to compare the fabric-elasticity relationships.
 A final limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated.

### 2021-09-30 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex

% An example for the ar2rc document class.
% Copyright (C) 2017 Martin Schroen
% Modifications Copyright (C) 2020 Kaishuo Zhang
%
% This program is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
%
% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.

% Use quote enviromnent for manuscript text

\documentclass{AR2RC}

\usepackage{siunitx}

\title{Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals}
\author{Mathieu Simon, Michael Indermaur, Denis Schenk, Seyedmahdi Hosseinitabatabaei, Bettina	M. Willie and Philippe Zysset}
\journal{Bone}
\doi{BONE-D-21-00758}

\begin{document}

\maketitle

\vspace{1em}The authors would like to thanks the reviewers for their time spent for this article review. The work they performed will allow to improve the quality of the present article and help the authors to transmit the knowledge resulting of their work with more clarity. For this, we answered at your comments as presented in this document.

\section{Reviewer \#1}

\RC This report analyzed three dimensional morphology and stiffness of trabecular bone in patients with Osteogenesis Imperfecta (OI), and compared them with those from healthy control group. 3D morphologies were extracted from HR-pQCT data; while stiffness information was obtained by numerical simulation techniques. This is an important research topic with potential applications in assessing bone fragility as a result of OI. However, this reviewer felt the manuscript did not provide sufficient information to clearly explain new progress and novelty.

\AR To explain new progress shown in the present article, the manuscript was modified at multiple places as follow:\par

Abstract, Line 23-24
\begin{quote}
	Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls \Add{because validity of these relationships remains unclear for OI bone and are crucial for bone strength assessment using hFE}.
\end{quote}

Abstract, Line 48-49
\begin{quote}
	 Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections. \Add{Thus, the hFE scheme used for bone strength assessment of healthy individuals is valid for OI individuals as well.}
\end{quote}

\subsection{Methods}

\RC The general methodology used in this paper followed closely a previously published paper by the same senior author (Zysset, ref 12) on human trabecular bone. Briefly, CT imaging of trabecular bone was followed by micro-FE modeling for stiffness (FE stiffness), and a fabric-elasticity model (Predicted stiffness) reported also by the same senior author (Zysset) in 1995 (ref 32 ). Other than applications of the protocols to OI patients, contributions to either imaging or the modeling techniques seem to be moderate.

\AR You have raised an important point here. The methodology used in the present paper is based on previous publications for trabecular bone analysis. With time, these different works led to the development of the hFE scheme for bone strength estimation. However, the work cited here (Zysset, ref 12), is based on \si{\micro}CT scan and was performed on human vertebrae, proximal femur and distal radius. It is the case for another previous work (Zysset, ref 24) where an analysis of the influence of boundary conditions type was performed. Again, stiffness tensor was computed from numerical simulations of trabecular bone from a \si{\micro}CT scans on the proximal femur. The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia 2) using HR-pQCT scans i.e. lower resolution scans 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals. More emphasis on these points is put with the following modifications of the manuscript:\par

Introduction, Line 44-45
\begin{quote}
	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
\end{quote}

\subsection{Article contribution}

\RC Key progress and contribution are in the conclusion that fabric-elasticity relationships in OI trabecular bone are similar compared to healthy bone, based on which one could use HR-pQCT based hFE outcomes for fracture risk assessment. This conclusion was made based on filtered data (Figure 7), and therefore is valid only when trabecular bone is homogeneous enough. When all data including those with low BV/TV are included, the correlation between the fitted stiffness tensor and the observed one is different between the OI group and the healthy group. In other words, the conclusion is no longer valid for trabecular bone with low BV/TV. In terms of fracture risk, trabecular bone with lower density (low BV/TV) in OI patients would pose higher fracture risk, but could not be accurately predicted with the proposed method. Therefore, progress related to OI bone seems to be limited.

\AR Explain filtering of low stiffness means filtering of low BV/TV...

\subsection{OI brittelness}

\RC OI affects the production of collagen Type I, and therefore the brittleness of the bone matrix (a composite of bone mineral and collagen protein). Current methodology seemed to be limited to elasticity theory. Its application to predicting bone fragility (toughness) in the case of OI needs has yet to be proved.

\AR Michi articles

\section{Reviewer \#2}

\RC The manuscript showed a mathematical investigation of the fabric-elasticity relationships in case of osteogenesis imperfecta and healthy trabecular bones, obtained from distal tibiae. The evidence suggested that the fabric-elasticity relationships of sufficiently homogeneous trabecular tissue with osteogenesis imperfecta and healthy trabecular tissue are similar.

\subsection{Clinical output}

\RC Methodologically, the paper is interesting and technically sound. It is clear and easy to follow. I do appreciate the direct comparison with the literature, and the sample size collected by the authors. On the other hand, as a Bone reader, I have found unusual reading this kind of engineering/mathematical paper here, and I suppose it is far from the typical audience of the journal. I understand there is a final clinical message, but the way it is reached could be misleading. I would have expected to read this kind of paper in a more engineering journal. I suggest to increase and clarify the possible clinical output of the study in order to increase the impact.

\AR Clarifications in text

\subsection{OI type differentiation}

\RC Another concern consists in the grouping of the bones with OI. Indeed, while the OI group was composed of patients with a diagnosis of OI Type I, III, and IV, the tibiae were merged in a single group (OI group), without any differentiation during the following steps of the study. In the literature and in the studies cited by the authors, the differences among bones with different types of OI resulted remarkable. Thus, I would like to find more details about the OI types in the linear regressions (differentiating the regressions for each type of OI type). Moreover, it would be interesting to see which is the OI type more influenced by the filtering. My idea is that bones with OI type I are "similar" to healthy group, they can easily respect the homogeneity condition, therefore they were not excluded by the filtering strategies. By contrast, the type IV may have a microstructure different from the healthy bones, and they can explain the lack of regression in the OI group. If this is the case, the final message should be mitigated.

\AR Change figures in linear regressions to highlight the similarities among the OI types

\subsection{Introduction}
Introduction: please, add statistics also about the epidemiology of OI in men and women. Moreover, add the lacks in the literature about the OI that this study could solve.

\subsection{Methods - Participants}
Could you add statistics about body anthropometry? I was wondering, why the groups were matched considering age and not height and weight? Could you explain your choice?

\subsection{Methods - HR-pQCT}
The authors used two different protocols to identify the region of interest in the tibiae. Considering that the same protocol could be applied on both groups (control and OI) I do not understand why two different protocols were used. There is a difficult in defining the dense structure in OI specimens? To clarify the manuscript, I suggest to label the most proximal stack in the control tibia as Stack n. 1

\subsection{Methods - Image analysis}
Pag 3, lines 34-35, left column. The stack was divided into two halves WITH A TRANSVERSE PLANE. Pag 3, lines 4-7, right column. Did you use any filter to reduce the noise? If yes, please specify the procedure. Pag 4, line 4. Does your concern about the rotation consist in the errors associated with interpolation?

\subsection{Methods - Linear Regression}
I would specify here that you are considering the entire sample.

\subsection{Methods - Data filtering}
You removed the regions that violated the assumption of homogeneity. Could this affect the final findings of the study? I mean, could this filtering remove the specimens with OI type IV?

\subsection{Resutlts - Morphological analysis}
Please, report here the distribution of OI types in the cases matched with controls.

\subsection{Table 2}
Make easier and more effective the comparison with the literature, indicating below the table the OI types analyzed in the other studies. The positioning on different lines, in the same cell, could not be caught.

\subsection{Results - Comparison with literature}
Could you perform a test to evaluate if differences exist among the different regressions?

\subsection{Discussion}
The discussion is interesting and full of details and rigorous considerations. However, it is based only on the engineering/mathematical approach used in the study, and far from the clinical audience. I would like to read more about the effects of the different OI types in the provided findings.

\end{document}

### 2021-10-01 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex
Response to reviewers draft beginning

@@
 	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
+\end{quote}
+
+Methods - HR-pQCT, Line 28-29
+\begin{quote}
+	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Brütisellen, Switzerland) was performed at the distal tibia on all patients from both groups.
+\Add{The distal tibia was chosen because numerous studies were performed on the distal radius [2, 5, 12, 14, 16, 31], some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) [1, 4, 9] but only one study with XtremeCTII was performed on the distal tibia [27].
+Good bone strength assessment of tibia is crucial for individuals in wheelchair.
+As their tibiae do not sustain the body weight anymore, remodeling takes place and low energy fracture can occur.}
 \end{quote}

### 2021-10-01 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex
Following of article revision

@@
 
-Abstract, Line 23-24
+Abstract, Lines 23-24
 \begin{quote}
@@
 
-Abstract, Line 48-49
+Abstract, Lines 48-49
 \begin{quote}
@@
 It is the case for another previous work (Zysset, ref 24) where an analysis of the influence of boundary conditions type was performed.
-Again, stiffness tensor was computed from numerical simulations of trabecular bone from a \si{\micro}CT scans on the proximal femur.
-The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia 2) using HR-pQCT scans i.e.
-lower resolution scans 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals.
+Again, stiffness tensor was computed from numerical simulations of trabecular bone from \si{\micro}CT scans on the proximal femur.
+The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia, 2) using HR-pQCT scans i.e.
+lower resolution scans than the two references cited here, 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals.
 More emphasis on these points is put with the following modifications of the manuscript:\par
 
-Introduction, Line 44-45
+Introduction, Lines 44-45
 \begin{quote}
@@
 
-Methods - HR-pQCT, Line 28-29
+Methods - HR-pQCT, Lines 28-29
 \begin{quote}
@@
 \Add{The distal tibia was chosen because numerous studies were performed on the distal radius [2, 5, 12, 14, 16, 31], some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) [1, 4, 9] but only one study with XtremeCTII was performed on the distal tibia [27].
-Good bone strength assessment of tibia is crucial for individuals in wheelchair.
+Accurate bone strength assessment of tibia is crucial for individuals in wheelchair.
 As their tibiae do not sustain the body weight anymore, remodeling takes place and low energy fracture can occur.}
+\end{quote}
+
+Results - Comparison with literature, Lines 61-62
+\begin{quote}
+	Table 4 shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femura studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m [12, 24]}.
+\end{quote}
+
+Discussion, Page 10, Lines 55-56, Left column
+\begin{quote}
+	To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
+\Add{These high numbers of participants allow to obtain strong results to compare then with literature.}
+\end{quote}
+
+Discussion, Page 10, Lines 38-40, Right column
+\begin{quote}
+	With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
+\Add{Thus, the same hFE analysis can be applied for bone strength assessment of OI diagnosed individuals or of healthy ones.
+Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity\Del{.
+Nevertheless}, the error created by such ROIs is negligible as this  concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.
+\end{quote}
+
+Discussion, Page 11, Lines 39-43, Right column
+
+\begin{quote}
+	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
+with a CV lower than 0.263 and were matched for BV/TV and DA.\Add{Therefore, standard hFE analysis used for bone strength estimation of healthy individuals can be used for OI diagnosed individuals as well.}
 \end{quote}

### 2021-10-03 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex
Response to reviewers draft following

@@
 \usepackage{siunitx}
+\usepackage{subcaption}
+\usepackage{hyperref}
+
+\hypersetup{hidelinks}
 
@@
 
-\AR Explain filtering of low stiffness means filtering of low BV/TV...
+\AR Effectively, the key contribution, the fact that fabric-elasticity relationships in OI trabecular bone are similar to healthy bone, is based on filtered data.
+Nevertheless, the filtering was performed according to a coefficient of variation (CV) threshold, i.e.
+a homogeneity threshold.
+Figure 6, see Figure \ref{CV_BVTV}, shows CV in relation to BV/TV and it can be seen from this Figure that ROIs presenting a high heterogeneity have a lower BV/TV.
+Moreover, this tendency is confirmed by the Pearson's correlation coefficient being strictly different from zero, meaning there is a monotonic relation between CV and BV/TV.
+The fact that we want to highlight here is that a high CV means a low BV/TV and therefore, a low stiffness since it is mainly driven by BV/TV.
+This is confirmed by comparing Figure 5 and 7, see Figure \ref{CV_Filtering}.
+Both Figures have the same logarithmic scales and the filtering removes mainly the ROIs with a stiffness orders of magnitudes lower than the stiffness of homogeneous ROIs.
+Thus, when hFE is performed for bone strength assessment, ROIs that are mostly empty (low BV/TV) do not participate significantly (orders of magnitude lower) to the whole structure strength meaning that the error performed on such ROI stiffness can be neglected.
+
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}[t]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]
+		{Pictures/03_CV_BVTV}
+		\caption{CV in relation to BV/TV.}
+		\label{CV_BVTV}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.5\textwidth}
+		\centering
+		\includegraphics[width=\textwidth, trim= 250 140 250 0]
+		{Pictures/R1_CVFiltering}
+		\caption{Linear regression of the complete data set, healthy and OI pooled together. Shape differs according to CV value of the ROI and color according to the BV/TV.}
+		\label{CV_Filtering}
+	\end{subfigure}
+	\caption{Relations between coefficient of variation (CV) and BV/TV}
+	\label{01_Contribution}
+\end{figure}
 
@@
 
-\AR Michi articles
+\AR Yes, OI is mainly characterized by impaired synthesis of type I collagen which compose the bone ECM and could therefore affects its toughness.
+To investigate this phenomenon, a first study was performed by Indermaur and al.
+(ref 15).
+They shown that the compressive strength of bone at the ECM level is related to degree of mineralization and that it is not reduced in OI compared to healthy individuals.
+A second study (not published yet) was performed to compare OI and healthy bone at the ECM level in tension.
+Their results shown no significant differences between the two phenotypes (i.e.
+healthy and OI).
+This, together with results of the present article, allow to conclude that fabric-strength relationships are similar between healthy and OI bone.
+This means that the main cause of bone fragility in OI is linked to the low bone mass.
 
@@
 
-\AR Clarifications in text
+\AR Thank you for this remark.
+The main reason for the technical details was the clarity of the methods used to compare OI with healthy trabecular bone.
+On the other hand, Bone include articles with technical components as well.
+To clarify the clinical significance of this study, the manuscript is modified as follow:
 

### 2021-10-05 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex
Response to reviewers draft following II

@@
 \usepackage{hyperref}
+\usepackage{array}
+\usepackage{multicol}
+\usepackage{rotating}
+\usepackage{multirow}
+\usepackage{float}
+\usepackage[justification=centering]{caption}
+\usepackage{tabularx}
+\usepackage{booktabs}
+
+\renewcommand{\arraystretch}{1.2}
 
@@
 
-\AR To explain new progress shown in the present article, the manuscript was modified at multiple places as follow:\par
+\AR To explain new progress shown in the present article, the manuscript abstract was modified as follow:\par
 
@@
 
-\subsection{Methods}
+Then, the manuscript is modified at different places to explain the new progress and put more emphasis on the clinical outcome, see following responses.
+
+\newpage
+\subsection{Article contribution}
 
@@
 
-Introduction, Lines 44-45
+Introduction, Page 2, Lines 44-45
 \begin{quote}
@@
 	To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
-\Add{These high numbers of participants allow to obtain strong results to compare then with literature.}
+\Add{These high numbers of participants allow to obtain strong results to compare with literature.}
 \end{quote}
 
+\newpage
 Discussion, Page 10, Lines 38-40, Right column
@@
 Discussion, Page 11, Lines 39-43, Right column
-
 \begin{quote}
@@
 
-\subsection{Article contribution}
+\subsection{Study conclusions}
 
@@
 
+\newpage
 \subsection{OI brittelness}
@@
 
+\newpage
 \section{Reviewer \#2}
@@
 
+Introduction, Page 2, Lines 28-31
+\begin{quote}
+	Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength [5].
+\Add{HR-pQCT-based FEA becomes more widely used in clinical studies [ref] and is currently approved by the American Food and Drug Administration (FDA).}
+\end{quote}
+
+Reference:\\
+Bergh, J.P.V.D., Szulc, P., Cheung, .A.M., Bouxsein, .M., Engelke, .K., Chapurlat, .R., Den, J.P.V., Nl, B.J., 2021.
+The clinical application of high-resolution peripheral computed tomography (HR-pQCT) in adults: state of the art and future directions.
+Osteoporosis Int 32, 1465–1485.
+URL:https://doi.org/10.1007/s00198-021-05999-z, doi:10.1007/s00198-021-05999-z/Published.\par
+
+As for Reviewer \#1, Comment 1.1\par
+
+Introduction, Page2, Lines 44-45
+\begin{quote}
+	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
+\end{quote}
+
+Discussion, Page 10, Lines 38-40, Right column
+\begin{quote}
+	With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
+\Add{Thus, the same hFE analysis can be applied for bone strength assessment of OI diagnosed individuals or of healthy ones.
+Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity\Del{.
+Nevertheless}, the error created by such ROIs is negligible as this  concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.
+\end{quote}
+
+Discussion, Page 11, Lines 39-43, Right column
+\begin{quote}
+	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
+with a CV lower than 0.263 and were matched for BV/TV and DA.\Add{Therefore, standard hFE analysis used for bone strength estimation of healthy individuals can be used for OI diagnosed individuals as well.}
+\end{quote}
+
 \subsection{OI type differentiation}
@@
 
-\AR Change figures in linear regressions to highlight the similarities among the OI types
+\AR Effectively, in the different studies compared to the present one for trabecular morphometry, some of them presented remarkable differences among the OI types.
+However, the objective here is to compare the effect of OI phenotype to healthy trabecular fabric-elasticity relationships.
+Nevertheless, Figure \ref{02_OITypes} shows the positions of the different types of OI in the 2 most explanatory plots i.e.
+the coefficient of variation (CV) in relation to BV/TV (Figure \ref{CV_BVTV2}) and the linear regression with the complete OI data set (Figure \ref{LinearRegression}).
+In both plots, ROIs of OI type III are in the average values of the OI group.
+After filtering, the relative number of OI type III removed is slightly higher than for the other types (about 80\% of OI type III compared to 60\% of OI type I and IV).
+As their are only 6 ROIs taken from the same OI type III individual, it is difficult to draw conclusion about it.
+However, the relations observed for healthy individuals are similar than the ones observed with the OI type I and IV.
+Thus, as for Reviewer \#1 comment 1.2, when the ROI BV/TV increases, stiffness increases as well and CV tends to decrease.
+Finally, ROIs with a sufficiently low CV have enough bone mass and fabric-elasticity relationships holds, both in healthy and OI trabecular bone.
 
+\begin{figure}[h!]
+	\centering
+	\begin{subfigure}[t]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]
+		{Pictures/R2_CV_BVTV}
+		\caption{CV in relation to BV/TV.}
+		\label{CV_BVTV2}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.475\textwidth}
+		\centering
+		\includegraphics[width=\textwidth, trim= 0 210 0 0]
+		{Pictures/R2_LinearRegression}
+		\caption{Linear regression of the complete OI data set.}
+		\label{LinearRegression}
+	\end{subfigure}
+	\caption{Relative positions of the different OI types for different analyses.}
+	\label{02_OITypes}
+\end{figure}
+
+\newpage
 \subsection{Introduction}
-Introduction: please, add statistics also about the epidemiology of OI in men and women. Moreover, add the lacks in the literature about the OI that this study could solve.
+\RC Introduction: please, add statistics also about the epidemiology of OI in men and women. Moreover, add the lacks in the literature about the OI that this study could solve.
+
+\AR For a bit of more completeness the following sentences were added in the manuscript:\par
+
+Introduction, Page 2, Lines 6-11
+\begin{quote}
+	OI can be categorized according to disease severity [23], where most familiar forms are:
+	\begin{itemize}
+		\item Type I: mild
+		\item Type II: perinatally lethal
+		\item Type III: most severe surviving form
+		\item Type IV: intermediate severity
+	\end{itemize}
+
+	\Add{OI occurs worldwide and without sex preference.
+OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% [ref1]}.
+Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels.
+Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals [9, 19, 28] \Add{but aBMD is limited for bone strength estimation [ref2]}.
+\end{quote}
+
+References\vspace{-1em}
+\begin{enumerate}
+	\item Martin, E., Shapiro, J.R., 2007. Osteogenesis imperfecta: Epidemiology and pathophysiology. Curr Osteoporos Rep 5, 91–97. URL: http://www.le.ac.uk/genetics/collagen/.
+	\item E, P., J, S.M., N, H.E., C, M.G., A, N.M., A, K.J., 2006.
+The population burden of fractures originates in women with osteopenia, not osteoporosis.
+Osteoporos Int 17, 1404–1409.
+doi:10.1007/s00198-006-0135-9.
+\end{enumerate}
+
+As for Reviewer \#1, Comment 1.1 and Comment 2.1\par
+
+Introduction, Page2, Lines 44-45
+\begin{quote}
+	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
+\end{quote}
 
 \subsection{Methods - Participants}
-Could you add statistics about body anthropometry? I was wondering, why the groups were matched considering age and not height and weight? Could you explain your choice?
+\RC Could you add statistics about body anthropometry? I was wondering, why the groups were matched considering age and not height and weight? Could you explain your choice?
+
+\AR Sorry but the anthropometric data are not available.
+Regarding the matching for trabecular architecture comparison, it was performed according to age and sex because their are the parameters influencing the most trabecular bone, namely the bone volume fraction (BV/TV).
+Matching the data sets according to these two covariables lead to reduce the bias between the two groups (healthy and OI).
 
 \subsection{Methods - HR-pQCT}
-The authors used two different protocols to identify the region of interest in the tibiae.
+\RC The authors used two different protocols to identify the region of interest in the tibiae.
 Considering that the same protocol could be applied on both groups (control and OI) I do not understand why two different protocols were used.
@@
 
+\AR The reason for the use of different protocols relies in the fact that the data are obtained from two different studies.
+Data of the healthy group provide from a study of reproducibility where an in-house protocol was applied for the stacks location.
+This protocol was designed such as it includes the standard clinical region.
+OI data provide from another study investigating treatment effect.
+For this study, it was decided to apply the standard scanning protocol.
+The possibility to analyze a similar anatomical region for the first time using HR-pQCT II was exploited as the data were available.
+As suggested, label of the most proximal stack in the control group is changed to n.1 to improve clarity.
+
 \subsection{Methods - Image analysis}
-Pag 3, lines 34-35, left column.
+\RC Page 3, lines 34-35, left column.
 The stack was divided into two halves WITH A TRANSVERSE PLANE.
@@
 
+\AR Yes a filtering is used in the standard image processing. The filtering procedure is added to the manuscript as shown below.\par
+
+Methods - Image analysis, Page 2, Line 62, Right column
+\begin{quote}
+	Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking)\Add{.
+Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed} and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}).
+\end{quote}
+
+Yes again, by keeping the ROI in its original configuration, errors due to interpolation are avoided.This is now precised in the manuscript.\par
+
+Methods - Image analysis, Page 4, Line 5-6, Left column
+\begin{quote}
+	Such rotation would potentially decrease image quality \Add{due to interpolation}.
+\end{quote}
+
 \subsection{Methods - Linear Regression}
-I would specify here that you are considering the entire sample.
+\RC I would specify here that you are considering the entire sample.
+
+\AR The manuscript is modified in consequence:\par
+
+Methods - Linear Regression, Page 4, Lines 47-48, Right column
+\begin{quote}
+	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model \Add{using the entire sample of both groups}.
+\end{quote}
 
 \subsection{Methods - Data filtering}
-You removed the regions that violated the assumption of homogeneity. Could this affect the final findings of the study? I mean, could this filtering remove the specimens with OI type IV?
+\RC You removed the regions that violated the assumption of homogeneity. Could this affect the final findings of the study? I mean, could this filtering remove the specimens with OI type IV?
 
-\subsection{Resutlts - Morphological analysis}
-Please, report here the distribution of OI types in the cases matched with controls.
+\AR The first part of this question is closely related to Comment 1.2 of Reviewer \# 1.
+we would suggest to refer to our answer to this comment.
+For the second part, we would suggest to look at the plots used to answer to your comment 2.2.
+The filtering do not remove completely a specific OI type.
+
+\subsection{Results - Morphological analysis}
+\RC Please, report here the distribution of OI types in the cases matched with controls.
+
+\AR To answer to your request, the manuscript is modified as follow:
+\begin{quote}
+	The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the matched healthy and OI individuals, respectively. \Add{The OI group matched with control was composed of nineteen type I and nine type IV.}
+\end{quote}
 
 \subsection{Table 2}
-Make easier and more effective the comparison with the literature, indicating below the table the OI types analyzed in the other studies.
+\RC Make easier and more effective the comparison with the literature, indicating below the table the OI types analyzed in the other studies.
 The positioning on different lines, in the same cell, could not be caught.
+
+\AR To increase the readability of the table, it was changed in the manuscript, see Table \ref{Table2}.
+
+\begin{sidewaystable*}
+	\centering
+	\caption{Summary of the tibia ROIs morphological analysis and comparison with literature.
+Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians.
+The study of Kocijan et al.
+[16] presents n.s.
+for non-significant p value test result.}
+	\label{Table2}
+	\resizebox{\textwidth}{!}{
+	\begin{tabular}{cccccccccc}
+		\toprule
+		\multirow{2}{*}{Variable} & \multirow{2}{*}{Group} & \multicolumn{2}{c}{Present study} & \multicolumn{2}{c}{Folkestad et al.
+[9]} & \multicolumn{2}{c}{Kocijan et al.
+[16]} & \multicolumn{2}{c}{Rolvien et al.
+[26]} \\
+		& & Values & p value & Values & p value & Values & p value & Values & p value \\
+		\midrule
+
+		\multirow{3}{*}{Age} & Healthy & 41 $\pm$ 14 &  & 54 (21-77) & & 44 (38-52) &  & 49 $\pm$ 16 &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{41 $\pm$ 15} &  & \multirow{2}{*}{53 (21-77)} &  & 42 (35-56)\textsuperscript{1} & & \multirow{2}{*}{46 $\pm$ 16} & \\
+		& & &  &  &  & 48 (35-58)\textsuperscript{2} & & & \\[3ex]
+
+		\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
+		& OI Type I & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114) & <0.0001 & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
+		&  OI Type III \& IV & & & & & 0.081 (0.056-0.092) & <0.0001 & & \\[3ex]
+
+		\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
+		& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
+		&  OI Type III \& IV & & & & & 0.89 (0.81-1.08) & <0.001 & & \\[3ex]
+
+		\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
+		& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
+		&  OI Type III \& IV & & & & & 0.078 (0.068-0.092) & n.s. & & \\[3ex]
+
+		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
+		& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
+		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+
+		\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
+		& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
+		&  OI Type III \& IV & & & & & 0.698 (0.511-0.890) & <0.001 & & \\[3ex]
+
+		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & - &  & - & & - &  \\
+		& OI Type I & \multirow{2}{*}{0.056 (0.015-0.079)} &  & - &  & - &  & - & \\
+		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+
+		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & - &  & - & & - &  \\
+		& OI Type I & \multirow{2}{*}{2.018 (1.901-2.158)} &  & - &  & - &  & - & \\
+		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+
+		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & - &  & - & & - &  \\
+		& OI Type I & \multirow{2}{*}{-1.178 $\pm$ 0.441} &  & - &  & - &  & - & \\
+		&  OI Type III \& IV & & & & & - &  & & \\
+
+		\bottomrule
+	\end{tabular}}
+\end{sidewaystable*}
 

### 2021-10-05 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Beginning of article revision
Response to reviewers draft following II

--- 2021-07-28 09_Article/Simon_2021_Bone_Draft_V03.tex
@@
 \usepackage{multicol}
+\usepackage[normalem]{ulem}
 
@@
 	\nonumnote{\textit{Abbreviations:\enspace}#1}}
+
+\newcommand{\Add}[1]{{\color{blue}{#1}}}
+
+\newcommand{\Del}[1]{{\color{red}\sout{#1}}}
 
@@
 
-Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels.
-Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018}.
+	\Add{OI occurs worldwide and without sex preference.
+OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% \cite{Martin2007}}.Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels.
+Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018} \Add{but aBMD is limited for bone strength estimation \cite{Pasco2006}}.
 \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well.
@@
 Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}.
-Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
+\Add{HR-pQCT-based FEA becomes more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
 High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}.
@@
 Each ROI had to contain trabecular bone, but no cortical bone.
-For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 3) to be at roughly the same anatomical location as for the OI group.
-Then for both the healthy group and the OI group, the stack was divided into two halves and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
+For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number \Del{3}\Add{1}) to be at roughly the same anatomical location as for the OI group.
+Then, for both the healthy group and the OI group, the stack was divided into two halves \Add{with a transverse plane} and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
 For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone.
@@
 \\
+
 	As the previous studies \cite{Folkestad2012,Kocijan2015,Rolvien2018} have the same age range as our matched groups, we can compare the morphological parameters.
@@
 \\
+
 	The linear regressions performed in this study on original data sets showed $R^2_{adj}$ and NE in the expected range (i.
@@
 \\
+
 	Figure \ref{02_CV_BVTV} presents the CV in relation to BV/TV.
@@
 \\
+
 	The linear regressions performed on the filtered data sets present a direct effect of filtering as the ROIs meeting the homogeneity assumption lead to better results compared to linear regression including ROIs with high CV values.
@@
 These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\
+
 	After BV/TV and DA matching, grouping the data sets together led to similar $R^2_{adj}$ and NE as for the individual filtered data sets.
@@
 Nevertheless, the error created by such ROIs is negligible as this concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.\\
+
 	Imposing $k$ and $l$ allows one to compare the other regressions parameters and estimate the effect of different image resolutions.
@@
 \\
+
 	The analysis of TMD gives interesting outputs as well.
@@
 \\
+
 	The main limitations of this study are the definition of "homogeneous enough" (i.e.

### 2021-10-05 03_Scripts/Response2Reviewer1.py

@@ -151,5 +151,3 @@ Full_LMM = smf.mixedlm("LogSxy ~ Sii + Sij + Sjj + LogBVTV + Logmxy - 1",
 # 05 Plot the results of the regression
 R2adj, NE = PlotRegressionResults(Full_LMM,FullSystem, Random=False)
 
-# 06 Set filter parameters
-CVLim = 0.263

### 2021-10-05 03_Scripts/Response2Reviewer2.py

@@ -1,0 +1,289 @@
+# 00 Initialization
+import os
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import statsmodels.formula.api as smf
+from scipy.stats.distributions import norm, t, f, chi2
+from scipy.stats import shapiro
+
+
+desired_width = 500
+pd.set_option('display.max_rows', 100)
+pd.set_option('display.max_columns', desired_width)
+pd.set_option('display.width', desired_width)
+np.set_printoptions(linewidth=desired_width,suppress=True,formatter={'float_kind':'{:3}'.format})
+plt.rc('font', size=12)
+
+
+## Define functions
+def PlotRegressionResults(Model, Data, Random=True, Colors=[(0,0,1),(0,1,0),(1,0,0)]):
+
+    ## Get data from the model
+    Y_Obs = np.exp(Model.model.endog)
+    Y_Fit = np.exp(Model.fittedvalues)
+    if not Random:
+        Y_Fit = np.exp(Model.predict())
+    N = int(Model.nobs)
+    C = np.matrix(Model.cov_params())
+    X = np.matrix(Model.model.exog)
+    NROIs = int(N/12)
+
+    if not C.shape[0] == X.shape[1]:
+        C = C[:-1,:-1]
+
+
+    ## Compute R2 and standard error of the estimate
+    E = np.log(Y_Obs) - np.log(Y_Fit)
+    RSS = np.sum(E ** 2)
+    SE = np.sqrt(RSS / Model.df_resid)
+    TSS = np.sum((Model.model.endog - Model.model.endog.mean()) ** 2)
+    RegSS = TSS - RSS
+    R2 = RegSS / TSS
+
+    ## Compute R2 adj and NE
+    R2adj = 1 - RSS/TSS * (12*N-1)/(12*N-X.shape[1]-1)
+
+    NE = np.array([])
+    for i in range(0,N,12):
+        ObservedTensor = Y_Obs[i:i+12]
+        PredictedTensor = Y_Fit[i:i+12]
+
+        Numerator = np.sum((ObservedTensor-PredictedTensor)**2)
+        Denominator = np.sum(ObservedTensor**2)
+
+        NE = np.append(NE,np.sqrt(Numerator/Denominator))
+
+
+    Line = np.linspace(min(Y_Obs.min(), Y_Fit.min()),
[…]

### 2021-10-06 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex
Response to reviewers draft following III

@@
 		\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
-		& OI Type I & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114) & <0.0001 & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
-		&  OI Type III \& IV & & & & & 0.081 (0.056-0.092) & <0.0001 & & \\[3ex]
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
+		& & & & & & 0.081 (0.056-0.092)\textsuperscript{2} & <0.0001\textsuperscript{2} & & \\[3ex]
 
 		\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
-		& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
-		&  OI Type III \& IV & & & & & 0.89 (0.81-1.08) & <0.001 & & \\[3ex]
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
+		& & & & & & 0.89 (0.81-1.08)\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]
 
 		\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
-		& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
-		&  OI Type III \& IV & & & & & 0.078 (0.068-0.092) & n.s. & & \\[3ex]
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090)\textsuperscript{1} & n.s.\textsuperscript{1} & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
+		& & & & & & 0.078 (0.068-0.092)\textsuperscript{2} & n.s.\textsuperscript{2} & & \\[3ex]
 
-		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
-		& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
-		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & \multirow{3}{*}{-} & & 0.409 $\pm$ 0.023 & 0.0003 \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
+		& & & & & & &  & & \\[3ex]
 
 		\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
-		& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
-		&  OI Type III \& IV & & & & & 0.698 (0.511-0.890) & <0.001 & & \\[3ex]
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504)\textsuperscript{1} & <0.0001\textsuperscript{1} & - & \\
+		& & & & & & 0.698 (0.511-0.890\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]
 
-		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & - &  & - & & - &  \\
-		& OI Type I & \multirow{2}{*}{0.056 (0.015-0.079)} &  & - &  & - &  & - & \\
-		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & \multirow{3}{*}{-} &  & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.056 (0.015-0.079)} & & & & &  & & \\
+		& & & & & & & & & \\[3ex]
 
-		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & - &  & - & & - &  \\
-		& OI Type I & \multirow{2}{*}{2.018 (1.901-2.158)} &  & - &  & - &  & - & \\
-		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{2.018 (1.901-2.158)} & & & & & & & \\
+		& & & & & & & & & \\[3ex]
 
-		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & - &  & - & & - &  \\
-		& OI Type I & \multirow{2}{*}{-1.178 $\pm$ 0.441} &  & - &  & - &  & - & \\
-		&  OI Type III \& IV & & & & & - &  & & \\
+		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{-1.178 $\pm$ 0.441} & & & & & & & \\
+		& & & & & & & & & \\
 
@@
 	\end{tabular}}
+	OI types included in the study: Present study, types I and IV; Folkestad et al. [9], type I; Kocijan et al. [16], 1) type I 2) type III and IV; Rolvien et al. [26], type I and IV.
 \end{sidewaystable*}

### 2021-10-06 09_Article/01_Revision1_Reviewers/Response to Reviewers_00.tex
Finish first draft of Response to Reviewers

--- 2021-10-05 09_Article/01_Revision1_Reviewers/Response to Reviewers.tex
@@
 	To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
-\Add{These high numbers of participants allow to obtain strong results to compare with literature.}
+\Add{These numbers of participants are higher than in the other studies in the literature comparing OI with healthy control [9,16,26].}
 \end{quote}
@@
 	With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-\Add{Thus, the same hFE analysis can be applied for bone strength assessment of OI diagnosed individuals or of healthy ones.
+\Add{Thus, the same hFE analysis can be applied for bone strength assessment of OI diagnosed individuals as for healthy ones.
 Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity\Del{.
@@
 Thus, as for Reviewer \#1 comment 1.2, when the ROI BV/TV increases, stiffness increases as well and CV tends to decrease.
-Finally, ROIs with a sufficiently low CV have enough bone mass and fabric-elasticity relationships holds, both in healthy and OI trabecular bone.
+Finally, ROIs with a sufficiently low CV have enough bone mass and fabric-elasticity relationships holds, both in healthy and OI, whatever the type, trabecular bone.
 
@@
 
-\AR Sorry but the anthropometric data are not available.
-Regarding the matching for trabecular architecture comparison, it was performed according to age and sex because their are the parameters influencing the most trabecular bone, namely the bone volume fraction (BV/TV).
+\AR We did not add other the anthropometric data because we thought they were not relevant for the present study and were not available for the control group.
+The idea of age and sex matching was to analyze similar healthy and OI population, as Folkestad et al.
+(ref 9), Kocijan et al.
+(ref 16) and Rolvien et al.
+(ref 26) did.
 Matching the data sets according to these two covariables lead to reduce the bias between the two groups (healthy and OI).
+The aim of this initial analysis was to confirm that our groups are as different from each other as what can be found in the literature and results obtained confirmed this hypothesis.
 
@@
 For this study, it was decided to apply the standard scanning protocol.
-The possibility to analyze a similar anatomical region for the first time using HR-pQCT II was exploited as the data were available.
+The possibility to analyze a similar anatomical region for the first time using tibiae HR-pQCT II scans was exploited as the data were available.
 As suggested, label of the most proximal stack in the control group is changed to n.1 to improve clarity.
@@
 
-\AR Yes a filtering is used in the standard image processing. The filtering procedure is added to the manuscript as shown below.\par
+\AR The transverse plane precision is added to the manuscript. Yes a filtering is used in the standard image processing. The filtering procedure is added to the manuscript as shown below.\par
 
@@
 
-		\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
-		& OI Type I & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114) & <0.0001 & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
-		&  OI Type III \& IV & & & & & 0.081 (0.056-0.092) & <0.0001 & & \\[3ex]
+		\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & & 0.14 $\pm$ 0.03 & & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.164 $\pm$ 0.079} & \multirow{2}{*}{<0.01} & \multirow{2}{*}{0.08 $\pm$ 0.03} & \multirow{2}{*}{<0.001} & 0.098 (0.088-0.114)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{0.095 $\pm$ 0.008} & \multirow{2}{*}{<0.0001} \\
+		& & & & & & 0.081 (0.056-0.092)\textsuperscript{2} & <0.0001\textsuperscript{2} & & \\[3ex]
 
-		\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
-		& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
-		&  OI Type III \& IV & & & & & 0.89 (0.81-1.08) & <0.001 & & \\[3ex]
+		\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & & 1.94 (1.74-2.13) & & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.650 $\pm$ 0.198} & \multirow{2}{*}{<0.001} & \multirow{2}{*}{1.30 (0.75-1.53)} & \multirow{2}{*}{<0.001} & 1.33 (1.07-1.55)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{1.428 $\pm$ 0.098} & \multirow{2}{*}{<0.0001} \\
+		& & & & & & 0.89 (0.81-1.08)\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]
 
-		\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
-		& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
-		&  OI Type III \& IV & & & & & 0.078 (0.068-0.092) & n.s. & & \\[3ex]
+		\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & & 0.07 (0.06-0.08) & & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.306 (0.292-0.331)} & \multirow{2}{*}{0.2} & \multirow{2}{*}{0.07 (0.06-0.08)} & \multirow{2}{*}{0.5} & 0.074 (0.064-0.090)\textsuperscript{1} & n.s.\textsuperscript{1} & \multirow{2}{*}{0.066 $\pm$ 0.004} & \multirow{2}{*}{0.046}\\
+		& & & & & & 0.078 (0.068-0.092)\textsuperscript{2} & n.s.\textsuperscript{2} & & \\[3ex]
 
-		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
-		& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
-		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+		\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & & 0.44 (0.39-0.51) & & \multirow{3}{*}{-} & & 0.409 $\pm$ 0.023 & \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{1.422 $\pm$ 0.694} & \multirow{2}{*}{0.01} & \multirow{2}{*}{0.68 (0.57-1.19)} & \multirow{2}{*}{<0.001} & &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \multirow{2}{*}{0.0003}\\
+		& & & & & & & & & \\[3ex]
 
-		\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
-		& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
-		&  OI Type III \& IV & & & & & 0.698 (0.511-0.890) & <0.001 & & \\[3ex]
+		\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & & 0.20 (0.16-0.25) & & 0.221 (0.170-0.242) & & - &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.631 $\pm$ 0.383} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.40 (0.31-1.11)} & \multirow{2}{*}{<0.001} & 0.382 (0.311-0.504)\textsuperscript{1} & <0.0001\textsuperscript{1} & - & \\
+		& & & & & & 0.698 (0.511-0.890\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]
 
-		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & - &  & - & & - &  \\
-		& OI Type I & \multirow{2}{*}{0.056 (0.015-0.079)} &  & - &  & - &  & - & \\
-		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+		\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & & \multirow{3}{*}{-} &  & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{0.056 (0.015-0.079)} & \multirow{2}{*}{<0.001} & & & &  & & \\
+		& & & & & & & & & \\[3ex]
 
-		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & - &  & - & & - &  \\
-		& OI Type I & \multirow{2}{*}{2.018 (1.901-2.158)} &  & - &  & - &  & - & \\
-		&  OI Type III \& IV & & & & & - &  & & \\[3ex]
+		\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{2.018 (1.901-2.158)} & \multirow{2}{*}{0.02} & & & & & & \\
+		& & & & & & & & & \\[3ex]
 
-		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & - &  & - & & - &  \\
-		& OI Type I & \multirow{2}{*}{-1.178 $\pm$ 0.441} &  & - &  & - &  & - & \\
-		&  OI Type III \& IV & & & & & - &  & & \\
+		\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+		& \multirow{2}{*}{OI} & \multirow{2}{*}{-1.178 $\pm$ 0.441} & \multirow{2}{*}{<0.0001} & & & & & & \\
+		& & & & & & & & & \\
 
@@
 	\end{tabular}}
+	OI types included in the study: Present study, types I and IV; Folkestad et al. [9], type I; Kocijan et al. [16], 1) type I 2) type III and IV; Rolvien et al. [26], type I and IV.
 \end{sidewaystable*}
 
+\newpage
 \subsection{Results - Comparison with literature}
-Could you perform a test to evaluate if differences exist among the different regressions?
+\RC Could you perform a test to evaluate if differences exist among the different regressions?
+
+\AR As the full data sets of the other studies are not available, it is not possible to perform a statistical test to see if the stiffness constants i.e.
+the slopes are different.
+Nevertheless, it is possible to perform statistical tests on BV/TV, DA and NE (for filtered data sets), assuming they were normally distributed.
+These tests will only provide insights about differences in the explanatory variables and the fit quality but not about differences in the slopes (stiffness constants).
+As said in the discussion, these differences provide from the different scan resolutions between the two studies.
 
 \subsection{Discussion}
-The discussion is interesting and full of details and rigorous considerations.
+\RC The discussion is interesting and full of details and rigorous considerations.
 However, it is based only on the engineering/mathematical approach used in the study, and far from the clinical audience.
@@
 
+\AR Thank you for your remark.
+The main idea of this study was to determine if standard hFE analyses could be performed on OI individuals.
+As shown in Figure \ref{02_OITypes}, there is no strict differences between OI type I and IV.
+Moreover, as filtering leaved only one ROI of OI type III, it is not possible to perform statistics about it.
+So the main conclusion of the present article is that fabric-elasticity relationships of trabecular bone are not affected by the OI phenotype, either it is type I, III or IV.
+This means that standard hFE analyses can be performed by clinicians even if the patient is OI diagnosed, whatever the type, because it will not have a significant difference on the bone strength assessment.
+
 \end{document}

### 2021-10-13 09_Article/01_Revision1_Reviewers/Response to Reviewers_01.tex
Finish Response to Reviewers and manuscript revision

--- 2021-10-06 09_Article/01_Revision1_Reviewers/Response to Reviewers_00.tex
@@
 
-\vspace{1em}The authors would like to thanks the reviewers for their time spent for this article review.
-The work they performed will allow to improve the quality of the present article and help the authors to transmit the knowledge resulting of their work with more clarity.
-For this, we answered at your comments as presented in this document.
+\vspace{1em}The authors would like to thank the reviewers for their time spent for this article.
+The comments helped us to improve the manuscript and to express the relevance and outcome of this work with more clarity.
+Answers are provided point by point in this document.
+In the manuscript, the added elements are written in blue and the canceled text is in red.
 
@@
 
-\AR To explain new progress shown in the present article, the manuscript abstract was modified as follow:\par
+\AR The main novelty resides in the application of our microFE homogenisation methodology to trabecular bone from OI patients.
+We hypothesised that the architectural quality of trabecular bone degrades in OI compared to healthy controls, i.e.
+lower apparent mechanical properties emerge for the same bone volume fraction and fabric.
+This was never investigated before and is relevant for the application of homogenised finite element analysis (hFEA) to bones from OI patients.
+In most standard FEA, the apparent mechanical properties of bone are derived from healthy tissue and would therefore overestimate bone stiffness or strength if these properties would be inferior in OI patients.
+Accordingly, the results of our study fill an important gap in current knowledge and provides a first important validation step towards the use of homogenised FEA in clinical studies involving OI patients.
+To explain this, the manuscript abstract was modified as follows:\par
 
@@
 \begin{quote}
-	Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls \Add{because validity of these relationships remains unclear for OI bone and are crucial for bone strength assessment using hFE}.
+	\Add{However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.}
+	Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
 \end{quote}
@@
 \begin{quote}
-	 Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections.
-\Add{Thus, the hFE scheme used for bone strength assessment of healthy individuals is valid for OI individuals as well.}
+	 Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
+\Del{Since highly heterogenous ROIs coincide with low BV/TV and can be handled smoothly by the fabric-elasticity relationships we expect them to play a minor role in hFE analysis of distal bone sections.} \Add{Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.}
 \end{quote}
@@
 
-\newpage
 \subsection{Article contribution}
@@
 
-\AR You have raised an important point here.
-The methodology used in the present paper is based on previous publications for trabecular bone analysis.
-With time, these different works led to the development of the hFE scheme for bone strength estimation.
-However, the work cited here (Zysset, ref 12), is based on \si{\micro}CT scan and was performed on human vertebrae, proximal femur and distal radius.
-It is the case for another previous work (Zysset, ref 24) where an analysis of the influence of boundary conditions type was performed.
-Again, stiffness tensor was computed from numerical simulations of trabecular bone from \si{\micro}CT scans on the proximal femur.
-The present study extend the current knowledge by 1) applying this hFE methodology at the distal tibia, 2) using HR-pQCT scans i.e.
-lower resolution scans than the two references cited here, 3) with an important control cohort of 120 healthy individuals and 4) demonstrating the applicability of the hFE scheme on OI diagnosed individuals.
-More emphasis on these points is put with the following modifications of the manuscript:\par
+\AR This article does not propose a novel methodology, but applies the developed technique to coarser HR-pQCT reconstructions of the distal tibia in order to clarify whether the fabric-elasticity relationships derived for healthy bone are also valid for OI bone.
+The following contributions can be listed:
+1) application to the distal tibia with comparison of densitometric and micro-architectural parameters, 2) use of HR-pQCT scans i.e.
+lower resolution scans than the two cited references, 3) an important control cohort of 120 healthy individuals, 4) lack of degradation in trabecular architecture in OI samples compared to controls and 5) validation of homogenised linear FEA to OI diagnosed individuals.
+The following modifications were undertaken: \par
 
@@
 \begin{quote}
-	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
+	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{ Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.}
 \end{quote}
@@
 	HR-pQCT imaging (XtremeCTII, SCANCO Medical, Brütisellen, Switzerland) was performed at the distal tibia on all patients from both groups.
-\Add{The distal tibia was chosen because numerous studies were performed on the distal radius [2, 5, 12, 14, 16, 31], some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) [1, 4, 9] but only one study with XtremeCTII was performed on the distal tibia [27].
-Accurate bone strength assessment of tibia is crucial for individuals in wheelchair.
-As their tibiae do not sustain the body weight anymore, remodeling takes place and low energy fracture can occur.}
+\Add{The distal tibia was chosen because numerous studies were performed on the distal radius [2, 5, 12, 14, 16, 31], some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) [4, 9] but only one study was performed on the distal tibia with XtremeCTII [27].}
 \end{quote}
@@
 \begin{quote}
-	Table 4 shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femura studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m [12, 24]}.
+	Table 4 shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femora studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m [12, 24]}.
 \end{quote}
@@
 	To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
-\Add{These numbers of participants are higher than in the other studies in the literature comparing OI with healthy control [9,16,26].}
+\Add{These numbers of participants are higher than in other studies comparing OI with healthy controls [9,16,26].}
 \end{quote}
 
-\newpage
 Discussion, Page 10, Lines 38-40, Right column
@@
 	With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-\Add{Thus, the same hFE analysis can be applied for bone strength assessment of OI diagnosed individuals as for healthy ones.
-Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity\Del{.
-Nevertheless}, the error created by such ROIs is negligible as this  concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.
+\Add{Thus, the same hFE analysis can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones.
+Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity \Del{.
+Nevertheless}, \Add{the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.}
 \end{quote}
@@
 \begin{quote}
-	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
-with a CV lower than 0.263 and were matched for BV/TV and DA.\Add{Therefore, standard hFE analysis used for bone strength estimation of healthy individuals can be used for OI diagnosed individuals as well.}
+	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough \Add{to trust the computational homogenisation methodology} (i.e.
+CV lower than 0.263), and were matched for BV/TV and DA.
+\Add{Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients as well.}
 \end{quote}
@@
 
-\AR Effectively, the key contribution, the fact that fabric-elasticity relationships in OI trabecular bone are similar to healthy bone, is based on filtered data.
-Nevertheless, the filtering was performed according to a coefficient of variation (CV) threshold, i.e.
+\AR Indeed, the demonstration that fabric-elasticity relationships in OI trabecular bone are similar to healthy bone, is based on filtered data.
+Nevertheless, the filtering was performed according to an objective coefficient of variation (CV) threshold, i.e.
 a homogeneity threshold.
-Figure 6, see Figure \ref{CV_BVTV}, shows CV in relation to BV/TV and it can be seen from this Figure that ROIs presenting a high heterogeneity have a lower BV/TV.
-Moreover, this tendency is confirmed by the Pearson's correlation coefficient being strictly different from zero, meaning there is a monotonic relation between CV and BV/TV.
-The fact that we want to highlight here is that a high CV means a low BV/TV and therefore, a low stiffness since it is mainly driven by BV/TV.
-This is confirmed by comparing Figure 5 and 7, see Figure \ref{CV_Filtering}.
-Both Figures have the same logarithmic scales and the filtering removes mainly the ROIs with a stiffness orders of magnitudes lower than the stiffness of homogeneous ROIs.
-Thus, when hFE is performed for bone strength assessment, ROIs that are mostly empty (low BV/TV) do not participate significantly (orders of magnitude lower) to the whole structure strength meaning that the error performed on such ROI stiffness can be neglected.
+The concept of fabric-elasticity requires minimal homogeneity of the ROI to have a biomechanical meaning and typically breaks down when only part of the ROI contains bone.
+In fact, the excluded data concern both healthy and OI ROIs and are not valid to make any comparisons.
+On the other hand, the empirical fabric-elasticity relationships are smooth with respect to volume fraction and fabric and were not found to induce deviations with respect to experimental results in our previous validation studies on the radius (ref 16) and the tibia (in preparation).
+This is likely due to the fact that heterogeneous ROIs have low (to extremely low) volume fractions.
+Theses ROIs not contribute significantly to the mechanical integrity of these bone sections.
 
@@
 \newpage
-\subsection{OI brittelness}
+\subsection{OI brittleness}
 
@@
 
-\AR Yes, OI is mainly characterized by impaired synthesis of type I collagen which compose the bone ECM and could therefore affects its toughness.
-To investigate this phenomenon, a first study was performed by Indermaur and al.
-(ref 15).
-They shown that the compressive strength of bone at the ECM level is related to degree of mineralization and that it is not reduced in OI compared to healthy individuals.
-A second study (not published yet) was performed to compare OI and healthy bone at the ECM level in tension.
-Their results shown no significant differences between the two phenotypes (i.e.
-healthy and OI).
-This, together with results of the present article, allow to conclude that fabric-strength relationships are similar between healthy and OI bone.
-This means that the main cause of bone fragility in OI is linked to the low bone mass.
+\AR In fact, our group just challenged this belief as OI bone matrix was found to be stiffer, stronger and no less brittle than healthy bone in biopsies from the iliac crest in both uniaxial compression (Indermaur et al., JBMR 2021) and tension (Indermaur et al., ASBMR 2021).
+Nevertheless, we agree with the reviewer that the present study does not formally prove that fabric-strength relationships are identical in OI and healthy trabecular bone.
+We did not run non-linear homogenisation analyses, but our previous experience allows to speculate that if the bone matrix properties are not degraded and the fabric-elasticity relationships are preserved, then the fabric-strength relationships are very likely to be preserved as well.
 
+%
+%
+%
+%
+%
+%
 \newpage
@@
 \AR Thank you for this remark.
-The main reason for the technical details was the clarity of the methods used to compare OI with healthy trabecular bone.
-On the other hand, Bone include articles with technical components as well.
+The main reason for the technical details was the clarity of the approach used to compare OI with healthy trabecular bone.
+On the other hand, Bone includes articles with advanced technical components as well.
 To clarify the clinical significance of this study, the manuscript is modified as follow:
+
+Abstract, Lines 23-24
+\begin{quote}
+	\Add{However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.}
+	Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
+\end{quote}
+
+Abstract, Lines 48-49
+\begin{quote}
+	 Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
+\Del{Since highly heterogenous ROIs coincide with low BV/TV and can be handled smoothly by the fabric-elasticity relationships we expect them to play a minor role in hFE analysis of distal bone sections.} \Add{Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.}
+\end{quote}
 
@@
 \begin{quote}
-	Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength [5].
-\Add{HR-pQCT-based FEA becomes more widely used in clinical studies [ref] and is currently approved by the American Food and Drug Administration (FDA).}
+	Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength [5] \Add{used for fracture risk estimation}.
+\Add{Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis}.
+Homogenized finite element (hFE) is based on \Add{local} BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort [23].
+High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius [31, 14, 2].
+\Add{Homogenised QCT-based FEA has become more widely used in clinical studies [ref] and is currently approved by the American Food and Drug Administration (FDA).}
 \end{quote}
@@
 \begin{quote}
-	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
-with a CV lower than 0.263 and were matched for BV/TV and DA.\Add{Therefore, standard hFE analysis used for bone strength estimation of healthy individuals can be used for OI diagnosed individuals as well.}
+	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough \Add{to trust the computational homogenisation methodology} (i.e.
+CV lower than 0.263), and were matched for BV/TV and DA.
+\Add{Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients as well.}
 \end{quote}
 
+\newpage
 \subsection{OI type differentiation}
@@
 After filtering, the relative number of OI type III removed is slightly higher than for the other types (about 80\% of OI type III compared to 60\% of OI type I and IV).
-As their are only 6 ROIs taken from the same OI type III individual, it is difficult to draw conclusion about it.
+As there are only 6 ROIs taken from the same OI type III individual, it is difficult to draw conclusion about it.
 However, the relations observed for healthy individuals are similar than the ones observed with the OI type I and IV.
@@
 
-\AR For a bit of more completeness the following sentences were added in the manuscript:\par
+\AR For completeness the following sentences were added in the manuscript:\par
 
@@
 \begin{quote}
-	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{Validation of similar fabric-elasticity relationships will allow to use HR-pQCT-based hFE for OI bone strength assessment with comparative confidence as for healthy bone.}
+	Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{ Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.}
 \end{quote}
 
+\newpage
 \subsection{Methods - Participants}
@@
 
-Yes again, by keeping the ROI in its original configuration, errors due to interpolation are avoided.This is now precised in the manuscript.\par
+Yes again, by keeping the ROI in its original configuration, errors due to interpolation are avoided. This is now precised in the manuscript.\par
 
@@
 \AR The first part of this question is closely related to Comment 1.2 of Reviewer \# 1.
-we would suggest to refer to our answer to this comment.
+We would suggest to refer to our answer to this comment.
 For the second part, we would suggest to look at the plots used to answer to your comment 2.2.
-The filtering do not remove completely a specific OI type.
+The filtering does not remove completely a specific OI type.
 
@@
 These tests will only provide insights about differences in the explanatory variables and the fit quality but not about differences in the slopes (stiffness constants).
-As said in the discussion, these differences provide from the different scan resolutions between the two studies.
+As said in the discussion, these differences derive from the different scan resolutions between the two studies.
 
@@
 Moreover, as filtering leaved only one ROI of OI type III, it is not possible to perform statistics about it.
-So the main conclusion of the present article is that fabric-elasticity relationships of trabecular bone are not affected by the OI phenotype, either it is type I, III or IV.
-This means that standard hFE analyses can be performed by clinicians even if the patient is OI diagnosed, whatever the type, because it will not have a significant difference on the bone strength assessment.
+However, this one ROI is not different to the other ROIs kept after filtering.
+So the main conclusion of the present article is that when the ROI has enough mass (BV/TV) for homogenization to make sense, the distribution of this mass (fabric-elasticity relationship) is similarly optimized either in OI (type I, III, IV) or healthy trabecular bone.
+This means that standard hFE analyses can be performed by clinicians even if the patient is OI diagnosed, whatever the type, because it will not have a significant difference on the bone stiffness assessment.
 

### 2021-10-13 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Finish Response to Reviewers and manuscript revision

@@
 The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
-Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
+\Add{However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.} Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
 
@@
 
-		Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
+		Compared to healthy controls, we found the OI samples to have significantly lower \Add{CV,} BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
 These results are in agreement with previous studies.
-The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range.
+\Del{The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range}.
+\Add{The stiffnesses of highly heterogeneous ROIs are randomly lower with respect to the fabric-elasticity relationships and reflects the limit of validity of the computational homogenisation methodology.}
+
 		The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups.
@@
 
-		In conclusion, despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the mechanical analysis.
-Since highly heterogenous ROIs coincide with very low BV/TV, we expect them to play a minor role in hFE analysis of distal bone sections.
+		In conclusion, \Add{OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls.} Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
+\Del{Since highly heterogenous ROIs coincide with low BV/TV and can be handled smoothly by the fabric-elasticity relationships we expect them to play a minor role in hFE analysis of distal bone sections.} \Add{Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.}
 	\end{abstract}
@@
 	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}.
-Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008}.
-\Add{HR-pQCT-based FEA becomes more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Homogenized finite element (hFE) is based on BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
+Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008} \Add{used for fracture risk estimation}.
+\Add{Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis}.
+Homogenized finite element (hFE) is based on \Add{local} BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
 High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}.
-Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
+\Add{Homogenised QCT-based FEA has become more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
 However, HR-pQCT-based hFE relies on fabric-elasticity relationships.
-Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.
+Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{ Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.}
 
@@
 	Switzerland) was performed at the distal tibia on all patients from both groups.
-Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
+\Add{The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}.} Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
 Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
@@
 	The HR-pQCT scans were evaluated using the manufacturer's standard protocol.
-Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking) and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}).
+Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking)\Add{.
+Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed} and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}).
 Then, mask segmented images were used for further analysis.\\
@@
 Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system.
-Such rotation would potentially decrease image quality.
+Such rotation would potentially decrease image quality \Add{due to interpolation}.
 The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
@@
 	\subsection{Linear Regression}
-	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model.
+	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model \Add{using the entire sample of both groups}.
 Standard linear models assume independent and identically distributed (iid) variables.
@@
 The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively.
-BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
+\Add{The OI group matched with control was composed of nineteen type I and nine type IV.} BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
 Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001.
@@
 Values are presented as mean $\pm$ standard deviation when statistical test is performed on the means or median (inter-quartile range) when test is on medians.
-The study of \cite{Kocijan2015} presents n.s.
+The study of Kocijan et al.
+[16] presents n.s.
 for non-significant p value test result.}
 		\label{Table2}
+		\resizebox{\textwidth}{!}{
 			\begin{tabular}{cccccccccc}
@@
 				\multirow{3}{*}{Age} & Healthy & 41 $\pm$ 14 &  & 54 (21-77) & & 44 (38-52) &  & 49 $\pm$ 16 &  \\
-			& OI Type I & \multirow{2}{*}{41 $\pm$ 15} &  & 53 (21-77) &  & 42 (35-56) & & \multirow{2}{*}{46 $\pm$ 16} & \\
-			&  OI Type III \& IV & &  &  &  & 48 (35-58) & & & \\[3ex]
-
-			\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & <0.01 & 0.14 $\pm$ 0.03 & <0.001 & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & <0.0001 \\
-			& OI Type I & \multirow{2}{*}{0.164 $\pm$ 0.079} &  & 0.08 $\pm$ 0.03 &  & 0.098 (0.088-0.114) & <0.0001 & \multirow{2}{*}{0.095 $\pm$ 0.008} & \\
-			&  OI Type III \& IV & & & & & 0.081 (0.056-0.092) & <0.0001 & & \\[3ex]
-
-			\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & <0.001 & 1.94 (1.74-2.13) & <0.001 & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & <0.0001 \\
-			& OI Type I & \multirow{2}{*}{0.650 $\pm$ 0.198} &  & 1.30 (0.75-1.53) &  & 1.33 (1.07-1.55) & <0.0001 & \multirow{2}{*}{1.428 $\pm$ 0.098} & \\
-			&  OI Type III \& IV & & & & & 0.89 (0.81-1.08) & <0.001 & & \\[3ex]
-
-			\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & 0.2 & 0.07 (0.06-0.08) & 0.5 & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & 0.046 \\
-			& OI Type I & \multirow{2}{*}{0.306 (0.292-0.331)} &  & 0.07 (0.06-0.08) &  & 0.074 (0.064-0.090) & n.s. & \multirow{2}{*}{0.066 $\pm$ 0.004} & \\
-			&  OI Type III \& IV & & & & & 0.078 (0.068-0.092) & n.s. & & \\[3ex]
-
-			\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & 0.01 & 0.44 (0.39-0.51) & <0.001 & - & & 0.409 $\pm$ 0.023 & 0.0003 \\
-			& OI Type I & \multirow{2}{*}{1.422 $\pm$ 0.694} &  & 0.68 (0.57-1.19) &  & - &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \\
-			&  OI Type III \& IV & & & & & - &  & & \\[3ex]
-
-			\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & 0.02 & 0.20 (0.16-0.25) & <0.001 & 0.221 (0.170-0.242) & & - &  \\
-			& OI Type I & \multirow{2}{*}{0.631 $\pm$ 0.383} &  & 0.40 (0.31-1.11) &  & 0.382 (0.311-0.504) & <0.0001 & - & \\
-			&  OI Type III \& IV & & & & & 0.698 (0.511-0.890) & <0.001 & & \\[3ex]
-
-			\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & <0.001 & - &  & - & & - &  \\
-			& OI Type I & \multirow{2}{*}{0.056 (0.015-0.079)} &  & - &  & - &  & - & \\
-			&  OI Type III \& IV & & & & & - &  & & \\[3ex]
-
-			\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & 0.02 & - &  & - & & - &  \\
-			& OI Type I & \multirow{2}{*}{2.018 (1.901-2.158)} &  & - &  & - &  & - & \\
-			&  OI Type III \& IV & & & & & - &  & & \\[3ex]
-
-			\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & <0.0001 & - &  & - & & - &  \\
-			& OI Type I & \multirow{2}{*}{-1.178 $\pm$ 0.441} &  & - &  & - &  & - & \\
-			&  OI Type III \& IV & & & & & - &  & & \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{41 $\pm$ 15} &  & \multirow{2}{*}{53 (21-77)} &  & 42 (35-56)\textsuperscript{1} & & \multirow{2}{*}{46 $\pm$ 16} & \\
+				& & &  &  &  & 48 (35-58)\textsuperscript{2} & & & \\[3ex]
+
+				\multirow{3}{*}{BV/TV} & Healthy & 0.222 $\pm$ 0.081 & & 0.14 $\pm$ 0.03 & & 0.141 (0.130-.0170) & & 0.162 $\pm$ 0.010 & \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.164 $\pm$ 0.079} & \multirow{2}{*}{<0.01} & \multirow{2}{*}{0.08 $\pm$ 0.03} & \multirow{2}{*}{<0.001} & 0.098 (0.088-0.114)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{0.095 $\pm$ 0.008} & \multirow{2}{*}{<0.0001} \\
+				& & & & & & 0.081 (0.056-0.092)\textsuperscript{2} & <0.0001\textsuperscript{2} & & \\[3ex]
+
+				\multirow{3}{*}{Tb.N.} & Healthy & 0.842 $\pm$ 0.144 & & 1.94 (1.74-2.13) & & 1.76 (1.59-2.08) & & 2.143 $\pm$ 0.089 & \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.650 $\pm$ 0.198} & \multirow{2}{*}{<0.001} & \multirow{2}{*}{1.30 (0.75-1.53)} & \multirow{2}{*}{<0.001} & 1.33 (1.07-1.55)\textsuperscript{1} & <0.0001\textsuperscript{1} & \multirow{2}{*}{1.428 $\pm$ 0.098} & \multirow{2}{*}{<0.0001} \\
+				& & & & & & 0.89 (0.81-1.08)\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]
+
+				\multirow{3}{*}{Tb.Th.} & Healthy & 0.301 (0.287-0.321) & & 0.07 (0.06-0.08) & & 0.081 (0.074-0.087) & & 0.075 $\pm$ 0.003 & \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.306 (0.292-0.331)} & \multirow{2}{*}{0.2} & \multirow{2}{*}{0.07 (0.06-0.08)} & \multirow{2}{*}{0.5} & 0.074 (0.064-0.090)\textsuperscript{1} & n.s.\textsuperscript{1} & \multirow{2}{*}{0.066 $\pm$ 0.004} & \multirow{2}{*}{0.046}\\
+				& & & & & & 0.078 (0.068-0.092)\textsuperscript{2} & n.s.\textsuperscript{2} & & \\[3ex]
+
+				\multirow{3}{*}{Tb.Sp.} & Healthy & 0.924 $\pm$ 0.257 & & 0.44 (0.39-0.51) & & \multirow{3}{*}{-} & & 0.409 $\pm$ 0.023 & \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{1.422 $\pm$ 0.694} & \multirow{2}{*}{0.01} & \multirow{2}{*}{0.68 (0.57-1.19)} & \multirow{2}{*}{<0.001} & &  & \multirow{2}{*}{0.727 $\pm$ 0.095} & \multirow{2}{*}{0.0003}\\
+				& & & & & & & & & \\[3ex]
+
+				\multirow{3}{*}{Tb.Sp.SD} & Healthy & 0.317 $\pm$ 0.136 & & 0.20 (0.16-0.25) & & 0.221 (0.170-0.242) & & - &  \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.631 $\pm$ 0.383} & \multirow{2}{*}{0.02} & \multirow{2}{*}{0.40 (0.31-1.11)} & \multirow{2}{*}{<0.001} & 0.382 (0.311-0.504)\textsuperscript{1} & <0.0001\textsuperscript{1} & - & \\
+				& & & & & & 0.698 (0.511-0.890\textsuperscript{2} & <0.001\textsuperscript{2} & & \\[3ex]
+
+				\multirow{3}{*}{SMI} & Healthy & 0.001 (-0.021-0.033) & & \multirow{3}{*}{-} &  & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{0.056 (0.015-0.079)} & \multirow{2}{*}{<0.001} & & & &  & & \\
+				& & & & & & & & & \\[3ex]
+
+				\multirow{3}{*}{DA} & Healthy & 1.992 (1.826-2.020) & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{2.018 (1.901-2.158)} & \multirow{2}{*}{0.02} & & & & & & \\
+				& & & & & & & & & \\[3ex]
+
+				\multirow{3}{*}{ln(CV)} & Healthy & -1.723 $\pm$ 0.344 & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} & & \multirow{3}{*}{-} &  \\
+				& \multirow{2}{*}{OI} & \multirow{2}{*}{-1.178 $\pm$ 0.441} & \multirow{2}{*}{<0.0001} & & & & & & \\
+				& & & & & & & & & \\
 
 				\bottomrule
-		\end{tabular}
+		\end{tabular}}
+		OI types included in the study: Present study, types I and IV; \citeauthor{Folkestad2012}\cite{Folkestad2012}, type I; \citeauthor{Kocijan2015}\cite{Kocijan2015}, 1) type I 2) type III and IV; \citeauthor{Rolvien2018}\cite{Rolvien2018}, type I and IV.
 	\end{sidewaystable*}
@@
 	\subsection{Comparison with literature}
-	Table \ref{Table4} shows results obtained compared to literature.
+	Table \ref{Table4} shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femora studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m \cite{Gross2013,Panyasantisuk2015}}.
 \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs.
@@
 To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
-\\
+\Add{These numbers of participants are higher than in other studies comparing OI with healthy controls \cite{Folkestad2012,Kocijan2015,Rolvien2018}}.\\
 
@@
 With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-In hFE analysis of bones, it is not possible to exclude part of the mesh because of high heterogeneity.
-Nevertheless, the error created by such ROIs is negligible as this concerns ROIs with extremely low BV/TV and therefore stiffness leading to a minor impact on the full model.\\
+\Add{Thus, the same hFE analysis can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones.
+Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity \Del{.
+Nevertheless}, \Add{the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.}\\
 
@@
 	The analysis of TMD gives interesting outputs as well.
-The t-test reveal a higher TMD in OI trabecular bone than in healthy condition, agreeing with \citeauthor{Boyde1999}\cite{Boyde1999}, with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
-The linear regression slope CI being strictly different from zero shows that a relation exists between TMD and BV/TV.
-This can have different origins.
-From a biological point of view, a mineralisation gradient develops between the core of the trabecula undergoing secondary mineralisation and the outer surface subjected to remodelling and deposition of younger, less mineralised matrix.
-As trabecular thickness decreases with BV/TV, this means that a lower BV/TV could lead to a lower mineralization of the core of the trabeculae.
-The second explanation for this slope comes from the scanning.
-Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV as the ratio of bone surface to bone volume BS/BV decreases.
+The t-test reveals a \Add{4\%} higher TMD in OI trabecular bone than in healthy condition, agreeing with \citeauthor{Boyde1999}\cite{Boyde1999}, with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
+The linear regression \Del{slope CI being strictly different from zero shows that a} \Add{gives a significant} relation \Del{exists} between TMD and BV/TV\Del{.
+This}\Add{ which} can have different origins.
+From a biological point of view, a mineralisation gradient develops between the core of a trabeculae \Del{undergoing secondary mineralisation} and the outer surface \Del{the outer surface subjected to remodelling and deposition of younger, less mineralised matrix}.
+\Add{The core undergoes secondary mineralisation and surface is subjected to remodelling and deposition of younger, less mineralised matrix.} As trabecular thickness decreases with BV/TV, this means that a lower BV/TV could lead to a lower mineralization of the core of the trabeculae.
+The second explanation \Del{for this slope} comes from the scanning.
+Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV as the ratio of bone surface to bone volume (BS/BV) decreases.
 The former (biological) explanation is expected to have a lower impact than the latter (scanning).
-Regarding the other coefficients of the linear regression, the CI of group variable excludes zero as well, meaning that the intercept depends on the group.
-This could be explained by the bisphosphonate treatment that OI patients receive.
-Effectively, bisphosphonates are aimed to slow down the remodelling process which lead to higher (secondary) mineralization of the bone.
-According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest adding a correction accounting for the TMD in FE simulations to catch the higher modulus, ultimate stress and post-yield behaviour of OI bone compared to healthy bone at the ECM level.
-Regarding the interaction between BV/TV and the group, the p-value obtained showed a high non-significance meaning that the slopes of both OI and healthy groups are the same.
-This could be visualised in Figure \ref{02_TMD} where the fitted lines are obtained using the fixed effects of the model and the group variable if fixed.
-In this plot, performed with the interaction term (BV/TV x Group) the fitted lines appear to be quasi-parallel.
-\\
+\Del{Regarding the other coefficients of the linear regression, the CI of group variable excludes zero as well, meaning that}\Add{Moreover,} the intercept \Add{of the linear regression} depends on the group \Add{i.e.
+TMD is still different between healthy and OI group.} This could be explained by the bisphosphonate treatment that OI patients receive.
+Indeed, bisphosphonates are aimed to slow down the remodelling process which lead to higher (secondary) mineralization of the bone.
+According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest adding a correction accounting for the TMD in FE simulations to catch the \Del{higher modulus, ultimate stress and post-yield}\Add{different} behaviour of OI bone compared to healthy bone at the ECM level.
+Regarding the \Del{interaction between BV/TV and the group, the p-value obtained showed a high non-significance meaning that the slopes of both OI and healthy groups are the same}\Add{effect of BV/TV over TMD, it is similar between the healthy and the OI group}.
+This could be visualised in Figure \ref{02_TMD} where the fitted lines \Add{of both groups} are obtained using the fixed effects of the model\Del{ and the group variable is fixed}.
+In this plot performed with the interaction term (BV/TV x Group), the fitted lines appear to be quasi-parallel.\\
 
@@
 low CV) and the fact that it is limited to XCTII reconstructions of the distal tibia.
-Moreover, having only one patient with OI type III where we could extract ROIs does not allows to perform statistics.
-%As those patients are less active, it could be interesting to analyse the impact of this condition on the weight bearing tibia.
-	ROI homogeneity is a necessary condition underlying homogenisation of mechanical properties and the use of fabric-elasticity relationships. %has an important impact on the analysis quality. 
-	As proposed earlier, ROI homogeneity could be assessed with an alternative ROI filtering and more efforts could be invested to improve the model for highly heterogeneous ROIs, but as it concerns mainly ROIs with extremely low stiffness, the impact on hFE models should remain negligible.
-	%if the proportion of such ROIs stays low.
+Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.ROI homogeneity\Add{, more specifically a balanced bone connectivity of the 6 faces of the ROI,} is a necessary condition underlying the computational homogenisation of mechanical properties, \Del{and the use of fabric-elasticity relation\-ships.
+As proposed earlier, ROI homogeneity could be asses\-sed with an alternative ROI filtering and more efforts could be invested to improve the model for highly heterogeneous ROIs, but as this concerns mainly ROIs with low stiffness, the impact on hFE models should remain negligible.} \Add{but when calibrated with homogeneous ROIs, fabric-elasticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-homogeneity and low-stiffness ROIs.
+In fact, these ROIs were not shown to compromise the performance of hFE models of the distal radius \cite{Varga2011, Hosseini2017, AriasMoreno2019}.}
+
 	A similar study could be performed on radius XCTII reconstructions to generalise the above findings.
@@
 
-	In conclusion, the OI trabecular bone from the distal tibia analysed in the present study had similar morphology compared to those reported in the literature.
-We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough i.e.
-with a CV lower than 0.263 and were matched for BV/TV and DA.
-	\citeauthor{Indermaur2021} \cite{Indermaur2021} found that the compressive behaviour of OI bone extracellular matrix is not inferior to healthy control at the ECM level.
+	In conclusion, OI trabecular bone from the distal tibia analysed in the present study had similar morphology compared to those reported in the literature.
+\Add{In particular, we confirmed a significantly lower volume fraction of OI trabecular bone compared to healthy controls that points towards a lower bone strength independently of potentially compromised bone matrix material properties.
+We also observed a higher tBMD}
+	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough \Add{to trust the computational homogenisation methodology} (i.e.
+CV lower than 0.263), and were matched for BV/TV and DA.
+\Add{Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients as well.} \Del{Indermaur et al.\cite{Indermaur2021} found that the compressive behaviour of OI bone extra\-cellular matrix is not inferior to healthy control at the ECM level.
 Similar results for tensile properties at the extracellular matrix level would indicate that fabric-strength relationships may be similar as well.
-Under this hypothesis, fragility of OI bone should mainly be explained by thin cortices and low BV/TV of the trabecular compartment.
+Under this hypothesis, fragility of OI bone should mainly be explained by thin cortices and low BV/TV of the trabecular compartment.}
 

### 2021-10-13 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Added comment on TMD and BV/TV influences on mechanics

@@
 \Add{In particular, we confirmed a significantly lower volume fraction of OI trabecular bone compared to healthy controls that points towards a lower bone strength independently of potentially compromised bone matrix material properties.
-We also observed a higher tBMD}
+We also observed a 4\% higher TMD in OI which suggest, according to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021}, an increase in ultimate stress about 2\%.
+%Figure7A y = 0.59x%
+	However, this light increase is largely overcome by the lower BV/TV observed (\textasciitilde 26\% less) in OI.
+According to the results of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015}, this lower volume fraction leads to a decrease about 45\% in stiffness.} %(74**2-100**2)/100**2
 	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough \Add{to trust the computational homogenisation methodology} (i.e.

### 2021-10-15 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript.tex
Finalize revision submission

--- 2021-10-13 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
@@
 The hFE scheme uses homogenised local bone volume fraction (BV/TV) and anisotropy information (fabric) to compute healthy bone strength within a reasonable computation time using fabric-elasticity relationships.
-\Add{However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.} Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
+However, it is unclear if these relationships quantified previously for healthy controls are valid for trabecular bone from OI patients.
+Thus, the aim of this study is to investigate fabric-elasticity relationships in OI trabecular bone compared to healthy controls.
 
@@
 
-		Compared to healthy controls, we found the OI samples to have significantly lower \Add{CV,} BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
+		Compared to healthy controls, we found the OI samples to have significantly lower CV, BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
 These results are in agreement with previous studies.
-\Del{The stiffness of ROIs from OI bone reached lower values compared to healthy controls and the multilinear fabric-elasticity fits tended to overestimate the stiffness in the lower range}.
-\Add{The stiffnesses of highly heterogeneous ROIs are randomly lower with respect to the fabric-elasticity relationships and reflects the limit of validity of the computational homogenisation methodology.}
+The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
 
@@
 
-		In conclusion, \Add{OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls.} Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
-\Del{Since highly heterogenous ROIs coincide with low BV/TV and can be handled smoothly by the fabric-elasticity relationships we expect them to play a minor role in hFE analysis of distal bone sections.} \Add{Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.}
+		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls.
+Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
+Accordingly, the elastic properties used for FEA of healthy bones are also valid for OI bones.
 	\end{abstract}
@@
 
-	\Add{OI occurs worldwide and without sex preference.
-OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% \cite{Martin2007}}.Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels.
-Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018} \Add{but aBMD is limited for bone strength estimation \cite{Pasco2006}}.
+	OI occurs worldwide and without sex preference.
+OI type I is estimated to occur in 45\% of the total OI cases, type II about 10\%, type III 25\% and type IV 20\% \cite{Martin2007}.Bone fragility in OI is complex and not totally understood, despite the investigations at different hierarchical levels.
+Multiple studies show that DXA areal bone mineral density (aBMD) tends to be lower in OI compared to healthy individuals \cite{Folkestad2012,Lindahl2015,Scheres2018} but aBMD is limited for bone strength estimation \cite{Pasco2006}.
 \citeauthor{Folkestad2012}\cite{Folkestad2012}, \citeauthor{Kocijan2015}\cite{Kocijan2015}, and \citeauthor{Rolvien2018}\cite{Rolvien2018} have shown that the microstructure is different as well.
@@
 	High resolution peripheral quantitative computed tomography (HR-pQCT) reconstructions allow \textit{in vivo} assessment of cortical and trabecular architecture and volumetric bone mineral density (BMD) at the distal radius and distal tibia \cite{Boutroy2005}.
-Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008} \Add{used for fracture risk estimation}.
-\Add{Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis}.
-Homogenized finite element (hFE) is based on \Add{local} BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
+Moreover, HR-pQCT reconstructions can be used for finite element analysis (FEA) to predict mechanical properties, such as bone stiffness and strength \cite{Boutroy2008} used for fracture risk estimation.
+Micro-finite element (\si{\micro}FE) analysis converts each voxel into an element and thus requires high performance computer for a bone section analysis.
+Homogenized finite element (hFE) is based on local BV/TV and anisotropy information (fabric) from HR-pQCT reconstructions and can be used to assess bone strength within reasonable computation effort \cite{Pahr2009}.
 High correlations were found between patient-specific hFE and mechanical compression experiments of freshly frozen human samples at the distal radius \cite{Varga2011,Hosseini2017,AriasMoreno2019}.
-\Add{Homogenised QCT-based FEA has become more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).} Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
+Homogenised QCT-based FEA has become more widely used in clinical studies \cite{VanDenBergh2021} and is currently approved by the American Food and Drug Administration (FDA).
+Thus, it could be legitimate to use hFE on OI affected bone, for estimation of bone strength and potentially associated fracture risk.
 However, HR-pQCT-based hFE relies on fabric-elasticity relationships.
-Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.\Add{ Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.}
+Therefore, the present study aims to compare trabecular bone microstructure of healthy and OI bone samples and to test the hypothesis of similar fabric-elasticity relationships.
+Similar fabric-elasticity relationships will allow to extend the application of HR-pQCT-based hFE of healthy bone to OI bone at least in the linear elastic regime.
 
@@
 	Switzerland) was performed at the distal tibia on all patients from both groups.
-\Add{The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}.} Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
-Namely, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
+The distal tibia was chosen because numerous studies were performed on the distal radius \cite{AriasMoreno2019,Boutroy2008,Gross2013,Hosseini2017,Kocijan2015,Varga2011}, some study performed analysis on both distal radius and tibia with XtremeCTI (XtremeCT) \cite{Boutroy2005,Folkestad2012} but only one study was performed on the distal tibia with XtremeCTII \cite{Schenk2020}.
+Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
+Name\-ly, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
 On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.
@@
 	The HR-pQCT scans were evaluated using the manufacturer's standard protocol.
-Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking)\Add{.
-Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed} and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}).
+Briefly, an automatic contouring algorithm was applied to define the periosteal contour of the tibia (masking).
+Then, a Gauss filtering (sigma = 0.8, support = 1 voxel) was performed and a threshold was applied for segmentation of cortical bone (450 mgHA/cm\textsuperscript{3}) and trabecular bone (320 mgHA/cm\textsuperscript{3}).
 Then, mask segmented images were used for further analysis.\\
@@
 Each ROI had to contain trabecular bone, but no cortical bone.
-For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number \Del{3}\Add{1}) to be at roughly the same anatomical location as for the OI group.
-Then, for both the healthy group and the OI group, the stack was divided into two halves \Add{with a transverse plane} and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
+For the healthy group, the ROIs were selected in the most proximal stack uniquely (see Figure \ref{01_Healthy} stack number 1) to be at roughly the same anatomical location as for the OI group.
+Then, for both the healthy group and the OI group, the stack was divided into two halves with a transverse plane and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
 For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone.
@@
 Unlike periodicity-compatible mixed uniform boundary conditions (PMUBCs), KUBCs do not require to rotate the ROI into the respective fabric coordinate system.
-Such rotation would potentially decrease image quality \Add{due to interpolation}.
+Such rotation would potentially decrease image quality due to interpolation.
 The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
@@
 	\subsection{Linear Regression}
-	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model \Add{using the entire sample of both groups}.
+	The orthotropic stiffness tensors obtained after transformation onto fabric coordinate system were then used to perform a multiple linear regression with the Zysset-Curnier model using the entire sample of both groups.
 Standard linear models assume independent and identically distributed (iid) variables.
@@
 The mean age was 41y $\pm$ 14y and 41y $\pm$ 15y for the mat\-ched healthy and OI individuals, respectively.
-\Add{The OI group matched with control was composed of nineteen type I and nine type IV.} BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
+The OI group matched with control was composed of nineteen type I and nine type IV.
+BV/TV of healthy individuals was higher than BV/TV of OI group with a difference 95\% CI of [0.016, 0.101] and p-value <0.01.
 Similarly, trabecular number was higher in the matched healthy group compared to the matched OI group with a difference 95\% CI of [0.099, 0.285] and a corresponding p-value <0.001.
@@
 	\subsection{Comparison with literature}
-	Table \ref{Table4} shows results obtained \Add{using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m} compared to \Del{literature} \Add{previous femora studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m \cite{Gross2013,Panyasantisuk2015}}.
+	Table \ref{Table4} shows results obtained using HR-pQCT tibiae scans with a resolution of 61 \si{\micro}m compared to previous femora studies using \si{\micro}CT with a resolution of 18\si{\micro}m coarsened to 36\si{\micro}m \cite{Gross2013,Panyasantisuk2015}.
 \citeauthor{Gross2013} \cite{Gross2013} has the larger number of ROIs.
@@
 To do this, the study included two groups of participants composed of 120 healthy control and 50 OI diagnosed patients respectively.
-\Add{These numbers of participants are higher than in other studies comparing OI with healthy controls \cite{Folkestad2012,Kocijan2015,Rolvien2018}}.\\
+These numbers of participants are higher than in other studies comparing OI with healthy controls \cite{Folkestad2012,Kocijan2015,Rolvien2018}.\\
 
@@
 With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-\Add{Thus, the same hFE analysis can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones.
-Although} in hFE analysis \Del{of bones,} it is not possible to exclude part of the mesh because of high heterogeneity \Del{.
-Nevertheless}, \Add{the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.}\\
+Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones.
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\\
 
@@
 Nevertheless, $l$ stays in the same range as for the two other studies \cite{Gross2013,Panyasantisuk2015}, meaning the relative weight of DA remains constant.
-On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.
-\\
+On the other hand, the higher $k$ highlights an increased relative weight of BV/TV.\\
 
 	The analysis of TMD gives interesting outputs as well.
-The t-test reveals a \Add{4\%} higher TMD in OI trabecular bone than in healthy condition, agreeing with \citeauthor{Boyde1999}\cite{Boyde1999}, with 95\% certainty and a very high significance level, even with similar BV/TV and DA.
-The linear regression \Del{slope CI being strictly different from zero shows that a} \Add{gives a significant} relation \Del{exists} between TMD and BV/TV\Del{.
-This}\Add{ which} can have different origins.
-From a biological point of view, a mineralisation gradient develops between the core of a trabeculae \Del{undergoing secondary mineralisation} and the outer surface \Del{the outer surface subjected to remodelling and deposition of younger, less mineralised matrix}.
-\Add{The core undergoes secondary mineralisation and surface is subjected to remodelling and deposition of younger, less mineralised matrix.} As trabecular thickness decreases with BV/TV, this means that a lower BV/TV could lead to a lower mineralization of the core of the trabeculae.
-The second explanation \Del{for this slope} comes from the scanning.
-Effectively, during scanning a phenomenon called partial volume effect occurs and its impact decreases with an increasing BV/TV as the ratio of bone surface to bone volume (BS/BV) decreases.
+The t-test reveals a 4\% higher TMD in OI trabecular bone than in healthy condition, agreeing with \citeauthor{Boyde1999}\cite{Boyde1999}, with 95\% certainty and a very high significance level, despite similar BV/TV and DA.
+The linear regression shows a significant relation between TMD and BV/TV, that is parallel in the healthy and OI groups (Figure \ref{02_TMD}).
+This can have two different origins.
+From a biological point of view, a mineralisation gradient develops between the core of a trabecula and the outer surface.
+The core undergoes secondary mineralisation while the surface is subjected to remodelling and deposition of younger, less mineralised matrix.
+An increase in BV/TV reduces the relative proportion of bone surface bone per unit bone volume (BS/BV), which would therefore raise the mean TMD.
+The second explanation comes from the partial volume effect of the scanner that reduces the intensity of surface voxels.
+Following the same argument, the reduced ratio of bone surface to bone volume (BS/BV) raises the mean TMD with increasing BV/TV.
 The former (biological) explanation is expected to have a lower impact than the latter (scanning).
-\Del{Regarding the other coefficients of the linear regression, the CI of group variable excludes zero as well, meaning that}\Add{Moreover,} the intercept \Add{of the linear regression} depends on the group \Add{i.e.
-TMD is still different between healthy and OI group.} This could be explained by the bisphosphonate treatment that OI patients receive.
-Indeed, bisphosphonates are aimed to slow down the remodelling process which lead to higher (secondary) mineralization of the bone.
-According to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021} these findings suggest adding a correction accounting for the TMD in FE simulations to catch the \Del{higher modulus, ultimate stress and post-yield}\Add{different} behaviour of OI bone compared to healthy bone at the ECM level.
-Regarding the \Del{interaction between BV/TV and the group, the p-value obtained showed a high non-significance meaning that the slopes of both OI and healthy groups are the same}\Add{effect of BV/TV over TMD, it is similar between the healthy and the OI group}.
-This could be visualised in Figure \ref{02_TMD} where the fitted lines \Add{of both groups} are obtained using the fixed effects of the model\Del{ and the group variable is fixed}.
-In this plot performed with the interaction term (BV/TV x Group), the fitted lines appear to be quasi-parallel.\\
+Nevertheless, the intercept of the linear regression depends on the group i.e.
+TMD is significantly higher in the OI group over the entire BV/TV range.
+This could be explained by the biphosphonate treatment that OI patients receive.
+Indeed, bisphosphonates slow down the remodelling process which lead to higher (secondary) mineralization of the bone.
+According to the recent results of \citeauthor{Indermaur2021}\cite{Indermaur2021}, our findings suggest adding a correction for TMD in FE simulations to account for the resulting increase in modulus, ultimate stress and post-yield properties of OI bone compared to healthy bone at the ECM level.\\
 
@@
 low CV) and the fact that it is limited to XCTII reconstructions of the distal tibia.
-Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.ROI homogeneity\Add{, more specifically a balanced bone connectivity of the 6 faces of the ROI,} is a necessary condition underlying the computational homogenisation of mechanical properties, \Del{and the use of fabric-elasticity relation\-ships.
-As proposed earlier, ROI homogeneity could be asses\-sed with an alternative ROI filtering and more efforts could be invested to improve the model for highly heterogeneous ROIs, but as this concerns mainly ROIs with low stiffness, the impact on hFE models should remain negligible.} \Add{but when calibrated with homogeneous ROIs, fabric-elasticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-homogeneity and low-stiffness ROIs.
-In fact, these ROIs were not shown to compromise the performance of hFE models of the distal radius \cite{Varga2011, Hosseini2017, AriasMoreno2019}.}
+Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.
+ROI homogeneity, more specifically a balanced bone connectivity of the 6 faces of the ROI, is a necessary condition underlying the computational homogenisation of mechanical properties and the results of heterogenous samples must be disregarded.
+However, when calibrated with homogeneous ROIs, fabric-elas\-ticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-homogeneity and low-stiffness ROIs.
+In fact, these low-property ROIs were not shown to compromise the performance of hFE models of the distal radius \cite{Varga2011, Hosseini2017, AriasMoreno2019}.\\
 
@@
 The section scanned for the control group was at a fixed distance, while the section scanned for the OI group was placed at a distance relative to the total length of the tibia.
-Nevertheless, as shown in Figure \ref{01_ClinicalSections}, the locations compare well and the ROIs were matched for BV/TV and DA to compare the fabric-elasticity relationships.
+Nevertheless, as shown in Figure \ref{01_ClinicalSections}, the locations compare well and the ROIs were matched for BV/TV and DA to compare the fabric-elas\-ticity relationships.
 A final limitation is that the measurements were performed on different HR-pQCT devices which were not cross-calibrated.
@@
 	In conclusion, OI trabecular bone from the distal tibia analysed in the present study had similar morphology compared to those reported in the literature.
-\Add{In particular, we confirmed a significantly lower volume fraction of OI trabecular bone compared to healthy controls that points towards a lower bone strength independently of potentially compromised bone matrix material properties.
-We also observed a higher tBMD}
-	We could not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when the ROIs were homogeneous enough \Add{to trust the computational homogenisation methodology} (i.e.
-CV lower than 0.263), and were matched for BV/TV and DA.
-\Add{Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients as well.} \Del{Indermaur et al.\cite{Indermaur2021} found that the compressive behaviour of OI bone extra\-cellular matrix is not inferior to healthy control at the ECM level.
-Similar results for tensile properties at the extracellular matrix level would indicate that fabric-strength relationships may be similar as well.
-Under this hypothesis, fragility of OI bone should mainly be explained by thin cortices and low BV/TV of the trabecular compartment.}
+In particular, we confirmed a significantly lower (-26\%) volume fraction of OI trabecular bone compared to healthy controls that points towards a substantial reduction in bone strength of 45\% when applying the well-established squared relationship with BV/\-TV \cite{Fleps2020}.
+	We also observed a 5\% higher TMD in OI % With mean of each group 4.3% with fitted line at mean BV/TV range 4.7% TMD increase in OI.
+	which suggests, according to the results of \citeauthor{Indermaur2021}\cite{Indermaur2021}, an increase in matrix ultimate stress of about 13\%.
+% Mean BV/TV:0.249 healthy: 576, OI: 603, Michi slope: y = -0.4 + 0.0011*x
+	However, this moderate increase in matrix properties is large\-ly overcompensated by the 45\% reduction in trabecular bone strength associated with BV/TV. % Using k=2: (74**2-100**2)/100**2
+	We did not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when only valid, homogenous ROIs were included and matched for both BV/TV and DA.
+Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients, potentially with a small correction for TMD.
 

### 2021-11-18 09_Article/01_Revision2_Reviewers/Response to Reviewers_00.tex
Draft of second revision

% An example for the ar2rc document class.
% Copyright (C) 2017 Martin Schroen
% Modifications Copyright (C) 2020 Kaishuo Zhang
%
% This program is free software: you can redistribute it and/or modify
% it under the terms of the GNU General Public License as published by
% the Free Software Foundation, either version 3 of the License, or
% (at your option) any later version.
%
% This program is distributed in the hope that it will be useful,
% but WITHOUT ANY WARRANTY; without even the implied warranty of
% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
% GNU General Public License for more details.
%
% You should have received a copy of the GNU General Public License
% along with this program.  If not, see <http://www.gnu.org/licenses/>.

% Use quote enviromnent for manuscript text

\documentclass{AR2RC}

\usepackage{siunitx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{array}
\usepackage{multicol}
\usepackage{rotating}
\usepackage{multirow}
\usepackage{float}
\usepackage[justification=centering]{caption}
\usepackage{tabularx}
\usepackage{booktabs}

\renewcommand{\arraystretch}{1.2}

\hypersetup{hidelinks}

\title{Fabric-Elasticity Relationships of Tibial Trabecular Bone are Similar in Osteogenesis Imperfecta and Healthy Individuals}
\author{Mathieu Simon, Michael Indermaur, Denis Schenk, Seyedmahdi Hosseinitabatabaei, Bettina	M. Willie and Philippe Zysset}
\journal{Bone}
\doi{BONE-D-21-00758-R1}

\begin{document}

\maketitle

\vspace{1em}Once again, the authors would like to thank the reviewers for helping us to improve our article. We do believe that their and our time spent working on this document will definitively improve its message to the scientific community. As for the first revision, answers to comments are provided point by point in this document. In the revised manuscript, the added elements are written in blue and the canceled text is in red.

\section{Density}
\RC The first highlight states that OI trabecular bone is more mineralized compared to healthy bone, but the first sentence in the last paragraph of the Abstract says that OI bone is significantly less dense than healthy controls. This is going to be confusing for many. You need to clarify, which is presumably because you’re referring to tissue density in one case and apparent density in the other.

\AR Yes, you are right. To improve clarity the highlight and the abstract were modified as follow:

Highlight 1
\begin{quote}
	Osteogenesis Imperfecta (OI) \Del{trabecular} bone \Add{tissue} is more mineralized compared to healthy
\end{quote}

Abstract, last paragraph, first sentence
\begin{quote}
	In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and \Del{less dense} \Add{have a lower BV/TV} than healthy controls
\end{quote}

\section{Groups age ranges}
\RC Your response to the reviewer (2.9) incorrectly states the mean and plus/minus age of both groups (at least comparing to the comments in Methods 2.1). Which are the correct numbers?

\AR Effectively the numbers are different but both are correct. On one hand, in Methods 2.1 the mean and plus/minus age of both groups refers to the entire cohort. On the other hand, the response to reviewer 2.9 refers to Results 3.1 after the sex and age matching of groups. We forgot to precise it in the first response to reviewers.

\section{Typo}
\RC Page 3, second column, 4th line: Please fix typo “no enough” for “not enough”.

\AR This was changed.

\section{Exponential notation}
\RC Page 9, second column, 11th line from bottom: Why express the modulus as an exponent as such? Is this a typo? Currently it is written “10\textsuperscript{0}”. Why not write 1 MPa?

\AR This format was chosen because the text refers to Figure 5b (see below) where the exponential scale is written in power of 10. To improve clarity, the text was changed as shown here:

Page 9, second column, Discussion, Lines 50-52
\begin{quote}
	The linear regression plot (Figure 5b) shows that when the stiffness term decreases to \Add{1 MPa (}10\textsuperscript{0}\Add{)} \Del{MPa} and lower, the linear regression tends to overestimate the stiffness.
\end{quote}

\begin{figure}[h!]
	\centering
	\includegraphics[width=0.5\textwidth]
	{Pictures/02_GR_OI_LMM}
	\caption{Linear regression results for original OI group using the fixed effects of the linear mixed-effect model.}
	\label{CV_BVTV2}
\end{figure}

\newpage
\section{Clinical refocusing}
\RC Both reviewers were concerned about the fit of the manuscript content with the audience of Bone. While I agree that the Bone audience is accustomed to technical papers, please consider further refocusing to the clinical aspect. I would question where it is necessary to explicitly include equations 2 and 3 (page 4) since those are presumably outlined in the cited reference. If needed at all, considering moving to Appendix. Furthermore, equation 7 is not explained and therefore its relevance (and frequent reference later in the Discussion) is not clear.

\AR We understand your concern and agree to further refocusing to the clinical aspect. Regarding equations 2 and 3, even if they are outlined in cited references, we think that they are necessary for technical completeness. On the other hand, as they are more related to technical aspect we moved them in Appendix, as suggested. The text citing equation 7 was modified to describe the term contained and the signification of this NE, see below. We hope that these firsts steps will help to decrease the technical difficulty and increase the clarity of the paper from a clinical point of view.

Page 5, second column, Methods 2.5.3, last paragraph
\begin{quote}
	\begin{equation}
	\text{NE} = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
	\label{Eq206}
	\end{equation}

	\Add{Where $\mathbb{S}_o$ and $\mathbb{S}_p$ are the observed and predicted stiffness tensors, respectively. This relative norm error allows to quantify the accuracy of the prediction.}
\end{quote}

\section{Results validity}
\RC Most importantly, both reviewers had concerns about the validity of the results across the entire range of trabecular bone microarchitecture encountered in OI, and in particular the case where filtering was required to exclude highly heterogeneous bone microarchitecture. While it is understood that the focus of the paper is to explore whether fabric-elasticity relationships in OI and healthy bone are similar (thus justifying why filtering was performed), it is concluded that this is “encouraging the use of HR-pQCT based hFE outcomes for fracture risk assessment.” The comment about future fracture prediction by hFE is of course very important clinically, but the messaging about how to use these results are still not clear. Both the Abstract and the Discussion should comment on the implications of the fabric-elastic relationship not being defined for highly heterogeneous bone. Firstly, given there is only one OI type III where ROIs could be extracted, does this mean hFE should be restricted to OI type I? What evidence is there that it can be employed for OI type III or IV? This needs to be clearly justified. Second, in the future, when significant heterogeneity is encountered in a bone, do you still recommend to use hFE? In other words, please comment whether the findings of skewness for high heterogeneity affects if/when hFE is appropriate, and if there are alternatives to hFE in that case.

\AR Higly heterogeneous bone is rare in healthy conditions leading to the observed good correlations between simulations and experiments in the work of Varga et al. 2011, Hosseini et al. 2017, and Arias-Moreno et al. 2019. However, high heterogeneity arise from low bone volume fraction as outlined in the plot relating CV and BV/TV. This fact is determinant. As low BV/TV will lead to low stiffness, the error performed on such ROI is negligible as it not have a significant contribution to the stiffness of the full tibial section. For example, the observed stiffness component $\lambda_{ii}$ of one highly heterogeneous ROI (worst case) is around 1 MPa and the predicted value is around 100 MPa. This error is relatively small in comparison with the homogeneous ROI presenting a value around 10 GPa. Thus, only homogeneous ROIs (i.e. ROIs with higher BV/TV) contribute significantly to the bone stiffness. Moreover, despite the contribution of trabecular bone, the cortical bone will contribute even more significantly to tibial stiffness. Cortical bone having a higher BV/TV than trabecular bone, the relative error is low. Another point is that ROIs here have 5.3mm side length, reducing this size will improve homogeneity of ROIs and thus reduce error of the stiffness estimation. To improve the clarity of the message regarding validity of hFE among the OI types or heterogeneous bone, the manuscript was modified as follow:

Abstract, paragraph 4, last sentence
\begin{quote}
	The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology. \Add{Nevertheless, the error performed on these highly heterogeneous ROIs is negligible in comparison with homogeneous ROIs which are order of magnitude stiffer.}
\end{quote}

Discussion, Page 10, Second Column, Line 30
\begin{quote}
	\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e. fabric-elasticity relationships) are independent of the OI type. Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix D.}
\end{quote}

Discussion, Page 10, second column, last paragraph, last sentence
\begin{quote}
	Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}. Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity), such as in OI tpye III.}
\end{quote}

\section{Supplementary material}
\RC It is suggested that Figure 2 in the response to reviewers be included in the manuscript (appendix or supplementary material).

\AR As required, the Figure was added in Appendix D

\end{document}

### 2021-11-18 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Draft of second revision

--- 2021-10-15 09_Article/01_Revision1_Reviewers/Simon_2021_Bone_Manuscript.tex
@@
 
-		Compared to healthy controls, we found the OI samples to have significantly lower CV, BV/TV and trabecular number (Tb.N.), significantly higher trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
+		Compared to healthy controls, we found the OI samples to have significantly lower BV/TV and trabecular number (Tb.N.), significantly higher CV, trabecular separation (Tb.Sp.) and trabecular separation standard deviation (Tb.Sp.SD), but no differences in trabecular thickness (Tb.Th.).
 These results are in agreement with previous studies.
 The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
+\Add{Nevertheless, the error performed on these highly heterogeneous ROIs is negligible in comparison with homogeneous ROIs which are order of magnitude stiffer.}
 
 		The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups.
-Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
-Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.
-
-		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and less dense than healthy controls.
+Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.\Add{ Moreover, the filtering step did not exclude a particular OI type.} Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.
+
+		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and \Del{less dense} \Add{have a lower BV/TV} than healthy controls.
 Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
@@
 Then, for both the healthy group and the OI group, the stack was divided into two halves with a transverse plane and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
-For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no enough trabecular bone.
+For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no\Add{t} enough trabecular bone.
 This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\
@@
 The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
-This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.\\
-
-	\begin{equation}
-	\begin{split}
-	&\mathbb{S}(\rho,\mathbf{M}) &=& \quad\sum_{i=1}^{3} \lambda_{ii} \mathbf{M}_i \otimes \mathbf{M}_i \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \lambda_{ij} \mathbf{M}_i \otimes \mathbf{M}_j \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \mu_{ij} \mathbf{M}_i \overline{\underline{\otimes}} \mathbf{M}_j \\
-	&\text{with} &\\
-	&\qquad\lambda_{ii} &=& \quad(\lambda_0 + 2\mu_0)\rho^k m_i^{2l} \\
-	&\qquad\lambda_{ij} &=& \quad\lambda_0' \rho^k m_i^{l} m_j^{l} \\
-	&\qquad\mu_{ij} &=& \quad\mu_0 \rho^k m_i^{l} m_j^{l} \\
-	\end{split}
-	\label{Eq202}
-	\end{equation}
-
-	Where $\otimes$ and $\overline{\underline{\otimes}}$ are the dyadic and symmetric product of second order tensors, respectively.
-To express the stiffness tensor obtained from the homogenisation process with the Zysset-Curnier model, it had to be transformed into the fabric coordinate system using a coordinates transformation formula (see Equation \ref{Eq203p}) and projected onto orthotropy, leading to 12 components.
-
-	\begin{equation}
-	S_{ijkl}' = Q_{im}Q_{jn}Q_{ko}Q_{lp} S_{mnop}
-	\label{Eq203p}
-	\end{equation}
-
-	Where $\mathbb{S}'$ and $\mathbb{S}$ are the transformed and the original stiffness tensor respectively and $\mathbf{Q}$ is the orthogonal matrix that maps the original coordinate system into the new one (fabric).
+This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV \Del{or $\rho$}, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$\Del{, as shown in Equation \ref{Eq202}}.
+\Add{This stiffness tensor is then transformed into the fabric coordinate system} and projected onto orthotropy, leading to 12 components.
+\Add{The technical details of the computations performed are presented in Appendix \ref{A0}.}\\
+
 	The Zysset-Curnier model is built with the assumption of orthotropy and homogeneity of the ROI.
@@
 Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient squared ($R^2$) and the standard error of the estimate (SE).
-Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}.
+Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}\Add{, respectively}.
 
@@
 	\end{equation}
+
+	\Add{Where $\mathbb{S}_o$ and $\mathbb{S}_p$ are the observed and predicted stiffness tensors, respectively. This relative norm error allows to quantify the accuracy of the prediction.}
 
@@
 These differences come from ROIs presenting a low stiffness.
-The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to $10^0$ MPa and lower, the linear regression tends to overestimate the stiffness.
+The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to \Add{1 MPa (}10\textsuperscript{0}\Add{)} \Del{MPa} and lower, the linear regression tends to overestimate the stiffness.
 This is because ROI stiffness is highly dependent on BV/TV values.
@@
 However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further.
-\\
+\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e.
+fabric-elasticity relationships) are independent of the OI type.
+Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix} \ref{A3}\Add{.} \\
 
@@
 With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones.
-Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\\
+Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}.
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity).}\\
 
@@
 	\appendix
+	\section{Technical Computations}\label{A0}
+
+	The Zysset-Curnier model \cite{Zysset1995} computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.
+
+	\begin{equation}
+	\begin{split}
+	&\mathbb{S}(\rho,\mathbf{M}) &=& \quad\sum_{i=1}^{3} \lambda_{ii} \mathbf{M}_i \otimes \mathbf{M}_i \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \lambda_{ij} \mathbf{M}_i \otimes \mathbf{M}_j \\ &&&+ \sum_{\substack{i,j=1\\i \neq j}}^{3} \mu_{ij} \mathbf{M}_i \overline{\underline{\otimes}} \mathbf{M}_j \\
+	&\text{with} &\\
+	&\qquad\lambda_{ii} &=& \quad(\lambda_0 + 2\mu_0)\rho^k m_i^{2l} \\
+	&\qquad\lambda_{ij} &=& \quad\lambda_0' \rho^k m_i^{l} m_j^{l} \\
+	&\qquad\mu_{ij} &=& \quad\mu_0 \rho^k m_i^{l} m_j^{l} \\
+	\end{split}
+	\label{Eq202}
+	\end{equation}
+
+	Where $\otimes$ and $\overline{\underline{\otimes}}$ are the dyadic and symmetric product of second order tensors, respectively.
+To express the stiffness tensor obtained from the homogenisation process with the Zysset-Curnier model, it had to be transformed into the fabric coordinate system using a coordinates transformation formula, see Equation \ref{Eq203p}.
+
+	\begin{equation}
+	S_{ijkl}' = Q_{im}Q_{jn}Q_{ko}Q_{lp} S_{mnop}
+	\label{Eq203p}
+	\end{equation}
+
+	Where $\mathbb{S}'$ and $\mathbb{S}$ are the transformed and the original stiffness tensor respectively and $\mathbf{Q}$ is the orthogonal matrix that maps the original coordinate system into the new one (fabric).
+After transformation, the stiffness tensor is projected onto orthotropy, leading to 12 components which can be represented in a 6x6 matrix as follow:
+
+	\begin{equation}
+	\mathbb{S} = 
+	\begin{pmatrix}
+		\lambda_{11} & \lambda_{12} & \lambda_{13} & 0 & 0 & 0 \\
+		\lambda_{21} & \lambda_{22} & \lambda_{23} & 0 & 0 & 0 \\
+		\lambda_{31} & \lambda_{32} & \lambda_{33} & 0 & 0 & 0 \\
+		0 & 0 & 0 & \mu_{23} & 0 & 0 \\
+		0 & 0 & 0 & 0 & \mu_{31} & 0 \\
+		0 & 0 & 0 & 0 & 0 & \mu_{12}
+	\end{pmatrix}
+	\label{Eq203x}
+	\end{equation}
+
 	\section{Linear Models}\label{A1}
@@
 	\clearpage
+	\section{OI Types Properties Distribution}\label{A3}
+
+	As the OI cohort is composed of different OI types, it was verified that the analyses performed where not sensitive to the type of OI.
+
+	\begin{figure}[h]
+		\centering
+		\begin{subfigure}[t]{0.45\textwidth}
+			\centering
+			\includegraphics[width=\textwidth]
+			{Pictures/R2_CV_BVTV}
+			\caption{CV in relation to BV/TV.}
+		\end{subfigure}
+		\hfill
+		\begin{subfigure}[t]{0.45\textwidth}
+			\centering
+			\includegraphics[width=\textwidth]
+			{Pictures/R2_LinearRegression}
+			\caption{Linear regression of the complete OI data set.}
+		\end{subfigure}
+		\hspace{-5mm}\caption{Relative positions of the different OI types for different analyses.}
+	\end{figure}
+
+	%\clearpage
 	% Loading bibliography database

### 2021-11-18 09_Article/01_Revision2_Reviewers/Response to Reviewers_00_bis.tex
Finish Response to Reviewers and manuscript revision

--- 2021-11-18 09_Article/01_Revision2_Reviewers/Response to Reviewers_00.tex
@@
 The text citing equation 7 was modified to describe the term contained and the signification of this NE, see below.
+Moreover, equations 1, 6, and 7 (for the fabric, the $R^2_{adj}$ and the norm error NE, respectively) were moved into Appendix A as well.
 We hope that these firsts steps will help to decrease the technical difficulty and increase the clarity of the paper from a clinical point of view.
@@
 \begin{quote}
-	\begin{equation}
-	\text{NE} = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
-	\label{Eq206}
-	\end{equation}
-
-	\Add{Where $\mathbb{S}_o$ and $\mathbb{S}_p$ are the observed and predicted stiffness tensors, respectively. This relative norm error allows to quantify the accuracy of the prediction.}
+	Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE)\Del{, see Equation 6 and 7}.
+\Add{This relative norm error allows to quantify the accuracy of the prediction.
+Details of the computation of these numbers are available in Appendix A.}
 \end{quote}

### 2021-11-18 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Finish Response to Reviewers and manuscript revision

@@
 The fabric tensor $\mathbf{M}$ was computed using mean intercept length (MIL) method \cite{Moreno2014}.
-It is a positive-definite second-order tensor computed as shown below.
-
-	\begin{equation}
-	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i} = \sum_{i=1}^{3}{m_i (\mathbf{m}_i \otimes \mathbf{m}_i})
-	\label{Eq201}
-	\end{equation}
-
-	where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ \cite{Cowin1985,Harrigan1985}.
-The fabric tensor is independent of BV/TV and normalised with $tr(\mathbf{M}) = 3$.
-The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.
+\Del{It is a positive-definite second-order tensor computed as shown below.}\\
+
+	\Del{where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ }\cite{Cowin1985,Harrigan1985}\Del{.
+The fabric tensor is independent of BV/TV and normalised with $tr(\mathbf{M}) = 3$.} The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.
 Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualisation of its fabric tensor.\\
@@
 Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient squared ($R^2$) and the standard error of the estimate (SE).
-Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE), see Equation \ref{Eq205} and \ref{Eq206}\Add{, respectively}.
-
-	\begin{equation}
-	R^2_{adj} = 1 - \frac{\mathrm{RSS}}{\mathrm{TSS}} \frac{(12n-1)}{(12n - p - 1)}
-	\label{Eq205}
-	\end{equation}
-
-	Where RSS is the residual sum of squares and TSS is the total sum of squares i.e. sum of the square of the observations y. $n$ is the number of ROIs and $p$ the number of parameters.
-
-	\begin{equation}
-	\text{NE} = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
-	\label{Eq206}
-	\end{equation}
-
-	\Add{Where $\mathbb{S}_o$ and $\mathbb{S}_p$ are the observed and predicted stiffness tensors, respectively. This relative norm error allows to quantify the accuracy of the prediction.}
+Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE)\Del{, see Equation \ref{Eq205} and \ref{Eq206}}.
+\Add{This relative norm error allows to quantify the accuracy of the prediction.
+Details of the computation of these numbers are available in Appendix \ref{A0}.}
 
@@
 	The Zysset-Curnier model \cite{Zysset1995} computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV or $\rho$, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$, as shown in Equation \ref{Eq202}.
+The fabric tensor $\mathbf{M}$ is a positive-definite second-order tensor computed as shown below.
+
+	\begin{equation}
+	\mathbf{M} = \sum_{i=1}^{3}{m_i \mathbf{M}_i} = \sum_{i=1}^{3}{m_i (\mathbf{m}_i \otimes \mathbf{m}_i})
+	\label{Eq201}
+	\end{equation}
+
+	Where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ \cite{Cowin1985,Harrigan1985}.
+The fabric tensor is independent of BV/TV and normalised with $tr(\mathbf{M}) = 3$.
+Then, it is used to compute the stiffness tensor.
 
@@
 	\end{equation}
+
+	Finally, after performing the linear regression, quality of the results is evaluated using the relative error of the norm of fourth-order tensors (NE) and the adjusted Pearson's correlation coefficient ($R^2_{adj}$).
+These coefficients are computed as follow:
+
+	\begin{equation}
+	\text{NE} = \sqrt{\frac{(\mathbb{S}_o - \mathbb{S}_p) :: (\mathbb{S}_o - \mathbb{S}_p)}{\mathbb{S}_o :: \mathbb{S}_o}}
+	\label{Eq206}
+	\end{equation}
+
+	Where $\mathbb{S}_o$ and $\mathbb{S}_p$ are the observed and predicted stiffness tensors, respectively.
+
+	\begin{equation}
+	R^2_{adj} = 1 - \frac{\mathrm{RSS}}{\mathrm{TSS}} \frac{(12n-1)}{(12n - p - 1)}
+	\label{Eq205}
+	\end{equation}
+
+	Where RSS is the residual sum of squares and TSS is the total sum of squares i.e. sum of the square of the observations y. $n$ is the number of ROIs and $p$ the number of parameters.
 

### 2021-11-21 09_Article/01_Revision2_Reviewers/Response to Reviewers_00.tex
Implement Philippe suggestions into manuscript

@@
 \vspace{1em}Once again, the authors would like to thank the reviewers for helping us to improve our article.
-We do believe that their and our time spent working on this document will definitively improve its message to the scientific community.
+We do believe that their time spent working on this document will definitively improve its message to the scientific community.
 As for the first revision, answers to comments are provided point by point in this document.
@@
 
-\AR Effectively the numbers are different but both are correct.
+\AR The numbers are indeed different but both are correct.
 On one hand, in Methods 2.1 the mean and plus/minus age of both groups refers to the entire cohort.
@@
 
-\AR This was changed.
+\AR This was corrected.
 
@@
 
-Page 9, second column, Discussion, Lines 50-52
+Page 9, second column, Discussion, lines 50-52
 \begin{quote}
@@
 Regarding equations 2 and 3, even if they are outlined in cited references, we think that they are necessary for technical completeness.
-On the other hand, as they are more related to technical aspect we moved them in Appendix, as suggested.
+On the other hand, as they are more related to technical aspect we moved equation 2 and 3 but also equations 1,6 and 7 in the Appendix, as proposed.
 The text citing equation 7 was modified to describe the term contained and the signification of this NE, see below.
-We hope that these firsts steps will help to decrease the technical difficulty and increase the clarity of the paper from a clinical point of view.
+We hope that these steps will help to decrease the technical difficulty and increase the clarity of the paper from a clinical point of view.
 
@@
 Both the Abstract and the Discussion should comment on the implications of the fabric-elastic relationship not being defined for highly heterogeneous bone.
-Firstly, given there is only one OI type III where ROIs could be extracted, does this mean hFE should be restricted to OI type I?
+
+\AR We believe we can resolve these concerns.
+The filtering is made due to the sampling of cubic ROIs for computational homogenisation and does not filter bone microstructure from another.
+In our study, high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI and correlates with low, sometimes extremely low BV/TV and stiffness.
+That is why many more OI than healthy ROIs are filtered out.
+The mentioned limitation resides in the sampling of homogenous cubes and the application of computational homogenisation, not in the fabric-elasticity relationship which application to heterogeneous ROIs is well-defined, smooth in BV/TV, smooth in fabric and independent of heterogeneity.
+Finally, the use of the fabric-elasticity relationship was validated experimentally also in osteoporotic bones where medullary cavities are also important.
+
+\newpage
+\RC Firstly, given there is only one OI type III where ROIs could be extracted, does this mean hFE should be restricted to OI type I?
 What evidence is there that it can be employed for OI type III or IV?
 This needs to be clearly justified.
-Second, in the future, when significant heterogeneity is encountered in a bone, do you still recommend to use hFE?
+
+\AR Although no further statistical analysis is meaningful due to the small numbers, ROIs from all OI types fit nicely on a unique fabric-elasticity relationship.
+The lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed.
+No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.
+
+\RC Second, in the future, when significant heterogeneity is encountered in a bone, do you still recommend to use hFE?
 In other words, please comment whether the findings of skewness for high heterogeneity affects if/when hFE is appropriate, and if there are alternatives to hFE in that case.
 
-\AR Higly heterogeneous bone is rare in healthy conditions leading to the observed good correlations between simulations and experiments in the work of Varga et al.
+\AR The fabric-elasticity relationships between healthy and OI ROIs were found to be equivalent from a few MPa up to several GPa.
+Any ROI stiffness below a few MPa is meaningless in a mechanical analysis because of being several orders of magnitude smaller than the other ROIs and the cortex.
+Accordingly, hFE can be applied to bone with large voids or cavities as the BV/TV at the edges are smeared over the neighboring elements.
+
+%Highly heterogeneous bone is rare in healthy conditions leading to the observed good correlations between simulations and experiments in the work of Varga et al.
 2011, Hosseini et al.
@@
 2019.
-However, high heterogeneity arise from low bone volume fraction as outlined in the plot relating CV and BV/TV.
 This fact is determinant.
-As low BV/TV will lead to low stiffness, the error performed on such ROI is negligible as it not have a significant contribution to the stiffness of the full tibial section.
-For example, the observed stiffness component $\lambda_{ii}$ of one highly heterogeneous ROI (worst case) is around 1 MPa and the predicted value is around 100 MPa.
-This error is relatively small in comparison with the homogeneous ROI presenting a value around 10 GPa.
-Thus, only homogeneous ROIs (i.e.
-ROIs with higher BV/TV) contribute significantly to the bone stiffness.
-Moreover, despite the contribution of trabecular bone, the cortical bone will contribute even more significantly to tibial stiffness.
-Cortical bone having a higher BV/TV than trabecular bone, the relative error is low.
-Another point is that ROIs here have 5.3mm side length, reducing this size will improve homogeneity of ROIs and thus reduce error of the stiffness estimation.
+As low BV/TV will lead to low stiffness, the influence of such ROIs is negligible as it does not have a significant contribution to the stiffness of the full tibial section.
+%For example, the observed stiffness component $\lambda_{ii}$ of one highly heterogeneous ROI (worst case) is around 1 MPa and the predicted value is around 100 MPa.
+This apparently large error remains relatively small in comparison with homogeneous ROIs presenting a value approaching 10 GPa.
+%We believe that the heterogeneous ROIs are meaningless in the hFE analysis of a whole tibial section. In other words, a potential difference between healthy and OI for these
+
+%Thus, only homogeneous ROIs (i.e.
+ROIs with higher BV/TV and connected to the surrounding bone) contribute significantly to the overall bone stiffness.
+%Moreover, despite the contribution of trabecular bone, the cortical bone will contribute even more significantly to tibial stiffness.
+Cortical bone having a higher BV/TV than trabecular bone, the relative error becomes even lower.
+%Another point is that ROIs here have 5.3mm side length.
+Since the element size is inferior to the ROI size, the heterogeneity is better resolved and the uncertainty affects a lower tissue volume.
+Finally, it is important to realise that the scatter associated with heterogeneous ROIs (as shown in Appendix) reflects a limitation of the homogenisation methodology to identify the fabric-elasticity relationship and is not a true error of the fabric-elasticity paradigm.
+The extrapolation of the fabric-elasticity relationship to heterogeneous ROIs that is made in hFE may be quite reasonable, but the actual uncertainty cannot be quantified with the available tools.
+
 To improve the clarity of the message regarding validity of hFE among the OI types or heterogeneous bone, the manuscript was modified as follow:
@@
 	The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
-\Add{Nevertheless, the error performed on these highly heterogeneous ROIs is negligible in comparison with homogeneous ROIs which are order of magnitude stiffer.}
+\Add{However, this limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs may be reasonable but can simply not be evaluated with the employed homogenisation methodology.
+Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have little influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.}
 \end{quote}
 
-Discussion, Page 10, Second Column, Line 30
+Discussion, page 9, first column, line 56-59
 \begin{quote}
-	\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e.
-fabric-elasticity relationships) are independent of the OI type.
+	The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
+\Add{This comes from the fact that high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI.}
+\end{quote}
+
+Discussion, page 9, second column, line 55-60
+\begin{quote}
+	Some ROIs with low BV/TV do not have every side of the cube connected by bone, leading to extremely low terms in the stiffness tensor (see Appendix C).
+Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure 5b).
+\Add{This represents a limitation of the homogenization methodology and t}\Del{T}herefore, a filtering \Add{according to the heterogeneity level} is indispensable to assess and compare fabric-elasticity relationships, as done by Panyasantisuk et al.
+[28].
+\end{quote}
+
+\newpage
+Discussion, page 10, second column, line 30
+\begin{quote}
+	\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the fabric-elasticity relationships are independent of the OI type.
 Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix D.}
@@
 
-Discussion, Page 10, second column, last paragraph, last sentence
+Discussion, page 10, second column, line 47-50
+\begin{quote}
+	These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.
+\Add{Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs where homogenization procedure is valid.}
+\end{quote}
+
+Discussion, page 10, second column, last paragraph, last sentence
 \begin{quote}
 	Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}.
-Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity), such as in OI tpye III.}
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness\Del{.}\Add{ order of magnitude lower than homogeneous ones and the cortex (< 1MPa compared to 10 GPa).
+Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone [35,17] where medullary cavities are important and therefore lead to locally high heterogeneity.}
+\end{quote}
+
+Discussion, page 11, second column, line 40-24
+\begin{quote}
+	Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.
+\Add{Nevertheless, the lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed.
+No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.}
 \end{quote}

### 2021-11-21 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript.tex
Implement Philippe suggestions into manuscript

--- 2021-11-18 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
@@
 The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
-\Add{Nevertheless, the error performed on these highly heterogeneous ROIs is negligible in comparison with homogeneous ROIs which are order of magnitude stiffer.}
+However, this limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs may be reasonable but can simply not be evaluated with the employed homogenisation methodology.
+Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have little influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.
 
 		The filtering of highly heterogeneous ROIs removed these low stiffness ROIs and led to similar correlation coefficients for both OI and healthy groups.
-Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.\Add{ Moreover, the filtering step did not exclude a particular OI type.} Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.
-
-		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and \Del{less dense} \Add{have a lower BV/TV} than healthy controls.
+Finally, the BV/TV and DA matched data revealed no significant differences in fabric-elasticity parameters between OI and healthy individuals.
+Moreover, the filtering step did not exclude a particular OI type.
+Compared to previous studies, the stiffness constants from the 61 \si{\micro}m resolution HR-pQCT ROIs were lower than for the 36 \si{\micro}m resolution \si{\micro}CT ROIs.
+
+		In conclusion, OI trabecular bone of the distal tibia was shown to be significantly more heterogeneous and have a lower BV/TV than healthy controls.
 Despite the reduced linear regression parameters found for HR-pQCT images, the fabric-elasticity relationships between OI and healthy individuals are similar when the trabecular bone ROIs are sufficiently homogeneous to perform the computational stiffness analysis.
@@
 Participants in the healthy group were scanned using an in-house protocol as described in \cite{Schenk2020}.
-Name\-ly, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
+Na\-me\-ly, the reference line was positioned at the proximal margin of the dense structure formed by the tibia plafond and three stacks were scanned proximal to this reference line, see Figure \ref{01_Healthy}.
 On the other hand, for member of the OI group, the reference line is placed at the subchondral endplate of the ankle joint.
@@
 Then, for both the healthy group and the OI group, the stack was divided into two halves with a transverse plane and the ROIs were selected to have the centers of three ROIs in the proximal half and three in the distal half, see Figure \ref{01_ROISelection}.
-For one individual diagnosed with OI type III it was not possible to extract any ROI as there was no\Add{t} enough trabecular bone.
+For one individual diagnosed with OI type III it was not possible to extract any ROI as there was not enough trabecular bone.
 This led to 720 healthy ROIs from 120 individuals and 294 OI ROIs from 49 individuals.\\
@@
 The fabric tensor $\mathbf{M}$ was computed using mean intercept length (MIL) method \cite{Moreno2014}.
-\Del{It is a positive-definite second-order tensor computed as shown below.}\\
-
-	\Del{where $m_i$ are the eigenvalues of $\mathbf{M}$ and $\mathbf{M}_i$ are the dyadic product of the corresponding eigenvectors $\mathbf{m}_i$ }\cite{Cowin1985,Harrigan1985}\Del{.
-The fabric tensor is independent of BV/TV and normalised with $tr(\mathbf{M}) = 3$.} The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.
+The fabric eigenvalues allow to compute the degree of anisotropy (DA) of the ROI by dividing the highest eigenvalue by the lowest one.
 Figure \ref{01_FabricExample} shows an example of a typical ROI with the visualisation of its fabric tensor.\\
@@
 The homogenization process allowed to calculate the components of the stiffness tensor and to calibrate the parameters of the Zysset-Curnier fabric-elasticity model \cite{Zysset1995}.
-This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV \Del{or $\rho$}, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$\Del{, as shown in Equation \ref{Eq202}}.
-\Add{This stiffness tensor is then transformed into the fabric coordinate system} and projected onto orthotropy, leading to 12 components.
-\Add{The technical details of the computations performed are presented in Appendix \ref{A0}.}\\
+This model computes the fourth order stiffness tensor $\mathbb{S}$ using the BV/TV, fabric tensor $\mathbf{M}$, three elasticity parameters $\lambda_0$, $\lambda_0$', and $\mu_0$, and two exponents, $k$ and $l$.
+This stiffness tensor is then transformed into the fabric coordinate system and projected onto orthotropy, leading to 12 components.
+The technical details of the computations performed are presented in Appendix \ref{A0}.\\
 
@@
 As this assumption was violated by the fact that six ROIs were analysed per individual, a linear mixed-effect model was preferred.
-This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-independence of ROIs from the same individual.
+This last model, shown in Equation \ref{Eq204} in Laird-Ware form \cite{Laird1982}, considered the non-in\-de\-pen\-dence of ROIs from the same individual.
 A more detailed form of this model is presented in Appendix \ref{A1}.
@@
 Linear regression quality for the TMD analysis was assessed using the Pearson correlation coefficient squared ($R^2$) and the standard error of the estimate (SE).
-Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE)\Del{, see Equation \ref{Eq205} and \ref{Eq206}}.
-\Add{This relative norm error allows to quantify the accuracy of the prediction.
-Details of the computation of these numbers are available in Appendix \ref{A0}.}
+Linear regression on Zysset-Curnier model was assessed using the adjusted Pearson correlation coefficient squared ($R^2_{adj}$) and relative error between the orthotropic observed and the predicted tensor using norm of fourth-order tensors (NE).
+This relative norm error allows to quantify the accuracy of the prediction.
+Details of the computation of these numbers are available in Appendix \ref{A0}.
 
@@
 The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
+This comes from the fact that high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI.
 Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
@@
 These differences come from ROIs presenting a low stiffness.
-The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to \Add{1 MPa (}10\textsuperscript{0}\Add{)} \Del{MPa} and lower, the linear regression tends to overestimate the stiffness.
+The linear regression plot (Figure \ref{02_OI}) shows that when the stiffness term decreases to 1 MPa (10\textsuperscript{0}) and lower, the linear regression tends to overestimate the stiffness.
 This is because ROI stiffness is highly dependent on BV/TV values.
@@
 Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
-Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
+This represents a limitation of the homogenization methodology and therefore, a filtering according to the heterogeneity level is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
 An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.
@@
 However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further.
-\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e.
+Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e.
 fabric-elasticity relationships) are independent of the OI type.
-Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix} \ref{A3}\Add{.} \\
+Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix \ref{A3}.
+\\
 
@@
 $R^2_{adj}$, NE, and the range of stiffness values are almost at the level of the healthy group.
-These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\
+These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.
+Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs where homogenization procedure is valid.\\
 
@@
 With this statistical non-significance of the groups and their low relative differences in the computed stiffness constants, it can be stated that: if trabecular bone is homogeneous enough, there is no reason to assume differences in fabric-elasticity relationships between healthy and OI trabecular bone.
-Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}.
-Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity).}\\
+Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones, regardless the OI type.
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness order of magnitude lower than homogeneous ones and the cortex (< 1MPa compared to 10 GPa).
+Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone \cite{Varga2011,Hosseini2017} where medullary cavities are important and therefore lead to locally high heterogeneity.\\
+
+	%\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity).}\\
 
@@
 Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.
+Nevertheless, the lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed.
+No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.
 ROI homogeneity, more specifically a balanced bone connectivity of the 6 faces of the ROI, is a necessary condition underlying the computational homogenisation of mechanical properties and the results of heterogenous samples must be disregarded.
-However, when calibrated with homogeneous ROIs, fabric-elas\-ticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-homogeneity and low-stiffness ROIs.
+However, when calibrated with homogeneous ROIs, fabric-elas\-ticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-ho\-mo\-ge\-nei\-ty and low-stiffness ROIs.
 In fact, these low-property ROIs were not shown to compromise the performance of hFE models of the distal radius \cite{Varga2011, Hosseini2017, AriasMoreno2019}.\\

### 2021-11-22 09_Article/01_Revision2_Reviewers/Response to Reviewers_00.tex
Implement Philippe suggestions into manuscript

@@
 \section{Density}
-\RC The first highlight states that OI trabecular bone is more mineralized compared to healthy bone, but the first sentence in the last paragraph of the Abstract says that OI bone is significantly less dense than healthy controls.
+\RC The first highlight states that OI trabecular bone is more mineralised compared to healthy bone, but the first sentence in the last paragraph of the Abstract says that OI bone is significantly less dense than healthy controls.
 This is going to be confusing for many.
@@
 Regarding equations 2 and 3, even if they are outlined in cited references, we think that they are necessary for technical completeness.
-On the other hand, as they are more related to technical aspect we moved equation 2 and 3 but also equations 1,6 and 7 in the Appendix, as proposed.
+On the other hand, as they are more related to technical aspect we moved equation 2 and 3 as proposed but also equations 1,6 and 7 for consistency in the Appendix.
 The text citing equation 7 was modified to describe the term contained and the signification of this NE, see below.
@@
 \AR We believe we can resolve these concerns.
-The filtering is made due to the sampling of cubic ROIs for computational homogenisation and does not filter bone microstructure from another.
+The filtering is made due to the sampling of relatively large cubic ROIs for computational homogenisation and does not filter bone microstructure from another.
 In our study, high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI and correlates with low, sometimes extremely low BV/TV and stiffness.
@@
 
-\AR The fabric-elasticity relationships between healthy and OI ROIs were found to be equivalent from a few MPa up to several GPa.
+\AR The fabric-elasticity relationships between healthy and OI ROIs were found to be equivalent from 8 \% to 45 \% of BV/TV and from a few MPa up to several GPa of stiffness.
 Any ROI stiffness below a few MPa is meaningless in a mechanical analysis because of being several orders of magnitude smaller than the other ROIs and the cortex.
-Accordingly, hFE can be applied to bone with large voids or cavities as the BV/TV at the edges are smeared over the neighboring elements.
+Accordingly, hFE can also be applied to bones with large voids or cavities as their geometry is resolved to the scale of the element size and the mechanical properties of their edges vanish smoothly with BV/TV.
+Validity of hFE analyses in such circumstances could also be established for vertebral bodies with metastatic defects Ref.
+
+Reference:\\
+Stadelmann, M.A., Schenk, D.E., Maquer, G., Lenherr, C., Buck,
+F.M., Bosshardt, D.D., Hoppe, S., Theumann, N., Alkalay, R.N.,
+Zysset, P.K., 2020. Conventional finite element models estimate
+the strength of metastatic human vertebrae despite alterations of the
+bone’s tissue and structure. Bone 141. doi:10.1016/j.bone.2020.115598.
 
@@
 
-To improve the clarity of the message regarding validity of hFE among the OI types or heterogeneous bone, the manuscript was modified as follow:
+To improve the clarity of the message regarding validity of hFE among the OI types or trabecular bone in the vicinity of holes, the manuscript was modified as follow:
 
@@
 	The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
-\Add{However, this limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs may be reasonable but can simply not be evaluated with the employed homogenisation methodology.
-Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have little influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.}
+\Add{This limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs is probably reasonable but can simply not be evaluated with the employed homogenisation methodology.
+Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have negligible influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.}
 \end{quote}
@@
 	The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
-\Add{This comes from the fact that high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI.}
+\Add{In fact, high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI compared to healthy controls.}
 \end{quote}
 
+\newpage
 Discussion, page 9, second column, line 55-60
@@
 Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure 5b).
-\Add{This represents a limitation of the homogenization methodology and t}\Del{T}herefore, a filtering \Add{according to the heterogeneity level} is indispensable to assess and compare fabric-elasticity relationships, as done by Panyasantisuk et al.
+\Del{Therefore}\Add{This represents a limitation of the homogenisation methodology and} a filtering \Add{according to the heterogeneity level becomes}\Del{is} indispensable to assess and compare fabric-elasticity relationships, as done by Panyasantisuk et al.
 [28].
@@
 
-\newpage
 Discussion, page 10, second column, line 30
@@
 	These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.
-\Add{Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs where homogenization procedure is valid.}
+\Add{Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs for which the selected homogenisation procedure is valid.}
 \end{quote}
@@
 	Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}.
-Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness\Del{.}\Add{ order of magnitude lower than homogeneous ones and the cortex (< 1MPa compared to 10 GPa).
-Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone [35,17] where medullary cavities are important and therefore lead to locally high heterogeneity.}
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness\Del{.}\Add{ orders of magnitude lower than the homogeneous ones or the cortex (< 1 MPa compared to 10 GPa).
+Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone [36,17] or vertebral bodies with metastatic defects [33] where high heterogeneity is present.}
 \end{quote}
@@
 	Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.
-\Add{Nevertheless, the lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed.
+\Add{Nevertheless, the lack of homogeneous type III ROIs is essentially due to the lack of trabecular bone in which case no relationship is needed.
 No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.}

### 2021-11-22 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript.tex
Implement Philippe suggestions into manuscript

@@
 The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
-However, this limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs may be reasonable but can simply not be evaluated with the employed homogenisation methodology.
-Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have little influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.
+This limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs is probably reasonable but can simply not be evaluated with the employed homogenisation methodology.
+Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have negligible influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.
 
@@
 The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
-This comes from the fact that high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI.
+In fact, high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI compared to healthy controls.
 Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
@@
 Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
-This represents a limitation of the homogenization methodology and therefore, a filtering according to the heterogeneity level is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
+This represents a limitation of the homogenisation methodology and a filtering according to the heterogeneity level becomes indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
 An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.
@@
 However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further.
-Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e.
-fabric-elasticity relationships) are independent of the OI type.
+Nevertheless, it should be noted that this relation between BV/TV and CV as well as the fabric-elasticity relationships are independent of the OI type.
 Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix \ref{A3}.
@@
 These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.
-Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs where homogenization procedure is valid.\\
+Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs for which the selected homogenisation procedure is valid.\\
 
@@
 Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones, regardless the OI type.
-Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness order of magnitude lower than homogeneous ones and the cortex (< 1MPa compared to 10 GPa).
-Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone \cite{Varga2011,Hosseini2017} where medullary cavities are important and therefore lead to locally high heterogeneity.\\
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness orders of magnitude lower than the homogeneous ones or the cortex (< 1 MPa compared to 10 GPa).
+Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone \cite{Varga2011,Hosseini2017} or vertebral bodies with metastatic defects \cite{Stadelmann2020} where high heterogeneity is present.\\
 
@@
 Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.
-Nevertheless, the lack of homogeneous type III ROIs is simply due to the lack of bone in which case no relationship is needed.
+Nevertheless, the lack of homogeneous type III ROIs is essentially due to the lack of trabecular bone in which case no relationship is needed.
 No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.
@@
 	We did not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when only valid, homogenous ROIs were included and matched for both BV/TV and DA.
-Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients, potentially with a small correction for TMD.
+Therefore, with the gained knowledge of this study, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients, potentially with a small correction for TMD.
 

### 2021-11-22 09_Article/01_Revision2_Reviewers/Simon_2021_Bone_Manuscript_Revised.tex
Implement Philippe suggestions into manuscript

@@
 The stiffnesses of highly heterogeneous ROIs were randomly lower with respect to the fabric-elasticity relationships, which reflects the limit of validity of the computational homogenisation methodology.
-\Add{Nevertheless, the error performed on these highly heterogeneous ROIs is negligible in comparison with homogeneous ROIs which are order of magnitude stiffer.}
+\Add{This limitation does not challenge the fabric-elasticity relationship, which extrapolation to heterogeneous ROIs is probably reasonable but can simply not be evaluated with the employed homogenisation methodology.
+Moreover, due to their low BV/TV, the potential (unknown) errors on these heterogeneous ROIs would have negligible influence on whole bone stiffness in comparison to homogeneous ROIs which are orders of magnitude stiffer.}
 
@@
 the group has no significant influence on the TMD versus BV/TV slope, the model was modified to add the interaction regressor (BV/TV x Group).
-The detailed linear systems for each model discussed here are available in Appendix \ref{A1} and a summary of the data sets used for the different methods is shown in Table \ref{Table1}.\\
+The detailed linear systems for each model discussed here are available in Appendix \ref{A1} and a summary of the data sets used for the different methods is shown in Table \ref{Table1}.
+\\
 
@@
 The CV shows that heterogeneity of OI trabecular bone is higher compared to healthy control and is consequently the more discriminant parameter.
-Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
+\Add{In fact, high heterogeneity is found at the edge of void spaces or medullary cavities that are much larger in OI compared to healthy controls.} Finally, the significant differences observed in BV/TV and DA, even with matched age and sex, justifies the choice of a variable matching for fabric-elasticity relationships analysis, because the linear regression must be performed on similar ranges to obtain comparable values.
 \\
@@
 Trying to homogenise such ROIs can lead to errors of multiple order of magnitude, as observed on the plot (Figure \ref{02_OI}).
-Therefore, a filtering is indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
+\Del{Therefore}\Add{This represents a limitation of the homogenisation methodology and} a filtering \Add{according to the heterogeneity level becomes}\Del{is} indispensable to assess and compare fabric-elasticity relationships, as done by \citeauthor{Panyasantisuk2015} \cite{Panyasantisuk2015}.
 An alternative to CV filtering for assessing the ROI heterogeneity could be to compute the proportion of the area filled by bone on each of the six faces of the ROI.
@@
 However, the value of this threshold results from an optimisation process in the study of \citeauthor{Panyasantisuk2015}\cite{Panyasantisuk2015} and could be investigated further.
-\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the homogenization process (i.e.
-fabric-elasticity relationships) are independent of the OI type.
-Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix} \ref{A3}\Add{.} \\
+\Add{Nevertheless, it should be noted that this relation between BV/TV and CV as well as the fabric-elasticity relationships are independent of the OI type.
+Distribution of OI types for CV in relation to BV/TV as well as the linear regression of the original data set are available in Appendix \ref{A3}.} \\
 
@@
 $R^2_{adj}$, NE, and the range of stiffness values are almost at the level of the healthy group.
-These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.\\
+These results give confidence to the filtering procedure and represent a first step in accepting the hypothesis of healthy and OI trabecular bone having the same fabric-elasticity relationships.
+\Add{Indeed, this confirms that filtering according to heterogeneity level allows to compare fabric-elasticity relationships of homogeneous ROIs for which the selected homogenisation procedure is valid.}\\
 
@@
 Thus, the same hFE methodology can be applied for bone stiffness assessment of OI diagnosed individuals as for healthy ones\Add{, regardless the OI type}.
-Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness.\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity).}\\
+Although in hFE analysis it is not possible to exclude part of the mesh because of high heterogeneity, the influence of such ROIs seems negligible on full bone models as the fabric-elasticity relationships can extrapolate their stiffness smoothly and as this concerns only ROIs with low BV/TV and stiffness\Del{.}\Add{ orders of magnitude lower than the homogeneous ones or the cortex (< 1 MPa compared to 10 GPa).
+Moreover, studies validated the use of fabric-elasticity relationships in osteoporotic bone \cite{Varga2011,Hosseini2017} or vertebral bodies with metastatic defects \cite{Stadelmann2020} where high heterogeneity is present.}\\
+	%\Add{ Moreover, ROI side length can be reduced to improve accuracy if necessary for trabecular bone presenting extremely low BV/TV (and thus high heterogeneity).}\\
 
@@
 Moreover, having only one patient with OI type III where we could extract ROIs does not allow to perform statistics on OI types.
-ROI homogeneity, more specifically a balanced bone connectivity of the 6 faces of the ROI, is a necessary condition underlying the computational homogenisation of mechanical properties and the results of heterogenous samples must be disregarded.
+\Add{Nevertheless, the lack of homogeneous type III ROIs is essentially due to the lack of trabecular bone in which case no relationship is needed.
+No outlier and no evidence of a different relationship was seen in our results for any OI type including the sample with OI type III.} ROI homogeneity, more specifically a balanced bone connectivity of the 6 faces of the ROI, is a necessary condition underlying the computational homogenisation of mechanical properties and the results of heterogenous samples must be disregarded.
 However, when calibrated with homogeneous ROIs, fabric-elas\-ticity relationships offer a smooth and reasonable extrapolation of stiffness properties for the concerned low-homogeneity and low-stiffness ROIs.
@@
 	However, this moderate increase in matrix properties is large\-ly overcompensated by the 45\% reduction in trabecular bone strength associated with BV/TV. % Using k=2: (74**2-100**2)/100**2
-	We did not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when only valid, homogenous ROIs were included and matched for both BV/TV and DA.
-Therefore, HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients, potentially with a small correction for TMD.
+	We did not find statistically significant differences in fabric-elasticity relationships between healthy and OI trabecular bone, when only valid ROIs were included and matched for both BV/TV and DA.
+Therefore, \Add{with the gained knowledge of this study,} HR-pQCT-based hFE analyses used for bone stiffness estimation of healthy individuals can be used for OI diagnosed patients, potentially with a small correction for TMD.
 
